---
title: "Annual Report Data Analysis - Question 3.14-PSC"
author: "Cristin Young"
date: "2025-05-13"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

#Read in data
```{r}
library(tidyverse)
library(janitor)
library(gt)

all_data <- read_csv("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
```

#3.14 To what extent did children/youth report reductions in overall distress after receiving the intervention?
PSC = 224
YOQ Caregiver = 292
YOQ Self = 345
```{r}
table(all_data$pre_assessment_name)
table(all_data$trimmed_pre_assessment_name)

youth <- all_data %>% 
  filter(client_type == "Youth (16-25)",
         trimmed_pre_assessment_name == "Pediatric")

child <- all_data %>% 
  filter(client_type == "Child (0-15)",
         trimmed_pre_assessment_name == "Pediatric")

child_youth <- all_data %>% 
  filter(client_type != "Parent/Caregiver",
         trimmed_pre_assessment_name == "Pediatric")
```

##Checking missing data

How much data is missing? 

- 91% (n=204/224) of rows have a pre but no post score

- 0 rows have a post score but no pre score

- 9% (n=21/224) of rows have both a pre and post score

```{r}
pre_no_post <- child_youth %>% 
  filter(is.na(trimmed_post_assessment_name))

#checking there's no opposite to above (yes post, no pre score)
post_no_pre <- child_youth %>% 
  filter(is.na(trimmed_pre_assessment_name) & !is.na(trimmed_post_assessment_name))

pre_post <- child_youth %>% 
  filter(!is.na(trimmed_pre_assessment_name) & !is.na(trimmed_post_assessment_name))
```

##do we have any duplicates?

Looks like they're reporting both caregiver AND self-assessment. Keeping only self-assessment right now.
```{r}
#how many duplicates are there?
n_occur <- data.frame(table(pre_post$client_id))
#how many times do they occur?
n_occur[n_occur$Freq > 1,]
```

#PSC cutoffs
https://tnaap.org/wp-content/uploads/2022/06/PSC-35-Scoring-Instructions.pdf 
- Lower score is better
- For children ages 4 and 5, the PSC score of 24 or higher suggests the presence of significant behavioral
or emotional problems. 
- For children ages 6 through 16, the cutoff score is 28 or higher.

#Creating positive outcome and cutoff variables
```{r}
pre_post <- pre_post %>% 
  mutate(positive_outcome = case_when(
    pre_score > post_score ~ "Yes",
    TRUE ~ "No"),
    child_cutoff = case_when(
      post_score >= 24 & client_type == "Child (0-15)" ~ 1,
      post_score < 24 & client_type == "Child (0-15)" ~ 0),
    youth_cutoff = case_when(
      post_score >= 28 & client_type == "Youth (16-25)" ~ 1,
      post_score < 28 & client_type == "Youth (16-25)" ~ 0
    )
  ) 
```

#how many positive outcomes?
20 / 21 = 95%
```{r}
table(pre_post$positive_outcome)
pre_post %>% 
  select(pre_score, post_score, positive_outcome)
```

#how many above clinical cutoff at discharge?
Clinical cutoff:
- For children ages 4 and 5, the PSC score of 24 or higher suggests the presence of significant behavioral
or emotional problems. 
    - one child was above clinical cutoff (25) at discharge
- For children ages 6 through 16, the cutoff score is 28 or higher.
    - 2 youth were above clinical cutoff (scores of 34 and 36) at discharge
  
```{r}
#child
 pre_post %>% 
  filter(child_cutoff == 1)

#youth
 pre_post %>% 
  filter(youth_cutoff == 1)
```

##mean and SD of pre and post scores
```{r}
#mean and sd of this data
summary_pre_score <- pre_post %>%
  summarise(
    mean_score = mean(pre_score, na.rm = TRUE),
    median_score = median(pre_score, na.rm = TRUE),
    sd_score = sd(pre_score, na.rm = TRUE)
  )

summary_pre_score
summary_pre_score %>% 
  knitr::kable() %>% 
  clipr::write_clip()

summary_post_score <- pre_post %>%
  summarise(
    mean_score = mean(post_score, na.rm = TRUE),
    median_score = median(post_score, na.rm = TRUE),
    sd_score = sd(post_score, na.rm = TRUE)
  )

summary_post_score
summary_post_score %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```

##do a t test on the data
```{r}
t.test(pre_post$pre_score, pre_post$post_score, paired = TRUE)
```
##plot pre vs post scores
```{r}
pre_post_long <- pre_post %>%
  ungroup() %>% 
  select(client_id, pre_score, post_score, pre_time_point, post_time_point) %>% 
  pivot_longer(
    cols = -client_id,
    names_to = c("timepoint", ".value"),
    names_sep = "_"
  )

pre_post_long$time <- factor(pre_post_long$time, levels = c("Initial", "Follow up 1", "Follow up 2"))

pre_post_long %>% 
  ggplot() + 
  geom_line(aes(time, score, group = client_id, color = factor(client_id)), show.legend = FALSE) 
```

