---
title: "Annual Report Data Analysis - Question 1"
author: "Cristin Young"
date: "2025-05-13"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

#Read in data
```{r}
library(tidyverse)
library(janitor)
library(gt)

demographic_data_filtered <- read_csv("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv", show_col_types = FALSE)
all_data <- read_csv("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
```
#1. Did the EBP/CDEP Workstream increase access to services for targeted populations?

##1.01 Did improved access to services affect all subpopulations?

Cross-tab of study population by relevant demographic data

Total number of clients served and by demographics

###race
```{r}
race_summary <- demographic_data_filtered %>%
  select(race) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

race_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()

all_data_race_summary <- all_data %>%
  select(race) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

all_data_race_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```

###ethnicity
```{r}
ethnicity_summary <- demographic_data_filtered %>%
  select(ethnicity) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

ethnicity_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()

all_data_ethnicity_summary <- all_data %>%
  select(ethnicity) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

all_data_ethnicity_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```

###gender identity
```{r}
gender_summary <- demographic_data_filtered %>%
  select(gender_identity) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

gender_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()

all_data_gender_summary <- all_data %>%
  select(gender_identity) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

all_data_gender_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```

###sexual orientation
```{r}
sex_summary <- demographic_data_filtered %>%
  select(sexual_orientation) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

sex_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()

all_data_sex_summary <- all_data %>%
  select(sexual_orientation) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

all_data_sex_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```

###insurance status
```{r}
insurance_summary <- demographic_data_filtered %>%
  select(insurance_status) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

insurance_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()

all_data_insurance_summary <- all_data %>%
  select(insurance_status) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  count(variable, value) %>%
  group_by(variable) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(variable, desc(n))

all_data_insurance_summary %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```


##1.02 How does access differ compared to previous access to services before EBP/CDEP funding? For each subpopulation?

Cross-tab of demographics and grant types: Total number of clients served and by demographic, and by whether grant was start up or expansion
```{r}
glimpse(all_data)
```

