---
title: "Annual Report Data Analysis - Question 2.01"
author: "Cristin Young and Tristan Burgess"
date: '2025-05-17'
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

This notebook answers Question 2.01: Did successful transition rates out of EBP/CDEP services differ by location? and Question 2.02: Did successful transition rates out of EBP/CDEP services differ by subpopulations?

#Read in data
```{r}
library(tidyverse)
library(janitor)
library(gt)
library(clipr)


#all_data <- read_csv("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
all_data <- read_csv("/Users/tristan/CWS Dropbox/Tristan Burgess/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
```


# Question 2.01: Did successful transition rates out of EBP/CDEP services differ by location?
THREE inclusion criteria:
 1. must have initial time point data (i.e., be enrolled in an intervention), 
 2. must have a valid reason for discharge selected at time of discharge (i.e., excludes those with "no reason selected"), 
 3. client must be attached to a specific grantee location

##Make a table of participation and completion rates by location
```{r}
data201 <- all_data %>% 
  filter(!is.na(location) & 
           !is.na(pre_score) & 
           !is.na(post_score) & 
           !is.nat(location) &
           !is.na(reason_for_discharge))

data201$positive_outcomes <- factor(data201$positive_outcomes, levels = c("No", "Yes"))
```

## % reported Mutually agreed upon cessation of Tx/successful completion at Discharge by location
```{r}
perc_sucessful_table <- data201 %>% 
  group_by(location) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )
              
perc_sucessful_table %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)
```

## % of those who met Tx goals (% successful completion + improved score) by location"
mutually agreed cessation of tx = 892
clinically referred out = 11
```{r}
data201_txgoals <- data201 %>% 
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)",
         positive_outcomes == "Yes")
table(data201_txgoals$location)

#or
p_yes = function(x) round(x[2]/sum(x), 2) # 1 in this line was doing p_no!

table <- addmargins(xtabs(data=data201, ~location + positive_outcomes, subset = data201$reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)"), FUN = list(p_yes))
table %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)
```


#Question 2.02: Did successful transition rates out of EBP/CDEP services differ by subpopulations?
look at data201 by demographics

##Race
```{r}
detailed_race <- data201 %>% 
  group_by(race) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

detailed_race %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$positive_outcomes, data201$race)

model <- glm(positive_outcomes ~ race, data = data201, family = "binomial")
summary(model)
table(data201$insurance_status)
```

##Race summary
Roll detailed race variables, based on first answer, up into: 
- Asian
- Black or African American
- Don't Know
- Hispanic or Latino
- Middle Eastern or North African
- Native American/American Indian or Alaska Native
- Native Hawaiian or Pacific Islander
- Not Applicable
- Prefer Not to Answer
- White
```{r}
#make new column of only the first race recorded
data201 <- data201 %>% 
  separate(race, into = c("first_race", NA), sep = ",", extra = "drop", remove = FALSE)

summary_race <- data201 %>% 
  group_by(first_race) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

summary_race %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$positive_outcomes, data201$first_race)

model <- glm(positive_outcomes ~ first_race, data = data201, family = "binomial")
summary(model)
table(data201$first_race)
```


##Ethnicity
```{r}
detailed_ethnicity <- data201 %>% 
  group_by(ethnicity) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

detailed_ethnicity %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$positive_outcomes, data201$ethnicity)
```

##Gender Identity
```{r}
gender_identity <- data201 %>% 
  group_by(gender_identity) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

gender_identity %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$gender_identity, data201$positive_outcomes)

model <- glm(positive_outcomes ~ gender_identity, data = data201, family = "binomial")
summary(model)
```
##Sexual orientation
```{r}
sex_orientation <- data201 %>% 
  group_by(sexual_orientation) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

sex_orientation %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$positive_outcomes, data201$sexual_orientation)
```

##Insurance status
```{r}
insurance <- data201 %>% 
  group_by(insurance_status) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

insurance %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$positive_outcomes, data201$insurance_status)

model <- glm(positive_outcomes ~ insurance_status, data = data201, family = "binomial")
summary(model)
```