---
title: "Quarterly Report Results"
author: "Cristin Young"
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    df-print: paged
    toc: true
    toc-depth: 3
    toc-location: left
    theme: cosmo
    embed-resources: TRUE
execute:
  echo: false
  warning: false
  message: false
---

::: {.callout-note title="TO DOs"}
[ ] anything <11 needs to be suppressed. If only one item is suppressed, also suppress the total (so people can't back calculate).
:::

## Period

2024-01-01 - 2026-03-31

## Setup

```{r}
#| label: setup
library(readr)
library(tidyverse)
library(haven)
library(janitor)

knitr::opts_chunk$set(rownames.print = FALSE)

# load in all data
discharge_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv") %>% 
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2026-03-31")

all_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv") %>% 
    filter(intervention_start >= "2024-01-01",
           intervention_start <= "2026-03-31")

discharge_data_distinct <- discharge_data %>% 
  distinct(client_system_id, .keep_all = TRUE)

all_data_distinct <- all_data %>% 
  distinct(client_system_id, .keep_all = TRUE)

cat("Unique providers (discharge data):", discharge_data |> distinct(provider_id) |> nrow(), "\n")
cat("Discharge data (distinct by client ID) rows; this is the denominator for most tables below:", discharge_data |> distinct(client_system_id, .keep_all = TRUE) |> nrow(), "\n")
cat("Discharge data rows (total):", nrow(discharge_data), "\n")
```

## Outcomes Data by Round

### Participants Served
**Data used**: Discharge data, which includes 382 clients that had more than one complete intervention.

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered only for `intervention_start >= "2024-01-01" & <= "2026-03-31"`
```{r}
#| label: tbl-participants-served-by-round

discharge_data %>% 
  tabyl(funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Improved outcomes
**Data used**: Discharge data

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered only for `intervention_start >= "2024-01-01" & <= "2026-03-31"`
```{r}
#| label: tbl-count-outcomes

nrow(discharge_data %>% filter(!is.na(positive_outcomes)))
```

#### Improved outcomes table
**Assumptions:** Needed to have an intervention start and end date and reason for discharge; data used: discharge_data
```{r}
#| label: tbl-improved-outcomes

discharge_data %>%
  filter(!is.na(intervention_start),
         !is.na(intervention_end),
         !is.na(reason_for_discharge)) %>% 
  tabyl(positive_outcomes) %>%               
  adorn_pct_formatting(digits = 1)
```

#### Improved outcomes by funding round
**Assumptions:** Needed to have an intervention start and end date and reason for discharge; data used: discharge_data
```{r}
#| label: tbl-positive-outcomes-by-round

discharge_data %>% 
  filter(!is.na(positive_outcomes)) %>% 
  tabyl(positive_outcomes, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 
```

## Satisfaction Surveys
Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree." The second is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed. High satisfaction on question 1 is defined as a score of 4 or 5.

**Assumptions:** Distinct by client system ID and joined to satisfaction survey data; must have an intervention named; data used: satisfaction survey and all data (demo+discharge joined)
```{r}
#| label: setup-satisfaction
#| include: false

sat_all <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/combined_satisfaction_clean.csv",
  show_col_types = FALSE
)

# Unified Q1 score — coalesce across three survey-specific Q1 columns
sat_all <- sat_all |>
  mutate(
    q1_score = coalesce(
      x1_overall_i_am_satisfied_with_the_services_my_child_received_score,
      x1_overall_i_am_satisfied_with_the_services_i_received_score,
      x1_i_like_the_services_that_i_received_here_score
    )
  )

all_data_distinct <- all_data |>
  distinct(client_system_id, .keep_all = TRUE) |> 
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  )

sat_all_data <- sat_all |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  ) |>
  left_join(all_data_distinct,
            by = c("provider", "provider_id", "provider_system_id", "client_system_id")) |>
  filter(!is.na(intervention_name))

score_cols <- names(sat_all_data)[str_detect(names(sat_all_data), "_score$")]

# Overall Q1 high satisfaction rate
pct_4_5 <- sat_all_data |>
  summarise(
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1)
  ) |>
  pull(pct_4_5)

# Overall combined survey high satisfaction rate (>= 80% of questions at 4 or 5)
df2 <- sat_all_data |>
  select(all_of(score_cols)) |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(everything(), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree    = sum(c_across(everything()) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  ungroup() |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )

df2
```
### Question 1 Satisfaction

Overall, `{r} pct_4_5`% of survey respondents were highly satisfied based on question 1. The following table shows high satisfaction by round:

```{r}
#| label: tbl-q1-satisfaction-round

sat_all_data |> 
  filter(!is.na(funding_round)) |> 
  group_by(funding_round) |> 
  summarise(
    n_surveys = n(),
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1),
    .groups = "drop"
  )
```

### Combined Survey Satisfaction

Overall, `{r} df2$pct_meeting_80pct`% of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed. The following table shows high combined satisfaction by round, for clients where both a client ID and intervention name were available:

**Assumptions:** Must have an intervention name; data used: satisfaction survey and all data (demo+discharge joined)
```{r}
#| label: tbl-combined-satisfaction-round

sat_all_data |>
  filter(!is.na(funding_round)) |>
  group_by(funding_round) |>
  select(funding_round, all_of(score_cols)) |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  mutate(across(all_of(score_cols), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(all_of(score_cols)))),
    n_agree    = sum(c_across(all_of(score_cols)) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )
```
## Discharge

### Discharge Reason 
**Assumptions:** Needed to have a reason for discharge; data used: discharge_data
```{r}
#| label: tbl-reason-for-discharge

discharge_data |>
  filter(!is.na(reason_for_discharge),
         reason_for_discharge != "No Reason Selected") |>
  tabyl(reason_for_discharge) |>
  adorn_pct_formatting(digits = 1)
```

### Discharge Reason by Round
The results below show % of successful discharges by round, once NA's are removed from the data.

**Assumptions:** Needed to have a reason for discharge and a funding round; data used: discharge data
```{r}
#| label: tbl-discharged-by-reason

discharge_data |>
 filter(!is.na(reason_for_discharge)) %>% 
  tabyl(reason_for_discharge, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() 
```

### Discharge Reason by Funding Round
**Assumptions:** Needed to have an intervention START date -- this is because the data is broken out by funding round, not the funding round in which the intervention ended; data used: discharge_data
```{r}
#| label: tbl-discharge-reason-by-round

discharge_data |>
  filter(!is.na(intervention_start)) |>
  tabyl(reason_for_discharge, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

## Client Type

### Client Type Overall
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have client_type
```{r}
#| label: tbl-client-type

.d <- all_data |> 
  filter(!is.na(client_type)) |> 
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

#cat(.d |> nrow(), "of", n_distinct(all_data$client_system_id), "clients have data\n")
cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Client Type by Funding Round
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have client type and funding round
```{r}
#| label: tbl-client-type-by-round

.d <- all_data |> 
  filter(!is.na(client_type),
         !is.na(funding_round)) |> 
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```


## Client Demographics

### Primary Language
**Assumptions:** Distinct by client system ID, Don't know, no primary language selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-primary-language

.d <- all_data |> 
  filter(
    primary_language != "Don't Know",
    primary_language != "No Primary Language Selected",
    primary_language != "Not Applicable",
    primary_language != "Prefer Not To Answer",
    !is.na(primary_language)) |> 
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(primary_language) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Housing Status
**Assumptions:** Distinct by client system ID, Don't know, no housing status selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-housing-status

.d <- all_data |> 
  filter(
    housing_status != "Don't Know",
    housing_status != "No Housing Status Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(housing_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(housing_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Client Type with Average Age
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-client-type-age

all_data_distinct <- all_data_distinct |>
  mutate(
    age_start_tx_num = case_when(
      age_at_start_of_treatment_in_years %in% c("No Age Provided", "", " ") ~ NA_character_,
      TRUE ~ age_at_start_of_treatment_in_years
    ) |> as.numeric()
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

tab_counts <- all_data_distinct |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

tab_age <- all_data_distinct |>
  group_by(client_type) |>
  summarise(
    mean_age   = round(mean(age_start_tx_num, na.rm = TRUE), 1),
    n_with_age = sum(!is.na(age_start_tx_num)),
    .groups = "drop"
  )

tab_counts |>
  left_join(tab_age, by = "client_type") |>
  select(client_type, n, percent, mean_age, n_with_age)
```

### Insurance Status
**Assumptions:** Distinct by client system ID, Don't know, no insurance status selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-insurance-status

.d <- all_data |> 
  filter(
    insurance_status != "Don't Know",
    insurance_status != "No Insurance Status Selected",
    insurance_status != "Not Applicable",
    insurance_status != "Prefer Not To Answer",
    !is.na(insurance_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(insurance_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Gender Identity
**Assumptions:** Distinct by client system ID, Don't know, no gender identity selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-gender-identity

.d <- all_data |> 
  filter(
    gender_identity != "Don't Know",
    gender_identity != "No Gender Identify Selected",
    gender_identity != "Not Applicable",
    gender_identity != "Prefer Not To Answer",
    !is.na(gender_identity)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(gender_identity) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Sexual Orientation
**Assumptions:** Distinct by client system ID, Don't know, no sexual orientation selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-sexual-orientation

.d <- all_data |> 
  filter(
    sexual_orientation != "Don't Know",
    sexual_orientation != "No Sexual Orientation Selected",
    sexual_orientation!= "Not Applicable",
    sexual_orientation != "Prefer Not To Answer",
    !is.na(sexual_orientation)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(sexual_orientation) |> 
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Race

Race is coded numerically as follows:

| Code | Race                                               |
|------|----------------------------------------------------|
| 1    | Asian                                              |
| 2    | Black or African American                          |
| 3    | Hispanic or Latino                                 |
| 4    | Middle Eastern or North African                    |
| 5    | Native American / American Indian or Alaska Native |
| 6    | Native Hawaiian or Pacific Islander                |
| 7    | White                                              |
| 88   | 2+ races selected                                  |
**Assumptions:** Distinct by client system ID, Don't know, no race selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-race

.d <- all_data |> 
  filter(
    race != "Don't Know",
    race != "No Race Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(code_num)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(code_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```

### Ethnicity

Ethnicity is coded numerically as follows:

| Code | Ethnicity               |
|------|-------------------------|
| 1    | Chinese                 |
| 2    | Asian Indian            |
| 3    | Filipino                |
| 4    | Vietnamese              |
| 5    | Korean                  |
| 6    | Japanese                |
| 7    | African American        |
| 8    | Jamaican                |
| 9    | Haitian                 |
| 10   | Nigerian                |
| 11   | Somali                  |
| 12   | Mexican                 |
| 13   | Puerto Rican            |
| 14   | Salvadoran              |
| 15   | Cuban                   |
| 16   | Dominican               |
| 17   | Guatemalan              |
| 18   | Native Hawaiian         |
| 19   | Samoan                  |
| 20   | Chamorro                |
| 21   | English                 |
| 22   | German                  |
| 23   | Irish                   |
| 24   | Italian                 |
| 25   | Polish                  |
| 26   | Scottish                |
| 88   | 2+ ethnicities selected |
**Assumptions:** Distinct by client system ID, Don't know, no ethnicity selected, not applicable, and prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)
```{r}
#| label: tbl-ethnicity

.d <- all_data |> 
   filter(
    race != "Don't Know",
    race != "No Ethnicity Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(ethn_num)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(ethn_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```


### Services in Preferred Language

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** The percent of clients that report services received in their preferred language out of the total that completed the satisfaction survey.

**Data used:** satisfacation survey and all data (demo+discharge joined)
```{r}
#| label: tbl-preferred-language

sat_all_data |>
  arrange(date_clean) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label))) |>
  rename(preferred_lang_response = were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |>
  tabyl(preferred_lang_response) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```
### Number (%) of cases for caregiver/parent highest level of education
**Assumptions:** Distinct by client system ID; must have highest level of education information available
```{r}
#| label: tbl-highest-caregiver-education

.d <- all_data |> 
  filter(!is.na(highest_caregiver_education_level)) %>%  
  distinct(client_system_id, .keep_all = TRUE)

.d |>
  tabyl(highest_caregiver_education_level) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

cat(sprintf("%.1f%% of clients have this data\n", 100 * nrow(.d) / n_distinct(all_data$client_system_id)))
```
