---
title: "Annual Report Results-12-08-2025"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Period

2024-01-01 - 2025-12-31

## Client_sat

**Definition:** Clients from the primary evaluation who indicate high satisfaction with services divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated timeframe.

**Calculation:** This KPI represents ***the percent*** of clients that report high satisfaction with services out of the total that completed the satisfaction survey by period(s). Clients must have a record in the eInsight platform with an intervention type and start date, have discharged from services (with an intervention end date), and have completed a satisfaction survey to be included in this count. Calculation assumes dates before 2024 or after 2025 are correct and any rows with these dates have not been revised.

```{r, client_sat}
#| echo: false
#| warning: false
#| message: false
#| include: false

# Produced by data cleaning/satisfaction_data_cleaning.qmd
# Contains all four survey types; date-cleaned; filtered to reporting period;
# includes year and quarter_label columns.
sat_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

combined_all <- read_csv(
  file.path(sat_dir, "combined_satisfaction_clean.csv"),
  show_col_types = FALSE
) %>%
  mutate(quarter_label = fct_inorder(quarter_label))
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree" on the satisfaction survey. The second definition is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed.

**Nitty gritty:** There are four satisfaction surveys: Caregiver (n=`{r} sum(combined_all$source == "caregiver")`), Child (13-17) (n=`{r} sum(combined_all$source == "child")`), Adult (18-59) (n=`{r} sum(combined_all$source == "adult")`), and Adult (60+) (n=`{r} sum(combined_all$source == "senior")`). A separate data frame was made for question 1, and high satisfaction was defined as the percent of individuals who answered a 4 or 5 for question 1.

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (\<= "2025-12-31").

#### Question 1 Satisfaction

The following table shows high survey satisfaction by year.

```{r}
#| echo: false
#| warning: false
#| message: false

# Question 1 satisfaction by year
pct_4_5_year <- combined_all %>% 
  filter(!is.na(year),!is.na(q2)) %>% 
  group_by(year) %>% 
  summarise(
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    x_4_5 = sum(q2 %in% c(4, 5), na.rm = TRUE),
    n_4_5 = length(q2)
  )

pct_4_5_year

# Question 1 satisfaction overall
pct_4_5_overall <- combined_all %>% 
  filter(!is.na(year),!is.na(q2)) %>% 
  summarise(
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    x_4_5 = sum(q2 %in% c(4, 5), na.rm = TRUE),
    n_4_5 = length(q2)
  )

pct_4_5_overall 
```

#### Combined Survey Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

summary_agree <- combined_all %>%
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) 

df2 <- summary_agree %>% 
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4,5), na.rm = TRUE),
    pct_agree = n_agree / n_answered
  ) %>%
  ungroup() %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df2

# CONFIRMED THIS WORKS THE SAME AS ABOVE
df3 <- summary_agree %>%   
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  rowwise() %>% 
  mutate(count_4 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "4", na.rm = TRUE), 
         count_5 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "5", na.rm = TRUE),
         n_4_5 = sum(c(count_4, count_5), na.rm = TRUE),
         n_total = sum(!is.na(c_across(starts_with("q")))),
         perc_4_5 = n_4_5 / n_total
         ) %>% 
  ungroup() 

df4 <- df3 %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(perc_4_5 >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df4
```

The following table shows high survey satisfaction results by quarter, for data where both `client_id` and `quarter_label` were available.

```{r}
#| echo: false
#| warning: false
#| message: false

summary_agree_quarter <- combined_all %>%
  filter(!is.na(quarter_label)) %>% 
  group_by(quarter_label) %>% 
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4, 5)),
    pct_agree = n_agree / n_answered
  ) %>%
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

summary_agree_quarter
```

## Client_end

**Definition:** Clients from the primary evaluation whose discharge reason is "mutually agreed cessation of treatment" for clients who discharged from an intervention during the designated time frame, divided by the total number of clients from the primary evaluation served and who have a discharge reason recorded.

**Calculation:** This KPI represents ***the percent*** of clients who have a disenrollment/discharge response as "mutually agreed cessation of treatment" out of the total number of clients with a discharge reason by period(s).

**Assumptions:** This is discharge data from EInsight, so can have clients in here multiple times (they can have multiple interventions, or the same intervention on multiple dates). I removed NA's from reason_for_discharge, but did not clean dates at all. I filtered by intervention_start date \>= "2024-01-1" and \<= "12-31-2025."

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

# number of rows with a reason for discharge
count_rows <- nrow(discharge_data_filtered %>% filter(!is.na(reason_for_discharge)))
count_rows

# number of rows with successful discharge
# table of positive outcomes 
tab3 <- discharge_data_filtered %>%
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2025-12-31") %>% 
  tabyl(reason_for_discharge) %>%               
  adorn_pct_formatting(digits = 2)

# count of successful discharge
yes_count <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(n)

# percent of successful discharge
yes_pct3 <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(valid_percent)

tab3
```

The results below show % of successful discharges by `intervention_end_quarter`, once NA's are removed from the data.

```{r}
#| echo: false
#| warning: false
#| message: false
# total table of successful discharges by quarter of intervention END
tab4q <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge),
         intervention_end >= "2024-01-01",
         intervention_end <= "2025-12-31") %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

tab4q %>% filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)")

discharge_data_filtered$intervention_end_year <- year(discharge_data_filtered$intervention_end)

tab4y <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge),
         intervention_end >= "2024-01-01",
         intervention_end <= "2025-12-31") %>% 
  tabyl(reason_for_discharge, intervention_end_year) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 

tab4y %>% filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)")

tab4 <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge),
         !is.na(intervention_name),
         !is.na(pre_measure_s),
         !is.na(post_measure_s),
         !is.na(location),
         intervention_start >= "2024-01-01",
         intervention_start <= "2025-12-31",
         positive_outcomes == "Yes") %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1)

tab4


tab4 %>% filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)")
```

## Client_lang

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** This KPI represents ***the percent*** of clients that report services received in their preferred language on the satisfaction survey out of the total that completed the satisfaction survey by period(s).

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (\<= "2025-12-31"). I filtered out NAs for both the quarter label and the question first before calculating %.

The following table shows the percent of clients who received services in their preferred language on the satisfaction survey by quarter:

```{r}
#| echo: false
#| warning: false
#| message: false
# combining all satisfaction survey data
client_lang_all <- bind_rows(
  caregiver %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "caregiver"),
  child %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "child"),
  adult %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "adult"),
  senior %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "senior"))

# first clean date column
client_lang_all <- client_lang_all %>% 
  mutate(
    date_clean = str_replace(date_completed, "^00", "20"), # replace 00xx with 20xx
    date_clean = str_replace(date_clean, "^29", "20"), # replace 29xx with 20xx
    date_clean = as.Date(date_clean)
  )

# add in a quarter column 
client_lang_all <- client_lang_all %>% 
  filter(!is.na(date_clean),
         date_clean <= "2025-12-31") %>% 
  mutate(quarter_label = paste0("Q", quarter(date_clean), " ", year(date_clean)),
       quarter_label = factor(quarter_label, levels = unique(quarter_label)) #to sort chronologically
  )

# add in a year column 
client_lang_all$year <- year(client_lang_all$date_clean)
  
tab_q <- client_lang_all %>% 
  filter(!is.na(quarter_label),
         !is.na(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response)) %>% 
  tabyl(quarter_label, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 

tab_q 

tab_y <- client_lang_all %>% 
  filter(!is.na(quarter_label),
         !is.na(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response)) %>% 
  tabyl(year, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 

tab_y 

tab <- client_lang_all %>% 
  #filter(!is.na(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response)) %>% 
  tabyl(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
  adorn_pct_formatting(digits = 1) 

tab
```

## 3.12 PSS Reductions in parent-related stress

**Definition:** PSS item(s) measuring reduction in parent/caregiver-related stress after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

**Calculation:** Sum of Q1--Q18 (reverse Q4, Q5, Q7, Q8); this KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

**Scoring:** To calculate a total PSS score, items 1, 2, 5, 6, 7, 8, 17, and 18 should be reverse scored as follows: (1=5) (2=4) (3=3) (4=2) (5=1). The PSS score is then obtained by summing across all items. Higher scores indicate higher levels of perceived stress. If follow-up is LOWER than initial, that indicates a positive outcome.

-   1 = Strongly disagree

-   2 = Disagree

-   3 = Undecided

-   4 = Agree

-   5 = Strongly Agree

**Assumptions:** Total score = NA if 2+ items are missing; used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 18 - 90

```{r}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/pss_data_cleaning.qmd
pss_paired <- read_csv(file.path(assess_dir, "pss_paired_clean.csv"), show_col_types = FALSE)

# summary stats
tab <- pss_paired %>% 
  filter(!is.na(positive_outcome)) %>% 
  tabyl(positive_outcome) %>% 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
pss_paired_long <- pss_paired %>%
  ungroup() %>% 
  select(client_system_id, initial_score, follow_up_score) %>% 
  filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) %>% 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) %>%
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
library(ggridges)
pss_plot <- ggplot(pss_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, PSS") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
pss_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_pss <- pss_paired_long %>% 
  group_by(score_type) %>% 
  filter(!is.na(score_value)) %>% 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) %>% 
  mutate(y_text = if_else(score_type == "Follow-up", 1.08, 1.085))

summary_stats_pss
clipr::write_clip(summary_stats_pss)

# Ridgeline plot
pss_ridge_plot <- ggplot(pss_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  #scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "PSS Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, PSS (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_pss, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.07,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_pss,
            aes(x = median_score, y = y_text, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

pss_ridge_plot
```

## 3.13 YOQ Reductions in overall distress

**Definition:** This KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

The YOQ total score reflects total distress in a client's life, with a high score reflecting increased distress related to experiencing a high number of symptoms, interpersonal difficulties, and decreased satisfaction and quality of life.

**Range:** 0 - 120

**Calculation:** Sum of 30 items (0-4 scale); invalid if 3+ items missing; total score only calculated if 28+ items completed; including those doing reliably better / total who took YOQ by quarter RCI = baseline score - follow-up score - RCI ≤ -10 doing reliably worse - RCI ≥ 10 doing reliably better

**Assumptions:** A positive outcome was if a client was doing reliably better as defined by the YOQ RCI. I used follow-up quarter as the date. I used the LAST follow-up date if there were \>1 assessments. I filtered out NA's from "reliably better"

### Caregiver

```{r}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/yoq_data_cleaning.qmd
yoq_caregiver_paired <- read_csv(file.path(assess_dir, "yoq_caregiver_paired_clean.csv"), show_col_types = FALSE)

# summary stats
tab <- yoq_caregiver_paired %>% 
  filter(!is.na(reliably_better),
         rci != 0) %>% 
  tabyl(reliably_better) %>% # true = doing reliably better
  adorn_pct_formatting(digits = 0) 

tab 

#reliably changed
reliable_change <- yoq_caregiver_paired %>% 
  filter(rci >= 10 | rci <= -10)

#no reliable change
no_reliable_change <- yoq_caregiver_paired %>%
  filter(rci > -10, rci < 10,
         rci != 0)  # strictly between -10 and 10

#these people did reliably better
pos_rci <- reliable_change %>% 
  filter(rci >= 10)

#these people did reliably worse
neg_rci <- reliable_change %>% 
  filter(rci <= -10)

#conditional color formatting
yoq_caregiver_paired$rci_col <- as.factor((yoq_caregiver_paired[,7] > 10 | yoq_caregiver_paired[,7] <= -10)*1)

yoq_caregiver_paired %>% 
  filter(rci != 0) %>% 
  ggplot(aes(rci)) + 
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 1) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Reliable Change Index",
       y = "",
       title = "Distribution of reliable change scores for YOQ Caregiver") +
  scale_fill_manual("Reliable Change Index",
                    labels = c("no meaningful change (n=77)", "meaningful change (n=114)"), 
                    values = c("#ABDDDE", "#046C9A"))
```

### Self

```{r}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/yoq_data_cleaning.qmd
yoq_self_paired <- read_csv(file.path(assess_dir, "yoq_self_paired_clean.csv"), show_col_types = FALSE)

# summary stats
tab <- yoq_self_paired %>% 
  filter(!is.na(reliably_better),
         rci != 0) %>% #grantees are copying initial scores, maybe?
  tabyl(reliably_better) %>% 
  adorn_pct_formatting(digits = 0)

tab

#reliably changed
reliable_change <- yoq_self_paired %>% 
  filter(rci >= 10 | rci <= -10)

#no reliable change
no_reliable_change <- yoq_self_paired %>%
  filter(rci > -10, rci < 10,
         rci != 0)  # strictly between -10 and 10

#these people did reliably better
pos_rci <- reliable_change %>% 
  filter(rci >= 10)

#these people did reliably worse
neg_rci <- reliable_change %>% 
  filter(rci <= -10)

#conditional color formatting
yoq_self_paired$rci_col <- as.factor((yoq_self_paired[,7] > 10 | yoq_self_paired[,7] <= -10)*1)

yoq_self_paired %>% 
  filter(rci != 0) %>% 
  ggplot(aes(rci)) + 
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 1) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Reliable Change Index",
       y = "",
       title = "Distribution of reliable change scores for YOQ Self") +
  scale_fill_manual("Reliable Change Index",
                    labels = c("no meaningful change (n=199)", "meaningful change (n=206)"), 
                    values = c("#ABDDDE", "#046C9A"))
```

## 3.14 Basis-24 Reductions in adult problematic symptoms

**Definition:** Basis-24 item(s) measuring reduction in parent/caregiver-related problematic symptoms after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

**Calculation:** Sum of Q1-Q24 (reverse Q4-Q9); this KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

**Scoring:** Each item is rated on a five-point scale from 0, indicating the lowest severity of symptoms, to 4, indicating greatest severity of symptoms. It is important to note that the **rating values for items 4 through 9 are reversed**. Subscale and overall mean scores also range from 0 to 4. The overall range for this 24-item questionnaire is 0-96, with **higher global scores indicating greater functional impairment (lower is better)**.

The six distinct subscales are calculated by summing up and then dividing by the number of non-missing items for that subscale. An overall mean score is also calculated. For example, the Depression/Functioning subscale consists of six items. If the respondent answers all items in this subscale, then the subscale score is the mean of all six items. If one item is omitted, then the subscale score is the sum rating of the items answered, divided by five.

There are no specified clinical cutoffs for the BASIS-24. Subjective interpretation of scores is made by the provider based on patient history and further investigation of reported symptoms.

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 0-96

```{r}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/basis24_data_cleaning.qmd
basis24_paired <- read_csv(file.path(assess_dir, "basis24_paired_clean.csv"), show_col_types = FALSE)

# summary stats
tab <- basis24_paired %>% 
  filter(!is.na(positive_outcome)) %>% 
  tabyl(positive_outcome) %>% 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
basis24_paired_long <- basis24_paired %>%
  ungroup() %>% 
  filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) %>% 
  select(client_system_id, initial_score, follow_up_score) %>% 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) %>%
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
basis24_plot <- ggplot(basis24_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, Basis-24") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
basis24_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_basis24 <- basis24_paired_long %>% 
  group_by(score_type) %>% 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) %>% 
  mutate(y_text = if_else(score_type == "Follow-up", 1.0375, 1.04))

summary_stats_basis24
#clipr::write_clip(summary_stats_basis24)

# Ridgeline plot
basis24_ridge_plot <- ggplot(basis24_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  #scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Basis-24 Total Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, Basis-24 (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_basis24, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.035,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_basis24,
            aes(x = median_score, y = y_text, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

basis24_ridge_plot
```

## 3.15 CPSS-5 Reductions in PTSD symptoms (youth)

**Definition:** CPSS-5 item(s) measuring reduction in child/youth report of reductions in PTSD symptoms/severity after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

The CPSS-5-I is a 27-item semi-structured interview that assesses PTSD DSM-5 diagnosis and symptom severity in the past month based on interviewer ratings. The CPSS-5-I assesses history of Criterion A traumatic experiences to identify an index trauma. Next, the 20 items assessing DSM-5 PTSD symptoms are presented. The interviewer rates each item from 0 (not at all) to 4 (6 or more times a week/almost always) based on the frequency and severity of the reported symptom experienced in the past month related to the index trauma. Finally, seven items assess impairment of endorsed symptoms on daily functioning pertinent to youth (e.g., fun things you want to do, doing your chores, relationships with your friends).

**Calculation:** Sum of Q1-Q20; this KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

**Scoring:** The total severity score ranges from 0 to 80 and is calculated by summing the ratings of the first 20 items. **Higher scores indicate greater severity of symptoms (lower scores are better).** A cut off score of 31 can be used for identifying a probable PTSD diagnosis in children.

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 0-80 \### Interviewer version

```{r}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/cpss5_data_cleaning.qmd
cpss_i_paired <- read_csv(file.path(assess_dir, "cpss_i_paired_clean.csv"), show_col_types = FALSE)

# summary stats
tab <- cpss_i_paired %>% 
  filter(!is.na(positive_outcome)) %>% 
  tabyl(positive_outcome) %>% 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
cpss_i_paired_long <- cpss_i_paired %>%
  ungroup() %>% 
    filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) %>% 
  select(client_system_id, initial_score, follow_up_score) %>% 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) %>%
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
cpss_i_plot <- ggplot(cpss_i_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, CPSS-5 Interviewer Version") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
cpss_i_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_cpss_i <- cpss_i_paired_long %>% 
  group_by(score_type) %>% 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) 

summary_stats_cpss_i
#clipr::write_clip(summary_stats_cpss_i)

# Ridgeline plot
cpss_i_ridge_plot <- ggplot(cpss_i_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  #scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "CPSS Interviewer Total Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, CPSS Interviewer (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_cpss_i, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.03,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_cpss_i,
            aes(x = median_score, y = 1.035, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

cpss_i_ridge_plot
```

### Self version

```{r}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/cpss5_data_cleaning.qmd
cpss_s_paired <- read_csv(file.path(assess_dir, "cpss_s_paired_clean.csv"), show_col_types = FALSE)

# summary stats
tab <- cpss_s_paired %>% 
  filter(!is.na(positive_outcome)) %>% 
  tabyl(positive_outcome) %>% 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
cpss_s_paired_long <- cpss_s_paired %>%
  ungroup() %>% 
     filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) %>% 
  select(client_system_id, initial_score, follow_up_score) %>% 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) %>%
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
cpss_s_plot <- ggplot(cpss_s_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, CPSS-5 Self Version") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
cpss_s_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_cpss_s <- cpss_s_paired_long %>% 
  group_by(score_type) %>% 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) 

summary_stats_cpss_s
#clipr::write_clip(summary_stats_cpss_s)

wilcox.test(cpss_s_paired$follow_up_score, cpss_s_paired$initial_score, paired = TRUE)

# Ridgeline plot
cpss_s_ridge_plot <- ggplot(cpss_s_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "CPSS Self Total Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, CPSS Self (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_cpss_s, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.03,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_cpss_s,
            aes(x = median_score, y = 1.035, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

cpss_s_ridge_plot
```

## Rikke's SPSS master file

**Access to services**

Among these clients/participants, the largest share (\[XX%\]) received services through a Round 2 program implementing \[EBP\], which \[brief context about the intervention—e.g., target population and core focus\].

Service reach was also concentrated among \[ENTITY TYPE: CBO\] (\[XX%\]) and in Region 8 (Los Angeles County; \[XX%\]).

At the same time, the Workstream served a diverse set of children, youth, and caregivers across provider types. For example, \[XX%\] of clients/participants received services through non-traditional behavioral health providers (e.g., hospitals/hospital systems; \[XX%\]) and tribal entities (\[XX%\]).

```{r}
#| echo: false
#| warning: false
#| message: false
library(haven)

# load in spss data
master_grantee_file <- haven::read_spss("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/MasterGranteeSpreadsheet_forBHESQuantTeam_1-20-26_NEW!.sav") %>% 
  zap_missing() %>%
  mutate(across(where(is.labelled), as.character)) %>%
  mutate(across(where(is.character), ~ na_if(str_trim(.x), ""))) %>% 
  rename_with(~ gsub("^@", "", .x)) %>% 
  janitor::clean_names() %>% 
  rename_with(~ gsub("^x", "", .x))

#join to all data
all_data_spss <- left_join(all_data, master_grantee_file, by = c("provider" = "1a_grantee_provider_name", "intervention_name" = "9a_ebpcde"))

#get #'s for report 
all_data_spss %>%
  filter(`4funding_round` == "Round 4") %>% 
  distinct(client_system_id, provider) %>%   # one row per client per round
  tabyl(provider) %>%
  adorn_totals("row") %>%
  adorn_rounding(digits = 2)

#join to just demographic data - need to add provider name and intervention first to help with join
key <- c("client_system_id", "provider_system_id", "provider_id")

discharge_small <- discharge_data_filtered %>%
  select(all_of(key), provider, intervention_name)

discharge_small %>%
  count(across(all_of(key)), name = "n") %>%
  filter(n > 1)

discharge_unique <- discharge_small %>%
  distinct(across(all_of(key)), .keep_all = TRUE)

demo_discharge <- left_join(demographic_data_filtered, discharge_unique, by = key)

demo_spss <- left_join(demo_discharge, master_grantee_file, by = c("provider_id" = "2grantee_provider_id"))

#get #'s on master spss file
demo_spss %>% 
  tabyl(`12a_grantee_entity_type_for_report`, `8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

demo_spss %>% 
  filter(`7a_grant_track` == "Expansion") %>% 
  tabyl(`12a_grantee_entity_type_for_report`, `8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

master_grantee_file %>% 
  #filter(`7a_grant_track` == "Expansion") %>% 
  tabyl(`1a_grantee_provider_name`, `8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

demo_spss %>% 
  tabyl(`12a_grantee_entity_type`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

demo_spss %>% 
  tabyl(`8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

test <- demo_spss %>% 
  filter(is.na(`8c_regions`))

table(test$provider_id)

master_grantee_file %>%
  count(`2grantee_provider_id`) %>%
  filter(n > 1)
```
