---
title: "Annual Report Results-12-08-2025"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Period

2024-01-01 - 2025-12-31

## Client_sat

**Definition:** Clients from the primary evaluation who indicate high satisfaction with services divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated timeframe.

**Calculation:** This KPI represents ***the percent*** of clients that report high satisfaction with services out of the total that completed the satisfaction survey by period(s). Clients must have a record in the eInsight platform with an intervention type and start date, have discharged from services (with an intervention end date), and have completed a satisfaction survey to be included in this count. Calculation assumes dates before 2024 or after 2025 are correct and any rows with these dates have not been revised.

```{r, client_sat}
#| echo: false
#| warning: false
#| message: false
#| include: false

# Produced by data cleaning/satisfaction_data_cleaning.qmd
# Contains all four survey types; date-cleaned; filtered to reporting period;
# includes year and quarter_label columns.
sat_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

combined_all <- read_csv(
  file.path(sat_dir, "combined_satisfaction_clean.csv"),
  show_col_types = FALSE
) %>%
  mutate(quarter_label = fct_inorder(quarter_label))
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree" on the satisfaction survey. The second definition is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed.

**Nitty gritty:** There are four satisfaction surveys: Caregiver (n=`{r} sum(combined_all$source == "caregiver")`), Child (13-17) (n=`{r} sum(combined_all$source == "child")`), Adult (18-59) (n=`{r} sum(combined_all$source == "adult")`), and Adult (60+) (n=`{r} sum(combined_all$source == "senior")`). A separate data frame was made for question 1, and high satisfaction was defined as the percent of individuals who answered a 4 or 5 for question 1.

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (\<= "2025-12-31").

#### Question 1 Satisfaction

The following table shows high survey satisfaction by year.

```{r}
#| echo: false
#| warning: false
#| message: false

# Question 1 satisfaction by year
pct_4_5_year <- combined_all %>% 
  filter(!is.na(year),!is.na(q2)) %>% 
  group_by(year) %>% 
  summarise(
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    x_4_5 = sum(q2 %in% c(4, 5), na.rm = TRUE),
    n_4_5 = length(q2)
  )

pct_4_5_year

# Question 1 satisfaction overall
pct_4_5_overall <- combined_all %>% 
  filter(!is.na(year),!is.na(q2)) %>% 
  summarise(
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    x_4_5 = sum(q2 %in% c(4, 5), na.rm = TRUE),
    n_4_5 = length(q2)
  )

pct_4_5_overall 
```

#### Combined Survey Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

summary_agree <- combined_all %>%
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) 

df2 <- summary_agree %>% 
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4,5), na.rm = TRUE),
    pct_agree = n_agree / n_answered
  ) %>%
  ungroup() %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df2

# CONFIRMED THIS WORKS THE SAME AS ABOVE
df3 <- summary_agree %>%   
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  rowwise() %>% 
  mutate(count_4 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "4", na.rm = TRUE), 
         count_5 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "5", na.rm = TRUE),
         n_4_5 = sum(c(count_4, count_5), na.rm = TRUE),
         n_total = sum(!is.na(c_across(starts_with("q")))),
         perc_4_5 = n_4_5 / n_total
         ) %>% 
  ungroup() 

df4 <- df3 %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(perc_4_5 >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df4
```

The following table shows high survey satisfaction results by quarter, for data where both `client_id` and `quarter_label` were available.

```{r}
#| echo: false
#| warning: false
#| message: false

summary_agree_quarter <- combined_all %>%
  filter(!is.na(quarter_label)) %>% 
  group_by(quarter_label) %>% 
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4, 5)),
    pct_agree = n_agree / n_answered
  ) %>%
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

summary_agree_quarter
```

## Client_end

**Definition:** Clients from the primary evaluation whose discharge reason is "mutually agreed cessation of treatment" for clients who discharged from an intervention during the designated time frame, divided by the total number of clients from the primary evaluation served and who have a discharge reason recorded.

**Calculation:** This KPI represents ***the percent*** of clients who have a disenrollment/discharge response as "mutually agreed cessation of treatment" out of the total number of clients with a discharge reason by period(s).

**Assumptions:** This is discharge data from EInsight, so can have clients in here multiple times (they can have multiple interventions, or the same intervention on multiple dates). I removed NA's from reason_for_discharge, but did not clean dates at all. I filtered by intervention_start date \>= "2024-01-1" and \<= "12-31-2025."

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

# number of rows with a reason for discharge
count_rows <- nrow(discharge_data_filtered %>% filter(!is.na(reason_for_discharge)))
count_rows

# number of rows with successful discharge
# table of positive outcomes 
tab3 <- discharge_data_filtered %>%
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2025-12-31") %>% 
  tabyl(reason_for_discharge) %>%               
  adorn_pct_formatting(digits = 2)

# count of successful discharge
yes_count <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(n)

# percent of successful discharge
yes_pct3 <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(valid_percent)

tab3
```

The results below show % of successful discharges by `intervention_end_quarter`, once NA's are removed from the data.

```{r}
#| echo: false
#| warning: false
#| message: false
# total table of successful discharges by quarter of intervention END
tab4q <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge),
         intervention_end >= "2024-01-01",
         intervention_end <= "2025-12-31") %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

tab4q %>% filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)")

discharge_data_filtered$intervention_end_year <- year(discharge_data_filtered$intervention_end)

tab4y <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge),
         intervention_end >= "2024-01-01",
         intervention_end <= "2025-12-31") %>% 
  tabyl(reason_for_discharge, intervention_end_year) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 

tab4y %>% filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)")

tab4 <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge),
         !is.na(intervention_name),
         !is.na(pre_measure_s),
         !is.na(post_measure_s),
         !is.na(location),
         intervention_start >= "2024-01-01",
         intervention_start <= "2025-12-31",
         positive_outcomes == "Yes") %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1)

tab4


tab4 %>% filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)")
```

## Rikke's SPSS master file

**Access to services**

Among these clients/participants, the largest share (\[XX%\]) received services through a Round 2 program implementing \[EBP\], which \[brief context about the intervention—e.g., target population and core focus\].

Service reach was also concentrated among \[ENTITY TYPE: CBO\] (\[XX%\]) and in Region 8 (Los Angeles County; \[XX%\]).

At the same time, the Workstream served a diverse set of children, youth, and caregivers across provider types. For example, \[XX%\] of clients/participants received services through non-traditional behavioral health providers (e.g., hospitals/hospital systems; \[XX%\]) and tribal entities (\[XX%\]).

```{r}
#| echo: false
#| warning: false
#| message: false
library(haven)

# load in spss data
master_grantee_file <- haven::read_spss("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Master Files BHES Internal/Master_WorkingFile_3-2-26.sav") %>% 
  zap_missing() %>%
  mutate(across(where(is.labelled), as.character)) %>%
  mutate(across(where(is.character), ~ na_if(str_trim(.x), ""))) %>% 
  rename_with(~ gsub("^@", "", .x)) %>% 
  janitor::clean_names() %>% 
  rename_with(~ gsub("^x", "", .x))

#join to all data
all_data_spss <- left_join(all_data, master_grantee_file, by = c("provider" = "1a_grantee_provider_name", "intervention_name" = "9a_ebpcde"))

#get #'s for report 
all_data_spss %>%
  filter(`4funding_round` == "Round 4") %>% 
  distinct(client_system_id, provider) %>%   # one row per client per round
  tabyl(provider) %>%
  adorn_totals("row") %>%
  adorn_rounding(digits = 2)

#join to just demographic data - need to add provider name and intervention first to help with join
key <- c("client_system_id", "provider_system_id", "provider_id")

discharge_small <- discharge_data_filtered %>%
  select(all_of(key), provider, intervention_name)

discharge_small %>%
  count(across(all_of(key)), name = "n") %>%
  filter(n > 1)

discharge_unique <- discharge_small %>%
  distinct(across(all_of(key)), .keep_all = TRUE)

demo_discharge <- left_join(demographic_data_filtered, discharge_unique, by = key)

demo_spss <- left_join(demo_discharge, master_grantee_file, by = c("provider_id" = "2grantee_provider_id"))

#get #'s on master spss file
demo_spss %>% 
  tabyl(`12a_grantee_entity_type_for_report`, `8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

demo_spss %>% 
  filter(`7a_grant_track` == "Expansion") %>% 
  tabyl(`12a_grantee_entity_type_for_report`, `8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

master_grantee_file %>% 
  #filter(`7a_grant_track` == "Expansion") %>% 
  tabyl(`1a_grantee_provider_name`, `8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

demo_spss %>% 
  tabyl(`12a_grantee_entity_type`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

demo_spss %>% 
  tabyl(`8c_regions`) %>% 
  adorn_totals("row") %>% 
  adorn_rounding(digits = 2)

test <- demo_spss %>% 
  filter(is.na(`8c_regions`))

table(test$provider_id)

master_grantee_file %>%
  count(`2grantee_provider_id`) %>%
  filter(n > 1)
```
