---
title: "Exploratory Descriptives CYBHI Discharges"
author: "Rebecca Heilman"
date: "2025-02-16"
output: html_document
---


aaksdfh;aksdjflakdsfjk;saljfkldafjklsdfj;klasjdfklfjk
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import File

Import the CYBHI discharge file (raw extract).

```{r}
library(readxl)
CYBHI_Discharges <- readxl::read_xlsx("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/Exports - Raw/2025_02/Discharge/CYBHI_EInsights Export_Discharge Summary Report_Export 02.14.2025.xlsx")
View(CYBHI_Discharges)
summary(CYBHI_Discharges)
```

## Modify values for NULL

Adjust the fields that have hyphens for null values and make them true Null values (defined by NA). Define the score fields as integers. Glimpse the data to review a sample of and ensure accurate mutations took place.

```{r}
library(dplyr)

CYBHI_Discharges <- CYBHI_Discharges %>%
  mutate(across(c(`Pre Score`, `Post Score`, `Positive Outcomes`), ~ ifelse(. == "-", NA, .))) %>% 
  mutate(across(c(`Pre Score`, `Post Score`, `Positive Outcomes`), as.integer)) %>% 
      glimpse()
```

## Create Measure Type and Time Point variables

Using the pre- and post-measure fields, develop a measure type and time point variable. First Clean string values as needed.

```{r}
library(tidyr) 
library(stringr)

CYBHI_Discharges <- CYBHI_Discharges %>%
  mutate(
    Pre_Measures_Clean = str_replace_all(
      `Pre Measure(s)`, "\\bIntial\\b", "Initial")) %>% 
  mutate(
    Post_Measures_Clean = str_replace_all(
      `Post Measure(s)`, "\\bUp\\b", "up")) 

```

```{r}

CYBHI_Discharges <- CYBHI_Discharges %>%
  mutate(
    Pre_Measures_Clean = str_squish(Pre_Measures_Clean),
    Post_Measures_Clean = str_squish(Post_Measures_Clean)
  )

replace_list <- c(
  "Strengths and Difficulties Questionnaire P4-10 Initial" = 
  "SDQ Parent or Teacher (age 4-10) Initial", 
  "Pediatric Symptom Checklist \\(PSC\\)" = 
  "PSC-35", 
  "Pediatric Symptom Checklist \\(PSC\\) Follow up 1" = 
  "PSC-35 Follow up 1")

CYBHI_Discharges <- CYBHI_Discharges %>%
  mutate(
    Pre_Measures_Clean = str_replace_all(Pre_Measures_Clean, replace_list),
    Post_Measures_Clean = str_replace_all(Post_Measures_Clean, replace_list)
  )
```

```{r}

unique(CYBHI_Discharges$Pre_Measures_Clean)

unique(CYBHI_Discharges$Post_Measures_Clean)
```

## Extract core measure names (removing "Initial", "Follow up X")

```{r}
CYBHI_Discharges <- CYBHI_Discharges %>%
  mutate(
    Pre_Time_Point = str_extract(
      `Pre_Measures_Clean`, "(Initial|Follow up \\d+)"),
    PreAssessment_Name = str_remove(
      `Pre_Measures_Clean`, "(Initial|Follow up \\d+)") %>% str_trim()
  ) %>% 
  mutate(
    Post_Time_Point = str_extract(
      `Post_Measures_Clean`, "(Initial|Follow up \\d+)"),
    PostAssessment_Name = str_remove(
      `Post_Measures_Clean`, "(Initial|Follow up \\d+)") %>% str_trim()
  )

```

## Check if Pre and Post Assessment Distinct Names Match

Using the new fields, create a flag to identify the matching vs non-matching assessment tools between initial (pre) and follow up time points (post).

```{r}
CYBHI_Discharges <- CYBHI_Discharges %>%
mutate(MatchDistinct = PreAssessment_Name == PostAssessment_Name)

```

## Create Clean Assessment Tool Name Fields - No Distinction between Type of Form

Create fields that categorize the assessment tools by name without distinguishing between the type of form used (Interviewer vs. self-report, child vs parent or caregiver administration, for example).

```{r}
CYBHI_Discharges <- CYBHI_Discharges %>%
  mutate(
    OverarchingPreAssessment = word(PreAssessment_Name, 1),
    OverarchingPostAssessment = word(PostAssessment_Name, 1)
  )
```

## Check if Pre and Post Assessment Nondistinct, Overarching Names Match

Using the new fields, create a flag to identify the matching vs non-matching assessment tools between initial (pre) and follow up time points (post).

```{r}
CYBHI_Discharges <- CYBHI_Discharges %>%
mutate(MatchOverarching = OverarchingPreAssessment == OverarchingPostAssessment)
```

## Create Summary Table showing counts by Pre and Post-Assessments for all cases

```{r}
summary_table_combo <- CYBHI_Discharges %>%
  group_by(PreAssessment_Name) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `PreAssessment_Name` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Create Summary Table showing counts by Pre and Post-Assessments for only matched cases (exact matches)

```{r}
summary_table_matches <- CYBHI_Discharges %>%
  filter(MatchDistinct == TRUE) %>% 
  group_by(PreAssessment_Name) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `PreAssessment_Name` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Create Summary Table showing counts by Pre and Post-Assessments - clear distinctions on assessment tool names

```{r}
summary_table_distinct <- CYBHI_Discharges %>%
  group_by(PreAssessment_Name, PostAssessment_Name) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `PreAssessment_Name` = "Total",
              `PostAssessment_Name` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Create Summary Table showing counts by Pre and Post-Assessments for all cases - Using Overarching Assessment Names

```{r}
summary_table_combo_overarching <- CYBHI_Discharges %>%
  group_by(OverarchingPreAssessment) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `OverarchingPreAssessment` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Create Summary Table showing counts by Pre and Post-Assessments for only matched cases (exact matches, based on overarching assessment names)

```{r}
summary_table_matches_overarching <- CYBHI_Discharges %>%
  filter(MatchOverarching == TRUE) %>% 
  group_by(OverarchingPreAssessment) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `OverarchingPreAssessment` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Create Summary Table showing counts by Pre and Post-Assessments - clear distinctions on assessment tool names - using Overarching assessment names

```{r}
summary_table_distinct_overarching <- CYBHI_Discharges %>%
  group_by(OverarchingPreAssessment, OverarchingPostAssessment) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `OverarchingPreAssessment` = "Total",
              `OverarchingPostAssessment` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Look at only those that are Follow up 1 time point

```{r}
summary_followup1_matches_overarching <- CYBHI_Discharges %>%
  filter(MatchOverarching == TRUE & Post_Time_Point == "Follow up 1") %>% 
  group_by(OverarchingPreAssessment) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `OverarchingPreAssessment` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Create a Summary Table by Funding Round, Assessment Type, and Time Point

```{r}
summRounds_followup1_matches_overarching <- CYBHI_Discharges %>%
  filter(MatchOverarching == TRUE & Post_Time_Point == "Follow up 1") %>% 
  group_by(`Funding Round`, OverarchingPreAssessment) %>%
  summarize(
    count_pre_scores = sum(!is.na(`Pre Score`)),   
    count_post_scores = sum(!is.na(`Post Score`)), 
    total_cases = n(),
    .groups = "drop"
  ) %>% 
  bind_rows(
    summarize(.,
              `OverarchingPreAssessment` = "Total",
              count_pre_scores = sum(count_pre_scores),
              count_post_scores = sum(count_post_scores),
              total_cases = sum(total_cases))
  )
```

## Save Combination Summary Tables and Matching Summary Tables

```{r}
library(openxlsx)

CYBHI_SummaryTables <- createWorkbook()

addWorksheet(CYBHI_SummaryTables, "AllCasesDistinct")
writeData(CYBHI_SummaryTables, "AllCasesDistinct", summary_table_combo)

addWorksheet(CYBHI_SummaryTables, "MatchedCasesDistinct")
writeData(CYBHI_SummaryTables, "MatchedCasesDistinct", summary_table_matches)

addWorksheet(CYBHI_SummaryTables, "AllCasesHiLevel")
writeData(CYBHI_SummaryTables, "AllCasesHiLevel", summary_table_combo_overarching)

addWorksheet(CYBHI_SummaryTables, "MatchedCasesHiLevel")
writeData(CYBHI_SummaryTables, "MatchedCasesHiLevel", summary_table_matches_overarching)

addWorksheet(CYBHI_SummaryTables, "FU1_MatchedCasesHiLevel")
writeData(CYBHI_SummaryTables, "FU1_MatchedCasesHiLevel", summary_followup1_matches_overarching)

addWorksheet(CYBHI_SummaryTables, "ByRoundFU1_MatchHiLev")
writeData(CYBHI_SummaryTables, "ByRoundFU1_MatchHiLev", summRounds_followup1_matches_overarching)

saveWorkbook(CYBHI_SummaryTables, "CYBHI_SummaryTables.xlsx", overwrite = TRUE)
```
