---
title: "NOMS_Visuals"
output: html_document
---

#each NOMS assessment topic area (except for the NOMS sub-scale used for evaluation study question 2.00) will have two visuals, which will show: 1) the averages and the distributions at both initial and discharge time points, and 2) the percentages of those who selected ratings of 4 or higher (indicating the positive outcome) across the sub-scale item(s) at initial and discharge time points.

#the first report out will use a ridgeline plot to visualize across the different versions (caregiver self-report, caregiver report on youth, and youth self-report) as well as overall (all versions): <https://kieranhealy.org/blog/archives/2021/12/19/comparing-distributions/>

#the second report out will use a dumbbell plot to visualize across the different versions (caregiver self-report, caregiver report on youth, and youth self-report) as well as overall (all versions): <https://r-graph-gallery.com/web-dumbbell-chart-with-a-gap-column.html>

# 

```{r}

library(ggplot2)
library(ggridges)
library(tidyr)
library(dplyr)
library(stringr)

# Step 1: Create Initial and Discharge in the dataframe
set.seed(1243098)
df <- df %>%
  mutate(
    pop_a_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))), # centered at 3, max/min of 1 and 5
    pop_a_discharge = pmax(1, pmin(5, rnorm(n(), 3.4, 0.6))),
    pop_b_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_b_discharge = pmax(1, pmin(5, rnorm(n(), 3.3, 0.6))),
    pop_c_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_c_discharge = pmax(1, pmin(5, rnorm(n(), 3.5, 0.6))),
    pop_total_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_total_discharge = pmax(1, pmin(5, rnorm(n(), 3.4, 0.6)))
  )

# Step 2: Pivot to long format
df_long <- df %>%
  pivot_longer(cols = ends_with(c("initial", "discharge")),
               names_to = c("subgroup", "timepoint"),
               names_pattern = "(.*)_(.*)") 

# Step 3: Rename groups for readability and order them
df_long$subgroup <- recode(df_long$subgroup,
                           "pop_a" = "Youth Self-Report",
                           "pop_b" = "Caregiver's Report on Youth/Child",
                           "pop_c" = "Caregiver Self-Report",
                           "pop_total" = "Total")

# Order the subgroups exactly as specified:
df_long$subgroup <- factor(df_long$subgroup, levels = c(
  "Youth Self-Report", 
  "Caregiver's Report on Youth/Child",
  "Caregiver Self-Report",
  "Total"
))

# Order the timepoints to render Discharge in front
df_long$timepoint <- factor(df_long$timepoint, levels = c("initial", "discharge"))

# Step 4: Plot
ggplot(df_long, aes(x = value, y = subgroup, fill = timepoint)) +
  geom_density_ridges(
    aes(color = timepoint),
    alpha = 0.7,
    scale = 0.9,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("initial" = "#8A8C8E", "discharge" = "#BBDB49"),
    breaks = c("initial", "discharge"),
    labels = c("Initial", "Discharge")
  ) +
  scale_color_manual(
    values = c("initial" = "#8A8C8E", "discharge" = "#BBDB49"),
    breaks = c("initial", "discharge"),
    labels = c("Initial", "Discharge")
  ) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 25)) +
  labs(
    x = "Measure: Mental Health (1-5 Likert Scale)",
    y = NULL,
    title = "Comparing Means of Initial and Discharge Time Points\nAcross NOMS Sub-scale by Subgroups"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    plot.title.position = "plot",
    axis.title.x = element_text(hjust = 0),
    axis.text.y = element_text(hjust = 0),
    legend.position = "bottom",
    legend.title = element_blank()
  )





```

```{r}

library(ggplot2)
library(ggridges)
library(tidyr)
library(dplyr)
library(stringr)
library(patchwork)

# Step 1: Create Initial and Discharge in the dataframe
set.seed(1243098)
df <- df %>%
  mutate(
    pop_a_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_a_discharge = pmax(1, pmin(5, rnorm(n(), 3.4, 0.6))),
    pop_b_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_b_discharge = pmax(1, pmin(5, rnorm(n(), 3.3, 0.6))),
    pop_c_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_c_discharge = pmax(1, pmin(5, rnorm(n(), 3.5, 0.6))),
    pop_total_initial = pmax(1, pmin(5, rnorm(n(), 3, 0.6))),
    pop_total_discharge = pmax(1, pmin(5, rnorm(n(), 3.4, 0.6)))
  )

# Step 2: Pivot to long format
df_long <- df %>%
  pivot_longer(cols = ends_with(c("initial", "discharge")),
               names_to = c("subgroup", "timepoint"),
               names_pattern = "(.*)_(.*)") 

# Step 3: Rename groups for readability and order them
df_long$subgroup <- recode(df_long$subgroup,
                           "pop_a" = "Youth Self-Report",
                           "pop_b" = "Caregiver's Report on Youth/Child",
                           "pop_c" = "Caregiver Self-Report",
                           "pop_total" = "Total")

# ✅ Order the subgroups exactly as you specified:
df_long$subgroup <- factor(df_long$subgroup, levels = c(
  "Total",
  "Caregiver Self-Report",
  "Caregiver's Report on Youth/Child",
  "Youth Self-Report"
))

# Order the timepoints to render Discharge in front
df_long$timepoint <- factor(df_long$timepoint, levels = c("initial", "discharge"))

# === 📌 Compute the Mean Difference and Order
mean_differences <- df_long %>%
  group_by(subgroup, timepoint) %>%
  summarise(mean_value = mean(value, na.rm = TRUE)) %>%
  pivot_wider(names_from = timepoint, values_from = mean_value) %>%
  mutate(Difference = discharge - initial)

# === 📌 Create the Gap Difference Table Plot
gap_table <- ggplot(mean_differences, aes(x = 1, y = subgroup)) +
  geom_text(aes(label = round(Difference, 2)), size = 3, hjust = 0) +  # 🔥 Smaller font for cells
  theme_void() +
  labs(title = "Mean Diff.\n(Discharge - Initial)") +  # 🔥 Updated the title
  theme(
    plot.title = element_text(hjust = 0.5, size = 8, margin = margin(t = 5, b = 5)), # 🔥 Smaller title
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.margin = margin(t = 40, r = 40, b = 40, l = 20)
  ) +
  scale_y_discrete(limits = c("Total", "Caregiver Self-Report", "Caregiver's Report on Youth/Child", "Youth Self-Report"))

# === 📌 Create the Ridgeline Plot with **Tight Label Alignment**
ridge_plot <- ggplot(df_long, aes(x = value, y = subgroup, fill = timepoint)) + 
  geom_density_ridges(
    aes(color = timepoint),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("initial" = "#8A8C8E", "discharge" = "#BBDB49"),
    breaks = c("initial", "discharge"),
    labels = c("Initial", "Discharge")
  ) +
  scale_color_manual(
    values = c("initial" = "#8A8C8E", "discharge" = "#BBDB49"),
    breaks = c("initial", "discharge"),
    labels = c("Initial", "Discharge")
  ) +
  scale_x_continuous(
    limits = c(1, 5),
    breaks = seq(1, 5, by = 1)
  ) +
  labs(
    x = "Measure: Mental Health (1-5 Likert Scale)",
    y = NULL,
    title = "Comparing Means of Initial and Discharge Time Points\nAcross NOMS Sub-scale by Subgroups"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_text(size = 10, angle = 0),
    plot.title.position = "plot",
    plot.title = element_text(size = 10, margin = margin(t = 20, b = 20)),
    axis.title.x = element_text(hjust = 0, size = 9),
    axis.text.y = element_text(hjust = 1, size = 9, margin = margin(r = 5)),
    axis.text.x = element_text(size = 9),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 9),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(20, 10, 20, 10)
  )

# === 📌 Combine the Plots and Adjust Layout
(ridge_plot | gap_table) + 
  plot_layout(widths = c(5, 1)) &
  theme(plot.margin = margin(10, 10, 10, 10))


```
