---
title: "Q1.01 Improved Access to Subpopulations"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Question 1.01 improved access to subpopulations

### Race

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by race.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and race is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer
1.  Asian
2.  Black or African American
3.  Hispanic or Latino
4.  Middle Eastern or North African
5.  Native American/American Indian or Alaska Native
6.  Native Hawaiian or Pacific Islander
7.  White
8.  2+ races selected

**Assumptions:** Clients had to have an intervention start, an intervention name, and have race characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)

# using all data to get quarters/interventions
all_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv")

all_data <- all_data |> 
  filter(!is.na(intervention_start), # has to have an intervention start date
         !is.na(intervention_name),
         intervention_start <= "2025-11-30") # has to have an intervention type named

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01") |> 
         #!is.na(code_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(code_num) |> 
  adorn_totals("row") |> 
  #adorn_percentages("row") |>
  adorn_pct_formatting(digits = 1)

tab 

tab <- all_data |> 
   filter(!is.na(client_type),
          !is.na(code_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(client_type, code_num) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

```

### Insurance status

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by insurance status.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and insurance is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer

1.  Private insurance

2.  Medi-Cal/Medi-Care

3.  No Insurance Status Selected

4.  No insurance/Indigent

5.  Other

6.  Veterans Benefits

**Assumptions:** Clients had to have an intervention start, an intervention name, and have insurance status characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| echo: false
#| warning: false
#| message: false

# Define coding for single responses
codebook_insurance <- c("Private Insurance" = 1, 
              "Medi-Cal/Medi-Care" = 2, 
              "No insurance/Indigent" = 3,
              "Other" = 4,
              "Veterans Benefits" = 5,
              "Don't Know" = 6,
              "Prefer Not To Answer" = 7,
              "Not Applicable" = 8, 
              "No Insurance Status Selected" = 9)

all_data <- all_data |> 
  mutate(
    # normalize text a bit
    insurance_status_clean = str_squish(insurance_status),
    # make 3 categories missing
    insurance_status_clean = if_else(
      insurance_status_clean %in% c("Don't Know", "Prefer Not To Answer",
                                    "Not Applicable", "No Insurance Status Selected") |
      insurance_status_clean == "",
      NA_character_,
      insurance_status_clean),
    # numeric code (NA if not in codebook or was set to NA above)
    insurance_num = unname(codebook_insurance[insurance_status]) |> 
      as.integer())
  

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2025-11-30") |> 
         # !is.na(insurance_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(insurance_status) |> 
  adorn_totals("row") |> 
  adorn_pct_formatting(digits = 2)

tab

tab <- all_data |> 
  filter(!is.na(client_type),
         !is.na(insurance_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(client_type, insurance_num) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 


```

### Sexual orientation

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by sexual orientation.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and sexual orientation is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer, no sexual orientation selected

1.  Asexual

2.  Bisexual

3.  Lesbian/Gay

4.  Other

5.  Pansexual

6.  Same Gender Loving

7.  Straight

**Assumptions:** Clients had to have an intervention start, an intervention name, and have sexual orientation characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| echo: false
#| warning: false
#| message: false

# Define coding for single responses
codebook_sexual_orientation <- c("Asexual" = 1, 
              "Bisexual" = 2, 
              "Lesbian/Gay" = 3,
              "Other" = 4,
              "Pansexual" = 5,
              "Same Gender Loving" = 6,
              "Straight" = 7)

all_data <- all_data |> 
  mutate(
    # normalize text a bit
    sexual_orientation_clean = str_squish(sexual_orientation),
    # make 3 categories missing
    sexual_orientation_clean = if_else(
      sexual_orientation_clean %in% c("Don't Know", 
                                      "No Sexual Orientation Selected",
                                      "Prefer Not To Answer",
                                    "Not Applicable") |
      sexual_orientation_clean == "",
      NA_character_,
      sexual_orientation_clean),
    # numeric code (NA if not in codebook or was set to NA above)
    sexual_orientation_num = unname(codebook_sexual_orientation[sexual_orientation_clean]) |> 
      as.integer())
  

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2025-11-30") |> 
         #!is.na(sexual_orientation_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(sexual_orientation) |> 
  adorn_totals("row") |> 
  adorn_pct_formatting(digits = 1)

tab 

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         !is.na(sexual_orientation_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(client_type, sexual_orientation_num) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

```

### Gender identity

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by gender identity.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and gender identity is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer

1.  Genderqueer

2.  Man/boy

3.  No Gender Identity Selected

4.  Non-binary

5.  Transgender man/boy

6.  Transgender woman/girl

7.  Two-spirit

8.  Woman/girl

**Assumptions:** Clients had to have an intervention start, an intervention name, and have gender identity characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| echo: false
#| warning: false
#| message: false

# Define coding for single responses
codebook_gender_identity <- c("Genderqueer" = 1, 
              "Man/boy" = 2, 
              "Non-binary" = 3,
              "Transgender man/boy" = 4,
              "Transgender woman/girl" = 5,
              "Two-spirit" = 6,
              "Woman/girl" = 7)

all_data <- all_data |> 
  mutate(
    # normalize text a bit
    gender_identity_clean = str_squish(gender_identity),
    # make 3 categories missing
    gender_identity_clean = if_else(
      gender_identity_clean %in% c("Don't Know", 
                                      "Prefer Not To Answer",
                                    "Not Applicable", "No Gender Identify Selected") |
      gender_identity_clean == "",
      NA_character_,
      gender_identity_clean),
    # numeric code (NA if not in codebook or was set to NA above)
    gender_identity_num = unname(codebook_gender_identity[gender_identity_clean]) |> 
      as.integer())

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2025-11-30") |> 
  #       gender_identity_num != 3) |> 
         #!is.na(gender_identity_num)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(gender_identity) |> 
  adorn_totals("row") |> 
  adorn_pct_formatting(digits = 1) 

tab 

```

## Preferred language

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by preferred language.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and preferred language is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer, no preferred language selected

1.  English

2.  Spanish

3.  Other

**Assumptions:** Clients had to have an intervention start, an intervention name, and have preferred language characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). Any language with \<200 rows of data is grouped into "Other" category

```{r}
#| echo: false
#| warning: false
#| message: false

all_data <- all_data |> 
  mutate(
    # normalize text a bit
    preferred_language_clean = str_squish(preferred_language),
    # make 3 categories missing
    preferred_language_clean = case_when(
      preferred_language_clean %in% c("Don't Know", 
                                      "Prefer Not To Answer",
                                      "Not Applicable", 
                                      "") ~ NA_character_,
      preferred_language_clean %in% c("English", "Spanish") ~ preferred_language_clean,
      TRUE ~ "Other"),
    # numeric code (NA if not in codebook or was set to NA above)
    preferred_language_num = case_when(
      preferred_language_clean == "English" ~ 1,
      preferred_language_clean == "Spanish" ~ 2,
      preferred_language_clean == "Other" ~ 3,
      TRUE ~ NA_integer_
    ))
  

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         !is.na(preferred_language_clean)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(intervention_start_quarter, preferred_language_clean) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

tab <- all_data |> 
  filter(!is.na(client_type),
         !is.na(preferred_language_clean)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(client_type, preferred_language_clean) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

```

## Client_groups: primary language

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by primary language.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and primary language is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer, no primary language selected

<!-- -->

1.  English

2.  Spanish

3.  Other

**Assumptions:** Clients had to have an intervention start, an intervention name, and have primary language characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). Any language with \<200 rows of data is grouped into "Other" category

```{r}
#| echo: false
#| warning: false
#| message: false

all_data <- all_data |> 
  mutate(
    # normalize text a bit
    primary_language_clean = str_squish(primary_language),
    # make 3 categories missing
    primary_language_clean = case_when(
      primary_language_clean %in% c("Don't Know", 
                                      "Prefer Not To Answer",
                                      "Not Applicable", 
                                      "") ~ NA_character_,
      primary_language_clean %in% c("English", "Spanish") ~ primary_language_clean,
      TRUE ~ "Other"),
    # numeric code (NA if not in codebook or was set to NA above)
    primary_language_num = case_when(
      primary_language_clean == "English" ~ 1,
      primary_language_clean == "Spanish" ~ 2,
      primary_language_clean == "Other" ~ 3,
      TRUE ~ NA_integer_
    ))
  

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         !is.na(primary_language_clean)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(intervention_start_quarter, primary_language_clean) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

tab <- all_data |> 
  filter(!is.na(client_type),
         !is.na(primary_language_clean)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(client_type, primary_language_clean) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 
```

## Client_groups: housing status

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by housing status.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and housing status is recoded as follows:

**Codebook:**

0.  NA, don't know, prefer not to answer, no housing status selected

<!-- -->

1.  Housed

2.  Unhoused

3.  Other (i.e. treatment facility)

**Assumptions:** Clients had to have an intervention start, an intervention name, and have housing status characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| echo: false
#| warning: false
#| message: false

all_data <- all_data |> 
  mutate(
    # normalize text a bit
    housing_status_clean = str_squish(housing_status),
    # make 3 categories missing
    housing_status_clean = case_when(
      housing_status_clean %in% c("Don't Know", 
                                      "Prefer Not To Answer",
                                      "Not Applicable", 
                                      "") ~ NA_character_,
      housing_status_clean %in% c("Housed", "Unhoused") ~ housing_status_clean,
      TRUE ~ "Other (i.e. treatment facility)"),
    # numeric code (NA if not in codebook or was set to NA above)
    housing_status_num = case_when(
      housing_status_clean == "Housed" ~ 1,
      housing_status_clean == "Unhoused" ~ 2,
      housing_status_clean == "Other (i.e. treatment facility)" ~ 3,
      TRUE ~ NA_integer_
    ))
  

tab <- all_data |> 
  filter(intervention_start >= "2024-01-01",
         !is.na(housing_status_clean)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(intervention_start_quarter, housing_status_clean) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

tab <- all_data |> 
  filter(!is.na(client_type),
         !is.na(housing_status_clean)) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  tabyl(client_type, housing_status_clean) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab 

```

