---
title: "Client Language"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

```{r setup}
#| echo: false
#| warning: false
#| message: false
# NOTE: requires caregiver/child/adult/senior — render after q1-00b_client-sat.qmd
library(tidyverse)
library(janitor)
library(openxlsx)
```


## Client_lang

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** This KPI represents ***the percent*** of clients that report services received in their preferred language on the satisfaction survey out of the total that completed the satisfaction survey by period(s).

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (\<= "2025-12-31"). I filtered out NAs for both the quarter label and the question first before calculating %.

The following table shows the percent of clients who received services in their preferred language on the satisfaction survey by quarter:

```{r}
#| echo: false
#| warning: false
#| message: false
# combining all satisfaction survey data
client_lang_all <- bind_rows(
  caregiver |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
    mutate(source = "caregiver"),
  child |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
    mutate(source = "child"),
  adult |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
    mutate(source = "adult"),
  senior |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
    mutate(source = "senior"))

# first clean date column
client_lang_all <- client_lang_all |> 
  mutate(
    date_clean = str_replace(date_completed, "^00", "20"), # replace 00xx with 20xx
    date_clean = str_replace(date_clean, "^29", "20"), # replace 29xx with 20xx
    date_clean = as.Date(date_clean)
  )

# add in a quarter column 
client_lang_all <- client_lang_all |> 
  filter(!is.na(date_clean),
         date_clean <= "2025-12-31") |> 
  mutate(quarter_label = paste0("Q", quarter(date_clean), " ", year(date_clean)),
       quarter_label = factor(quarter_label, levels = unique(quarter_label)) #to sort chronologically
  )

# add in a year column 
client_lang_all$year <- year(client_lang_all$date_clean)
  
tab_q <- client_lang_all |> 
  filter(!is.na(quarter_label),
         !is.na(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response)) |> 
  tabyl(quarter_label, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab_q 

tab_y <- client_lang_all |> 
  filter(!is.na(quarter_label),
         !is.na(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response)) |> 
  tabyl(year, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
  adorn_totals("row") |> 
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() 

tab_y 

tab <- client_lang_all |> 
  #filter(!is.na(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response)) |> 
  tabyl(were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |> 
  adorn_pct_formatting(digits = 1) 

tab

openxlsx::write.xlsx(tab_q,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Client_Lang.xlsx",sheetName="lang_quarterly")
openxlsx::write.xlsx(tab_y,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Client_Lang.xlsx",sheetName="lang_yearly",append=T)
openxlsx::write.xlsx(tab,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Client_Lang.xlsx",sheetName="lang_overall",append=T)
```

