---
title: "Client Satisfaction"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Client_sat

**Definition:** Clients from the primary evaluation who indicate high satisfaction with services divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated timeframe.

**Calculation:** This KPI represents ***the percent*** of clients that report high satisfaction with services out of the total that completed the satisfaction survey by period(s). Clients must have a record in the eInsight platform with an intervention type and start date, have discharged from services (with an intervention end date), and have completed a satisfaction survey to be included in this count. Calculation assumes dates before 2024 or after 2025 are correct and any rows with these dates have not been revised.

```{r client_sat}
#| echo: false
#| warning: false
#| message: false
#| include: false
library(readr)
library(tidyverse)
library(janitor)
library(openxlsx)
library(zoo)

# caregiver
caregiver_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-caregiver.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(caregiver_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(caregiver_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
caregiver <- caregiver_raw[-c(1, 2), ]
names(caregiver) <- new_names
caregiver <- caregiver |>
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# child
child_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-child_13-17.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE)
h1 <- unlist(child_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(child_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
child <- child_raw[-c(1, 2), ]
names(child) <- new_names
child <- child |>
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# adult
adult_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_18-59.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE)
h1 <- unlist(adult_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(adult_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
adult <- adult_raw[-c(1, 2), ]
names(adult) <- new_names
adult <- adult |>
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# senior
senior_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_60_plus.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(senior_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(senior_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
senior <- senior_raw[-c(1, 2), ]
names(senior) <- new_names
senior <- senior |>
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree" on the satisfaction survey. The second definition is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed.

**Nitty gritty:** There are four satisfaction surveys: Caregiver (n=`{r} nrow(caregiver)`), Child (13-17) (n=`{r} nrow(child)`), Adult (18-59) (n=`{r} nrow(adult)`), and Adult (60+) (n=`{r} nrow(senior)`). A separate data frame was made for question 1, and high satisfaction was defined as the percent of individuals who answered a 4 or 5 for question 1.

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (\<= "2025-12-31").

#### Question 1 Satisfaction

```{r client_sat_q1}
#| echo: false
#| warning: false
#| message: false
#| include: false
combined_q1 <- bind_rows(
  caregiver |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) |> 
    rename(q1 = matches("^x1_")) |> 
    mutate(source = "caregiver"),
  child |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) |> 
    rename(q1 = matches("^x1_")) |> 
    mutate(source = "child"),
  adult |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) |> 
    rename(q1 = matches("^x1_")) |> 
    mutate(source = "adult"),
  senior |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) |> 
    rename(q1 = matches("^x1_")) |> 
    mutate(source = "senior"))
```

The following table shows high survey satisfaction by year.

```{r}
#| echo: false
#| warning: false
#| message: false

# combining all satisfaction survey data
combined_all <- bind_rows(
  caregiver |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) |> 
    rename(q = matches("^x")) |> 
    mutate(source = "caregiver"),
  child |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) |> 
    rename(q = matches("^x")) |> 
    mutate(source = "child"),
  adult |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) |> 
    rename(q = matches("^x")) |> 
    mutate(source = "adult"),
  senior |> 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) |> 
    rename(q = matches("^x")) |> 
    mutate(source = "senior"))

# first clean date column
#unique(combined_all$date_completed)
combined_all <- combined_all |> 
  mutate(
    date_clean = str_replace(date_completed, "^00", "20"), # replace 00xx with 20xx
    date_clean = str_replace(date_clean, "^29", "20"), # replace 29xx with 20xx
    date_clean = as.Date(date_clean)
  )
#unique(combined_all$date_clean)

# add in a quarter column
combined_all <- combined_all |>
  filter(!is.na(date_clean),
         date_clean <= "2025-12-30") |>
  mutate(quarter_label = paste0("Q", quarter(date_clean), " ", year(date_clean)),
       quarter_label = factor(quarter_label, levels = unique(quarter_label)) #to sort chronologically
  )

#add in a year column
combined_all$year <- year(combined_all$date_clean)

# Question 1 satisfaction by year
pct_4_5_year <- combined_all |> 
  filter(!is.na(year),!is.na(q2)) |> 
  group_by(year) |> 
  summarise(
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    x_4_5 = sum(q2 %in% c(4, 5), na.rm = TRUE),
    n_4_5 = length(q2)
  )

pct_4_5_year

# Question 1 satisfaction overall
pct_4_5_overall <- combined_all |> 
  filter(!is.na(year),!is.na(q2)) |> 
  summarise(
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    x_4_5 = sum(q2 %in% c(4, 5), na.rm = TRUE),
    n_4_5 = length(q2)
  )

pct_4_5_overall 

openxlsx::write.xlsx(pct_4_5_year,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Question 1 Satisfaction.xlsx",sheetName="by year")

openxlsx::write.xlsx(pct_4_5_overall,"/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Question 1 Satisfaction.xlsx",sheetName="overall",append=T)

```

#### Combined Survey Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

summary_agree <- combined_all |>
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) |>
  mutate(across(everything(), ~ as.numeric(.))) 

df2 <- summary_agree |> 
  mutate(across(everything(), ~ na_if(., 0))) |> 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4,5), na.rm = TRUE),
    pct_agree = n_agree / n_answered
  ) |>
  ungroup() |> 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df2

# CONFIRMED THIS WORKS THE SAME AS ABOVE
df3 <- summary_agree |>   
  mutate(across(everything(), ~ na_if(., 0))) |> 
  rowwise() |> 
  mutate(count_4 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "4", na.rm = TRUE), 
         count_5 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "5", na.rm = TRUE),
         n_4_5 = sum(c(count_4, count_5), na.rm = TRUE),
         n_total = sum(!is.na(c_across(starts_with("q")))),
         perc_4_5 = n_4_5 / n_total
         ) |> 
  ungroup() 

df4 <- df3 |> 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(perc_4_5 >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df4

openxlsx::write.xlsx(df2,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Combined Survey Satisfaction.xlsx",sheetName="df2")

openxlsx::write.xlsx(df4,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Combined Survey Satisfaction.xlsx",sheetName="df4",append=T)
```

The following table shows high survey satisfaction results by quarter, for data where both `client_id` and `quarter_label` were available.

```{r}
#| echo: false
#| warning: false
#| message: false

summary_agree_quarter <- combined_all |>
  filter(!is.na(quarter_label)) |> 
  group_by(quarter_label) |> 
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) |>
  mutate(across(everything(), ~ as.numeric(.))) |> 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4, 5)),
    pct_agree = n_agree / n_answered
  ) |>
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

summary_agree_quarter

openxlsx::write.xlsx(summary_agree_quarter,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Combined Survey Satisfaction.xlsx",sheetName="quarterly",append=T)
```

