---
title: "3.03/3.04 NOMS Q1: Improved Mental Health"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## 3.03/3.04 NOMS Q1: Improved mental health

**Scoring:** An **increase** in score indicates a positive outcome. There are no benchmarks or clinical cutoffs established for NOMS scoring - if the follow-up assessment score is higher than the initial score, a positive outcome is indicated.

**Assumptions:** I filtered out dates before 2024-01-01 and after the date the code was run. I filtered data that's \>3 days from initial assessment and paired to the LAST follow-up. I filtered positive outcomes that had NA's before computing %.

### Child

Range: 1-5

Sum = Q1.

Use the `pairs_primary_long` data frame to compute scores. The following had a positive outcome by follow-up quarter:

```{r reading_in_data}
#| echo: false
#| warning: false
#| message: false
#| include: false
library(readr)
library(tidyverse)
library(readxl)
library(zoo)
library(janitor)
library(scales)

# dealing with crazy header rows
noms_child_raw <- read_xlsx("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/NOMS-child.xlsx", na = c("-", "", "NA", "N/A"), col_names = FALSE) 
h1 <- unlist(noms_child_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(noms_child_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
noms_child <- noms_child_raw[-c(1, 2), ]
names(noms_child) <- new_names
noms_child <- noms_child |>
  select(where(~ !all(is.na(.))))

#make sure columns are numeric
#add a total score column
#add a column for if there was any missing data, in case we want to exclude those rows later
#make sure it's date formatted and get rid of the weird dates before 2024 for now
noms_child <- noms_child |> 
  mutate(across(ends_with("_score"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(any_missing_q1x = any(is.na(c_across(matches("^(q1|x).*_score$")))),
         total_score = sum(c_across(matches("^(q1|x).*_score$")), na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  filter(date_completed <= Sys.Date(),     #date of running the code
         date_completed >= "2024-01-01")  |> #date to filter out random issue dates   
  filter(date_completed <= "2025-12-31")

#pivot long to get individual question scores too
noms_child_long <- noms_child |> 
  pivot_longer(
    cols = matches("^(q1|x).*_score$"),
    names_to = "question",
    values_to = "score"
  ) |> 
  mutate(score = as.numeric(score))

#pair by system_id and question (because the data is in a long format)
paired_outcomes_child <- noms_child_long |> 
  group_by(system_id, question) |> 
  arrange(date_completed, .by_group = TRUE) |> 
  mutate(
    days_since_initial = as.integer(date_completed - first(date_completed)),
    initial_date = first(date_completed),
    initial_score = first(score)
  ) |> 
  # flagging follow-ups
  mutate(
    is_within3 = days_since_initial <= 3 & days_since_initial >= 0,
    is_after3 = days_since_initial >3
  )

# paired data
pairs_primary_long <- paired_outcomes_child |> 
  filter(is_after3 == TRUE) |> #filter to data that's after 3 days from initial
  slice_tail(n = 1) |>  #pairing to the LAST follow-up
  transmute(
    system_id, question,
    initial_date,
    initial_score,
    follow_up_date = date_completed,
    follow_up_score = score,
    days_since_initial
  )
```

```{r}
#| echo: false 
#| warning: false 
#| message: false 

pairs_primary_long_q1 <- pairs_primary_long |> filter(question == "q1_how_would_you_rate_your_child_s_overall_mental_health_right_now_score")

# compute differences in f/u and initial scores
score_diff_child_q1 <- pairs_primary_long_q1 |> 
  mutate(score_change = follow_up_score - initial_score, 
         positive_outcome = ifelse(follow_up_score > initial_score, 1, 0), 
         initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
         follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)))

# summary stats
tab <- score_diff_child_q1 |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0)

tab
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
q1_child_score_long <- pairs_primary_long_q1 |>
  ungroup() |> 
  filter(!is.na(system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) |> 
  select(system_id, initial_score, follow_up_score) |> 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
child_q1_plot <- ggplot(q1_child_score_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Mental Health Score (1-5)", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, NOMS Child Q1") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5))  
  
child_q1_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_child_score_type <- q1_child_score_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  )

summary_stats_child_score_type

# Ridgeline plot
ridge_plot_child <- ggplot(q1_child_score_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Mental Health Score (1-5 Likert Scale)", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, NOMS Child") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_child_score_type, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.05,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_child_score_type,
            aes(x = median_score, y = 1.062, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

ridge_plot_child
```

Dot and arrow plot

```{r}
#| echo: false
#| warning: false
#| message: false
dot_and_arrow_plot <- score_diff_child_q1 |> 
  filter(!is.na(initial_score),
         !is.na(follow_up_score),
         !is.na(positive_outcome)) |> 
  mutate(direction = case_when(
    positive_outcome == 1 ~ "Improved",
    positive_outcome == 0 ~ "Declined",
    TRUE ~ "No Change"
  )) |> 
  filter(!is.na(direction),
         direction != "No Change") |> 
ggplot(aes(y = system_id)) +
  geom_segment(
    aes(
      x = initial_score, xend = follow_up_score,
      yend = system_id, color = direction
    ),
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  geom_point(aes(x = initial_score), size = 1, color = "grey30") +
  geom_point(aes(x = follow_up_score), size = 1, color = "grey30") +
  scale_color_manual(
    values = c("Improved" = "forestgreen",
               "Declined" = "firebrick",
               "No Change" = "gray50")
  ) +
  labs(
    x = "Score",
    y = "Client",
    title = "Change in direction of Q1 score: Child"
  ) +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     axis.text.y = element_blank(),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5))  
  
dot_and_arrow_plot
```

#### Summary stats

The percent of those who selected ratings of 4 or higher (excellent or very good) across Q1 at initial and discharge time points were:

```{r}
#| echo: false
#| warning: false
#| message: false
# % of those who selected 4 or higher on question 1
tab5 <- pairs_primary_long_q1 |> 
  ungroup() |> 
  summarise(
    initial_4_5 = round(100 * (mean(initial_score[!is.na(initial_score)]  %in% c(4,5), na.rm = TRUE)), 2),
    followup_4_5 = round(100 * (mean(follow_up_score[!is.na(follow_up_score)]  %in% c(4,5), na.rm = TRUE)), 2)
  )

tab5
```

### Self

Range: 1-5

Sum = Q1

Use the `pairs_primary_long_self` data frame to compute scores. The following had a positive outcome by follow-up quarter:

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
noms_self_raw <- read_xlsx("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/NOMS-self_report.xlsx", na = c("-", "", "NA", "N/A"), col_names = FALSE)
h1 <- unlist(noms_self_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(noms_self_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
noms_self <- noms_self_raw[-c(1, 2), ]
names(noms_self) <- new_names
noms_self <- noms_self |>
  select(where(~ !all(is.na(.))))

#make sure columns are numeric
#add a total score column
#add a column for if there was any missing data, in case we want to exclude those rows later
#make sure it's date formatted and get rid of the weird dates before 2024 for now
noms_self <- noms_self |> 
  mutate(across(ends_with("_score"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(any_missing_q1x = any(is.na(c_across(matches("^(q1|x).*_score$")))),
         total_score = sum(c_across(matches("^(q1|x).*_score$")), na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  filter(date_completed <= Sys.Date(),     #date of running the code
         date_completed >= "2024-01-01") |> #date to filter out random issue dates   
  filter(date_completed <= "2025-12-31")

#pivot long to get individual question scores too
noms_self_long <- noms_self |> 
  pivot_longer(
    cols = matches("^(q1|x).*_score$"),
    names_to = "question",
    values_to = "score"
  ) |> 
  mutate(score = as.numeric(score))

#pair per question
paired_outcomes_self <- noms_self_long |> 
  group_by(client_system_id, question) |> 
  arrange(date_completed, .by_group = TRUE) |> 
  mutate(
    days_since_initial = as.integer(date_completed - first(date_completed)),
    initial_date = first(date_completed),
    initial_score = first(score)
  ) |> 
  # flagging follow-ups
  mutate(
    is_within3 = days_since_initial <= 3 & days_since_initial >= 0,
    is_after3 = days_since_initial >3
  )

# paired data
pairs_primary_long_self <- paired_outcomes_self |> 
  filter(is_after3 == TRUE) |> #filtered for assessments >3 days from initial
  slice_tail(n = 1) |>  #paired on LAST assessment
  transmute(
    client_system_id, question,
    initial_date,
    initial_score,
    follow_up_date = date_completed,
    follow_up_score = score,
    days_since_initial
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q1_self <- pairs_primary_long_self |> 
  filter(question == "q1_how_would_you_rate_your_overall_mental_health_right_now_score")

# compute differences in f/u and initial scores
score_diff_self_q1 <- pairs_primary_long_q1_self |> 
  mutate(score_change = follow_up_score - initial_score,
         positive_outcome = ifelse(follow_up_score > initial_score, 1, 0),
         initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
         follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)))

# summary stats
tab <- score_diff_self_q1 |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
q1_self_score_long <- pairs_primary_long_q1_self |>
  ungroup() |> 
  filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) |> 
  select(client_system_id, initial_score, follow_up_score) |> 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
self_q1_plot <- ggplot(q1_self_score_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Mental Health Score (1-5)", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, NOMS Self Q1") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5))  
  
self_q1_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_self_score_type <- q1_self_score_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  )

summary_stats_self_score_type

# Ridgeline plot
ridge_plot_self <- ggplot(q1_self_score_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Mental Health Score (1-5 Likert Scale)", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, NOMS Self") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) 

ridge_plot_self
```

Dot and arrow plot

```{r}
#| echo: false
#| warning: false
#| message: false

dot_and_arrow_plot_self <- score_diff_self_q1 |> 
  filter(!is.na(initial_score),
         !is.na(follow_up_score),
         !is.na(positive_outcome)) |> 
  mutate(direction = case_when(
    positive_outcome == 1 ~ "Improved",
    positive_outcome == 0 ~ "Declined",
    TRUE ~ "No Change"
  )) |> 
  filter(!is.na(direction),
         direction != "No Change") |> 
  ggplot(aes(y = client_system_id)) + 
  geom_segment(
    aes(
      x = initial_score, xend = follow_up_score,
      yend = client_system_id, color = direction
    ),
    alpha = 0.7,
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  geom_point(aes(x = initial_score), size = 1, color = "grey30") +
  geom_point(aes(x = follow_up_score), size = 1, color = "grey30") +
  scale_color_manual(
    values = c("Improved" = "forestgreen",
               "Declined" = "firebrick",
               "No Change" = "gray50")
  ) +
  labs(
    x = "Score",
    y = "Client",
    title = "Change in direction of Q1 score: Self"
  ) +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     axis.text.y = element_blank(),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
```

#### Summary stats

The percent of those who selected ratings of 4 or higher (excellent or very good) across Q1 at initial and discharge time points were:

```{r}
#| echo: false
#| warning: false
#| message: false
# % of those who selected 4 or higher on question 1
tab5 <- pairs_primary_long_q1_self |> 
  ungroup() |> 
  summarise(
    initial_4_5 = round(100 * (mean(initial_score[!is.na(initial_score)]  %in% c(4,5), na.rm = TRUE)), 2),
    followup_4_5 = round(100 * (mean(follow_up_score[!is.na(follow_up_score)]  %in% c(4,5), na.rm = TRUE)), 2)
  )

tab5
```

### Summary stats overall

```{r}
#| echo: false
#| warning: false
#| message: false

# % of those who selected 4 or higher on question 1
df <- tribble(
  ~group, ~ initial, ~followup,
  "Child", 25.15, 53.46,
  "Self", 30.23, 46
) 

dot_and_arrow_plot_4_5 <- ggplot(df, aes(y = group)) + 
  geom_segment(aes(x = initial, xend = followup, yend = group),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "gray40", linewidth = 0.8) + 
  #initial
  geom_point(aes(x = initial, color = "Initial"), size = 12, show.legend = TRUE) + 
  geom_text(aes(x = initial, label = paste0(round(initial), "%")),
            size = 3.5, color = "grey20") +
  #followup
  geom_point(aes(x = followup, color = "Follow-up"), size = 12, show.legend = TRUE)+
  geom_text(aes(x = followup, label = paste0(round(followup), "%")),
            size = 3.5, color = "grey20") +
  scale_x_continuous(
    limits = c(0, 100),
    labels = label_percent(scale = 1)
  ) +
  scale_color_manual(
    name = "Timepoint",
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49")
  )+
  labs(
    x = "Score",
    y = "",
    title = "Change in direction of scores for those who selected 4 or higher on NOMS Q1 by subgroup") + 
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.text = element_text(size = 8),
     legend.position =c(0.98, 0.98),
     legend.justification = c("right", "top"),
     plot.margin = margin(5, 5, 5, 5))  

dot_and_arrow_plot_4_5
```

