---
title: "3.07/3.08 NOMS Q3: Improved Well-Being"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## 3.07/3.08 NOMS Q3: Improved well-being

### Child

Range: 7-35

Sum = Q3a+...+Q3g

Use the `pairs_primary_long` data frame to compute scores. The following had a positive outcome by follow-up quarter:

```{r}
#| echo: false
#| warning: false
#| message: false
library(scales)
pairs_primary_long_q3 <- pairs_primary_long |> 
  filter(str_detect(question, "^x3.*_score$"))

# compute differences in f/u and initial scores
score_diff_child_q3 <- pairs_primary_long_q3 |> 
  filter(!is.na(initial_score),
         !is.na(follow_up_score), 
         !is.na(system_id)) |> 
  group_by(system_id) |> 
  summarise(initial_total = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            score_change = followup_total - initial_total,
            positive_outcome = ifelse(followup_total > initial_total, 1, 0),
            initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
            follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))) |> 
  distinct(system_id, .keep_all = TRUE) |> 
  filter(between(initial_total, 7, 35),
         between(followup_total, 7, 35))  

# summary stats
tab <- score_diff_child_q3 |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 
  
tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false

q3_child_score_long <- score_diff_child_q3 |>
  ungroup() |> 
  select(system_id, initial_total, followup_total) |> 
  pivot_longer(
    cols = c(initial_total, followup_total),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total" = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
child_q3_plot <- ggplot(q3_child_score_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Improved Well-Being Score (7-35)", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, NOMS Child Q3") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5))  
  
child_q3_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_child_q3 <- q3_child_score_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) |> 
  mutate(y_text = if_else(score_type == "Initial", 1.09, 1.095))

summary_stats_child_q3

# Ridgeline plot
q3_ridge_plot_child <- ggplot(q3_child_score_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Improved Well-Being Score (1-5 Likert Scale)", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, NOMS Child") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_child_q3, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.08,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_child_q3,
            aes(x = median_score, y = y_text,
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

q3_ridge_plot_child
```

#### Summary stats

The percent of those who had total scores 28 or higher on Q2 at initial and discharge time points were:

Change in direction of total scores for those who scored 28 or higher on NOMS Q2 by subgroup"

```{r}
#| echo: false
#| warning: false
#| message: false
# % of those who selected 4 or higher on question 1
tab5 <- score_diff_child_q3 |> 
  ungroup() |> 
  summarise(
     initial_28 = round(100 * (mean(initial_total[!is.na(initial_total)] >= 28, na.rm = TRUE)), 2),
    followup_28 = round(100 * (mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE)), 2)
  )

tab5
```

### Self

Range: 7-35

Sum = Q3a+...+Q3g

Use the `pairs_primary_long_self` data frame to compute scores. The following had a positive outcome by follow-up quarter:

```{r}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q3_self <- pairs_primary_long_self |> 
  filter(str_detect(question, "^x3.*_score$"))

# compute differences in f/u and initial scores
score_diff_self_q3 <- pairs_primary_long_q3_self |> 
  filter(!is.na(initial_score),
         !is.na(follow_up_score),
         !is.na(client_system_id)) |> 
  group_by(client_system_id) |> 
  summarise(initial_total = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            score_change = followup_total - initial_total,
            positive_outcome = ifelse(followup_total > initial_total, 1, 0),
            initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
            follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))) |> 
  distinct(client_system_id, .keep_all = TRUE) |> 
  filter(between(initial_total, 7, 35),
         between(followup_total, 7, 35))

# summary stats
tab <- score_diff_self_q3 |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
q3_self_score_long <- score_diff_self_q3 |>
  ungroup() |> 
  select(client_system_id, initial_total, followup_total) |> 
  pivot_longer(
    cols = c(initial_total, followup_total),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total" = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
self_q3_plot <- ggplot(q3_self_score_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Improved Well-Being Score (7-35)", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, NOMS Self Q3") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5))  
  
self_q3_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_self_q3 <- q3_self_score_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) |> 
  mutate(y_text = if_else(score_type == "Initial", 1.085, 1.09))

summary_stats_self_q3

# Ridgeline plot
q3_ridge_plot_self <- ggplot(q3_self_score_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Measure: Improved Well-Being Score (1-5 Likert Scale)", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, NOMS Self") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_self_q3, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.08,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_self_q3,
            aes(x = median_score, y = 1.09, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
            vjust = 2, hjust = 0.5, size = 2)

q3_ridge_plot_self
```

#### Summary stats

The percent of those who selected ratings of 4 or higher (excellent or very good) across Q1 at initial and discharge time points were:

```{r}
#| echo: false
#| warning: false
#| message: false
# % of those who selected 4 or higher on question 1
tab5 <- score_diff_self_q3 |> 
  ungroup() |> 
  summarise(
     initial_28 = round(100 * (mean(initial_total[!is.na(initial_total)] >= 28, na.rm = TRUE)), 2),
    followup_28 = round(100 * (mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE)), 2)
  )

tab5
```

#### Summary stats overall

```{r}
#| echo: false
#| warning: false
#| message: false

# % of those who selected 4 or higher on question 1
dfq3 <- tribble(
  ~group, ~ initial, ~followup,
  "Child", 45.61, 60.14,
  "Self", 33.98, 44.48
)

dot_and_arrow_plot_4_5_q3 <- ggplot(dfq3, aes(y = group)) + 
  geom_segment(aes(x = initial, xend = followup, yend = group),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "gray40", linewidth = 0.8) + 
  #initial
  geom_point(aes(x = initial, color = "Initial"), size = 12, show.legend = TRUE) + 
  geom_text(aes(x = initial, label = paste0(round(initial), "%")),
            size = 3.5, color = "grey20") +
  #followup
  geom_point(aes(x = followup, color = "Follow-up"), size = 12, show.legend = TRUE)+
  geom_text(aes(x = followup, label = paste0(round(followup), "%")),
            size = 3.5, color = "grey20") +
  scale_x_continuous(
    limits = c(0, 100),
    labels = label_percent(scale = 1)
  ) +
  scale_color_manual(
    name = "Timepoint",
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49")
  )+
  labs(
    x = "Score",
    y = "",
    title = "Change in direction of total scores for those who scored 28 or higher on NOMS Q3 by subgroup") + 
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.text = element_text(size = 8),
     legend.position = "inside",
    legend.justification = c("left", "top"),
     plot.margin = margin(5, 5, 5, 5))  

dot_and_arrow_plot_4_5_q3
```

