---
title: "3.12 PSS: Reductions in Parent-Related Stress"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## 3.12 PSS Reductions in parent-related stress

**Definition:** PSS item(s) measuring reduction in parent/caregiver-related stress after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

**Calculation:** Sum of Q1--Q18 (reverse Q4, Q5, Q7, Q8); this KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

**Scoring:** To calculate a total PSS score, items 1, 2, 5, 6, 7, 8, 17, and 18 should be reverse scored as follows: (1=5) (2=4) (3=3) (4=2) (5=1). The PSS score is then obtained by summing across all items. Higher scores indicate higher levels of perceived stress. If follow-up is LOWER than initial, that indicates a positive outcome.

-   1 = Strongly disagree

-   2 = Disagree

-   3 = Undecided

-   4 = Agree

-   5 = Strongly Agree

**Assumptions:** Total score = NA if 2+ items are missing; used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 18 - 90

```{r}
#| echo: false
#| warning: false
#| message: false
library(ggridges)

raw_pss <- read_xlsx("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/PSS.xlsx", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(raw_pss[1, ], use.names = FALSE)           # question text row
h2 <- unlist(raw_pss[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
pss <- raw_pss[-c(1, 2), ]
names(pss) <- new_names
pss <- pss |>
  select(where(~ !all(is.na(.)))) |> 
  filter(date_completed <= "2025-12-31")

#reverse scored q's
rev_idx <- c(1, 2, 5, 6, 7, 8, 17, 18)

#score cols
score_cols <- names(pss)[endsWith(names(pss), "_score")]

#make sure columns are numeric
#add a total score column, which = NA if 2+ items are missing and reverse scoring correct items
#make sure it's date formatted
pss_scores <- pss |> 
  mutate(across(ends_with("_score"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(
    missing_count = sum(is.na(c_across(all_of(score_cols)))),
    total_score = {
      vals <- c_across(all_of(score_cols))
      vals[rev_idx] <- 6 - vals[rev_idx]
      sum(vals, na.rm = TRUE)
      }) |> 
  ungroup() |> 
  mutate(
    total_score = ifelse(missing_count >= 2, NA_real_, total_score),
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  select(client_system_id, collection, date_completed, total_score)
  
pss_paired <- pss_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments = n(),
    initial_date  = first(date_completed),
    initial_score = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed), as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score), NA_real_)
  ) |>
  ungroup() |> 
  filter(
    !is.na(follow_up_date)
  ) |> 
  mutate(
    positive_outcome = ifelse(follow_up_score < initial_score, 1, 0), 
    score_change = follow_up_score - initial_score,
    initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)))
    
# summary stats
tab <- pss_paired |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
pss_paired_long <- pss_paired |>
  ungroup() |> 
  select(client_system_id, initial_score, follow_up_score) |> 
  filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) |> 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
pss_plot <- ggplot(pss_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, PSS") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
pss_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_pss <- pss_paired_long |> 
  group_by(score_type) |> 
  filter(!is.na(score_value)) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) |> 
  mutate(y_text = if_else(score_type == "Follow-up", 1.08, 1.085))

summary_stats_pss
# clipr::write_clip(summary_stats_pss)

# Ridgeline plot
pss_ridge_plot <- ggplot(pss_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  #scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "PSS Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, PSS (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_pss, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.07,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_pss,
            aes(x = median_score, y = y_text, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

pss_ridge_plot
```

