---
title: "3.14 Basis-24: Reductions in Adult Problematic Symptoms"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(readxl)
library(tidyverse)
library(janitor)
library(zoo)
library(ggridges)
```


## 3.14 Basis-24 Reductions in adult problematic symptoms

**Definition:** Basis-24 item(s) measuring reduction in parent/caregiver-related problematic symptoms after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

**Calculation:** Sum of Q1-Q24 (reverse Q4-Q9); this KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

**Scoring:** Each item is rated on a five-point scale from 0, indicating the lowest severity of symptoms, to 4, indicating greatest severity of symptoms. It is important to note that the **rating values for items 4 through 9 are reversed**. Subscale and overall mean scores also range from 0 to 4. The overall range for this 24-item questionnaire is 0-96, with **higher global scores indicating greater functional impairment (lower is better)**.

The six distinct subscales are calculated by summing up and then dividing by the number of non-missing items for that subscale. An overall mean score is also calculated. For example, the Depression/Functioning subscale consists of six items. If the respondent answers all items in this subscale, then the subscale score is the mean of all six items. If one item is omitted, then the subscale score is the sum rating of the items answered, divided by five.

There are no specified clinical cutoffs for the BASIS-24. Subjective interpretation of scores is made by the provider based on patient history and further investigation of reported symptoms.

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 0-96

```{r}
#| echo: false
#| warning: false
#| message: false
raw_basis24 <- read_xlsx("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/Basis-24.xlsx", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE)
h1 <- unlist(raw_basis24[1, ], use.names = FALSE)           # question text row
h2 <- unlist(raw_basis24[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
basis24 <- raw_basis24[-c(1, 2), ]
names(basis24) <- new_names
basis24 <- basis24 |>
  select(where(~ !all(is.na(.))))

#reverse scored q's
rev_idx <- c(4, 5, 6, 7, 8, 9)

#score cols
score_cols <- names(basis24)[endsWith(names(basis24), "_score")]

#make sure columns are numeric
#add a total score column, which = NA if 2+ items are missing and reverse scoring correct items
#make sure it's date formatted
basis24_scores <- basis24 |> 
  mutate(across(ends_with("_score"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(
    missing_count = sum(is.na(c_across(all_of(score_cols)))),
    total_score = {
      vals <- c_across(all_of(score_cols))
      vals[rev_idx] <- 6 - vals[rev_idx]
      sum(vals, na.rm = TRUE)
      }) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  select(client_system_id, collection, date_completed, total_score)
  
basis24_paired <- basis24_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments = n(),
    initial_date  = first(date_completed),
    initial_score = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed), as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score), NA_real_)
  ) |>
  ungroup() |> 
  filter(
    !is.na(follow_up_date)
  ) |> 
  mutate(
    positive_outcome = ifelse(follow_up_score < initial_score, 1, 0), 
    score_change = follow_up_score - initial_score,
    initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)))
    
# summary stats
tab <- basis24_paired |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
basis24_paired_long <- basis24_paired |>
  ungroup() |> 
  filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) |> 
  select(client_system_id, initial_score, follow_up_score) |> 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
basis24_plot <- ggplot(basis24_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, Basis-24") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
basis24_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_basis24 <- basis24_paired_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) |> 
  mutate(y_text = if_else(score_type == "Follow-up", 1.0375, 1.04))

summary_stats_basis24

# Ridgeline plot
basis24_ridge_plot <- ggplot(basis24_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  #scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Basis-24 Total Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, Basis-24 (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_basis24, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.035,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_basis24,
            aes(x = median_score, y = y_text, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

basis24_ridge_plot
```

