---
title: "3.13 YOQ: Reductions in Overall Distress"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(janitor)
```


## 3.13 YOQ Reductions in overall distress

**Definition:** This KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

The YOQ total score reflects total distress in a client's life, with a high score reflecting increased distress related to experiencing a high number of symptoms, interpersonal difficulties, and decreased satisfaction and quality of life.

**Range:** 0 - 120

**Calculation:** Sum of 30 items (0-4 scale); invalid if 3+ items missing; total score only calculated if 28+ items completed; including those doing reliably better / total who took YOQ by quarter RCI = baseline score - follow-up score - RCI ≤ -10 doing reliably worse - RCI ≥ 10 doing reliably better

**Assumptions:** A positive outcome was if a client was doing reliably better as defined by the YOQ RCI. I used follow-up quarter as the date. I used the LAST follow-up date if there were \>1 assessments. I filtered out NA's from "reliably better"

### Caregiver

```{r}
#| echo: false
#| warning: false
#| message: false
raw_yoq_caregiver <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/YOQ-caregiver.csv", na = c("-", "", "NA", "N/A")) |> janitor::clean_names()

#make sure columns are numeric
#add a column for if there was any missing data, in case we want to exclude those rows later
#add a total score column, which = NA if 3+ items are missing
#make sure it's date formatted
yoq_caregiver <- raw_yoq_caregiver |> 
  mutate(across(matches("^x"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(
    missing_count = sum(is.na(c_across(starts_with("x")))),
    total_score = ifelse(missing_count >= 3, NA_real_, sum(c_across(starts_with("x")), na.rm = TRUE))) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  select(client_system_id, collection, date_completed, total_score)

yoq_caregiver_paired <- yoq_caregiver |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments = n(),
    initial_date  = first(date_completed),
    initial_score = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed), as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score), NA_real_)
  ) |>
  ungroup() |> 
  filter(
    !is.na(follow_up_date)
  ) |> 
  mutate(
    rci = initial_score - follow_up_score, 
    initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
            follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)),
    reliably_better = rci >= 10)
    
# summary stats
tab <- yoq_caregiver_paired |> 
  filter(!is.na(reliably_better),
         rci != 0) |> 
  tabyl(reliably_better) |> # true = doing reliably better
  adorn_pct_formatting(digits = 0) 

tab 

#reliably changed
reliable_change <- yoq_caregiver_paired |> 
  filter(rci >= 10 | rci <= -10)

#no reliable change
no_reliable_change <- yoq_caregiver_paired |>
  filter(rci > -10, rci < 10,
         rci != 0)  # strictly between -10 and 10

#these people did reliably better
pos_rci <- reliable_change |> 
  filter(rci >= 10)

#these people did reliably worse
neg_rci <- reliable_change |> 
  filter(rci <= -10)

#conditional color formatting
yoq_caregiver_paired$rci_col <- as.factor((yoq_caregiver_paired[,7] > 10 | yoq_caregiver_paired[,7] <= -10)*1)

yoq_caregiver_paired |> 
  filter(rci != 0) |> 
  ggplot(aes(rci)) + 
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 1) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Reliable Change Index",
       y = "",
       title = "Distribution of reliable change scores for YOQ Caregiver") +
  scale_fill_manual("Reliable Change Index",
                    labels = c("no meaningful change (n=77)", "meaningful change (n=114)"), 
                    values = c("#ABDDDE", "#046C9A"))
```

### Self

```{r}
#| echo: false
#| warning: false
#| message: false
raw_yoq_self <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/YOQ-self_report.csv", na = c("-", "", "NA", "N/A")) |> janitor::clean_names()

#make sure columns are numeric
#add a column for if there was any missing data, in case we want to exclude those rows later
#add a total score column, which = NA if 3+ items are missing
#make sure it's date formatted
yoq_self <- raw_yoq_self |> 
  mutate(across(matches("^x"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(
    missing_count = sum(is.na(c_across(starts_with("x")))),
    total_score = ifelse(missing_count >= 3, NA_real_, sum(c_across(starts_with("x")), na.rm = TRUE))) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  select(client_system_id, collection, date_completed, total_score)

yoq_self_paired <- yoq_self |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments = n(),
    initial_date  = first(date_completed),
    initial_score = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed), as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score), NA_real_)
  ) |>
  ungroup() |> 
  filter(
    !is.na(follow_up_date)
  ) |> 
  mutate(
    rci = initial_score - follow_up_score, 
    initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
            follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)),
    reliably_better = rci >= 10)
    
# summary stats
tab <- yoq_self_paired |> 
  filter(!is.na(reliably_better),
         rci != 0) |> #grantees are copying initial scores, maybe?
  tabyl(reliably_better) |> 
  adorn_pct_formatting(digits = 0)

tab

#reliably changed
reliable_change <- yoq_self_paired |> 
  filter(rci >= 10 | rci <= -10)

#no reliable change
no_reliable_change <- yoq_self_paired |>
  filter(rci > -10, rci < 10,
         rci != 0)  # strictly between -10 and 10

#these people did reliably better
pos_rci <- reliable_change |> 
  filter(rci >= 10)

#these people did reliably worse
neg_rci <- reliable_change |> 
  filter(rci <= -10)

#conditional color formatting
yoq_self_paired$rci_col <- as.factor((yoq_self_paired[,7] > 10 | yoq_self_paired[,7] <= -10)*1)

yoq_self_paired |> 
  filter(rci != 0) |> 
  ggplot(aes(rci)) + 
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 1) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Reliable Change Index",
       y = "",
       title = "Distribution of reliable change scores for YOQ Self") +
  scale_fill_manual("Reliable Change Index",
                    labels = c("no meaningful change (n=199)", "meaningful change (n=206)"), 
                    values = c("#ABDDDE", "#046C9A"))
```

