---
title: "3.15 CPSS-5: Reductions in PTSD Symptoms (Youth)"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(janitor)
library(zoo)
library(ggridges)
```


## 3.15 CPSS-5 Reductions in PTSD symptoms (youth)

**Definition:** CPSS-5 item(s) measuring reduction in child/youth report of reductions in PTSD symptoms/severity after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

The CPSS-5-I is a 27-item semi-structured interview that assesses PTSD DSM-5 diagnosis and symptom severity in the past month based on interviewer ratings. The CPSS-5-I assesses history of Criterion A traumatic experiences to identify an index trauma. Next, the 20 items assessing DSM-5 PTSD symptoms are presented. The interviewer rates each item from 0 (not at all) to 4 (6 or more times a week/almost always) based on the frequency and severity of the reported symptom experienced in the past month related to the index trauma. Finally, seven items assess impairment of endorsed symptoms on daily functioning pertinent to youth (e.g., fun things you want to do, doing your chores, relationships with your friends).

**Calculation:** Sum of Q1-Q20; this KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

**Scoring:** The total severity score ranges from 0 to 80 and is calculated by summing the ratings of the first 20 items. **Higher scores indicate greater severity of symptoms (lower scores are better).** A cut off score of 31 can be used for identifying a probable PTSD diagnosis in children.

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 0-80 \### Interviewer version

```{r}
#| echo: false
#| warning: false
#| message: false
cpss_i_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/CPSS-5-Interviewer.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE)
h1 <- unlist(cpss_i_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(cpss_i_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
cpss_i <- cpss_i_raw[-c(1, 2), ]
names(cpss_i) <- new_names
cpss_i <- cpss_i |>
  select(where(~ !all(is.na(.)))) |> 
  filter(date_completed <= "2025-12-31")

#score cols
score_cols <- names(cpss_i)[endsWith(names(cpss_i), "_score")]
clinical_cutoff_i <- 31

#make sure columns are numeric
#add a total score column
#make sure it's date formatted
cpss_i_scores <- cpss_i |> 
  mutate(across(ends_with("_score"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(
    missing_count = sum(is.na(c_across(all_of(score_cols)))),
    total_score = rowSums(across(all_of(score_cols)), na.rm = TRUE),
    clinically_relevant = if_else(
        !is.na(total_score) & total_score >= clinical_cutoff_i, TRUE, FALSE)) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  select(client_system_id, collection, date_completed, total_score)
  
cpss_i_paired <- cpss_i_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments = n(),
    initial_date  = first(date_completed),
    initial_score = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed), as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score), NA_real_)
  ) |>
  ungroup() |> 
  filter(
    !is.na(follow_up_date)
  ) |> 
  mutate(
    positive_outcome = ifelse(follow_up_score < initial_score, 1, 0), 
    score_change = follow_up_score - initial_score,
    initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)))
    
# summary stats
tab <- cpss_i_paired |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
cpss_i_paired_long <- cpss_i_paired |>
  ungroup() |> 
    filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) |> 
  select(client_system_id, initial_score, follow_up_score) |> 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
cpss_i_plot <- ggplot(cpss_i_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, CPSS-5 Interviewer Version") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
cpss_i_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_cpss_i <- cpss_i_paired_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) 

summary_stats_cpss_i

# Ridgeline plot
cpss_i_ridge_plot <- ggplot(cpss_i_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  #scale_x_continuous(limits = c(0,6), breaks = c(1, 2, 3, 4, 5)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "CPSS Interviewer Total Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, CPSS Interviewer (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_cpss_i, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.03,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_cpss_i,
            aes(x = median_score, y = 1.035, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

cpss_i_ridge_plot
```

### Self version

```{r}
#| echo: false
#| warning: false
#| message: false
cpss_s_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/CPSS-5-self.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE)
h1 <- unlist(cpss_s_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(cpss_s_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
cpss_s <- cpss_s_raw[-c(1, 2), ]
names(cpss_s) <- new_names
cpss_s <- cpss_s |>
  select(where(~ !all(is.na(.)))) |> 
  filter(date_completed <= "2025-12-31")

#score cols
score_cols <- names(cpss_s)[endsWith(names(cpss_s), "_score")]
clinical_cutoff_s <- 31

#make sure columns are numeric
#add a total score column
#make sure it's date formatted
cpss_s_scores <- cpss_s |> 
  mutate(across(ends_with("_score"), ~ as.numeric(.))) |> 
  rowwise() |> 
  mutate(
    missing_count = sum(is.na(c_across(all_of(score_cols)))),
    total_score = rowSums(across(all_of(score_cols)), na.rm = TRUE),
    clinically_relevant = if_else(
        !is.na(total_score) & total_score >= clinical_cutoff_s, TRUE, FALSE)) |> 
  ungroup() |> 
  mutate(
    date_completed = as.Date(date_completed),
    row_id = row_number()                # stable row id for disambiguation
  ) |> 
  select(client_system_id, collection, date_completed, total_score)
  
cpss_s_paired <- cpss_s_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments = n(),
    initial_date  = first(date_completed),
    initial_score = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed), as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score), NA_real_)
  ) |>
  ungroup() |> 
  filter(
    !is.na(follow_up_date)
  ) |> 
  mutate(
    positive_outcome = ifelse(follow_up_score < initial_score, 1, 0), 
    score_change = follow_up_score - initial_score,
    initial_quarter = paste0("Q", quarter(initial_date), " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)))
    
# summary stats
tab <- cpss_s_paired |> 
  filter(!is.na(positive_outcome)) |> 
  tabyl(positive_outcome) |> 
  adorn_pct_formatting(digits = 0) 

tab 
```

Plot it.

```{r}
#| echo: false
#| warning: false
#| message: false
cpss_s_paired_long <- cpss_s_paired |>
  ungroup() |> 
     filter(!is.na(client_system_id), 
         !is.na(initial_score), 
         !is.na(follow_up_score)) |> 
  select(client_system_id, initial_score, follow_up_score) |> 
  pivot_longer(
    cols = c(initial_score, follow_up_score),
    names_to = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score" = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type,
                        levels = c("Initial", "Follow-up"))
  )

# Plot
cpss_s_plot <- ggplot(cpss_s_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    #binwidth = 1, 
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Score", 
       y = "Count",
       title = "Distribution of Initial and Follow-up Scores, CPSS-5 Self Version") +
  theme_minimal()+
  theme(
     axis.title = element_text(size = 9),
     axis.text = element_text(size = 9),
     panel.grid.major = element_blank(),
     panel.grid.minor.x = element_blank(),
     axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
     legend.position = "inside",
     legend.justification = c("left", "top"),
     legend.title = element_blank(),
     legend.text = element_text(size = 8),
     plot.margin = margin(5, 5, 5, 5)) 
  
cpss_s_plot
```

Ridgeline plot

```{r}
# summary stats by score type
summary_stats_cpss_s <- cpss_s_paired_long |> 
  group_by(score_type) |> 
  summarise(
    mean_score = mean(score_value, na.rm = TRUE), 
    median_score = median(score_value, na.rm = TRUE),
    sd_score = sd(score_value, na.rm = TRUE), 
    n = sum(!is.na(score_value))
  ) 

summary_stats_cpss_s

wilcox.test(cpss_s_paired$follow_up_score, cpss_s_paired$initial_score, paired = TRUE)

# Ridgeline plot
cpss_s_ridge_plot <- ggplot(cpss_s_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha = 0.7,
    scale = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "CPSS Self Total Score", 
       y = NULL,
       title = "Distribution of Initial and Follow-up Scores, CPSS Self (lower is better)") +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x = element_text(size = 9, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position = "inside",
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5),
    plot.title = element_text(size = 11)
  ) + 
  geom_segment(data = summary_stats_cpss_s, 
               aes(x = median_score, y = 1,
                   xend = median_score, yend = 1.03,
                   color = "grey30"), linetype = "dotted") + 
  geom_text(data = summary_stats_cpss_s,
            aes(x = median_score, y = 1.035, 
                label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
                vjust = 2, hjust = 0.5, size = 2)

cpss_s_ridge_plot
```

