---
title: "Rikke's SPSS Master File"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Rikke's SPSS master file

**Access to services**

Among these clients/participants, the largest share (\[XX%\]) received services through a Round 2 program implementing \[EBP\], which \[brief context about the intervention—e.g., target population and core focus\].

Service reach was also concentrated among \[ENTITY TYPE: CBO\] (\[XX%\]) and in Region 8 (Los Angeles County; \[XX%\]).

At the same time, the Workstream served a diverse set of children, youth, and caregivers across provider types. For example, \[XX%\] of clients/participants received services through non-traditional behavioral health providers (e.g., hospitals/hospital systems; \[XX%\]) and tribal entities (\[XX%\]).

```{r setup-dependencies}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(janitor)
library(haven)

# discharge_data_filtered is used later in this chunk; reload it here
# since this file was split from the original annual_report_results.qmd
discharge_data_filtered <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv")
discharge_data_filtered$intervention_start_year <- year(discharge_data_filtered$intervention_start)
discharge_data_filtered <- discharge_data_filtered |>
  filter(!is.na(intervention_start),
         !is.na(intervention_name))

# TODO: demographic_data_filtered is referenced at the bottom of this file
# but was not defined in the original QMD either. Load or define it here
# once the correct source is identified.
```

```{r}
#| echo: false
#| warning: false
#| message: false

# load in spss data
master_grantee_file <- haven::read_spss("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/MasterGranteeSpreadsheet_forBHESQuantTeam_1-20-26_NEW!.sav") |> 
  zap_missing() |>
  mutate(across(where(is.labelled), as.character)) |>
  mutate(across(where(is.character), ~ na_if(str_trim(.x), ""))) |> 
  rename_with(~ gsub("^@", "", .x)) |> 
  janitor::clean_names() |> 
  rename_with(~ gsub("^x", "", .x))

#join to all data
all_data_spss <- left_join(all_data, master_grantee_file, by = c("provider" = "1a_grantee_provider_name", "intervention_name" = "9a_ebpcde"))

#get #'s for report 
all_data_spss |>
  filter(`4funding_round` == "Round 4") |> 
  distinct(client_system_id, provider) |>   # one row per client per round
  tabyl(provider) |>
  adorn_totals("row") |>
  adorn_rounding(digits = 2)

#join to just demographic data - need to add provider name and intervention first to help with join
key <- c("client_system_id", "provider_system_id", "provider_id")

discharge_small <- discharge_data_filtered |>
  select(all_of(key), provider, intervention_name)

discharge_small |>
  count(across(all_of(key)), name = "n") |>
  filter(n > 1)

discharge_unique <- discharge_small |>
  distinct(across(all_of(key)), .keep_all = TRUE)

demo_discharge <- left_join(demographic_data_filtered, discharge_unique, by = key)

demo_spss <- left_join(demo_discharge, master_grantee_file, by = c("provider_id" = "2grantee_provider_id"))

#get #'s on master spss file
demo_spss |> 
  tabyl(`12a_grantee_entity_type_for_report`, `8c_regions`) |> 
  adorn_totals("row") |> 
  adorn_rounding(digits = 2)

demo_spss |> 
  filter(`7a_grant_track` == "Expansion") |> 
  tabyl(`12a_grantee_entity_type_for_report`, `8c_regions`) |> 
  adorn_totals("row") |> 
  adorn_rounding(digits = 2)

master_grantee_file |> 
  #filter(`7a_grant_track` == "Expansion") |> 
  tabyl(`1a_grantee_provider_name`, `8c_regions`) |> 
  adorn_totals("row") |> 
  adorn_rounding(digits = 2)

demo_spss |> 
  tabyl(`12a_grantee_entity_type`) |> 
  adorn_totals("row") |> 
  adorn_rounding(digits = 2)

demo_spss |> 
  tabyl(`8c_regions`) |> 
  adorn_totals("row") |> 
  adorn_rounding(digits = 2)

test <- demo_spss |> 
  filter(is.na(`8c_regions`))

table(test$provider_id)

master_grantee_file |>
  count(`2grantee_provider_id`) |>
  filter(n > 1)
```
