---
title: "Annual Report Data Analysis - Question 3.16"
author: "Cristin Young"
date: "2025-05-15"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

This notebook answers Question 3.16: To what extent did children/youth report reductions in PTSD symptoms/PTSD symptom severity after the intervention?

#Read in data
```{r}
library(tidyverse)
library(janitor)
library(gt)

all_data <- read_csv("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
```

#3.16 To what extent did children/youth report reductions in PTSD symptoms/PTSD symptom severity after the intervention?
CPSS-5: 119
CPSS-5 (interviewer): 42
CATS (caregiver, ages 3-6): 11
CATS-2 (self-report, 7-17): 30

Assumptions: combining each into CPSS and CATS, respectively

##CPSS-5
Scores for the first part of the scale range from 0-80, higher scores indicating greater symptom severity. A cutoff score of 31 or higher in part one of the CPSS-5 is recommended for identifying probable PTSD.
```{r}
table(all_data$pre_assessment_name)
cpss <- all_data %>% 
  filter(trimmed_pre_assessment_name == "CPSS" | trimmed_pre_assessment_name == "CPSS-5")
table(cpss$client_type, useNA = "ifany")

cpss_youth <- cpss %>% 
  filter(client_type == "Youth (16-25)") 

cpss_child <- cpss %>% 
  filter(client_type == "Child (0-15)")

cpss_child_youth <- cpss %>% 
  filter(client_type != "Parent/Caregiver") 
```

##CATS
NOTE: NOT USING CATS UNTIL WE GET BETTER AGE RANGE DATA FROM EINSIGHT
Ages 3-6: The total symptom score is calculated by summing up the items 1-16 (possible range = 0-48). We recommend to use a cut-off ≥ 16 as indication of a clinically relevant level of symptoms in preschool children.
Ages 7-17: A total symptom score is calculated by summing up the raw scores of items 1-20 (possible range = 0-60). We recommend to use a cut-off ≥ 21 as indication of a clinically relevant level of symptoms.
```{r}
cats <- all_data %>% 
  filter(trimmed_pre_assessment_name == "CATS" | trimmed_pre_assessment_name == "CATS-2")
```


##Checking missing data

How much data is missing? 

- 60% (n=91/161) of rows have a pre but no post score

- 0 rows have a post score but no pre score

- 40% (n=60/151) of rows have both a pre and post score

```{r}
pre_no_post <- cpss %>% 
  filter(is.na(trimmed_post_assessment_name))

#checking there's no opposite to above (yes post, no pre score)
post_no_pre <- cpss %>% 
  filter(is.na(trimmed_pre_assessment_name) & !is.na(trimmed_post_assessment_name))

pre_post <- cpss %>% 
  filter(!is.na(trimmed_pre_assessment_name) & !is.na(trimmed_post_assessment_name))
```

##Do we have any duplicates?
Had a second follow-up - keeping only second follow-up
```{r}
#how many duplicates are there?
n_occur <- data.frame(table(pre_post$client_id))
#how many times do they occur?
n_occur[n_occur$Freq > 1,]

pre_post <- pre_post %>% 
  filter(!(client_id == "7406" & post_time_point == "Follow up 1"))
```

#CPSS cutoffs
- Lower score is better
- A cutoff score of 31 or higher in part one of the CPSS-5 is recommended for identifying probable PTSD.

##Creating positive outcome and cutoff variables
```{r}
pre_post <- pre_post %>% 
  mutate(positive_outcome = case_when(
    pre_score > post_score ~ "Yes",
    TRUE ~ "No"),
    ptsd_cutoff = case_when(
      post_score >= 31 ~ 1,
      post_score < 31 ~ 0))
```

#how many positive outcomes?
50 / 59 = 85%
Only 3 / 9 without positive outcomes had >3 difference in scores 
```{r}
table(pre_post$positive_outcome)
pre_post %>% 
  filter(positive_outcome == "No") %>% 
  select(pre_score, post_score, positive_outcome)
```

#how many above clinical cutoff at discharge?
Clinical cutoff:
15 individuals above clinical cutoff at discharge, spread between 31 and 64.  
```{r}
clin_cutoff <- pre_post %>% 
  filter(ptsd_cutoff == 1) %>% 
  select(post_score)
hist(clin_cutoff$post_score)

pre_post %>% 
  filter(ptsd_cutoff == 0)
```

##mean and SD of pre and post scores
```{r}
#mean and sd of this data
summary_pre_score <- pre_post %>%
  summarise(
    mean_score = mean(pre_score, na.rm = TRUE),
    median_score = median(pre_score, na.rm = TRUE),
    sd_score = sd(pre_score, na.rm = TRUE)
  )

summary_pre_score
summary_pre_score %>% 
  knitr::kable() %>% 
  clipr::write_clip()

summary_post_score <- pre_post %>%
  summarise(
    mean_score = mean(post_score, na.rm = TRUE),
    median_score = median(post_score, na.rm = TRUE),
    sd_score = sd(post_score, na.rm = TRUE)
  )

summary_post_score
summary_post_score %>% 
  knitr::kable() %>% 
  clipr::write_clip()
```

##do a t test on the data
```{r}
t.test(pre_post$pre_score, pre_post$post_score, paired = TRUE)
```