---
title: "YOQ 30.2 - Data Quality Scoring Check"
author: "Cristin Young"
date: "2025-04-25"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---
TO DO
[ ] add duplicate into issues gsheet from Lauren
[ ] check no meaningful change people --> is there a difference for these people here? 
    - is it only in certain orgs? 
    - what's the breakdown across orgs?
[ ] RCI --> stratify by above clinical cutoff at pre score (at intake)
[ ] RCI --> stratify by above vs below clinical cutoff



This code uses the discharge export file found here: https://www.dropbox.com/scl/fi/h5dbek0m613ly6nribzou/Discharge-Exploration.xlsx?rlkey=a7edwvrhfcifvnep1omy4mygd&dl=0 

And checks: 

1. That a decrease in score is a positive outcome
2. Whether a reliable change occurred (+/- 10 points change in score)

#**Read in data**

- Filter pre-measure to YOQ
```{r, include= FALSE}
library(tidyverse)
discharge_data <- readxl::read_xlsx("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/Exploration Set (edited)/2025_04/Discharge Exploration-all data.xlsx", na = c("NA", "-", ""))
yoq <- discharge_data %>% 
  filter(`Pre Measure(s)` == "YOQ 30.2 Caregiver Initial" |
         `Pre Measure(s)` == "YOQ 30.2 Self Report Initial")
```

#**Missing data**

How much data is missing? 

- 10% (482/4682) of rows are YOQ data

- 75% (n=363) of rows have a pre but no post score

- 0 rows have a post score but no pre score

- 25% (n=119) of rows have both a pre and post score
```{r, include = FALSE}
pre_no_post <- yoq %>% 
  filter(is.na(`Post Score`))

#checking there's no opposite to above (yes post, no pre score)
post_no_pre <- yoq %>% 
  filter(is.na(`Pre Score`) & !is.na(`Post Score`))

pre_post <- yoq %>% 
  filter(!is.na(`Pre Score`) & !is.na(`Post Score`))
```

#**Do we have any duplicates?**

Looks like one person (ID: 13042) took both versions of the YOQ (why??)
We don't have to exclude the other two in this analysis but should in overall, because it's the same score for two enrollments, effectively
```{r}
#how many duplicates are there?
n_occur <- data.frame(table(pre_post$`Client ID`))
#how many times do they occur?
n_occur[n_occur$Freq > 1,]
#what's the data? 
duplicates <- pre_post[pre_post$`Client ID` %in% n_occur$Var1[n_occur$Freq > 1],]
duplicates
```

#**Is a decrease in score always scored as "positive outcome"?**

Thankfully, yes. No wrong scoring here!
```{r}
pre_post %>% 
  mutate(result = case_when(
    `Pre Score` < `Post Score` & `Positive Outcomes` == "Yes" ~ FALSE,
    TRUE ~ TRUE
  )) %>% 
  filter(result == FALSE)
```

#**How many positive outcomes are there?**

64% have positive outcomes
```{r}
table(pre_post$`Positive Outcomes`)
77/119
```

#**Clinical cutoff**
##How many people were above the clinical cutoff at discharge?

Clinical cutoff:

  - Self-report: ≥ 30 --> n=16 (from n = 28)
  - Parent/guardian: ≥ 29 --> n=13 (from n = 20)
```{r}
caregiver <- pre_post %>% 
  filter(`Pre Measure(s)` == "YOQ 30.2 Caregiver Initial",
         `Post Score` >= 29)

self <- pre_post %>% 
  filter(`Pre Measure(s)` == "YOQ 30.2 Self Report Initial",
         `Post Score` >= 30)

test <- pre_post %>% 
  filter(`Pre Measure(s)` == "YOQ 30.2 Caregiver Initial",
        `Pre Score` >= 29)
```

#**RCI**

Let's create a reliable change index to see how many really matter
```{r, include = FALSE}
pre_post <- pre_post %>% 
  mutate(rci = `Pre Score` - `Post Score`)
```


**How many of these scores changed for the better or worse?**

RCI ≤-10 or ≥10 is a reliable change

- 51% (39/77) saw reliable change
- 43% (n=33) did reliably better
- 8% (n=6) did reliably worse
```{r, include = FALSE}
reliable_change <- pre_post %>% 
  filter(rci >= 10 | rci <= -10)

#these people did reliably better
pos_rci <- reliable_change %>% 
  filter(rci >= 10)

#these people did reliably worse
neg_rci <- reliable_change %>% 
  filter(rci <= -10)
```

#**Some plots for fun**

What's the spread?
```{r}
#conditional color formatting
pre_post$rci_col <- as.factor((pre_post[,17] > 10 | pre_post[,17] <= -10)*1)

pre_post %>% 
  ggplot(aes(rci)) + 
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 4) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))+
  labs(x = "Reliable Change Index") +
  scale_fill_manual("Reliable Change Index",
                    labels = c("no meaningful change (n=38)", "meaningful change (n=39)"), 
                    values = c("#ABDDDE", "#046C9A"))
```


