---
title: "CDEP Report"
author: "Cristin Young"
date: "`r Sys.Date()`"
format:
  html:
    df-print: paged
    toc: true
    toc-depth: 3
    toc-location: left
    theme: cosmo
    embed-resources: TRUE
execute:
  echo: false
  warning: false
  message: false
---

::: {.callout-note title="TO DOs"}
- get the correct grantee_IDs and double check results match Rikke's hand-curated file from eInsight
- filter for grantees that have multiple interventions and ONE grant
- clinical summary: had intervention (should be discharge_data)
:::

## CDEPs

81 CDEPs are the focus of this report.

Notes:

- so if it doesn't match, it's because it's NOT A CDEP
- match by funding round too
- how many CDEPs have more than one grant? 
- go SOLELY based on Rikke's crosswalk and not match first. 
- ok, so the mismatched ones are all ones that have multiple grants and they're putting the wrong thing in the dropdown

## Setup

```{r}
#| label: setup

library(readr)
library(tidyverse)
library(janitor)
library(haven)
library(stringr)

knitr::opts_chunk$set(rownames.print = FALSE)

#discharge_data    <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv")
#all_data          <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv")
#demographic_data  <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv")

#cdep_sav <- read_sav("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Master Files BHES Internal/Other File Versions/CDEPS_3-18-26.sav")

# intervention_name_crosswalk <- read_csv(
#   "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/intervention_name_review_RA_3-19-26.csv",
#   show_col_types = FALSE
# )

# intervention_name_crosswalk <- read_csv("/Users/cristin/Downloads/intervention_name_crosswalk_updated_RA.csv",
#   show_col_types = FALSE
# )  
# 
# # Clean crosswalk: remove data-entry artifacts, coerce provider_id to character
# crosswalk_clean <- intervention_name_crosswalk |>
#   filter(
#     !is.na(discharge_data_intervention_name),
#     discharge_data_intervention_name != "added this row",
#     !is.na(cdep_sav_intervention_name)
#   ) |>
#   mutate(provider_id = str_pad(as.character(provider_id), width = 3, pad = "0")) |>
#   distinct(
#     provider_id,
#     intervention_name          = discharge_data_intervention_name,
#     cdep_sav_intervention_name
#   )
# 
# # Step 1: direct matches — discharge intervention_name == cdep_sav @9aEBPCDEP
# direct_pairs <- discharge_data |>
#   distinct(provider_id, intervention_name) |>
#   inner_join(cdep_lookup, by = c("provider_id", "intervention_name" = "sav_name")) |>
#   distinct(provider_id, intervention_name)
# 
# # Step 2: crosswalk-bridged matches for names that don't align
# crosswalk_pairs <- discharge_data |>
#   distinct(provider_id, intervention_name) |>
#   anti_join(direct_pairs, by = c("provider_id", "intervention_name")) |>
#   inner_join(crosswalk_clean, by = c("provider_id", "intervention_name")) |>
#   inner_join(cdep_lookup, by = c("provider_id", "cdep_sav_intervention_name" = "sav_name")) |>
#   distinct(provider_id, intervention_name)
# 
# cdep_intervention_pairs <- bind_rows(direct_pairs, crosswalk_pairs)
# 
# # Pass 1: unambiguous corrections — one discharge name maps to exactly one correct name
# name_correction <- intervention_name_crosswalk |>
#   filter(
#     !is.na(discharge_data_intervention_name),
#     discharge_data_intervention_name != "added this row",
#     !is.na(correct_intervention_name),
#     correct_intervention_name != ""
#   ) |>
#   mutate(provider_id = str_pad(as.character(provider_id), width = 3, pad = "0")) |>
#   group_by(provider_id, discharge_data_intervention_name) |>
#   filter(n_distinct(correct_intervention_name) == 1) |>
#   ungroup() |>
#   distinct(
#     provider_id,
#     intervention_name         = discharge_data_intervention_name,
#     correct_intervention_name
#   )
# 
# # Pass 2: funding-round-disambiguated corrections for remaining ambiguous cases
# # (cases where the same discharge name maps to different interventions across funding rounds)
# ambig_correction_by_round <- intervention_name_crosswalk |>
#   filter(
#     !is.na(discharge_data_intervention_name),
#     discharge_data_intervention_name != "added this row",
#     !is.na(correct_intervention_name),
#     correct_intervention_name != ""
#   ) |>
#   mutate(provider_id = str_pad(as.character(provider_id), width = 3, pad = "0")) |>
#   group_by(provider_id, discharge_data_intervention_name) |>
#   filter(n_distinct(correct_intervention_name) > 1) |>
#   ungroup() |>
#   mutate(funding_round = str_extract(spss_grant_id, "(?<=-)(\\d)")) |>
#   group_by(provider_id, discharge_data_intervention_name, funding_round) |>
#   filter(n_distinct(correct_intervention_name) == 1) |>
#   ungroup() |>
#   distinct(
#     provider_id,
#     intervention_name     = discharge_data_intervention_name,
#     funding_round,
#     correct_intervention_name
#   )
# 
# apply_name_corrections <- function(df) {
#   df |>
#     left_join(name_correction,
#               by = c("provider_id", "intervention_name")) |>
#     left_join(ambig_correction_by_round,
#               by = c("provider_id", "intervention_name", "funding_round"),
#               suffix = c("_base", "_round")) |>
#     mutate(intervention_name = coalesce(correct_intervention_name_base,
#                                         correct_intervention_name_round,
#                                         intervention_name)) |>
#     select(-correct_intervention_name_base, -correct_intervention_name_round)
# }
# 
# discharge_cdep <- discharge_data |>
#   semi_join(cdep_intervention_pairs, by = c("provider_id", "intervention_name")) |>
#   apply_name_corrections()
# 
# # demographic data has no intervention_name — filter by provider_id only
# demo_cdep <- demographic_data |>
#   filter(provider_id %in% cdep_lookup$provider_id)
# 
# all_data_cdep <- all_data |>
#   semi_join(cdep_intervention_pairs, by = c("provider_id", "intervention_name")) |>
#   apply_name_corrections()

# using Rikke's files 
discharge_data_cdep    <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/CDEP Files for Cristin _March 25 2026/Final File/Excel Files/DischargeExport_CDEPs_March 25 2026_Final.csv", na = c("-", "", "NA", "N/A")) |> clean_names() |> 
    mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id),
    provider_id        = as.character(provider_id)
  )

clinical_summary_cdep  <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/CDEP Files for Cristin _March 25 2026/Final File/Excel Files/ClinicalSummary_CDEPS_File1_March 25, 2026_FINAL.csv", na = c("-", "", "NA", "N/A")) |> clean_names()

demographic_data_cdep  <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/CDEP Files for Cristin _March 25 2026/Final File/Excel Files/ClientDemographis_CDEPS_March 25_Final.csv", na = c("-", "", "NA", "N/A")) |> clean_names() |> 
  rename(client_system_id = client_id) |> 
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id),
    provider_id        = as.character(provider_id)
  )

# need to clean demographic and discharge data and recode race, etc. 
codebook_race <- c(
  "Asian"                                                = 1,
  "Black or African American"                            = 2,
  "Hispanic or Latino"                                   = 3,
  "Middle Eastern or North African"                      = 4,
  "Native American/American Indian or Alaska Native"     = 5,
  "Native Hawaiian or Pacific Islander"                  = 6,
  "White"                                                = 7
)

demographic_data_cdep <- demographic_data_cdep |>
  mutate(
    cleaned_race = str_remove_all(
      race,
      "\\b(Don't Know|Prefer Not To Answer|Not Applicable)\\b"
    ) |>
      str_replace_all(",+", ",") |>       # collapse extra commas left by removal
      str_replace_all("^,|,$", "") |>     # trim any leading/trailing commas
      na_if("")                            # empty string → NA
  ) |>
  mutate(
    recoded_race = case_when(
      is.na(cleaned_race)              ~ NA,   # no valid response
      str_detect(cleaned_race, ",")    ~ 2L,   # multiracial
      TRUE                             ~ 1L    # single race
    ),
    code_num = case_when(
      is.na(cleaned_race)              ~ NA,
      str_detect(cleaned_race, ",")    ~ 88L,  # multiracial code
      TRUE ~ unname(codebook_race[cleaned_race])
    )
  )

codebook_eth <- c(
  "Chinese" = 1, "Asian Indian" = 2, "Filipino" = 3, "Vietnamese" = 4,
  "Korean" = 5, "Japanese" = 6, "African American" = 7, "Jamaican" = 8,
  "Haitian" = 9, "Nigerian" = 10, "Somali" = 11, "Mexican" = 12,
  "Puerto Rican" = 13, "Salvadoran" = 14, "Cuban" = 15, "Dominican" = 16,
  "Guatemalan" = 17, "Native Hawaiian" = 18, "Samoan" = 19, "Chamorro" = 20,
  "English" = 21, "German" = 22, "Irish" = 23, "Italian" = 24, "Polish" = 25,
  "Scottish" = 26, "2+ ethnicities selected" = 88
)

demographic_data_cdep <- demographic_data_cdep |>
  mutate(
    cleaned_ethn = str_remove_all(
      ethnicity,
      "\\b(Don't Know|Prefer Not To Answer|Not Applicable)\\b"
    ) |>
      str_replace_all(",+", ",") |>
      str_replace_all("^,|,$", "") |>
      na_if("")
  ) |>
  mutate(
    recoded_ethn = case_when(
      is.na(cleaned_ethn)              ~ NA,
      str_detect(cleaned_ethn, ",")    ~ 2L,
      TRUE                             ~ 1L
    ),
    ethn_num = case_when(
      is.na(cleaned_ethn)              ~ NA,
      str_detect(cleaned_ethn, ",")    ~ 88L,
      TRUE ~ unname(codebook_eth[cleaned_ethn])
    )
  )

discharge_data_cdep <- discharge_data_cdep |>
  mutate(
    funding_round = case_when(
      provider_id == "604" & funding_round == "No Funding Round Selected" ~ "2",
      provider_id == "508" & funding_round == "No Funding Round Selected" ~ "2",
      provider_id == "607" & funding_round == "No Funding Round Selected" ~ "1",
      provider_id == "573"                                                 ~ "2",
      provider_id == "709" & funding_round == "No Funding Round Selected" ~ "1",
      TRUE ~ funding_round
    )
  )


discharge_data_cdep <- discharge_data_cdep |>
  mutate(
    pre_time_point      = str_extract(pre_measures, "(Initial|(?i)follow up \\d+)"),
    pre_assessment_name = str_remove(pre_measures, "(Initial|(?i)follow up \\d+)") |>
      str_trim()
  ) |>
  mutate(
    post_time_point      = str_extract(post_measures, "(Initial|(?i)follow up \\d+)"),
    post_assessment_name = str_remove(post_measures, "(Initial|(?i)follow up \\d+)") |>
      str_trim()
  ) |>
  mutate(
    # Exact match check
    MatchDistinct = pre_assessment_name == post_assessment_name,
    # First-word match — handles cases where "SDQ" appears as "SDQ Strengths & Difficulties..."
    trimmed_pre_assessment_name  = word(pre_assessment_name, 1),
    trimmed_post_assessment_name = word(post_assessment_name, 1),
    trimmed_match = trimmed_pre_assessment_name == trimmed_post_assessment_name
  ) 

all_data_cdep <- left_join(discharge_data_cdep, demographic_data_cdep, by = c("client_system_id", "provider_id", "provider_system_id"))

cat("Unique providers (clinical summary):", clinical_summary_cdep |> distinct(provider_id) |> nrow(), "\n")
cat("Unique providers (discharge data):", discharge_data_cdep |> distinct(provider_id) |> nrow(), "\n")
cat("Discharge data (distinct by client ID) rows; this is the denominator for most tables below:", discharge_data_cdep |> distinct(client_system_id, .keep_all = TRUE) |> nrow(), "\n")
cat("Discharge data rows (total):", nrow(discharge_data_cdep), "\n")
cat("Clinical summary (with collection, distinct by client ID) rows:", clinical_summary_cdep |> filter(!is.na(collection)) |>  distinct(client_system_id, .keep_all = TRUE) |> nrow(), "\n")
cat("Clients reached (demographic data, distinct by client ID):", demographic_data_cdep |> distinct(client_system_id, .keep_all = TRUE) |> nrow(), "\n")
```

## Results

### Enrollment

#### Grantees by Funding Round
**data used:** discharge_data
```{r}
#| label: tbl-grantees-by-round

discharge_data_cdep |>
  distinct(provider, .keep_all = TRUE) |>
  tabyl(funding_round) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Clients Enrolled by Funding Round
**Assumptions:** Needed to have an intervention start date; data used: discharge_data
```{r}
#| label: tbl-enrolled-by-round

discharge_data_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(funding_round) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Clients Enrolled by Intervention
**Assumptions:** Needed to have an intervention start date; data used: discharge_data
```{r}
#| label: tbl-enrolled-by-intervention

discharge_data_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(intervention_name) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Clients Enrolled by Intervention and Funding Round
**Assumptions:** Needed to have an intervention start date; data used: discharge_data
```{r}
#| label: tbl-enrolled-by-intervention-round

discharge_data_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(intervention_name, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

### Discharge

#### Discharged by Intervention
**Assumptions:** Needed to have an intervention end date; data used: discharge_data
```{r}
#| label: tbl-discharged-by-intervention

discharge_data_cdep |>
  filter(!is.na(intervention_end)) |>
  tabyl(intervention_name) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Discharged by Reason
**Assumptions:** Needed to have an intervention end date; data used: discharge; data
```{r}
#| label: tbl-discharged-by-reason

discharge_data_cdep |>
  filter(!is.na(intervention_end)) |>
  tabyl(reason_for_discharge) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Discharge Reason by Funding Round
**Assumptions:** Needed to have an intervention START date -- this is because the data is broken out by funding round, not the funding round in which the intervention ended; data used: discharge_data
```{r}
#| label: tbl-discharge-reason-by-round

discharge_data_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(reason_for_discharge, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Discharge Reason by Intervention
**Assumptions:** Needed to have an intervention start date and a reason for discharge; data used: discharge_data
```{r}
#| label: tbl-discharge-reason-by-intervention

discharge_data_cdep |>
  filter(!is.na(intervention_start),
         !is.na(reason_for_discharge)) |>
  tabyl(reason_for_discharge, intervention_name) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

### Client Type

#### Client Type Overall
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-client-type

all_data_cdep |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Client Type by Funding Round
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-client-type-by-round

all_data_cdep |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Client Type by Intervention
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-client-type-by-intervention

all_data_cdep |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(intervention_name, client_type) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

## Client Demographics

### Primary Language
**Assumptions:** Distinct by client system ID, Don't know, no primary language selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-primary-language

all_data_cdep |>
  filter(
    primary_language != "Don't Know",
    primary_language != "No Primary Language Selected",
    primary_language != "Not Applicable",
    primary_language != "Prefer Not To Answer",
    !is.na(primary_language)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(primary_language) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Housing Status
**Assumptions:** Distinct by client system ID, Don't know, no housing status selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-housing-status

all_data_cdep |>
  filter(
    housing_status != "Don't Know",
    housing_status != "No Housing Status Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(housing_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(housing_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Client Type with Average Age
**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-client-type-age

all_data_cdep_distinct <- all_data_cdep |>
  mutate(
    age_start_tx_num = case_when(
      ageat_startof_treatmentin_years %in% c("No Age Provided", "", " ") ~ NA_character_,
      TRUE ~ ageat_startof_treatmentin_years
    ) |> as.numeric()
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

tab_counts <- all_data_cdep_distinct |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

tab_age <- all_data_cdep_distinct |>
  group_by(client_type) |>
  summarise(
    mean_age   = round(mean(age_start_tx_num, na.rm = TRUE), 1),
    n_with_age = sum(!is.na(age_start_tx_num)),
    .groups = "drop"
  )

tab_counts |>
  left_join(tab_age, by = "client_type") |>
  select(client_type, n, percent, mean_age, n_with_age)
```

### Insurance Status
**Assumptions:** Distinct by client system ID, Don't know, no insurance status selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-insurance-status

all_data_cdep |>
  filter(
    insurance_status != "Don't Know",
    insurance_status != "No Insurance Status Selected",
    insurance_status != "Not Applicable",
    insurance_status != "Prefer Not To Answer",
    !is.na(insurance_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(insurance_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Gender Identity
**Assumptions:** Distinct by client system ID, Don't know, no gender identity selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-gender-identity

all_data_cdep |>
   filter(
    gender_identity != "Don't Know",
    gender_identity != "No Gender Identify Selected",
    gender_identity != "Not Applicable",
    gender_identity != "Prefer Not To Answer",
    !is.na(gender_identity)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(gender_identity) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Sexual Orientation
**Assumptions:** Distinct by client system ID, Don't know, no sexual orientation selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-sexual-orientation

all_data_cdep |>
   filter(
    sexual_orientation != "Don't Know",
    sexual_orientation != "No Sexual Orientation Selected",
    sexual_orientation!= "Not Applicable",
    sexual_orientation != "Prefer Not To Answer",
    !is.na(sexual_orientation)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(sexual_orientation, client_type) |> 
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Race

Race is coded numerically as follows:

| Code | Race                                               |
|------|----------------------------------------------------|
| 1    | Asian                                              |
| 2    | Black or African American                          |
| 3    | Hispanic or Latino                                 |
| 4    | Middle Eastern or North African                    |
| 5    | Native American / American Indian or Alaska Native |
| 6    | Native Hawaiian or Pacific Islander                |
| 7    | White                                              |
| 88   | 2+ races selected                                  |
**Assumptions:** Distinct by client system ID, Don't know, no race selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-race

all_data_cdep |>
  filter(
    race != "Don't Know",
    race != "No Race Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(code_num)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(code_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Ethnicity

Ethnicity is coded numerically as follows:

| Code | Ethnicity               |
|------|-------------------------|
| 1    | Chinese                 |
| 2    | Asian Indian            |
| 3    | Filipino                |
| 4    | Vietnamese              |
| 5    | Korean                  |
| 6    | Japanese                |
| 7    | African American        |
| 8    | Jamaican                |
| 9    | Haitian                 |
| 10   | Nigerian                |
| 11   | Somali                  |
| 12   | Mexican                 |
| 13   | Puerto Rican            |
| 14   | Salvadoran              |
| 15   | Cuban                   |
| 16   | Dominican               |
| 17   | Guatemalan              |
| 18   | Native Hawaiian         |
| 19   | Samoan                  |
| 20   | Chamorro                |
| 21   | English                 |
| 22   | German                  |
| 23   | Irish                   |
| 24   | Italian                 |
| 25   | Polish                  |
| 26   | Scottish                |
| 88   | 2+ ethnicities selected |
**Assumptions:** Distinct by client system ID, Don't know, no ethnicity selected, not applicable, and prefer not to answer all grouped as NA (so total n with NA = 8105); data used: all_data (demo + discharge data); must have intervention name
```{r}
#| label: tbl-ethnicity

all_data_cdep |>
   filter(
    race != "Don't Know",
    race != "No Ethnicity Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(ethn_num)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(ethn_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

## Satisfaction Surveys
**Assumptions:** Distinct by client system ID and joined to satisfaction survey data; must have an intervention named; data used: satisfaction survey and all data (demo+discharge joined)
```{r}
#| label: setup-satisfaction
#| include: false

sat_all <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/combined_satisfaction_clean.csv",
  show_col_types = FALSE
)

# Unified Q1 score — coalesce across three survey-specific Q1 columns
sat_all <- sat_all |>
  mutate(
    q1_score = coalesce(
      x1_overall_i_am_satisfied_with_the_services_my_child_received_score,
      x1_overall_i_am_satisfied_with_the_services_i_received_score,
      x1_i_like_the_services_that_i_received_here_score
    )
  )

all_data_cdep_distinct <- all_data_cdep |>
  distinct(client_system_id, .keep_all = TRUE) 

sat_cdep <- sat_all |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  ) |>
  left_join(all_data_cdep_distinct,
            by = c("provider", "provider_id", "provider_system_id", "client_system_id")) |>
  filter(!is.na(intervention_name))

score_cols <- names(sat_cdep)[str_detect(names(sat_cdep), "_score$")]

# Overall Q1 high satisfaction rate
pct_4_5 <- sat_cdep |>
  summarise(
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1)
  ) |>
  pull(pct_4_5)

# Overall combined survey high satisfaction rate (>= 80% of questions at 4 or 5)
df2 <- sat_cdep |>
  select(all_of(score_cols)) |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(everything(), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree    = sum(c_across(everything()) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  ungroup() |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )

df2
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree." The second is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed. High satisfaction on question 1 is defined as a score of 4 or 5.

### Question 1 Satisfaction

Overall, `{r} pct_4_5`% of survey respondents were highly satisfied based on question 1. The following table shows high satisfaction by intervention:

```{r}
#| label: tbl-q1-satisfaction

sat_cdep |>
  group_by(intervention_name) |>
  summarise(
    n_total = sum(!is.na(q1_score)),
    n_4_5   = sum(q1_score %in% c(4, 5), na.rm = TRUE),
    pct_4_5 = round(100 * n_4_5 / n_total, 1)
  )
```

### Combined Survey Satisfaction

Overall, `{r} df2$pct_meeting_80pct`% of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed. The following table shows high combined satisfaction by intervention, for clients where both a client ID and intervention name were available:
**Assumptions:** Must have an intervention name; data used: satisfaction survey and all data (demo+discharge joined)
```{r}
#| label: tbl-combined-satisfaction

sat_cdep |>
  filter(!is.na(intervention_name)) |>
  group_by(intervention_name) |>
  select(intervention_name, all_of(score_cols)) |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  mutate(across(all_of(score_cols), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(all_of(score_cols)))),
    n_agree    = sum(c_across(all_of(score_cols)) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )
```

### Services in Preferred Language

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** The percent of clients that report services received in their preferred language out of the total that completed the satisfaction survey.

**Data used:** satisfacation survey and all data (demo+discharge joined)
```{r}
#| label: tbl-preferred-language

sat_cdep |>
  arrange(date_clean) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label))) |>
  rename(preferred_lang_response = were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |>
  tabyl(preferred_lang_response) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

## Participant Outcomes

### Improved Outcomes

```{r}
#| label: setup-outcomes
#| include: false

assessment_rules <- tibble::tribble(
  ~trimmed_pre_assessment_name, ~higher_is_better, ~min_change,
  "Basis-24",  FALSE, 0.001, # weighted 0–4 scale; any positive change counts
  "CATS",      FALSE, 1,
  "GAD-7",     FALSE, 1,
  "NOMS",      TRUE,  1,
  "Pediatric", FALSE, 1,
  "PHQ-9",     FALSE, 1,
  "PSS",       FALSE, 1,
  "SDQ",       FALSE, 1,
  "YOQ",       FALSE, 1
)

discharge_cdep_outcome <- discharge_data_cdep |>
  left_join(assessment_rules, by = "trimmed_pre_assessment_name") |>
  mutate(
    raw_change    = post_score - pre_score,
    better_change = if_else(higher_is_better, raw_change, -raw_change),
    min_change    = coalesce(min_change, 0),
    outcome = case_when(
      is.na(pre_score) | is.na(post_score) ~ NA_character_,
      better_change >=  min_change ~ "improved",
      better_change <= -min_change ~ "worsened",
      TRUE                         ~ "no_change"
    ),
    outcome = factor(outcome, levels = c("worsened", "no_change", "improved"))
  )
```

#### Outcome Overall
**Assumptions:** Must have an intervention name and an outcome; data used: discharge_data
```{r}
#| label: tbl-outcome-overall

discharge_cdep_outcome |>
  tabyl(outcome) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Outcome by Intervention
**Assumptions:** Must have an intervention name and an outcome; data used: discharge_data
```{r}
#| label: tbl-outcome-by-intervention

discharge_cdep_outcome |>
  filter(!is.na(intervention_name), !is.na(outcome)) |>
  tabyl(intervention_name, outcome) |>
  adorn_totals("row") |>
  adorn_totals("col") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Outcome by Funding Round
**Assumptions:** Must have an intervention name and an outcome; data used: discharge_data
```{r}
#| label: tbl-outcome-by-round

discharge_cdep_outcome |>
  tabyl(outcome, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Outcome by Assessment Measure
**Assumptions:** Must have an intervention name and an outcome; data used: discharge_data
```{r}
#| label: tbl-outcome-by-measure

discharge_cdep_outcome |>
  filter(!is.na(outcome)) |>
  tabyl(post_assessment_name, outcome) |>
  adorn_totals("row") |>
  adorn_totals("col") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

### Treatment Goals Met

Two criteria must be met for a client to have met treatment goals: (1) a discharge reason of "Mutually agreed to cessation of treatment" and (2) an improvement according to pre- and post-measures collected (e.g., NOMS sub-scales). Improvement is determined by "improved" in the outcomare variable, compared with "no change" or "worsened."
**Assumptions:** Must have an intervention name and an outcome; data used: discharge_data
```{r}
#| label: tbl-treatment-goals

discharge_cdep_outcome |>
  filter(outcome %in% c("improved")) |>
  tabyl(reason_for_discharge) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```
