---
title: "CDEP Report"
author: "Cristin Young"
date: "2026-03-18"
format:
  html:
    df-print: paged
    toc: true
    toc-depth: 3
    toc-location: left
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

::: {.callout-note title="TO DOs"}
-   
:::

## CDEPs

81 CDEPs are the focus of this report.

## Setup

```{r}
#| label: setup

library(readr)
library(tidyverse)
library(janitor)
library(haven)

knitr::opts_chunk$set(rownames.print = FALSE)

discharge_data    <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv")
all_data          <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv")
demographic_data  <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv")

cdep_sav <- read_sav("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Master Files BHES Internal/Other File Versions/CDEPS_3-18-26.sav")

cdep_lookup <- cdep_sav |>
  select(provider_id = `@2GranteeProviderID`, sav_name = `@9aEBPCDEP`)

# intervention_name_crosswalk <- read_csv(
#   "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/intervention_name_review_RA_3-19-26.csv",
#   show_col_types = FALSE
# )

intervention_name_crosswalk <- read_csv("/Users/cristin/Downloads/intervention_name_crosswalk_updated_RA.xlsx",
  show_col_types = FALSE
)

write_csv(intervention_name_crosswalk, "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/intervention_name_review_more.csv")

# Clean crosswalk: remove data-entry artifacts, coerce provider_id to character
crosswalk_clean <- intervention_name_crosswalk |>
  filter(
    !is.na(discharge_data_intervention_name),
    discharge_data_intervention_name != "added this row",
    !is.na(cdep_sav_intervention_name)
  ) |>
  mutate(provider_id = str_pad(as.character(provider_id), width = 3, pad = "0")) |>
  distinct(
    provider_id,
    intervention_name          = discharge_data_intervention_name,
    cdep_sav_intervention_name
  )

# Step 1: direct matches — discharge intervention_name == cdep_sav @9aEBPCDEP
direct_pairs <- discharge_data |>
  distinct(provider_id, intervention_name) |>
  inner_join(cdep_lookup, by = c("provider_id", "intervention_name" = "sav_name")) |>
  distinct(provider_id, intervention_name)

# Step 2: crosswalk-bridged matches for names that don't align
crosswalk_pairs <- discharge_data |>
  distinct(provider_id, intervention_name) |>
  anti_join(direct_pairs, by = c("provider_id", "intervention_name")) |>
  inner_join(crosswalk_clean, by = c("provider_id", "intervention_name")) |>
  inner_join(cdep_lookup, by = c("provider_id", "cdep_sav_intervention_name" = "sav_name")) |>
  distinct(provider_id, intervention_name)

cdep_intervention_pairs <- bind_rows(direct_pairs, crosswalk_pairs)

discharge_cdep <- discharge_data |>
  semi_join(cdep_intervention_pairs, by = c("provider_id", "intervention_name"))

# demographic data has no intervention_name — filter by provider_id only
demo_cdep <- demographic_data |>
  filter(provider_id %in% cdep_lookup$provider_id)

all_data_cdep <- all_data |>
  semi_join(cdep_intervention_pairs, by = c("provider_id", "intervention_name"))
```

## Results

### Enrollment

#### Grantees by Funding Round

```{r}
#| label: tbl-grantees-by-round

discharge_cdep |>
  distinct(provider, .keep_all = TRUE) |>
  tabyl(funding_round) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Clients Enrolled by Funding Round

```{r}
#| label: tbl-enrolled-by-round

discharge_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(funding_round) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Clients Enrolled by Intervention

```{r}
#| label: tbl-enrolled-by-intervention

discharge_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(intervention_name) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Clients Enrolled by Intervention and Funding Round

```{r}
#| label: tbl-enrolled-by-intervention-round

discharge_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(intervention_name, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

### Discharge

#### Discharged by Intervention

```{r}
#| label: tbl-discharged-by-intervention

discharge_cdep |>
  filter(!is.na(intervention_end)) |>
  tabyl(intervention_name) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Discharged by Reason

```{r}
#| label: tbl-discharged-by-reason

discharge_cdep |>
  filter(!is.na(intervention_end)) |>
  tabyl(reason_for_discharge) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Discharge Reason by Funding Round

```{r}
#| label: tbl-discharge-reason-by-round

discharge_cdep |>
  filter(!is.na(intervention_start)) |>
  tabyl(reason_for_discharge, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Discharge Reason by Intervention

```{r}
#| label: tbl-discharge-reason-by-intervention

discharge_cdep |>
  filter(!is.na(intervention_start),
         !is.na(reason_for_discharge)) |>
  tabyl(reason_for_discharge, intervention_name) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

### Client Type

#### Client Type Overall

```{r}
#| label: tbl-client-type

demo_cdep |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Client Type by Funding Round

```{r}
#| label: tbl-client-type-by-round

all_data_cdep |>
  filter(!is.na(intervention_start)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Client Type by Intervention

```{r}
#| label: tbl-client-type-by-intervention

all_data_cdep |>
  filter(!is.na(intervention_start)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(intervention_name, client_type) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

## Client Demographics

### Primary Language

```{r}
#| label: tbl-primary-language

demo_cdep |>
  filter(
    housing_status != "Don't Know",
    housing_status != "No Primary Language Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(primary_language)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(primary_language) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Housing Status

```{r}
#| label: tbl-housing-status

demo_cdep |>
  filter(
    housing_status != "Don't Know",
    housing_status != "No Housing Status Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(housing_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(housing_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Client Type with Average Age

```{r}
#| label: tbl-client-type-age

demo_cdep_distinct <- demo_cdep |>
  mutate(
    age_start_tx_num = case_when(
      age_at_start_of_treatment_in_years %in% c("No Age Provided", "", " ") ~ NA_character_,
      TRUE ~ age_at_start_of_treatment_in_years
    ) |> as.numeric()
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

tab_counts <- demo_cdep_distinct |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

tab_age <- demo_cdep_distinct |>
  group_by(client_type) |>
  summarise(
    mean_age   = round(mean(age_start_tx_num, na.rm = TRUE), 1),
    n_with_age = sum(!is.na(age_start_tx_num)),
    .groups = "drop"
  )

tab_counts |>
  left_join(tab_age, by = "client_type") |>
  select(client_type, n, percent, mean_age, n_with_age)
```

### Insurance Status

```{r}
#| label: tbl-insurance-status

demo_cdep |>
  filter(
    insurance_status != "Don't Know",
    insurance_status != "No Insurance Status Selected",
    insurance_status != "Not Applicable",
    insurance_status != "Prefer Not To Answer",
    !is.na(insurance_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(insurance_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Gender Identity

```{r}
#| label: tbl-gender-identity

demo_cdep |>
   filter(
    housing_status != "Don't Know",
    housing_status != "No Gender Identify Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(gender_identity)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(gender_identity) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Sexual Orientation

```{r}
#| label: tbl-sexual-orientation

demo_cdep |>
   filter(
    housing_status != "Don't Know",
    housing_status != "No Sexual Orientation Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(sexual_orientation)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(sexual_orientation) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Race

Race is coded numerically as follows:

| Code | Race                                               |
|------|----------------------------------------------------|
| 1    | Asian                                              |
| 2    | Black or African American                          |
| 3    | Hispanic or Latino                                 |
| 4    | Middle Eastern or North African                    |
| 5    | Native American / American Indian or Alaska Native |
| 6    | Native Hawaiian or Pacific Islander                |
| 7    | White                                              |
| 88   | 2+ races selected                                  |

```{r}
#| label: tbl-race

demo_cdep |>
  filter(
    race != "Don't Know",
    race != "No Race Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(race)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(code_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Ethnicity

Ethnicity is coded numerically as follows:

| Code | Ethnicity               |
|------|-------------------------|
| 1    | Chinese                 |
| 2    | Asian Indian            |
| 3    | Filipino                |
| 4    | Vietnamese              |
| 5    | Korean                  |
| 6    | Japanese                |
| 7    | African American        |
| 8    | Jamaican                |
| 9    | Haitian                 |
| 10   | Nigerian                |
| 11   | Somali                  |
| 12   | Mexican                 |
| 13   | Puerto Rican            |
| 14   | Salvadoran              |
| 15   | Cuban                   |
| 16   | Dominican               |
| 17   | Guatemalan              |
| 18   | Native Hawaiian         |
| 19   | Samoan                  |
| 20   | Chamorro                |
| 21   | English                 |
| 22   | German                  |
| 23   | Irish                   |
| 24   | Italian                 |
| 25   | Polish                  |
| 26   | Scottish                |
| 88   | 2+ ethnicities selected |

```{r}
#| label: tbl-ethnicity

all_data_cdep |>
   filter(
    race != "Don't Know",
    race != "No Ethnicity Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(ethnicity)
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(ethn_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

## Satisfaction Surveys

```{r}
#| label: setup-satisfaction
#| include: false

sat_all <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/combined_satisfaction_clean.csv",
  show_col_types = FALSE
)

# Unified Q1 score — coalesce across three survey-specific Q1 columns
sat_all <- sat_all |>
  mutate(
    q1_score = coalesce(
      x1_overall_i_am_satisfied_with_the_services_my_child_received_score,
      x1_overall_i_am_satisfied_with_the_services_i_received_score,
      x1_i_like_the_services_that_i_received_here_score
    )
  )

# Join with CDEP discharge data to add intervention_name, then filter to CDEPs
all_data_cdep_distinct <- all_data_cdep |>
  distinct(client_system_id, .keep_all = TRUE) |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  )

sat_cdep <- sat_all |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  ) |>
  left_join(all_data_cdep_distinct,
            by = c("provider", "provider_id", "provider_system_id", "client_system_id")) |>
  filter(!is.na(intervention_name))

score_cols <- names(sat_cdep)[str_detect(names(sat_cdep), "_score$")]

# Overall Q1 high satisfaction rate
pct_4_5 <- sat_cdep |>
  summarise(
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1)
  ) |>
  pull(pct_4_5)

# Overall combined survey high satisfaction rate (>= 80% of questions at 4 or 5)
df2 <- sat_cdep |>
  select(all_of(score_cols)) |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(everything(), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree    = sum(c_across(everything()) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  ungroup() |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree." The second is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed.

There are four satisfaction surveys: Caregiver (n=`{r} sum(sat_all$source == "caregiver")`), Child 13–17 (n=`{r} sum(sat_all$source == "child")`), Adult 18–59 (n=`{r} sum(sat_all$source == "adult")`), and Adult 60+ (n=`{r} sum(sat_all$source == "senior")`). High satisfaction on question 1 is defined as a score of 4 or 5.

### Question 1 Satisfaction

Overall, `{r} pct_4_5`% of survey respondents were highly satisfied based on question 1. The following table shows high satisfaction by intervention:

```{r}
#| label: tbl-q1-satisfaction

sat_cdep |>
  group_by(intervention_name) |>
  summarise(
    n_total = sum(!is.na(q1_score)),
    n_4_5   = sum(q1_score %in% c(4, 5), na.rm = TRUE),
    pct_4_5 = round(100 * n_4_5 / n_total, 1)
  )
```

### Combined Survey Satisfaction

Overall, `{r} df2$pct_meeting_80pct`% of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed. The following table shows high combined satisfaction by intervention, for clients where both a client ID and intervention name were available:

```{r}
#| label: tbl-combined-satisfaction

sat_cdep |>
  filter(!is.na(intervention_name)) |>
  group_by(intervention_name) |>
  select(intervention_name, all_of(score_cols)) |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  mutate(across(all_of(score_cols), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(all_of(score_cols)))),
    n_agree    = sum(c_across(all_of(score_cols)) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )
```

### Services in Preferred Language

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** The percent of clients that report services received in their preferred language out of the total that completed the satisfaction survey, by period.

```{r}
#| label: tbl-preferred-language

sat_cdep |>
  arrange(date_clean) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label))) |>
  rename(preferred_lang_response = were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |>
  tabyl(preferred_lang_response) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

## Participant Outcomes

### Improved Outcomes

```{r}
#| label: setup-outcomes
#| include: false

assessment_rules <- tibble::tribble(
  ~trimmed_pre_assessment_name, ~higher_is_better, ~min_change,
  "Basis-24",  FALSE, 0.001, # weighted 0–4 scale; any positive change counts
  "CATS",      FALSE, 1,
  "GAD-7",     FALSE, 1,
  "NOMS",      TRUE,  1,
  "Pediatric", FALSE, 1,
  "PHQ-9",     FALSE, 1,
  "PSS",       FALSE, 1,
  "SDQ",       FALSE, 1,
  "YOQ",       FALSE, 1
)

discharge_cdep_outcome <- discharge_cdep |>
  left_join(assessment_rules, by = "trimmed_pre_assessment_name") |>
  mutate(
    raw_change    = post_score - pre_score,
    better_change = if_else(higher_is_better, raw_change, -raw_change),
    min_change    = coalesce(min_change, 0),
    outcome = case_when(
      is.na(pre_score) | is.na(post_score) ~ NA_character_,
      better_change >=  min_change ~ "improved",
      better_change <= -min_change ~ "worsened",
      TRUE                         ~ "no_change"
    ),
    outcome = factor(outcome, levels = c("worsened", "no_change", "improved"))
  )
```

#### Outcome Overall

```{r}
#| label: tbl-outcome-overall

discharge_cdep_outcome |>
  tabyl(outcome) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

#### Outcome by Intervention

```{r}
#| label: tbl-outcome-by-intervention

discharge_cdep_outcome |>
  filter(!is.na(intervention_start), !is.na(outcome)) |>
  tabyl(intervention_name, outcome) |>
  adorn_totals("row") |>
  adorn_totals("col") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Outcome by Funding Round

```{r}
#| label: tbl-outcome-by-round

discharge_cdep_outcome |>
  tabyl(outcome, funding_round) |>
  adorn_totals("row") |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

#### Outcome by Assessment Measure

```{r}
#| label: tbl-outcome-by-measure

discharge_cdep_outcome |>
  filter(!is.na(outcome)) |>
  tabyl(post_assessment_name, outcome) |>
  adorn_totals("row") |>
  adorn_totals("col") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns()
```

### Treatment Goals Met

Two criteria must be met for a client to have met treatment goals: (1) a discharge reason of "Mutually agreed to cessation of treatment" and (2) an improvement according to pre- and post-measures collected (e.g., NOMS sub-scales). A simple positive difference between pre- and post-scores determines improvement.

```{r}
#| label: tbl-treatment-goals

discharge_cdep_outcome |>
  filter(outcome %in% c("no_change", "improved")) |>
  tabyl(reason_for_discharge) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```
