
---
title: "Discharge data vs. paired assessment audit"
author: "Cristin Young"
date: "2026-03-17"
format:
  html:
    df-print: paged
    toc: true
    toc-float: true
---

::: {.callout-note}
This report was generated using AI under general human direction. The contents have been reviewed and validated by Cristin Young on March 17, 2026. 
:::

## Background

This report compares two sources of outcome data for CYBHI EBP/CDEP clients:

1. **`discharge_data`** — produced by `report_data_cleaning.Rmd` from eInsight's raw Discharge Report export. One row per client per assessment. Contains `positive_outcomes` ("Yes"/"No"), which is **eInsight's own built-in outcome flag** — it is not computed by the team's code.

2. **Cleaned paired assessment files** — produced by the individual `*_data_cleaning.qmd` scripts (one per instrument). One row per client with a matched initial and follow-up assessment. Contains `positive_outcome` ("better"/"no change"/"worse"), computed by the team as a three-level variable based on first and last assessment scores.

The key question is: **do these two outcome definitions agree, and where they diverge, why?**

---

```{r load-packages}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(janitor)
library(zoo)
library(gt)
```

```{r load-data}
#| echo: false
#| message: false
out_dir    <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments"

discharge_data <- read_csv(
  file.path(out_dir, "discharge_data_clean.csv"),
  show_col_types = FALSE
) |>
  mutate(across(c(pre_completion_date, post_completion_date), as.Date))

pss_paired    <- read_csv(file.path(out_dir, "pss_paired_clean.csv"),          show_col_types = FALSE)
basis24_paired <- read_csv(file.path(out_dir, "basis24_paired_clean.csv"),     show_col_types = FALSE)
phq9_paired   <- read_csv(file.path(out_dir, "phq9_paired_clean.csv"),         show_col_types = FALSE)
gad7_paired   <- read_csv(file.path(out_dir, "gad7_paired_clean.csv"),         show_col_types = FALSE)
cpss_i_paired <- read_csv(file.path(out_dir, "cpss_i_paired_clean.csv"),       show_col_types = FALSE)
cpss_s_paired <- read_csv(file.path(out_dir, "cpss_s_paired_clean.csv"),       show_col_types = FALSE)
yoq_c_paired  <- read_csv(file.path(out_dir, "yoq_caregiver_paired_clean.csv"),show_col_types = FALSE)
yoq_s_paired  <- read_csv(file.path(out_dir, "yoq_self_paired_clean.csv"),     show_col_types = FALSE)

# Raw files for the deep-dive sections
read_assessment_csv <- function(path, na_vals = c("-", "", "NA", "N/A")) {
  raw <- read_csv(path, na = na_vals, col_names = FALSE, show_col_types = FALSE,
                  col_types = cols(.default = "c"))
  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()
  new_names <- ifelse(is.na(h1_ff) | h1_ff == "", h2,
                      paste(h1_ff, h2, sep = " - ")) |>
    janitor::make_clean_names()
  df <- raw[-c(1, 2), ]
  names(df) <- new_names
  df |> select(where(~ !all(is.na(.))))
}

gad7_full  <- read_assessment_csv(file.path(assess_dir, "GAD-7.csv"))
cpss_s_raw <- read_assessment_csv(file.path(assess_dir, "CPSS-5-self.csv"))

resp_cols_gad <- names(gad7_full)[
  endsWith(names(gad7_full), "_response") &
  !str_detect(names(gad7_full), "if_you_checked")
]
resp_cols_cpss_s <- names(cpss_s_raw)[endsWith(names(cpss_s_raw), "_score")]
```

---

## Section 1: Match rates

For each instrument, we compare the number of clients in `discharge_data` to the number in the cleaned paired file, and report how many match on `client_system_id`.

We also separate out how many discharge rows have a post-assessment recorded — clients with no post cannot appear in the paired file by definition, so the share of discharge clients *with* a post is the more meaningful denominator for evaluating match rates.

```{r match-rates}
paired_list <- list(
  PSS        = list(pattern = "PSS",                   paired = pss_paired),
  `BASIS-24` = list(pattern = "Basis-24",              paired = basis24_paired),
  `PHQ-9`    = list(pattern = "PHQ-9",                 paired = phq9_paired),
  `GAD-7`    = list(pattern = "GAD-7",                 paired = gad7_paired),
  `CPSS-5-I` = list(pattern = "CPSS 5 - Interviewer",  paired = cpss_i_paired),
  `CPSS-5-S` = list(pattern = "CPSS-5",                paired = cpss_s_paired),
  `YOQ-C`    = list(pattern = "YOQ.*Caregiver",        paired = yoq_c_paired),
  `YOQ-S`    = list(pattern = "YOQ.*Self",             paired = yoq_s_paired)
)

match_summary <- imap_dfr(paired_list, \(spec, label) {
  dis <- discharge_data |>
    filter(str_detect(pre_measure_s, spec$pattern))
  dis_with_post <- dis |> filter(!is.na(post_completion_date))

  tibble(
    Instrument              = label,
    `Discharge rows`        = n_distinct(dis$client_system_id),
    `With post-assessment`  = n_distinct(dis_with_post$client_system_id),
    `In paired file`        = n_distinct(spec$paired$client_system_id),
    `Matched (ID overlap)`  = length(intersect(dis$client_system_id,
                                                spec$paired$client_system_id)),
    `% discharge matched`   = round(100 * `Matched (ID overlap)` / `Discharge rows`)
  )
})

match_summary |>
  gt() |>
  tab_header(title = "Client ID match rates: discharge_data vs. paired assessment files") |>
  tab_style(
    style = cell_fill(color = "#fff3cd"),
    locations = cells_body(rows = Instrument == "BASIS-24")
  )
```

Most of the low match rates are **explained by missing post-assessments** in `discharge_data`. For example, PSS has `r match_summary |> filter(Instrument == "PSS") |> pull("Discharge rows")` discharge clients but only `r match_summary |> filter(Instrument == "PSS") |> pull("With post-assessment")` have a recorded post — closely matching the `r match_summary |> filter(Instrument == "PSS") |> pull("In paired file")` clients in the paired file.

::: {.callout-warning}
## BASIS-24 client ID mismatch

BASIS-24 is an outlier: only **`r match_summary |> filter(Instrument == "BASIS-24") |> pull("Matched (ID overlap)")`** out of **`r match_summary |> filter(Instrument == "BASIS-24") |> pull("Discharge rows")`** discharge clients match the paired file, despite both files containing roughly the same number of clients (~94 and ~99 respectively).

The 47 paired-file clients that *do* appear in `discharge_data` are recorded there under **NOMS Self-Report**, not BASIS-24. This is not a pairing or date issue — it is a near-complete client population mismatch between the two sources. The root cause (different extraction periods, a data entry discrepancy, or a systematic ID issue) needs to be investigated before BASIS-24 outcomes can be used in reporting.
:::

---

## Section 2: Score agreement

For clients that match on `client_system_id`, we compare `pre_score`/`post_score` in `discharge_data` to `initial_score`/`follow_up_score` in the paired file. Scores are near-identical when clients do match, confirming that the two sources draw from the same underlying assessments.

```{r score-agreement}
score_summary <- imap_dfr(paired_list, \(spec, label) {
  dis <- discharge_data |>
    filter(str_detect(pre_measure_s, spec$pattern), !is.na(pre_completion_date)) |>
    distinct(client_system_id, .keep_all = TRUE) |>
    select(client_system_id, pre_score, post_score)

  paired_cols <- intersect(
    c("client_system_id", "initial_score", "follow_up_score"),
    names(spec$paired)
  )

  joined <- inner_join(dis, spec$paired |> select(all_of(paired_cols)),
                       by = "client_system_id") |>
    mutate(
      pre_diff  = pre_score  - initial_score,
      post_diff = post_score - follow_up_score
    )

  tibble(
    Instrument        = label,
    `n matched`       = nrow(joined),
    `Pre: mean diff`  = round(mean(joined$pre_diff,  na.rm = TRUE), 3),
    `Pre: % exact`    = round(100 * mean(joined$pre_diff  == 0, na.rm = TRUE)),
    `Post: mean diff` = round(mean(joined$post_diff, na.rm = TRUE), 3),
    `Post: % exact`   = round(100 * mean(joined$post_diff == 0, na.rm = TRUE))
  )
})

score_summary |>
  gt() |>
  tab_header(title = "Score agreement for matched clients (discharge vs. paired file)") |>
  tab_spanner(label = "Pre-assessment score", columns = c("Pre: mean diff", "Pre: % exact")) |>
  tab_spanner(label = "Post-assessment score", columns = c("Post: mean diff", "Post: % exact")) |>
  tab_style(
    style = cell_fill(color = "#fff3cd"),
    locations = cells_body(rows = Instrument == "BASIS-24")
  )
```

For all instruments except BASIS-24, mean score differences are essentially zero and exact matches are high (typically 90–100%). This confirms the two sources are recording the same assessments for the clients that do overlap.

---

## Section 3: `positive_outcomes` vs. `positive_outcome`

`discharge_data$positive_outcomes` ("Yes"/"No") comes from eInsight's raw export and reflects eInsight's internal outcome logic. The paired files' `positive_outcome` ("better"/"no change"/"worse") is the team's own calculation based on first and last assessment scores. For instruments with both variables, we cross-tabulate them.

```{r outcome-crosstab}
has_po <- names(paired_list)[
  sapply(paired_list, \(s) "positive_outcome" %in% names(s$paired))
]

crosstab <- imap_dfr(paired_list[has_po], \(spec, label) {
  dis <- discharge_data |>
    filter(str_detect(pre_measure_s, spec$pattern), !is.na(pre_completion_date),
           !is.na(positive_outcomes)) |>
    distinct(client_system_id, .keep_all = TRUE) |>
    select(client_system_id, positive_outcomes)

  inner_join(dis,
             spec$paired |>
               select(client_system_id, positive_outcome) |>
               mutate(paired_outcome = case_when(
                 positive_outcome ==  1 ~ "better",
                 positive_outcome ==  0 ~ "no change",
                 positive_outcome == -1 ~ "worse",
                 .default = NA_character_
               )),
             by = "client_system_id") |>
    filter(!is.na(paired_outcome)) |>
    count(positive_outcomes, paired_outcome) |>
    group_by(positive_outcomes) |>
    mutate(pct = round(100 * n / sum(n))) |>
    ungroup() |>
    mutate(instrument = label)
})

crosstab |>
  select(instrument, positive_outcomes, paired_outcome, n, pct) |>
  arrange(instrument, positive_outcomes, paired_outcome) |>
  gt(groupname_col = "instrument") |>
  tab_header(
    title    = "Cross-tabulation: discharge positive_outcomes vs. paired positive_outcome",
    subtitle = "For matched clients with post-assessments in both sources"
  ) |>
  cols_label(
    positive_outcomes = "Discharge (Yes/No)",
    paired_outcome    = "Paired file (better/no change/worse)",
    n                 = "n",
    pct               = "%"
  ) |>
  tab_style(
    style = cell_fill(color = "#d4edda"),
    locations = cells_body(
      rows = positive_outcomes == "Yes" & paired_outcome == "better"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3cd"),
    locations = cells_body(
      rows = positive_outcomes == "No" & paired_outcome == "better"
    )
  )
```

Two clear patterns emerge:

**"Yes" is reliable.** Across every instrument, 100% of "Yes" clients in `discharge_data` also score as "better" in the paired file. The two sources fully agree on positive outcomes.

**"No" is not equivalent to "worse".** "No" captures "worse" and "no change" as expected, but it also includes a substantial share of clients whose scores *did* improve by the paired file's definition — most notably GAD-7 (46% of "No" clients) and CPSS-5-S (58% of "No" clients). This means counting "No" as "negative outcome" in reporting will over-count poor outcomes.

---

## Section 4: GAD-7 deep dive

For GAD-7, 46% of clients discharge labelled "No" actually improved. Yet for the sample cases examined, the discharge pre/post dates and scores matched the paired file's initial/follow-up dates and scores exactly — both sources are using the same assessment instances.

```{r gad7-score-changes}
gad7_changes <- discharge_data |>
  filter(str_detect(pre_measure_s, "GAD-7"), !is.na(post_completion_date)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  mutate(score_change = post_score - pre_score)   # negative = improved

gad7_changes |>
  group_by(positive_outcomes) |>
  summarise(
    n      = n(),
    min    = min(score_change,    na.rm = TRUE),
    `25th` = quantile(score_change, .25, na.rm = TRUE),
    median = median(score_change, na.rm = TRUE),
    `75th` = quantile(score_change, .75, na.rm = TRUE),
    max    = max(score_change,    na.rm = TRUE)
  ) |>
  gt() |>
  tab_header(
    title    = "GAD-7 score change distribution by discharge outcome",
    subtitle = "Negative values = improvement (lower anxiety)"
  )
```

```{r gad7-change-overlap}
gad7_changes |>
  count(positive_outcomes, score_change) |>
  pivot_wider(names_from = positive_outcomes, values_from = n, values_fill = 0) |>
  arrange(score_change) |>
  gt() |>
  tab_header(
    title = "GAD-7: count of clients by score change and discharge outcome",
    subtitle = "Both 'Yes' and 'No' appear at the same change values (e.g., −1, −2, −3)"
  ) |>
  cols_label(score_change = "Score change (post − pre)")
```

The distributions overlap at changes of −1, −2, and −3. This rules out a simple minimum-change threshold as the explanation. Combined with the fact that dates and scores match between the two sources, this suggests **eInsight's internal logic for GAD-7 uses a definition of "positive outcome" that differs from simple pre-post improvement** — but the specific rule is not visible from the export data alone.

---

## Section 5: CPSS-5-S deep dive

For CPSS-5-S, 58% of discharge "No" clients scored as "better" in the paired file. Unlike GAD-7, the mechanism here is traceable: discharge is often using a **different pre-assessment instance** than the paired file.

```{r cpss-s-pre-date}
cpss_s_scores_raw <- cpss_s_raw |>
  mutate(
    client_system_id = as.numeric(client_system_id),
    date_completed   = as.Date(date_completed),
    total_score = rowSums(across(all_of(resp_cols_cpss_s), as.numeric), na.rm = TRUE)
  ) |>
  select(client_system_id, date_completed, total_score) |>
  arrange(client_system_id, date_completed)

cpss_s_no_better_ids <- discharge_data |>
  filter(str_detect(pre_measure_s, "CPSS-5"), !is.na(post_completion_date),
         positive_outcomes == "No") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  inner_join(cpss_s_paired |> select(client_system_id, positive_outcome),
             by = "client_system_id") |>
  filter(positive_outcome == 1) |>
  pull(client_system_id)

cpss_s_earliest <- cpss_s_scores_raw |>
  group_by(client_system_id) |>
  slice_min(date_completed, n = 1, with_ties = FALSE) |>
  rename(earliest_date = date_completed, earliest_score = total_score)

cpss_s_no_better_detail <- discharge_data |>
  filter(str_detect(pre_measure_s, "CPSS-5"),
         !is.na(post_completion_date),
         positive_outcomes == "No",
         client_system_id %in% cpss_s_no_better_ids) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  select(client_system_id,
         dis_pre_date = pre_completion_date, dis_pre = pre_score,
         dis_post_date = post_completion_date, dis_post = post_score) |>
  left_join(cpss_s_earliest, by = "client_system_id") |>
  mutate(
    pre_date_matches_earliest = dis_pre_date == earliest_date,
    days_later = as.integer(dis_pre_date - earliest_date),
    score_change_discharge = dis_post - dis_pre
  )

cpss_s_no_better_detail |>
  count(pre_date_matches_earliest) |>
  mutate(pct = round(100 * n / sum(n))) |>
  gt() |>
  tab_header(
    title = "CPSS-5-S 'No but better': does discharge pre-date match earliest raw assessment?"
  ) |>
  cols_label(pre_date_matches_earliest = "Pre-date matches earliest assessment",
             n = "n", pct = "%")
```

```{r cpss-s-days-later}
cpss_s_no_better_detail |>
  filter(!pre_date_matches_earliest) |>
  summarise(
    n         = n(),
    min_days  = min(days_later,  na.rm = TRUE),
    mean_days = round(mean(days_later, na.rm = TRUE)),
    max_days  = max(days_later,  na.rm = TRUE)
  ) |>
  gt() |>
  tab_header(
    title    = "CPSS-5-S: how much later is the discharge pre-date vs. earliest raw assessment?",
    subtitle = "For the 17 clients where dates do not match"
  ) |>
  cols_label(n = "n clients", min_days = "Min days later",
             mean_days = "Mean days later", max_days = "Max days later")
```

In 17 of 22 cases (77%), the discharge pre-assessment date is **later** than the earliest assessment in the raw CPSS-5 file — on average `r cpss_s_no_better_detail |> filter(!pre_date_matches_earliest) |> summarise(m = round(mean(days_later, na.rm=TRUE))) |> pull(m)` days later. The scores at these two time points often happen to be the same (because the client scored identically on both dates), but the discharge record misses the earlier improvement that the raw file captures.

This confirms that for CPSS-5-S, discharge_data's `positive_outcomes` is based on a **designated "Initial" assessment label in eInsight** rather than the chronologically first assessment. When clients were assessed before their official "Initial" collection point, that earlier assessment is invisible to the discharge calculation.

---

## Section 6: Summary and implications

```{r summary-table}
tribble(
  ~Finding, ~Implication,
  '"Yes" reliably = "better" across all instruments (100%)',
  'Clients flagged "Yes" in discharge_data can be treated as confirmed positive outcomes; the two sources fully agree.',

  '"No" conflates "worse", "no change", and some "better" cases',
  'Using "No" as a proxy for "negative outcome" in reporting will over-count poor outcomes. The three-level positive_outcome variable from the paired files gives a more precise breakdown.',

  'GAD-7: 46% of "No" clients improved; overlap in score changes at −1 to −3',
  'The mechanism is unclear — eInsight\'s internal GAD-7 outcome logic is not transparent from the export data. Do not use discharge positive_outcomes as the primary outcome variable for GAD-7.',

  'CPSS-5-S: 58% of "No" clients improved; discharge uses a later pre-assessment in 77% of cases',
  'eInsight assigns the "Initial" collection label as the pre-assessment, which may not be the client\'s chronologically first score. The paired file captures the earlier assessment and therefore detects more improvement. This is a systematic difference in pairing logic, not a data quality error.',

  'BASIS-24: near-total client ID mismatch (4% overlap)',
  'The two sources appear to be tracking largely different client populations. BASIS-24 outcomes cannot be cross-validated until the source of the ID discrepancy is identified.'
) |>
  gt() |>
  tab_header(title = "Summary of findings and implications for reporting") |>
  cols_width(Finding ~ pct(40), Implication ~ pct(60)) |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_column_labels()
  )
```
