---
title: "Quarterly Report Results-10-20-2025"
author: "Cristin Young"
format: docx
editor: visual
---

## Outcomes Data by Round

|  |  |  |  |  |  |  |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| Round | Theme | Participants Served | Improved Outcomes | Satisfaction | Successful Discharges | Quote |
| 1  | Parent/Caregivers Programs and Practices   |   |   |   |   |   |
| 2  | Trauma-Informed Programs and Practices   |   |   |   |   |   |
| 3  | Early Childhood Wraparound Services   |   |   |   |   |   |
| 4  | Youth-Driven Programs   |   |   |   |   |   |
| 5  | Early Intervention Programs and Practices   |   |   |   |   |   |

### Participants Served

**Data used**: Discharge data, which includes 338 clients that had more than one complete intervention.

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in

```{r}
#| echo: false
#| warning: false
#| message: false
library(readr)
discharge_data_filtered <- read_csv("/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv")
table(discharge_data_filtered$funding_round)
```

### Improved outcomes
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
library(janitor)
library(tidyverse)
# number of rows in discharge data 
nrow(discharge_data_filtered)

# number of rows with positive_outcomes != NA
count_pos <- nrow(discharge_data_filtered %>% filter(!is.na(positive_outcomes)))
count_pos

# table of positive outcomes 
tab <- discharge_data_filtered %>%
  tabyl(positive_outcomes) %>%               
  adorn_pct_formatting(digits = 1)
tab

# percent of positive outcomes
yes_pct <- tab %>%
  filter(positive_outcomes == "Yes") %>%
  pull(valid_percent)
```
Out of `{r} nrow(discharge_data_filtered)` rows of clients in EInsight's discharge data dashboard, `{r} count_pos` had both a pre- and post-score. Of those clients with complete intervention data in EInsight, `{r} yes_pct` had a positive outcome.

The results below show % of positive outcomes by round, once NA's are removed from the data.

**Nitty gritty**: Positive outcome was defined as \[ this needs to be confirmed with Rebecca/Steve; this is a column from EInsight and I believe is just defined as "Is the post-score lower/higher than the pre-score, based on the relevant assessment scoring metrics given to Steve's team"\].

```{r}
#| echo: false
#| warning: false
#| message: false
# table of positive outcomes by funding round
tab2 <- discharge_data_filtered %>% 
  filter(!is.na(positive_outcomes)) %>% 
  tabyl(positive_outcomes, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 
tab2
```

### Satisfaction

```{r}

```

### Successful discharges
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# number of rows with a reason for discharge
count_rows <- nrow(discharge_data_filtered %>% filter(!is.na(reason_for_discharge)))

# number of rows with successful discharge
# table of positive outcomes 
tab3 <- discharge_data_filtered %>%
  tabyl(reason_for_discharge) %>%               
  adorn_pct_formatting(digits = 0)
tab3

# count of successful discharge
yes_count <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(n)
yes_count

# percent of successful discharge
yes_pct3 <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(valid_percent)
yes_pct3
```
Successful discharges were defined as those clients that had "Mutually agreed cessation of treatment (Successful Completion)" as the `reason_for_discharge` in the discharge data from EInsight. Of `{r} nrow(discharge_data_filtered)` rows of discharge data, `{r} count_rows` rows had a reason for discharge stated. Of individuals with a reason for discharge stated, `{r} yes_count` (`{r} yes_pct3`) were successfully discharged.

The results below show % of successful discharges by round, once NA's are removed from the data.
```{r}
# total table of successful discharges by funding round
tab4 <- discharge_data_filtered %>% 
  filter(!is.na(reason_for_discharge)) %>% 
  tabyl(reason_for_discharge, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 
tab4
```
