---
title: "BASIS-24 Assessment Data Cleaning"
author: "Cristin Young"
date: "2026-03-09"
format:
  html:
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** BASIS-24 assessment data for analysis and reporting. It ingests the raw eInsight export, processes the two-row hierarchical headers, applies reverse scoring to items 4–9, computes total scores, and pairs each client's initial assessment with their last follow-up.

**Data source:** `Basis-24.csv` — eInsight Individual Assessments export.

**Scoring:**

-   24 items, each rated 0–4.
-   Items **4 through 9 are reverse scored** using the formula `4 − raw_score`.
-   Total score is a **weighted sum**: `sum(weight_i × rating_i) / (1 − sum of weights for missing items)`. Weights for Q1–Q24 are defined in the scoring chunk below and sum to 1.
-   Total score range: 0–4. Higher scores indicate greater functional impairment.
-   **A lower follow-up score indicates a positive outcome.**
-   `total_score` is set to `NA` if fewer than 23 of 24 items are answered (more than 1 item missing).

The items comprising each subscale are as follows: - Depression/Functioning: 1,2,3,9,10,12 - Relationships: 4,5,6,7,8 - Self-Harm: 11,20 - Emotional Lability: 13,18,19 - Psychosis: 14,15,16,17 - Substance Abuse: 21,22,23,24

**Pairing logic:** Clients are grouped by `client_system_id`. The first assessment is the initial; the last assessment (if ≥2 exist) is the follow-up. Clients with only one assessment are excluded.

**Output:**

| File | Contents |
|----|----|
| `basis24_paired_clean.csv` | One row per client: initial/follow-up scores, `positive_outcome`, `score_change`, quarter labels |

------------------------------------------------------------------------

## Open Issues / Action Items

------------------------------------------------------------------------

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(zoo)
```

------------------------------------------------------------------------

# Helper Function

```{r header-helper}
read_assessment_csv <- function(path, na_vals = c("-", "", "NA", "N/A", "0")) {
  raw <- read_csv(path, na = na_vals, col_names = FALSE, show_col_types = FALSE,
                  col_types = cols(.default = "c"))

  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()

  new_names <- ifelse(
    is.na(h1_ff) | h1_ff == "",
    h2,
    paste(h1_ff, h2, sep = " - ")
  )
  new_names <- janitor::make_clean_names(new_names)

  df <- raw[-c(1, 2), ]
  names(df) <- new_names

  df |> select(where(~ !all(is.na(.))))
}
```

------------------------------------------------------------------------

# 1. Read and Score

```{r basis24-read}

# "0" is a valid rating on the 0–4 scale — do not treat it as NA
basis24_raw <- read_assessment_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/Basis-24.csv",
  na_vals = c("-", "", "NA", "N/A")
)

# Reverse-scored item positions (1-indexed within score columns)
rev_idx    <- c(4, 5, 6, 7, 8, 9)
score_cols <- names(basis24_raw)[endsWith(names(basis24_raw), "_score")]

# Official BASIS-24 item weights (Q1–Q24); sum to 1
basis24_weights <- c(
  0.069588, 0.106112, 0.070040, 0.025082, 0.027949, 0.029729,
  0.026978, 0.024052, 0.056511, 0.094193, 0.048262, 0.058638,
  0.038422, 0.016425, 0.022986, 0.024274, 0.030429, 0.064966,
  0.058904, 0.052358, 0.017836, 0.013450, 0.010047, 0.012771
)

# Weighted score formula: sum(w_i * x_i) / (1 - sum of weights for missing items)
weighted_basis24 <- function(vals, weights) {
  missing_wt <- sum(weights[is.na(vals)])
  sum(weights * vals, na.rm = TRUE) / (1 - missing_wt)
}

basis24_scores <- basis24_raw |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  mutate(across(all_of(score_cols[rev_idx]), ~ 4 - .x)) |>
  rowwise() |>
  mutate(
    missing_count = sum(is.na(c_across(all_of(score_cols)))),
    total_score   = if_else(
      missing_count <= 1,
      weighted_basis24(c_across(all_of(score_cols)), basis24_weights),
      NA_real_
    )
  ) |>
  ungroup() |>
  mutate(
    date_completed = as.Date(date_completed),
    row_id         = row_number()
  ) |>
  select(client_system_id, collection, date_completed, total_score) |>
  filter(
    date_completed >= "2024-01-01",
    date_completed <= Sys.Date()
  )

cat("BASIS-24 rows after scoring:", nrow(basis24_scores), "\n")
cat("Unique clients:             ", n_distinct(basis24_scores$client_system_id), "\n")
```

------------------------------------------------------------------------

# 2. Pair Initial and Follow-up

```{r basis24-pair}
basis24_paired <- basis24_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments   = n(),
    initial_date    = first(date_completed),
    initial_score   = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed),  as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score),     NA_real_)
  ) |>
  ungroup() |>
  filter(!is.na(follow_up_date)) |>
  mutate(
    # Lower follow-up = positive outcome (lower impairment is better)
    # 1 = improved, 0 = no change, -1 = worsened
    positive_outcome = case_when(
      follow_up_score <  initial_score ~ 1L,
      follow_up_score == initial_score ~ 0L,
      follow_up_score >  initial_score ~ -1L
    ),
    score_change      = follow_up_score - initial_score,
    initial_quarter   = paste0("Q", quarter(initial_date),    " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))
  )

cat("BASIS-24 paired clients:", nrow(basis24_paired), "\n")
```

------------------------------------------------------------------------

# 3. Quick Check

```{r basis24-check}
basis24_paired |>
  summarise(
    n_paired          = n(),
    n_improved        = sum(positive_outcome ==  1, na.rm = TRUE),
    n_no_change       = sum(positive_outcome ==  0, na.rm = TRUE),
    n_worsened        = sum(positive_outcome == -1, na.rm = TRUE),
    pct_improved      = round(100 * n_improved  / n_paired, 1),
    n_missing_outcome = sum(is.na(positive_outcome))
  )
```

------------------------------------------------------------------------

# 4. Write Clean CSV

```{r write-csv}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(basis24_paired, file.path(out_dir, "basis24_paired_clean.csv"))

cat("CSV written successfully.\n")
```
