---
title: "BASIS-24 Assessment Data Cleaning"
author: "Cristin Young"
date: "2026-03-09"
format:
  html:
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** BASIS-24 assessment data for analysis and reporting. It ingests the raw eInsight export, processes the two-row hierarchical headers, applies reverse scoring to items 4–9, computes total scores, and pairs each client's initial assessment with their last follow-up.

**Data source:** `Basis-24.xlsx` — eInsight Individual Assessments export.

**Scoring:**

-   24 items, each rated 0–4.
-   Items **4 through 9 are reverse scored** using the formula `6 − raw_score`.
-   Total score range: 0–96. Higher scores indicate greater functional impairment.
-   **A lower follow-up score indicates a positive outcome.**
-   Note: The existing code does not apply a missing-item threshold for `total_score` (unlike PSS/YOQ). The sum is computed with `na.rm = TRUE` regardless of how many items are missing.

**Pairing logic:** Clients are grouped by `client_system_id`. The first assessment is the initial; the last assessment (if ≥2 exist) is the follow-up. Clients with only one assessment are excluded.

**Output:**

| File | Contents |
|----|----|
| `basis24_paired_clean.csv` | One row per client: initial/follow-up scores, `positive_outcome`, `score_change`, quarter labels |

------------------------------------------------------------------------

## Open Issues / Action Items

-   [ ] No missing-item threshold is applied for BASIS-24 (unlike PSS which uses `NA` if 2+ missing). Confirm whether a threshold should be added.
-   [ ] No date range filter is applied (unlike NOMS). Confirm whether a period filter is needed.

------------------------------------------------------------------------

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(zoo)
```

------------------------------------------------------------------------

# Helper Function

```{r header-helper}
read_assessment_excel <- function(path, na_vals = c("-", "", "NA", "N/A", "0")) {
  raw <- read_xlsx(path, na = na_vals, col_names = FALSE)

  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()

  new_names <- ifelse(
    is.na(h1_ff) | h1_ff == "",
    h2,
    paste(h1_ff, h2, sep = " - ")
  )
  new_names <- janitor::make_clean_names(new_names)

  df <- raw[-c(1, 2), ]
  names(df) <- new_names

  df |> select(where(~ !all(is.na(.))))
}
```

------------------------------------------------------------------------

# 1. Read and Score

```{r basis24-read}
basis24_raw <- read_assessment_excel(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/Basis-24.xlsx"
)

# Reverse-scored item positions (1-indexed within score columns)
rev_idx    <- c(4, 5, 6, 7, 8, 9)
score_cols <- names(basis24_raw)[endsWith(names(basis24_raw), "_score")]

basis24_scores <- basis24_raw |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  rowwise() |>
  mutate(
    missing_count = sum(is.na(c_across(all_of(score_cols)))),
    total_score = {
      vals          <- c_across(all_of(score_cols))
      vals[rev_idx] <- 6 - vals[rev_idx]
      sum(vals, na.rm = TRUE)
    }
  ) |>
  ungroup() |>
  mutate(
    date_completed = as.Date(date_completed),
    row_id         = row_number()
  ) |>
  select(client_system_id, collection, date_completed, total_score)

cat("BASIS-24 rows after scoring:", nrow(basis24_scores), "\n")
cat("Unique clients:             ", n_distinct(basis24_scores$client_system_id), "\n")
```

------------------------------------------------------------------------

# 2. Pair Initial and Follow-up

```{r basis24-pair}
basis24_paired <- basis24_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments   = n(),
    initial_date    = first(date_completed),
    initial_score   = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed),  as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score),     NA_real_)
  ) |>
  ungroup() |>
  filter(!is.na(follow_up_date)) |>
  mutate(
    # Lower follow-up = positive outcome (lower impairment is better)
    positive_outcome  = if_else(follow_up_score < initial_score, 1L, 0L),
    score_change      = follow_up_score - initial_score,
    initial_quarter   = paste0("Q", quarter(initial_date),    " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))
  )

cat("BASIS-24 paired clients:", nrow(basis24_paired), "\n")
```

------------------------------------------------------------------------

# 3. Quick Check

```{r basis24-check}
basis24_paired |>
  summarise(
    n_paired          = n(),
    n_positive        = sum(positive_outcome == 1, na.rm = TRUE),
    pct_positive      = round(100 * n_positive / n_paired, 1),
    n_missing_outcome = sum(is.na(positive_outcome))
  )
```

------------------------------------------------------------------------

# 4. Write Clean CSV

```{r write-csv}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(basis24_paired, file.path(out_dir, "basis24_paired_clean.csv"))

cat("CSV written successfully.\n")
```
