---
title: "CPSS-5 Assessment Data Cleaning"
author: "Cristin Young"
date: "2026-03-09"
format:
  html:
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** Child PTSD Symptom Scale, 5th Edition (CPSS-5) assessment data for analysis and reporting. It ingests two raw eInsight exports (interviewer and self-report versions), processes the two-row hierarchical headers, computes total scores, and pairs each client's initial assessment with their last follow-up.

**Data sources:** `CPSS-5-Interviewer.csv` and `CPSS-5-self.csv` — eInsight Individual Assessments exports. Both are CSV files with two-row hierarchical headers.

**Scoring:**

-   Sum of items Q1–Q20 only (the first 20 score columns).
-   Total score range: 0–80. Higher scores indicate greater PTSD symptom severity.
-   A clinical cutoff score of ≥31 can be used to identify a probable PTSD diagnosis.
-   **A lower follow-up score indicates a positive outcome.**
-   No missing-item threshold is applied (total computed with `na.rm = TRUE`).

| Symptom Severity | Score Range |
|------------------|-------------|
| Minimal          | 0–10        |
| Mild             | 11–20       |
| Moderate         | 21–40       |
| Severe           | 41–60       |
| Very Severe      | 61–80       |

**Pairing logic:** Clients are grouped by `client_system_id`. The first assessment is the initial; the last assessment (if ≥2 exist) is the follow-up. Clients with only one assessment are excluded.

**Outputs:**

| File | Contents |
|----|----|
| `cpss_i_paired_clean.csv` | One row per client: initial/follow-up scores, `positive_outcome`, `score_change`, quarter labels — interviewer version |
| `cpss_s_paired_clean.csv` | Same, self-report version |

------------------------------------------------------------------------

## Open Issues / Action Items

-   [ ] Not sure how to handle NAs

------------------------------------------------------------------------

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(zoo)
```

------------------------------------------------------------------------

# Helper Function

Both CPSS-5 files are CSVs with two-row headers. We use a `read_csv`-based helper (unlike NOMS/PSS/BASIS-24 which use Excel files).

```{r header-helper}
# Reads a CPSS-5 CSV with two-row hierarchical headers.
# Returns a cleaned data frame with combined column names and empty columns removed.
read_assessment_csv_headers <- function(path,
                                        na_vals = c("-", "", "NA", "N/A", "0")) {
  raw <- read_csv(path, na = na_vals, col_names = FALSE, show_col_types = FALSE)

  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()

  new_names <- ifelse(
    is.na(h1_ff) | h1_ff == "",
    h2,
    paste(h1_ff, h2, sep = " - ")
  )
  new_names <- janitor::make_clean_names(new_names)

  df <- raw[-c(1, 2), ]
  names(df) <- new_names

  df |> select(where(~ !all(is.na(.))))
}
```

------------------------------------------------------------------------

# Helper: Score and Pair

Both versions share identical scoring and pairing logic after header processing.

```{r pair-helper}
score_and_pair_cpss <- function(df, clinical_cutoff = 31) {
  score_cols <- names(df)[endsWith(names(df), "_score")]

  scored <- df |>
    #filter(date_completed <= "2025-12-31") |>
    mutate(across(all_of(score_cols), as.numeric)) |>
    rowwise() |>
    mutate(
      missing_count       = sum(is.na(c_across(all_of(score_cols)))),
      total_score         = rowSums(across(all_of(score_cols)), na.rm = TRUE),
      clinically_relevant = if_else(
        !is.na(total_score) & total_score >= clinical_cutoff, TRUE, FALSE
      )
    ) |>
    ungroup() |>
    mutate(
      date_completed = as.Date(date_completed),
      row_id         = row_number()
    ) |>
    select(client_system_id, collection, date_completed, total_score) |>
    filter(
      date_completed >= "2024-01-01",
      date_completed <= Sys.Date()
    )

  scored |>
    group_by(client_system_id) |>
    arrange(date_completed, .by_group = TRUE) |>
    summarise(
      n_assessments   = n(),
      initial_date    = first(date_completed),
      initial_score   = first(total_score),
      follow_up_date  = if_else(n_assessments >= 2, last(date_completed),  as.Date(NA)),
      follow_up_score = if_else(n_assessments >= 2, last(total_score),     NA_real_)
    ) |>
    ungroup() |>
    filter(!is.na(follow_up_date)) |>
    mutate(
      # Lower follow-up = positive outcome (lower symptom severity is better)
      # 1 = improved, 0 = no change, -1 = worsened
      positive_outcome = case_when(
        follow_up_score <  initial_score ~ 1L,
        follow_up_score == initial_score ~ 0L,
        follow_up_score >  initial_score ~ -1L
      ),
      score_change      = follow_up_score - initial_score,
      initial_quarter   = paste0("Q", quarter(initial_date),    " ", year(initial_date)),
      follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date)),
      initial_severity  = case_when(
        initial_score <= 10 ~ "Minimal",
        initial_score <= 20 ~ "Mild",
        initial_score <= 40 ~ "Moderate",
        initial_score <= 60 ~ "Severe",
        initial_score <= 80 ~ "Very Severe"
      ) |> factor(levels = c("Minimal", "Mild", "Moderate", "Severe", "Very Severe")),
      follow_up_severity = case_when(
        follow_up_score <= 10 ~ "Minimal",
        follow_up_score <= 20 ~ "Mild",
        follow_up_score <= 40 ~ "Moderate",
        follow_up_score <= 60 ~ "Severe",
        follow_up_score <= 80 ~ "Very Severe"
      ) |> factor(levels = c("Minimal", "Mild", "Moderate", "Severe", "Very Severe"))
    )
}
```

------------------------------------------------------------------------

# 1. CPSS-5 Interviewer Version

```{r cpss-i}
cpss_i_raw <- read_assessment_csv_headers(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/CPSS-5-Interviewer.csv"
)

cpss_i_paired <- score_and_pair_cpss(cpss_i_raw)

cat("CPSS-5 Interviewer paired clients:", nrow(cpss_i_paired), "\n")

cpss_i_paired |>
  summarise(
    n_paired          = n(),
    n_improved        = sum(positive_outcome ==  1, na.rm = TRUE),
    n_no_change       = sum(positive_outcome ==  0, na.rm = TRUE),
    n_worsened        = sum(positive_outcome == -1, na.rm = TRUE),
    pct_improved      = round(100 * n_improved / n_paired, 1),
    n_missing_outcome = sum(is.na(positive_outcome))
  )
```

------------------------------------------------------------------------

# 2. CPSS-5 Self-Report Version

```{r cpss-s}
cpss_s_raw <- read_assessment_csv_headers(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/CPSS-5-self.csv"
)

cpss_s_paired <- score_and_pair_cpss(cpss_s_raw)

cat("CPSS-5 Self paired clients:", nrow(cpss_s_paired), "\n")

cpss_s_paired |>
  summarise(
    n_paired          = n(),
    n_improved        = sum(positive_outcome ==  1, na.rm = TRUE),
    n_no_change       = sum(positive_outcome ==  0, na.rm = TRUE),
    n_worsened        = sum(positive_outcome == -1, na.rm = TRUE),
    pct_improved      = round(100 * n_improved / n_paired, 1),
    n_missing_outcome = sum(is.na(positive_outcome))
  )
```

------------------------------------------------------------------------

# 3. Write Clean CSVs

```{r write-csvs}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(cpss_i_paired, file.path(out_dir, "cpss_i_paired_clean.csv"))
write_csv(cpss_s_paired, file.path(out_dir, "cpss_s_paired_clean.csv"))

cat("CSVs written successfully.\n")
```
