---
title: "NOMS Assessment Data Cleaning"
author: "Cristin Young"
date: "2026-03-09"
format:
  html:
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** NOMS (National Outcome Measures) assessment data for analysis and reporting. It ingests two raw eInsight exports (child and self-report versions), processes the two-row hierarchical headers, coerces scores to numeric, computes total scores, filters to the reporting period, pivots to long format, and pairs each client's initial assessment with their last follow-up (>3 days after initial).

**Data sources (eInsight Individual Assessments exports):**

- **NOMS-child.xlsx** — Caregiver-completed NOMS for child clients.
- **NOMS-self_report.xlsx** — Client self-report NOMS.

Each file has a two-row header: row 1 contains question text, row 2 contains subheader labels (e.g., "Score", "Comments").

**Scoring:** Range 1–5 per question. An increase in follow-up score vs. initial indicates a positive outcome. Four questions are scored:

- Q1: Overall mental health (single item, range 1–5)
- Q2: Daily life functioning (7 items, range 7–35)
- Q3: Well-being (6 items, range 6–30)
- Q4: Social connectedness (6 items, range 6–30)

**Pairing logic:** Rows are grouped by client ID and question. The first assessment is the initial; the last assessment at least 3 days later is the follow-up. Clients without a qualifying follow-up are excluded from the paired output.

**Outputs:**

| File | Contents |
|------|----------|
| `noms_child_pairs_clean.csv` | Long-format paired assessments (one row per client × question), child version |
| `noms_self_pairs_clean.csv` | Long-format paired assessments, self-report version |

---

## Open Issues / Action Items

- [ ] The 3-day threshold for "follow-up" is a pragmatic cutoff. Confirm this is the intended definition.
- [ ] Date filter lower bound is `2024-01-01`; confirm this is the right start of the reporting window.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(zoo)
```

---

# Helper Function

```{r header-helper}
# Reads an eInsight assessment Excel export with two-row headers.
# Returns a cleaned data frame with combined column names and empty columns removed.
read_assessment_excel <- function(path, na_vals = c("-", "", "NA", "N/A")) {
  raw <- read_xlsx(path, na = na_vals, col_names = FALSE)

  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()

  new_names <- ifelse(
    is.na(h1_ff) | h1_ff == "",
    h2,
    paste(h1_ff, h2, sep = " - ")
  )
  new_names <- janitor::make_clean_names(new_names)

  df <- raw[-c(1, 2), ]
  names(df) <- new_names

  df |> select(where(~ !all(is.na(.))))
}
```

---

# 1. NOMS Child

## 1.1 Read and Score

```{r noms-child-read}
noms_child <- read_assessment_excel(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/NOMS-child.xlsx"
) |>
  mutate(across(ends_with("_score"), as.numeric)) |>
  rowwise() |>
  mutate(
    any_missing_q1x = any(is.na(c_across(matches("^(q1|x).*_score$")))),
    total_score     = sum(c_across(matches("^(q1|x).*_score$")), na.rm = TRUE)
  ) |>
  ungroup() |>
  mutate(
    date_completed = as.Date(date_completed),
    row_id         = row_number()
  ) |>
  filter(
    date_completed >= "2024-01-01",
 #   date_completed <= "2025-12-31",
    date_completed <= Sys.Date()
  )

cat("NOMS Child rows after cleaning:", nrow(noms_child), "\n")
cat("Unique clients:                 ", n_distinct(noms_child$system_id), "\n")
```

## 1.2 Pivot Long

```{r noms-child-long}
noms_child_long <- noms_child |>
  pivot_longer(
    cols      = matches("^(q1|x).*_score$"),
    names_to  = "question",
    values_to = "score"
  ) |>
  mutate(score = as.numeric(score))
```

## 1.3 Pair Initial and Follow-up

```{r noms-child-pair}
pairs_primary_long <- noms_child_long |>
  group_by(system_id, question) |>
  arrange(date_completed, .by_group = TRUE) |>
  mutate(
    days_since_initial = as.integer(date_completed - first(date_completed)),
    initial_date       = first(date_completed),
    initial_score      = first(score),
    is_after3          = days_since_initial > 3
  ) |>
  filter(is_after3) |>
  slice_tail(n = 1) |>
  transmute(
    system_id,
    question,
    initial_date,
    initial_score,
    follow_up_date  = date_completed,
    follow_up_score = score,
    days_since_initial
  )

cat("NOMS Child paired rows: ", nrow(pairs_primary_long), "\n")
cat("Unique clients paired:  ", n_distinct(pairs_primary_long$system_id), "\n")
cat("Questions covered:      ", n_distinct(pairs_primary_long$question), "\n")
```

---

# 2. NOMS Self-Report

## 2.1 Read and Score

```{r noms-self-read}
noms_self <- read_assessment_excel(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/NOMS-self_report.xlsx"
) |>
  mutate(across(ends_with("_score"), as.numeric)) |>
  rowwise() |>
  mutate(
    any_missing_q1x = any(is.na(c_across(matches("^(q1|x).*_score$")))),
    total_score     = sum(c_across(matches("^(q1|x).*_score$")), na.rm = TRUE)
  ) |>
  ungroup() |>
  mutate(
    date_completed = as.Date(date_completed),
    row_id         = row_number()
  ) |>
  filter(
    date_completed >= "2024-01-01",
#    date_completed <= "2025-12-31",
    date_completed <= Sys.Date()
  )

cat("NOMS Self rows after cleaning:", nrow(noms_self), "\n")
cat("Unique clients:               ", n_distinct(noms_self$client_system_id), "\n")
```

## 2.2 Pivot Long

```{r noms-self-long}
noms_self_long <- noms_self |>
  pivot_longer(
    cols      = matches("^(q1|x).*_score$"),
    names_to  = "question",
    values_to = "score"
  ) |>
  mutate(score = as.numeric(score))
```

## 2.3 Pair Initial and Follow-up

```{r noms-self-pair}
pairs_primary_long_self <- noms_self_long |>
  group_by(client_system_id, question) |>
  arrange(date_completed, .by_group = TRUE) |>
  mutate(
    days_since_initial = as.integer(date_completed - first(date_completed)),
    initial_date       = first(date_completed),
    initial_score      = first(score),
    is_after3          = days_since_initial > 3
  ) |>
  filter(is_after3) |>
  slice_tail(n = 1) |>
  transmute(
    client_system_id,
    question,
    initial_date,
    initial_score,
    follow_up_date  = date_completed,
    follow_up_score = score,
    days_since_initial
  )

cat("NOMS Self paired rows: ", nrow(pairs_primary_long_self), "\n")
cat("Unique clients paired: ", n_distinct(pairs_primary_long_self$client_system_id), "\n")
cat("Questions covered:     ", n_distinct(pairs_primary_long_self$question), "\n")
```

---

# 3. Write Clean CSVs

```{r write-csvs}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(pairs_primary_long,      file.path(out_dir, "noms_child_pairs_clean.csv"))
write_csv(pairs_primary_long_self, file.path(out_dir, "noms_self_pairs_clean.csv"))

cat("CSVs written successfully.\n")
```
