---
title: "PSS Assessment Data Cleaning"
author: "Cristin Young"
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** Parenting Stress Scale (PSS) assessment data for analysis and reporting. It ingests the raw eInsight export, processes the two-row hierarchical headers, maps text responses to numeric values, applies reverse scoring, computes total scores, and pairs each client's initial assessment with their last follow-up.

**Data source:** `PSS.csv` — eInsight Individual Assessments export.

**Scoring:**

- 18 items. Responses are stored as text and mapped to numeric values: Strongly Disagree = 1, Disagree = 2, Undecided = 3, Agree = 4, Strongly Agree = 5.
- Items 1, 2, 5, 6, 7, 8, 17, and 18 are **reverse scored** using the formula `6 − raw_score`.
- Total score range: 18–90. Higher scores indicate greater perceived stress.
- **A lower follow-up score indicates a positive outcome.**
- Total score is set to `NA` if 2 or more items are missing.

**Pairing logic:** Clients are grouped by `client_system_id`. The first assessment is the initial; the last assessment (if ≥2 assessments exist) is the follow-up. Clients with only one assessment are excluded.

**Output:**

| File | Contents |
|----|----|
| `pss_paired_clean.csv` | One row per client: initial/follow-up scores, `positive_outcome`, `score_change`, quarter labels |

------------------------------------------------------------------------

## Open Issues / Action Items

- [ ] The `missing_count` threshold of 2 for treating total score as `NA` — confirm this is the correct PSS rule.

------------------------------------------------------------------------

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(zoo)
```

------------------------------------------------------------------------

# Helper Function

```{r header-helper}
# Reads an eInsight assessment CSV export with two-row headers.
# Returns a cleaned data frame with combined column names and empty columns removed.
read_assessment_csv <- function(path, na_vals = c("-", "", "NA", "N/A")) {
  raw <- read_csv(path, na = na_vals, col_names = FALSE, show_col_types = FALSE,
                  col_types = cols(.default = "c"))

  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()

  new_names <- ifelse(
    is.na(h1_ff) | h1_ff == "",
    h2,
    paste(h1_ff, h2, sep = " - ")
  )
  new_names <- janitor::make_clean_names(new_names)

  df <- raw[-c(1, 2), ]
  names(df) <- new_names

  df |> select(where(~ !all(is.na(.))))
}
```

------------------------------------------------------------------------

# 1. Read and Score

```{r pss-read}
pss_raw <- read_assessment_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/PSS.csv"
)

# Reverse-scored item positions (1-indexed within response columns)
rev_idx       <- c(1, 2, 5, 6, 7, 8, 17, 18)
response_cols <- names(pss_raw)[endsWith(names(pss_raw), "_response")]

# Map text responses to numeric values
response_map <- c(
  "Strongly disagree" = 1,
  "Disagree"          = 2,
  "Undecided"         = 3,
  "Agree"             = 4,
  "Strongly agree"    = 5
)

pss_scores <- pss_raw |>
  mutate(across(all_of(response_cols), ~ response_map[.])) |>
  rowwise() |>
  mutate(
    missing_count = sum(is.na(c_across(all_of(response_cols)))),
    total_score = {
      vals          <- c_across(all_of(response_cols))
      vals[rev_idx] <- 6 - vals[rev_idx]
      sum(vals, na.rm = TRUE)
    }
  ) |>
  ungroup() |>
  mutate(
    total_score    = if_else(missing_count >= 2, NA_real_, total_score),
    date_completed = as.Date(date_completed),
    row_id         = row_number()
  ) |>
  select(client_system_id, collection, date_completed, total_score) |>
  filter(
    date_completed >= "2024-01-01",
    date_completed <= Sys.Date()
  )

cat("PSS rows after scoring:", nrow(pss_scores), "\n")
cat("Unique clients:        ", n_distinct(pss_scores$client_system_id), "\n")
```

------------------------------------------------------------------------

# 2. Pair Initial and Follow-up

```{r pss-pair}
pss_paired <- pss_scores |>
  group_by(client_system_id) |>
  arrange(date_completed, .by_group = TRUE) |>
  summarise(
    n_assessments   = n(),
    initial_date    = first(date_completed),
    initial_score   = first(total_score),
    follow_up_date  = if_else(n_assessments >= 2, last(date_completed),  as.Date(NA)),
    follow_up_score = if_else(n_assessments >= 2, last(total_score),     NA_real_)
  ) |>
  ungroup() |>
  filter(!is.na(follow_up_date)) |>
  mutate(
    # Lower follow-up score = positive outcome (lower stress is better)
    # 1 = improved, 0 = no change, -1 = worsened
    positive_outcome = case_when(
      follow_up_score <  initial_score ~ 1L,
      follow_up_score == initial_score ~ 0L,
      follow_up_score >  initial_score ~ -1L
    ),
    score_change      = follow_up_score - initial_score,
    initial_quarter   = paste0("Q", quarter(initial_date),    " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))
  )

cat("PSS paired clients:", nrow(pss_paired), "\n")
```

------------------------------------------------------------------------

# 3. Quick Check

```{r pss-check}
pss_paired |>
  summarise(
    n_paired          = n(),
    n_improved        = sum(positive_outcome ==  1, na.rm = TRUE),
    n_no_change       = sum(positive_outcome ==  0, na.rm = TRUE),
    n_worsened        = sum(positive_outcome == -1, na.rm = TRUE),
    pct_improved      = round(100 * n_improved / n_paired, 1),
    n_missing_outcome = sum(is.na(positive_outcome))
  )
```

------------------------------------------------------------------------

# 4. Write Clean CSV

```{r write-csv}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(pss_paired, file.path(out_dir, "pss_paired_clean.csv"))

cat("CSV written successfully.\n")
```
