---
title: "Report Data Cleaning"
author: "Cristin Young"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    df_print: paged
    toc: true
    toc_float: true
  pdf_document: default
  html_notebook: default
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** program data for analysis and reporting. It ingests two raw exports from the eInsight clinical management platform, cleans and recodes each, joins them together, and writes three clean CSVs used by downstream analysis scripts.

**Data sources (eInsight exports):**

- **Discharge Report** — One row per client assessment pair (pre/post). Contains intervention dates, pre/post measure names, and funding round.
- **Client Demographics** — One row per client (ideally). Contains race, ethnicity, language, age, provider, and other client characteristics.

**Outputs:**

| File | Contents |
|------|----------|
| `demographic_data_clean.csv` | Deduplicated, recoded demographic data |
| `discharge_data_clean.csv` | Cleaned discharge/assessment data |
| `demographic_discharge_clean.csv` | Joined dataset for outcome analyses |

---

## Open Issues / Action Items

- [ ] Confirm we are filtering to the **last follow-up time point** (filter `post_time_point == ??`). Which follow-up number should be treated as "final" for a given client?
- [ ] `nrow(dup_clients_diff_time) + nrow(dup_clients_same_time)` should equal `nrow(dup_clients)` but does not — investigate why some duplicate clients fall into neither category.
- [ ] Clients with the same `client_system_id` but different `assessment_date` values need special handling (see `client_system_id == 3664` as an example). Clarify whether mismatched dates reflect linked assessment + intervention records, or data entry errors.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(gt)
library(readxl)
library(openxlsx)
```

---

# 1. Read in Raw Data

Both files are pulled from the eInsight export folder in the shared Dropbox. The `na` argument ensures that common "missing" strings are treated as `NA` from the start. Column names are cleaned to snake_case via `janitor::clean_names()`. Date columns in the discharge data are stored as numeric by Excel and need explicit conversion.

```{r read-data}
discharge_data <- read_xlsx(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Discharge Report Export/2026-05-07-discharge_report.xlsx",
  na = c("-", "", "NA", "N/A")
) |>
  janitor::clean_names()

# These date columns come in as Excel serial numbers and must be converted
discharge_data <- discharge_data |>
  mutate(
    pre_completion_date = case_when(
      grepl("^[0-9]+\\.?[0-9]*$", pre_completion_date) ~
        janitor::convert_to_date(as.numeric(pre_completion_date)),
      TRUE ~ mdy(pre_completion_date)
    ),
    # Fix 2-digit years parsed as e.g. year 25 instead of 2025
    pre_completion_date = if_else(
      year(pre_completion_date) < 100,
      pre_completion_date + years(2000),
      pre_completion_date
    )
  )

discharge_data$pre_completion_date  <- janitor::convert_to_date(discharge_data$pre_completion_date)
discharge_data$post_completion_date <- janitor::convert_to_date(discharge_data$post_completion_date)

demographic_data <- readxl::read_xlsx(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Admin Report Exports/2026-04-21-client_demographics.xlsx",
  na = c("-", "", "NA", "N/A")
) |>
  janitor::clean_names()

# Quick row/ID counts
cat("Discharge rows:          ", nrow(discharge_data), "\n")
cat("Demographic rows:        ", nrow(demographic_data), "\n")
cat("Unique demographic IDs:  ", n_distinct(demographic_data$client_system_id), "\n")
```

---

# 2. Clean Demographic Data

## 2.1 Remove Demo/Test Accounts

eInsight includes test provider and client accounts with "DEMO" in their IDs. These are not real clients and must be excluded before any analysis.

```{r demo-filter-demographics}
demographic_data_filtered <- demographic_data |>
  filter(!if_any(c(provider_id, client_system_id), ~ str_detect(as.character(.x), "DEMO")))
```

## 2.2 Identify and Handle Duplicate Client IDs

The demographics export should have one row per client, but some clients appear more than once (e.g., due to selecting multiple languages or data entry errors). We identify duplicates, export them for review, then drop them from the analysis dataset for now.

```{r demo-duplicates}
# Identify which client IDs appear more than once
duplicates <- demographic_data_filtered |>
  group_by(client_system_id) |>
  filter(n() > 1) |>
  tally()

cat("Client IDs with duplicates:", nrow(duplicates), "\n")
cat("Total duplicated rows:     ", sum(duplicates$n), "\n")

# Inspect which columns differ across duplicate rows for the same client
demographic_data_filtered |>
  group_by(client_system_id) |>
  filter(n() > 1) |>
  summarise(
    across(everything(), ~ {
      u <- unique(.x)
      if (length(u) > 1) paste(u, collapse = ", ") else NA_character_
    }),
    .groups = "drop"
  ) |>
  select(client_system_id, where(~ any(!is.na(.))))
```

```{r demo-save-duplicates}
# Export the full duplicate records for manual review
duplicate_list <- demographic_data_filtered |>
  filter(client_system_id %in% duplicates$client_system_id)

write_csv(
  duplicate_list,
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_duplicate_list.csv"
)
```

```{r demo-drop-duplicates}
# Drop all rows for clients with any duplicate — conservative approach until
# a resolution strategy is decided for these records
demographic_data_filtered <- demographic_data_filtered |>
  filter(!(client_system_id %in% duplicates$client_system_id))

# Confirm: should now be TRUE (unique IDs)
length(unique(demographic_data_filtered$client_system_id)) == nrow(demographic_data_filtered)
```

## 2.3 Recode Race

Race is stored as a free-text, comma-separated field. We clean it in two passes:

1. Strip non-informative responses ("Don't Know", "Prefer Not To Answer", "Not Applicable").
2. Recode to a numeric summary:
   - `NA` — no valid response after stripping
   - `1` — exactly one valid race reported
   - `2` — two or more valid races reported (multiracial)
   - `code_num` assigns each single-race response a numeric code, or `88` for multiracial.

```{r recode-race}
codebook_race <- c(
  "Asian"                                                = 1,
  "Black or African American"                            = 2,
  "Hispanic or Latino"                                   = 3,
  "Middle Eastern or North African"                      = 4,
  "Native American/American Indian or Alaska Native"     = 5,
  "Native Hawaiian or Pacific Islander"                  = 6,
  "White"                                                = 7
)

demographic_data_filtered <- demographic_data_filtered |>
  mutate(
    cleaned_race = str_remove_all(
      race,
      "\\b(Don't Know|Prefer Not To Answer|Not Applicable)\\b"
    ) |>
      str_replace_all(",+", ",") |>       # collapse extra commas left by removal
      str_replace_all("^,|,$", "") |>     # trim any leading/trailing commas
      na_if("")                            # empty string → NA
  ) |>
  mutate(
    recoded_race = case_when(
      is.na(cleaned_race)              ~ NA,   # no valid response
      str_detect(cleaned_race, ",")    ~ 2L,   # multiracial
      TRUE                             ~ 1L    # single race
    ),
    code_num = case_when(
      is.na(cleaned_race)              ~ NA,
      str_detect(cleaned_race, ",")    ~ 88L,  # multiracial code
      TRUE ~ unname(codebook_race[cleaned_race])
    )
  )
```

## 2.4 Recode Ethnicity

Same two-pass approach as race. Ethnicity is also stored as a comma-separated string. Non-informative values are removed and responses are recoded to numeric codes; `88` = 2+ ethnicities selected.

```{r recode-ethnicity}
codebook_eth <- c(
  "Chinese" = 1, "Asian Indian" = 2, "Filipino" = 3, "Vietnamese" = 4,
  "Korean" = 5, "Japanese" = 6, "African American" = 7, "Jamaican" = 8,
  "Haitian" = 9, "Nigerian" = 10, "Somali" = 11, "Mexican" = 12,
  "Puerto Rican" = 13, "Salvadoran" = 14, "Cuban" = 15, "Dominican" = 16,
  "Guatemalan" = 17, "Native Hawaiian" = 18, "Samoan" = 19, "Chamorro" = 20,
  "English" = 21, "German" = 22, "Irish" = 23, "Italian" = 24, "Polish" = 25,
  "Scottish" = 26, "2+ ethnicities selected" = 88
)

demographic_data_filtered <- demographic_data_filtered |>
  mutate(
    cleaned_ethn = str_remove_all(
      ethnicity,
      "\\b(Don't Know|Prefer Not To Answer|Not Applicable)\\b"
    ) |>
      str_replace_all(",+", ",") |>
      str_replace_all("^,|,$", "") |>
      na_if("")
  ) |>
  mutate(
    recoded_ethn = case_when(
      is.na(cleaned_ethn)              ~ NA,
      str_detect(cleaned_ethn, ",")    ~ 2L,
      TRUE                             ~ 1L
    ),
    ethn_num = case_when(
      is.na(cleaned_ethn)              ~ NA,
      str_detect(cleaned_ethn, ",")    ~ 88L,
      TRUE ~ unname(codebook_eth[cleaned_ethn])
    )
  )
```

---

# 3. Clean Discharge Data

## 3.1 Remove Demo/Test Accounts

Same as demographics — filter out any rows where the provider or client ID contains "DEMO".

```{r demo-filter-discharge}
discharge_data_filtered <- discharge_data |>
  filter(!if_any(c(provider_id, client_system_id), ~ str_detect(as.character(.x), "DEMO"))) |>
  janitor::clean_names()

# Should be FALSE — one client can have multiple assessment rows
length(unique(discharge_data_filtered$client_system_id)) == nrow(discharge_data_filtered)
```

## 3.2 Inspect Duplicate Client Records

Unlike demographics, it is expected for a client to appear more than once in the discharge report (e.g., multiple interventions or follow-ups). This chunk documents which columns vary across rows for the same client.

```{r discharge-duplicates}
discharge_data_filtered |>
  group_by(client_system_id) |>
  summarise(n = n()) |>
  filter(n > 1)

# Which fields differ across rows for the same client?
discharge_data_filtered |>
  group_by(client_system_id) |>
  filter(n() > 1) |>
  summarise(
    across(everything(), ~ {
      u <- unique(.x)
      if (length(u) > 1) paste(u, collapse = ", ") else NA_character_
    }),
    .groups = "drop"
  ) |>
  select(client_system_id, where(~ any(!is.na(.))))
```

## 3.3 Parse Pre/Post Measure Names

The `pre_measure_s` and `post_measure_s` fields combine the time point (e.g., "Initial", "Follow Up 1") and assessment name (e.g., "SDQ") into one string. We split these into separate columns so they can be matched and filtered independently.

A secondary trimmed name (first word only) is created because some providers abbreviate assessment names differently in the pre vs. post fields.

```{r parse-measure-names}
# Check raw values before parsing
unique(discharge_data_filtered$pre_measure_s)
unique(discharge_data_filtered$post_measure_s)
```

```{r clean-measure-names}
discharge_data_filtered <- discharge_data_filtered |>
  mutate(
    pre_time_point      = str_extract(pre_measure_s, "(Initial|(?i)follow up \\d+)"),
    pre_assessment_name = str_remove(pre_measure_s, "(Initial|(?i)follow up \\d+)") |>
      str_trim()
  ) |>
  mutate(
    post_time_point      = str_extract(post_measure_s, "(Initial|(?i)follow up \\d+)"),
    post_assessment_name = str_remove(post_measure_s, "(Initial|(?i)follow up \\d+)") |>
      str_trim()
  ) |>
  mutate(
    # Exact match check
    MatchDistinct = pre_assessment_name == post_assessment_name,
    # First-word match — handles cases where "SDQ" appears as "SDQ Strengths & Difficulties..."
    trimmed_pre_assessment_name  = word(pre_assessment_name, 1),
    trimmed_post_assessment_name = word(post_assessment_name, 1),
    trimmed_match = trimmed_pre_assessment_name == trimmed_post_assessment_name
  )

# Rows where even the trimmed names don't match — review these manually
discharge_data_filtered |>
  filter(trimmed_match == FALSE)
```

## 3.4 Fix Missing Funding Rounds

Some providers did not select a funding round in eInsight ("No Funding Round Selected"). Based on program records, we manually assign the correct round for known providers.

| Provider ID | Organization                   | Correct Round |
|-------------|--------------------------------|---------------|
| 604         | Monterey Health Dept           | 2             |
| 508         | Children of the Rainbow        | 2             |
| 607         | Youth for Change               | 1             |
| 573         | Homeless Children's Network    | 2             |
| 709         | Placer County Office of Education | 1          |

```{r fix-funding-rounds}
discharge_data_filtered_clean <- discharge_data_filtered |>
  mutate(
    funding_round = case_when(
      provider_id == "604" & funding_round == "No Funding Round Selected" ~ "2",
      provider_id == "508" & funding_round == "No Funding Round Selected" ~ "2",
      provider_id == "607" & funding_round == "No Funding Round Selected" ~ "1",
      provider_id == "573"                                                 ~ "2",
      provider_id == "709" & funding_round == "No Funding Round Selected" ~ "1",
      TRUE ~ funding_round
    )
  )
```

---

# 4. EDA — Clients with Multiple Interventions

Some clients appear in the discharge data more than once because they participated in more than one intervention (or had multiple follow-up time points). Understanding this structure is important before analysis, because rows are not independent.

## 4.1 Identify Clients with >1 Complete Intervention

"Complete" here means the row has both a pre AND a post measure (i.e., the client has a pre–post pair, not just a baseline or a lone follow-up).

```{r dup-complete-interventions}
# Summarize what varies across rows for each client with >1 complete pair
dup_client_system_id <- discharge_data_filtered |>
  filter(!is.na(pre_measure_s) & !is.na(post_measure_s)) |>
  group_by(client_system_id) |>
  filter(n() > 1) |>
  summarise(
    across(everything(), ~ {
      u <- unique(.x)
      if (length(u) > 1) paste(u, collapse = ", ") else NA_character_
    }),
    .groups = "drop"
  ) |>
  select(client_system_id, where(~ any(!is.na(.))))

# Pull all rows for those clients
dup_clients <- discharge_data_filtered |>
  filter(
    client_system_id %in% dup_client_system_id$client_system_id,
    !is.na(pre_measure_s) & !is.na(post_measure_s)
  )

cat("Clients with >1 complete pre-post pair:", n_distinct(dup_clients$client_system_id), "\n")
```

## 4.2 Concurrent Interventions (Same Pre-Completion Date)

Clients who had two or more different assessments completed on the same date — indicating they were enrolled in multiple interventions simultaneously.

```{r dup-same-time}
dup_clients_same_time <- dup_clients |>
  group_by(client_system_id, pre_completion_date) |>
  summarise(
    across(pre_measure_s, ~ {
      u <- unique(.x)
      if (length(u) > 1) paste(u, collapse = ", ") else NA_character_
    }),
    .groups = "drop"
  ) |>
  filter(!is.na(pre_measure_s))

# Full rows for those clients
dup_clients_same_time_data <- discharge_data_filtered |>
  filter(client_system_id %in% dup_clients_same_time$client_system_id)

cat("Clients with concurrent interventions:", nrow(dup_clients_same_time), "\n")
```

## 4.3 Sequential Interventions (Different Start Dates)

Clients who completed interventions at genuinely different time points — i.e., they started a new intervention after finishing a prior one.

```{r dup-diff-time}
dup_clients_diff_time <- dup_clients |>
  group_by(client_system_id) |>
  filter(n_distinct(intervention_start, na.rm = TRUE) > 1) |>
  summarise(
    n_intervention_start    = n_distinct(intervention_start, na.rm = TRUE),
    intervention_start_dates = paste(sort(unique(intervention_start)), collapse = ", "),
    pre_measure_s_values = {
      vals <- unique(na.omit(pre_measure_s))
      if (length(vals) == 0) NA_character_ else paste(vals, collapse = ", ")
    },
    .groups = "drop"
  )

dup_clients_diff_time_data <- discharge_data_filtered |>
  filter(client_system_id %in% dup_clients_diff_time$client_system_id)

cat("Clients with sequential interventions:", nrow(dup_clients_diff_time), "\n")

# Note: same-time + diff-time should sum to dup_clients; if not, some clients
# fall into neither category — this is an open issue to resolve.
cat("Same-time + diff-time total:", nrow(dup_clients_diff_time) + nrow(dup_clients_same_time), "\n")
cat("Total dup_clients:          ", n_distinct(dup_clients$client_system_id), "\n")
```

---

# 5. Join Demographic and Discharge Data

Left join from discharge data onto demographics so we retain all assessment rows. Clients without a demographics record will have `NA` in demographic columns but remain in the dataset.

ID columns need to be the same type (character) before joining; they can come in as numeric from Excel.

```{r join-data}
demographic_data_filtered <- demographic_data_filtered |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  )

discharge_data_filtered_clean <- discharge_data_filtered_clean |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  )

all_data <- left_join(
  discharge_data_filtered_clean,
  demographic_data_filtered,
  by = c("client_system_id", "provider_id", "provider_system_id")
)
```

---

# 6. Quick Counts

```{r counts}
cat("Unique clients in demographics (served):      ", n_distinct(demographic_data_filtered$client_system_id), "\n")
cat("Unique clients in joined dataset (outcomes):  ", n_distinct(all_data$client_system_id), "\n")
```

---

# 7. Write Clean CSVs

```{r write-csvs}
write_csv(
  all_data,
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv"
)

write_csv(
  demographic_data_filtered,
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv"
)

write_csv(
  discharge_data_filtered_clean,
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv"
)

cat("CSVs written successfully.\n")
```
