---
title: "Satisfaction Survey Data Cleaning"
author: "Cristin Young"
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** satisfaction survey data for analysis and reporting. It ingests four raw exports from the eInsight clinical management platform (one per respondent type), processes the two-row hierarchical headers, removes test/demo accounts, and writes one cleaned CSV per survey type for use by downstream analysis scripts.

**Data sources (eInsight exports):**

- **Satisfaction Survey — Caregiver** — Respondents who are caregivers of a child receiving services.
- **Satisfaction Survey — Child (13–17)** — Respondents who are youth ages 13–17.
- **Satisfaction Survey — Adult (18–59)** — Respondents who are adults ages 18–59.
- **Satisfaction Survey — Adult (60+)** — Respondents who are older adults age 60 and above.

Each file has a two-row header: row 1 contains question text (with merged cells exported as `NA`), and row 2 contains subheader labels (e.g., "Rating", "Comments"). These are combined into clean column names using `zoo::na.locf()` to forward-fill the question text across merged cells.

**Outputs:**

| File | Contents |
|----|----|
| `satisfaction_caregiver_clean.csv` | Cleaned caregiver survey responses |
| `satisfaction_child_clean.csv` | Cleaned child (13–17) survey responses |
| `satisfaction_adult_clean.csv` | Cleaned adult (18–59) survey responses |
| `satisfaction_senior_clean.csv` | Cleaned adult (60+) survey responses |
| `combined_satisfaction_clean.csv` | All four surveys combined, with cleaned dates, period filter, `year`, and `quarter_label` — **primary input for downstream analysis** |

------------------------------------------------------------------------

## Open Issues / Action Items

------------------------------------------------------------------------

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(zoo)
```

------------------------------------------------------------------------

# Helper Function

The header-processing logic is identical across all four survey files. We define a helper function to avoid repeating it four times.

```{r header-helper}
# Reads a satisfaction survey CSV with two-row headers and returns a clean
# data frame with combined column names and empty columns removed.
# Rows marked as DEMO are filtered out.
read_satisfaction_survey <- function(path, na_vals = c("-", "", "NA", "N/A")) {
  raw <- read_csv(path, na = na_vals, col_names = FALSE, show_col_types = FALSE,
                  col_types = cols(.default = "c"))

  h1    <- unlist(raw[1, ], use.names = FALSE)
  h2    <- unlist(raw[2, ], use.names = FALSE)
  h1_ff <- na.locf(h1, na.rm = FALSE) |> str_squish()

  new_names <- ifelse(
    is.na(h1_ff) | h1_ff == "",
    h2,
    paste(h1_ff, h2, sep = " - ")
  )
  new_names <- janitor::make_clean_names(new_names)

  df <- raw[-c(1, 2), ]
  names(df) <- new_names

  df |>
    select(where(~ !all(is.na(.)))) |>
    filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))
}
```

------------------------------------------------------------------------

# 1. Read and Clean Survey Files

## 1.1 Caregiver

```{r read-caregiver}
caregiver <- read_satisfaction_survey(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-caregiver.csv"
)

cat("Caregiver rows:", nrow(caregiver), "\n")
```

## 1.2 Child (13–17)

```{r read-child}
child <- read_satisfaction_survey(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-child_13-17.csv"
)

cat("Child rows:", nrow(child), "\n")
```

## 1.3 Adult (18–59)

```{r read-adult}
adult <- read_satisfaction_survey(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_18-59.csv"
)

cat("Adult rows:", nrow(adult), "\n")
```

## 1.4 Adult (60+)

```{r read-senior}
senior <- read_satisfaction_survey(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_60_plus.csv"
)

cat("Senior rows:", nrow(senior), "\n")
```

------------------------------------------------------------------------

# 2. Quick Checks

Row counts and a check for any remaining DEMO records after filtering.

```{r checks}
surveys <- list(caregiver = caregiver, child = child, adult = adult, senior = senior)

purrr::imap_dfr(surveys, ~ {
  tibble(
    survey       = .y,
    n_rows       = nrow(.x),
    n_demo       = sum(str_detect(as.character(.x$provider_id), "DEMO"), na.rm = TRUE),
    n_missing_id = sum(is.na(.x$client_system_id))
  )
})
```

------------------------------------------------------------------------

# 3. Create Combined Dataset

Bind all four survey types into a single data frame, clean malformed dates, filter to the reporting period, and add `quarter_label` and `year` columns. The `source` column identifies which survey each row came from.

**Date cleaning rationale:** Some date_completed values have year prefixes of `00` or `29` (e.g., "0024-03-15" or "2924-03-15"), likely from data entry errors in eInsight. These are corrected to `20` before converting to Date.

**Period filter:** Rows with no valid date, with `date_clean < "2024-01-01"`, or with `date_clean > today` are excluded.

```{r create-combined}
combined_all <- bind_rows(
  caregiver |> mutate(source = "caregiver"),
  child     |> mutate(source = "child"),
  adult     |> mutate(source = "adult"),
  senior    |> mutate(source = "senior")
) |>
  mutate(
    date_clean = str_replace(date_completed, "^00", "20"),
    date_clean = str_replace(date_clean,     "^29", "20"),
    date_clean = as.Date(date_clean)
  ) |>
  filter(
    !is.na(date_clean),
    date_clean >= "2024-01-01",
    date_clean <= Sys.Date()
  ) |>
  mutate(
    year          = year(date_clean),
    quarter_label = paste0("Q", quarter(date_clean), " ", year(date_clean))
  ) |>
  # Sort chronologically so quarter_label order is preserved when read back from CSV
  arrange(date_clean)

cat("Combined rows:", nrow(combined_all), "\n")
cat("Date range:   ", format(min(combined_all$date_clean, na.rm = TRUE)), "to",
                      format(max(combined_all$date_clean, na.rm = TRUE)), "\n")
cat("Missing dates:", sum(is.na(combined_all$date_clean)), "\n")
```

------------------------------------------------------------------------

# 4. Write Clean CSVs

Individual survey files are written for reference. The combined file is the primary input for downstream analysis.

```{r write-csvs}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(caregiver,    file.path(out_dir, "satisfaction_caregiver_clean.csv"))
write_csv(child,        file.path(out_dir, "satisfaction_child_clean.csv"))
write_csv(adult,        file.path(out_dir, "satisfaction_adult_clean.csv"))
write_csv(senior,       file.path(out_dir, "satisfaction_senior_clean.csv"))
write_csv(combined_all, file.path(out_dir, "combined_satisfaction_clean.csv"))

cat("CSVs written successfully.\n")
```
