---
title: "YOQ Assessment Data Cleaning"
author: "Cristin Young"
date: "2026-03-09"
format:
  html:
    df-print: paged
    toc: true
    toc-float: true
---

## Project Overview

This notebook prepares **CYBHI EBP/CDEP** Youth Outcome Questionnaire (YOQ) assessment data for analysis and reporting. It ingests two raw eInsight exports (caregiver and self-report versions), converts item scores to numeric, computes total scores, pairs initial and follow-up assessments, and computes the Reliable Change Index (RCI).

**Data sources:** `YOQ-caregiver.csv` and `YOQ-self_report.csv` — eInsight Individual Assessments exports. These files use standard single-row headers (`janitor::clean_names()` is sufficient; no two-row header processing needed).

**Scoring:**

- 30 items, each rated 0–4.
- Total score range: 0–120. Higher scores indicate greater distress.
- Total score is set to `NA` if 3 or more items are missing.
- **RCI = initial score − follow-up score.** RCI ≥ 10 = "doing reliably better."

**Pairing logic:** Clients are grouped by `client_system_id`. The first assessment is the initial; the last assessment (if ≥2 exist) is the follow-up. Clients with only one assessment are excluded.

**Outputs:**

| File | Contents |
|------|----------|
| `yoq_caregiver_paired_clean.csv` | One row per client: initial/follow-up scores, `rci`, `reliably_better`, quarter labels |
| `yoq_self_paired_clean.csv` | Same, self-report version |

---

## Open Issues / Action Items

- [ ] No date range filter is applied to YOQ (unlike NOMS which filters to 2024–2025). Confirm whether a period filter is needed here.
- [ ] The `rci_col` variable in the main analysis uses `yoq_caregiver_paired[,7]` (column 7 by position) to identify reliable changers. This is fragile — consider whether to add an explicit named column here for robustness.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
```

---

# Helper Function

Both YOQ files share identical scoring and pairing logic. We define a single helper to avoid duplication.

```{r yoq-helper}
# Reads a YOQ CSV, scores items, and returns a wide paired data frame.
clean_yoq <- function(path) {
  raw <- read_csv(path, na = c("-", "", "NA", "N/A"), show_col_types = FALSE) |>
    janitor::clean_names()

  scored <- raw |>
    mutate(across(matches("^x"), as.numeric)) |>
    rowwise() |>
    mutate(
      missing_count = sum(is.na(c_across(starts_with("x")))),
      total_score   = if_else(
        missing_count >= 3,
        NA_real_,
        sum(c_across(starts_with("x")), na.rm = TRUE)
      )
    ) |>
    ungroup() |>
    mutate(
      date_completed = as.Date(date_completed),
      row_id         = row_number()
    ) |>
    select(client_system_id, collection, date_completed, total_score)

  scored |>
    group_by(client_system_id) |>
    arrange(date_completed, .by_group = TRUE) |>
    summarise(
      n_assessments   = n(),
      initial_date    = first(date_completed),
      initial_score   = first(total_score),
      follow_up_date  = if_else(n_assessments >= 2, last(date_completed),  as.Date(NA)),
      follow_up_score = if_else(n_assessments >= 2, last(total_score),     NA_real_)
    ) |>
    ungroup() |>
    filter(!is.na(follow_up_date)) |>
    mutate(
      # RCI = baseline − follow-up; positive = doing better (lower distress)
      rci              = initial_score - follow_up_score,
      reliably_better  = rci >= 10,
      initial_quarter  = paste0("Q", quarter(initial_date),    " ", year(initial_date)),
      follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))
    )
}
```

---

# 1. YOQ Caregiver

```{r yoq-caregiver}
yoq_caregiver_paired <- clean_yoq(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/YOQ-caregiver.csv"
)

cat("YOQ Caregiver paired clients:", nrow(yoq_caregiver_paired), "\n")

yoq_caregiver_paired |>
  filter(!is.na(reliably_better), rci != 0) |>
  summarise(
    n             = n(),
    n_better      = sum(reliably_better),
    pct_better    = round(100 * n_better / n, 1)
  )
```

---

# 2. YOQ Self-Report

```{r yoq-self}
yoq_self_paired <- clean_yoq(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Individual Assessments/YOQ-self_report.csv"
)

cat("YOQ Self paired clients:", nrow(yoq_self_paired), "\n")

yoq_self_paired |>
  filter(!is.na(reliably_better), rci != 0) |>
  summarise(
    n          = n(),
    n_better   = sum(reliably_better),
    pct_better = round(100 * n_better / n, 1)
  )
```

---

# 3. Write Clean CSVs

```{r write-csvs}
out_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

write_csv(yoq_caregiver_paired, file.path(out_dir, "yoq_caregiver_paired_clean.csv"))
write_csv(yoq_self_paired,      file.path(out_dir, "yoq_self_paired_clean.csv"))

cat("CSVs written successfully.\n")
```
