---
title: "DHCS Dyadic Care Model Report-02-10-2026"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Dyadic Care Data

Interventions are:

-   HealthySteps

-   Parent-Child Interaction Therapy (PCIT)

-   Attachment and Biobehavioral Catch-up (ABC)

-   Child-Parent Psychotherapy (CPP)

### Participants Served

**Data used**: Discharge data for above four interventions, which includes XX clients that had more than one complete intervention.

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered for `intervention_name ==` as above.

**UPDATE:**

-   funding rounds for discharge data

-   Update EVERYTHING with date_completed_response

## Setup

```{r}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(janitor)
library(openxlsx)

# reading in data
discharge_data_filtered <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv") 

all_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv") 

demographic_data_filtered <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv") 

# narrowing to desired grantees

## for those with interventions
cdeps <- c("HealthySteps", "PCIT (Parent-Child Interaction Therapy)", "CPP (Child Parent Psychotherapy)", "ABC (Attachment and Biobehavioral Catch-Up)")

discharge_cdep <- discharge_data_filtered %>% 
  filter(intervention_name %in% cdeps)

## get their provider ids because the demo data doesn't have provider names
discharge_cdep_ids <- discharge_cdep %>% 
  distinct(provider_id)

## add in 470, 076, 190
discharge_cdep_ids <- discharge_cdep_ids %>% 
  add_row(provider_id = c("470", "076", "190"))

## filter demo data for cdep ids
demo_cdep <- demographic_data_filtered %>% 
  filter(provider_id %in% discharge_cdep_ids$provider_id)

all_data_cdep <- all_data %>% 
  filter(provider_id %in% discharge_cdep_ids$provider_id,
         intervention_name %in% cdeps)
```

## Results

### Number and % enrolled

```{r}
#| echo: false
#| warning: false
#| message: false

# grantees 
discharge_cdep %>% 
  distinct(provider, .keep_all = TRUE) %>% 
  tabyl(funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

#by funding round
discharge_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
  tabyl(funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

#by intervention
discharge_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
  tabyl(intervention_name) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

#by intervention and funding round
discharge_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
  tabyl(intervention_name, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()
```

### Number and % discharged

```{r}
#| echo: false
#| warning: false
#| message: false

#by intervention
discharge_cdep %>% 
  filter(!is.na(intervention_end)) %>% 
  tabyl(intervention_name) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

#by discharge reason
discharge_cdep %>% 
  filter(!is.na(intervention_end)) %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

# by discharge reason and funding round
discharge_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
   # filter(!is.na(reason_for_discharge),
   #     reason_for_discharge != "No Reason Selected",
   #    intervention_end <= "2025-12-31") %>%
  tabyl(reason_for_discharge, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()

# by discharge reason and intervention
discharge_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
    filter(!is.na(reason_for_discharge)) %>% 
   #     reason_for_discharge != "No Reason Selected",
   #    intervention_end <= "2025-12-31") %>%
  tabyl(reason_for_discharge, intervention_name) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()
```

### Number (%) of client type categories

```{r}
#| echo: false
#| warning: false
#| message: false

# by client_type
demo_cdep %>% 
  #filter(!is.na(client_type)) %>% 
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(client_type) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

# by intervention and funding round
all_data_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(client_type, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()

# by intervention and client type
all_data_cdep %>% 
  filter(!is.na(intervention_start)) %>% 
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(client_type, intervention_name) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()
```

## Extras

### Primary Language

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
demo_cdep %>% 
  # filter(primary_language != "Don't Know",
  #        primary_language != "No Primary Language Selected",
  #        primary_language != "Not Applicable",
  #        primary_language != "Prefer Not To Answer",
  #        !is.na(primary_language)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(primary_language) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Housing Status

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
demo_cdep %>% 
  filter(housing_status != "Don't Know",
         housing_status != "No Housing Status Selected",
         housing_status != "Not Applicable",
         housing_status != "Prefer Not To Answer",
         !is.na(housing_status)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(housing_status) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Number (%) of client type categories with average age for each category

```{r}
#| echo: false
#| warning: false
#| message: false

# first get client type counts
tab_counts <- demo_cdep %>%
  #filter(client_type != "No Client Type Selected",
   #      !is.na(client_type)) %>% 
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(client_type) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1)

tab_counts

# then average age for each category
demo_cdep_distinct <- demo_cdep %>%
  mutate(
    age_start_tx_num = case_when(
      age_at_start_of_treatment_in_years %in% c("No Age Provided", "", " ") ~ NA_character_,
      TRUE ~ age_at_start_of_treatment_in_years
    ) %>% as.numeric()) %>% 
  distinct(client_system_id, .keep_all = TRUE) 
    
tab_age <- demo_cdep_distinct %>%
  group_by(client_type) %>%
  summarise(
    mean_age = round(mean(age_start_tx_num, na.rm = TRUE), 1),
    n_with_age = sum(!is.na(age_start_tx_num)),
    .groups = "drop"
  )

tab_age

# create final table
final_table <- tab_counts %>%
  left_join(tab_age, by = "client_type") %>%
  select(
    client_type,
    n,
    percent,
    mean_age,
    n_with_age
  )

final_table

# client type by intervention
demo_cdep %>%
  #filter(client_type != "No Client Type Selected",
   #      !is.na(client_type)) %>% 
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(client_type) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1)
```

### Insurance Status

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
demo_cdep %>% 
  filter(insurance_status != "Don't Know",
         insurance_status != "No Insurance Status Selected",
         insurance_status != "Not Applicable",
         insurance_status != "Prefer Not To Answer",
         !is.na(insurance_status)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(insurance_status) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Gender Identity

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
demo_cdep %>% 
  # filter(gender_identity != "Don't Know",
  #        gender_identity != "No Gender Identify Selected",
  #        gender_identity != "Not Applicable",
  #        gender_identity != "Prefer Not To Answer",
  #        !is.na(gender_identity)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(gender_identity) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Sexual Orientation

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
demo_cdep %>% 
 # filter(sexual_orientation != "Don't Know",
 #        sexual_orientation != "No Sexual Orientation Selected",
 #        sexual_orientation != "Not Applicable",
 #        sexual_orientation != "Prefer Not To Answer",
 #        !is.na(sexual_orientation)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(sexual_orientation) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Race

1.  Asian

2.  Black or African American

3.  Hispanic or Latino

4.  Middle Eastern or North African

5.  Native American/American Indian or Alaska Native

6.  Native Hawaiian or Pacific Islander

7.  White

8.  2+ races selected

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
demo_cdep %>% 
filter(race != "Don't Know",
       race != "No Race Selected",
       race != "Not Applicable",
       race != "Prefer Not To Answer",
       !is.na(race)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(code_num) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Ethnicity

1.  Chinese
2.  Asian Indian
3.  Filipino
4.  Vietnamese
5.  Korean
6.  Japanese
7.  African American
8.  Jamaican
9.  Haitian
10. Nigerian
11. Somali
12. Mexican
13. Puerto Rican
14. Salvadoran
15. Cuban
16. Dominican
17. Guatemalan
18. Native Hawaiian
19. Samoan
20. Chamorro
21. English
22. German
23. Irish
24. Italian
25. Polish
26. Scottish
27. 2+ ethnicities selected = 88

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_cdep %>% 
# filter(ethnicity != "Don't Know",
#        ethnicity != "No Ethnicity Selected",
#        ethnicity != "Not Applicable",
#        ethnicity != "Prefer Not To Answer",
#        !is.na(ethnicity)) %>%
  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(ethn_num) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

## Satisfaction survey stuff

\% percentage satisfied by intervention - Q1 and whole survey

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
library(zoo)

# caregiver
caregiver_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-caregiver.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(caregiver_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(caregiver_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
caregiver <- caregiver_raw[-c(1, 2), ]
names(caregiver) <- new_names
caregiver <- caregiver %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# child
child_raw <- read.csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-child_13-17.csv", na = c("-", "", "NA", "N/A", "0"), header = FALSE) 
h1 <- unlist(child_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(child_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
child <- child_raw[-c(1, 2), ]
names(child) <- new_names
child <- child %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# adult
adult_raw <- read.csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_18-59.csv", na.strings = c("-", "", "NA", "N/A", "0"), skip = 0, header = FALSE)
h1 <- unlist(adult_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(adult_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
adult <- adult_raw[-c(1, 2), ]
names(adult) <- new_names
adult <- adult %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# senior
senior_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_60_plus.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(senior_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(senior_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
senior <- senior_raw[-c(1, 2), ]
names(senior) <- new_names
senior <- senior %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree" on the satisfaction survey. The second definition is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed.

**Nitty gritty:** There are four satisfaction surveys: Caregiver (n=`{r} nrow(caregiver)`), Child (13-17) (n=`{r} nrow(child)`), Adult (18-59) (n=`{r} nrow(adult)`), and Adult (60+) (n=`{r} nrow(senior)`). A separate data frame was made for question 1, and high satisfaction was defined as the percent of individuals who answered a 4 or 5 for question 1.

### Question 1 Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
combined_q1 <- bind_rows(
  caregiver %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "caregiver"),
  child %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "child"),
  adult %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "adult"),
  senior %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "senior"))

all_data_cdep_distinct <- all_data_cdep %>% 
  distinct(client_system_id, .keep_all = TRUE) 

all_data_cdep_distinct <- all_data_cdep_distinct %>% 
  mutate(provider_system_id = as.character(provider_system_id),
         client_system_id = as.character(client_system_id))

combined_q1 <- combined_q1 %>% 
  mutate(provider_system_id = as.character(provider_system_id),
         client_system_id = as.character(client_system_id))

combined_q1 <- left_join(combined_q1, all_data_cdep_distinct, by = c("provider", "provider_id", "provider_system_id", "client_system_id")) 
combined_q1 <- combined_q1 %>% 
  filter(!is.na(intervention_name))
```

The following is the % of high satisfaction by intervention:

```{r}
#| echo: false
#| warning: false
#| message: false

combined_q1 %>%
  group_by(intervention_name) %>%
  summarise(
    n_total = sum(!is.na(q12)),
    n_4_5 = sum(q12 %in% c(4, 5), na.rm = TRUE),
    pct_4_5 = round((100 * n_4_5 / n_total), 1)
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
pct_4_5 <- combined_q1 %>%
  summarise(
    pct_4_5 = round((100 * sum(q12 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q12))), 1)
  ) %>%
  pull(pct_4_5)

pct_4_5
```

Overall, `{r} pct_4_5`% of survey respondents were highly satisfied based on question 1 answers. The following table shows high survey satisfaction by round.

```{r}
#| echo: false
#| warning: false
#| message: false
library(readxl)
library(janitor)

# combining all satisfaction survey data
combined_all <- bind_rows(
  caregiver %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "caregiver"),
  child %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "child"),
  adult %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "adult"),
  senior %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "senior"))

combined_all <- combined_all %>% 
  mutate(provider_system_id = as.character(provider_system_id),
         client_system_id = as.character(client_system_id))

combined_all <- left_join(combined_all, all_data_cdep_distinct, by = c("provider", "provider_id", "provider_system_id", "client_system_id")) 

combined_all <- combined_all %>% 
  filter(!is.na(intervention_name))
```

### Combined Survey Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

summary_agree <- combined_all %>%
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) 

df2 <- summary_agree %>% 
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4,5), na.rm = TRUE),
    pct_agree = n_agree / n_answered
  ) %>%
  ungroup() %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df2
```

Overall, `{r} df4$pct_meeting_80pct`% of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed, for those who completed a survey on or before February 10, 2026.

The following table shows high survey satisfaction results by funding round, for data where both client_id and `intervention_name` were available.

```{r}
#| echo: false
#| warning: false
#| message: false
summary_agree_intervention <- combined_all %>%
  group_by(intervention_name) %>% 
  filter(!is.na(intervention_name)) %>% 
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4, 5)),
    pct_agree = n_agree / n_answered
  ) %>%
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

summary_agree_intervention
```

### Preferred Language by Satisfaction Survey

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** This KPI represents ***the percent*** of clients that report services received in their preferred language on the satisfaction survey out of the total that completed the satisfaction survey by period(s).

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (\<= "2026-02-10"). I filtered out NAs for both the quarter label and the question first before calculating %.

The following table shows the percent of clients who received services in their preferred language on the satisfaction survey by quarter:

```{r}
#| echo: false
#| warning: false
#| message: false
# combining all satisfaction survey data
client_lang_all <- bind_rows(
  caregiver %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "caregiver"),
  child %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "child"),
  adult %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "adult"),
  senior %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) %>% 
    mutate(source = "senior"))

# first clean date column
client_lang_all <- client_lang_all %>% 
  mutate(
    date_clean = str_replace(date_completed, "^00", "20"), # replace 00xx with 20xx
    date_clean = str_replace(date_clean, "^29", "20"), # replace 29xx with 20xx
    date_clean = as.Date(date_clean)
  )

# add in a quarter column 
client_lang_all <- client_lang_all %>% 
  filter(!is.na(date_clean),
         date_clean <= Sys.Date()) %>% 
  mutate(quarter_label = paste0("Q", quarter(date_clean), " ", year(date_clean)),
       quarter_label = factor(quarter_label, levels = unique(quarter_label)) #to sort chronologically
  )

client_lang_all <- client_lang_all %>% 
  mutate(provider_system_id = as.character(provider_system_id),
         client_system_id = as.character(client_system_id))

client_lang_all <- left_join(client_lang_all, all_data_cdep_distinct, by = c("provider", "provider_id", "provider_system_id", "client_system_id")) 

client_lang_all <- client_lang_all %>% 
  filter(!is.na(intervention_name))

table(client_lang_all$were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response, useNA = "ifany")
```

## Participant outcomes

### Number and % improved outcome

```{r}
#| echo: false
#| warning: false
#| message: false

# make improved outcome variable
assessment_rules <- tibble::tribble(
  ~trimmed_pre_assessment_name, ~higher_is_better, ~min_change,
  "CATS",          FALSE,            1,
  "GAD-7",         FALSE,            1,
  "NOMS",          TRUE,             1,
  "Pediatric",     FALSE,            1,
  "PHQ-9",         FALSE,            1,
  "PSS",           FALSE,            1,
  "SDQ",           FALSE,            1,
  "YOQ",           FALSE,            1
)

# create sign change where positive always means better
discharge_cdep_outcome <- discharge_cdep %>%
  left_join(assessment_rules, by = "trimmed_pre_assessment_name") %>%
  mutate(
    raw_change = post_score - pre_score,
    better_change = if_else(higher_is_better, raw_change, -raw_change),
    min_change = coalesce(min_change, 0),  # default if you didn’t set one
    outcome = case_when(
      is.na(pre_score) | is.na(post_score) ~ NA_character_,
      better_change >=  min_change ~ "improved",
      better_change <= -min_change ~ "worsened",
      TRUE                         ~ "no_change"
    ),
    outcome = factor(outcome, levels = c("worsened", "no_change", "improved"))
  )

# checks
discharge_cdep_outcome %>% 
  count(trimmed_pre_assessment_name, outcome)
discharge_cdep_outcome %>% 
  group_by(trimmed_pre_assessment_name) %>% summarise(missing_rules = sum(is.na(higher_is_better)))

table(discharge_cdep_outcome$outcome)

# by outcome
discharge_cdep_outcome %>% 
 # distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(outcome) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

# by intervention and outcome
discharge_cdep_outcome %>% 
  filter(!is.na(intervention_start),
         !is.na(outcome)) %>% 
         #!is.na(outcome)) %>% 
#  distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(intervention_name, outcome) %>% 
  adorn_totals("row") %>% 
  adorn_totals("col") %>% 
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()

  # by outcome and funding round
discharge_cdep_outcome %>% 
 # filter(!is.na(intervention_start)) %>% 
 # distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(outcome, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()

# by outcome and measure
discharge_cdep_outcome %>% 
  filter(!is.na(outcome)) %>% 
  #distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(post_assessment_name, outcome) %>% 
  adorn_totals("row") %>% 
  adorn_totals("col") %>% 
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()
```

## % of clients who met treatment goals

Two criteria must be met for a client/individual to have ‘met treatment goals’: 1) a discharge reason of Mutually agreed to cessation of treatment and 2) an improvement according to pre and post measures collected (i.e., the NOMS sub-scales). For this metric, a simple difference between the pre and post score (in the positive direction depending on the tool) is used to determine a positive outcome.

```{r}
#| echo: false
#| warning: false
#| message: false
#| 
discharge_cdep_outcome %>% 
    filter(outcome == "no_change" | outcome == "improved") %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

```
