---
title: "Quarterly Report Results-1-12-2026"
author: "Cristin Young"
format: 
  docx:
    page-width: 7.5
editor: visual
---

## Outcomes Data by Round

### Participants Served

**Data used**: Discharge data, which includes 382 clients that had more than one complete intervention.

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered only for `intervention_start <= "2025-12-31"`

**External report to DHCS**: by round,

-   [x] \# of clients served;

-   [x] percent of cases with improved outcomes;

-   [x] percent with satisfaction with services;

-   [x] percent with successful discharges.

```{r}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(haven)

# load in all data
discharge_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv") %>% 
  filter(intervention_start <= "2025-12-31")
all_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv") %>% 
    filter(intervention_start >= "2024-01-01",
           intervention_start <= "2025-12-31")

discharge_data_distinct <- discharge_data %>% 
  filter(intervention_start <= "2025-12-31") %>% 
  distinct(client_system_id, .keep_all = TRUE)

all_data_distinct <- all_data %>% 
  filter(intervention_start >= "2024-01-01",
    intervention_start <= "2025-12-31") %>% 
  distinct(client_system_id, .keep_all = TRUE)

#number of grantees
all_data %>%
  summarise(n_distinct = n_distinct(provider_system_id))

# primary language
# housing status
# client type
# average age by client type

# load in spss data
master_grantee_file <- haven::read_spss("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/MasterGranteeSpreadsheet_forBHESQuantTeam_1-20-26_NEW!.sav") %>%
  zap_missing() %>%
  mutate(across(where(is.labelled), as.character)) %>%
  mutate(across(where(is.character), ~ na_if(str_trim(.x), ""))) %>%
  rename_with(~ gsub("^@", "", .x)) %>%
  janitor::clean_names() %>%
  rename_with(~ gsub("^x", "", .x))

# join to all data
# all_data_spss <- left_join(all_data, master_grantee_file, by = c("provider" = "1a_grantee_provider_name", "provider_id" = "2grantee_provider_id", "intervention_name" = "9ebpcdep"))
# 
# participants served
part_served <- table(discharge_data$funding_round, useNA = "ifany")
part_served
```

### Improved outcomes

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
library(janitor)
# number of rows in discharge data 
nrow(discharge_data)

# number of rows with positive_outcomes != NA
count_outcomes <- nrow(discharge_data %>% filter(!is.na(positive_outcomes)))
count_outcomes

# table of positive outcomes 
tab <- discharge_data %>%
  filter(!is.na(intervention_start),
         !is.na(intervention_end),
         !is.na(reason_for_discharge)) %>% 
  tabyl(positive_outcomes) %>%               
  adorn_pct_formatting(digits = 1)
tab

For rows that had valid data for: reason for discharge, intervention name, pre and post measures, and location, started an intervention between 01-01-2024 and 12-31-2025, and had a positive outcome defined as improvement in pre and post measures collected.

# percent of positive outcomes
yes_pct <- tab %>%
  filter(positive_outcomes == "Yes") %>%
  pull(valid_percent)
```

Out of `{r} nrow(all_data_spss)` rows of clients with an intervention start date prior to 11-30-2025, `{r} count_pos` had both a pre- and post-score. Of those clients with complete intervention data in EInsight, `{r} yes_pct` had a positive outcome.

The results below show % of positive outcomes by round, once NA's are removed from the data.

**Nitty gritty**: Positive outcome was defined as \[ this needs to be confirmed with Rebecca/Steve; this is a column from EInsight and I believe is just defined as "Is the post-score lower/higher than the pre-score, based on the relevant assessment scoring metrics given to Steve's team"\].

```{r}
#| echo: false
#| warning: false
#| message: false
# table of positive outcomes by funding round
tab2 <- discharge_data %>% 
  filter(!is.na(positive_outcomes)) %>% 
  tabyl(positive_outcomes, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() 
tab2
```

### Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
library(zoo)

# caregiver
caregiver_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-caregiver.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(caregiver_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(caregiver_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
caregiver <- caregiver_raw[-c(1, 2), ]
names(caregiver) <- new_names
caregiver <- caregiver %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# child
child_raw <- read.csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-child_13-17.csv", na = c("-", "", "NA", "N/A", "0"), header = FALSE) 
h1 <- unlist(child_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(child_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
child <- child_raw[-c(1, 2), ]
names(child) <- new_names
child <- child %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# adult
adult_raw <- read.csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_18-59.csv", na.strings = c("-", "", "NA", "N/A", "0"), skip = 0, header = FALSE)
h1 <- unlist(adult_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(adult_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
adult <- adult_raw[-c(1, 2), ]
names(adult) <- new_names
adult <- adult %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))

# senior
senior_raw <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/eInsight Exports/Satisfaction Survey Export/Satisfaction_survey-adult_60_plus.csv", na = c("-", "", "NA", "N/A", "0"), col_names = FALSE) 
h1 <- unlist(senior_raw[1, ], use.names = FALSE)           # question text row
h2 <- unlist(senior_raw[2, ], use.names = FALSE)           # subheaders
h1_ff <- na.locf(h1, na.rm = FALSE)                 # carries last non-NA to the right
h1_ff <- str_squish(h1_ff)
new_names <- ifelse(
  is.na(h1_ff) | h1_ff == "",
  h2,                                  # keep subheader only
  paste(h1_ff, h2, sep = " - ")        # combine when applicable
)
new_names <- janitor::make_clean_names(new_names)
senior <- senior_raw[-c(1, 2), ]
names(senior) <- new_names
senior <- senior %>%
  filter(!if_any(c(provider_id), ~ str_detect(as.character(.x), "DEMO")))
```

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree" on the satisfaction survey. The second definition is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed.

**Nitty gritty:** There are four satisfaction surveys: Caregiver (n=`{r} nrow(caregiver)`), Child (13-17) (n=`{r} nrow(child)`), Adult (18-59) (n=`{r} nrow(adult)`), and Adult (60+) (n=`{r} nrow(senior)`). A separate data frame was made for question 1, and high satisfaction was defined as the percent of individuals who answered a 4 or 5 for question 1.

#### Question 1 Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
combined_q1 <- bind_rows(
  caregiver %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "caregiver"),
  child %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "child"),
  adult %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "adult"),
  senior %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x1_")) %>% 
    rename(q1 = matches("^x1_")) %>% 
    mutate(source = "senior"))

combined_q1 <- combined_q1 %>% 
  filter(date_completed <= "2025-12-31")
```

The following is the % of high satisfaction by source:

```{r}
#| echo: false
#| warning: false
#| message: false

combined_q1 %>%
  group_by(source) %>%
  summarise(
    n_total = sum(!is.na(q12)),
    n_4_5 = sum(q12 %in% c(4, 5), na.rm = TRUE),
    pct_4_5 = round((100 * n_4_5 / n_total), 1)
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
pct_4_5 <- combined_q1 %>%
  summarise(
    pct_4_5 = round((100 * sum(q12 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q12))), 1)
  ) %>%
  pull(pct_4_5)

pct_4_5
```

Overall, `{r} pct_4_5`% of survey respondents were highly satisfied based on question 1 answers. The following table shows high survey satisfaction by round.

```{r}
#| echo: false
#| warning: false
#| message: false
library(readxl)
library(janitor)
# reading in provider data for funding round. this doesn't have client_system_id though
# provider_demo <- readxl::read_xlsx("/Users/cristin/Dropbox/Data/eInsight Exports/Admin Report Exports/2025-10-29-provider_demographics.xlsx", na = c("-", "", "NA", "N/A")) %>% 
#   janitor::clean_names() %>% 
#   select(provider, provider_id, provider_system_id, funding_round)

# # getting funding round via discharge data because this has client_system_id
# funding <- discharge_data_filtered %>% 
#   select(provider, provider_id, provider_system_id, client_system_id, funding_round) %>% 
#   #distinct(provider_id, provider, provider_system_id, funding_round, .keep_all = TRUE) %>% 
#   filter(!is.na(funding_round))
#          #!str_detect(funding_round, ","))
# funding %>% filter(str_detect(funding_round, ","))
# 
# provider_unique <- provider_demo %>%
#   distinct(provider_id, provider, provider_system_id, funding_round, .keep_all = TRUE) %>% 
#   filter(!is.na(funding_round),
#          !str_detect(funding_round, ","))

# combining all satisfaction survey data
combined_all <- bind_rows(
  caregiver %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "caregiver"),
  child %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "child"),
  adult %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "adult"),
  senior %>% 
    select(provider, provider_id, provider_system_id, client_system_id, collection, cps_id, 
           date_completed, matches("^x")) %>% 
    rename(q = matches("^x")) %>% 
    mutate(source = "senior"))
combined_all <- combined_all %>% 
  filter(date_completed <= "2025-12-31") 

# get funding round data
combined_round <- combined_all %>% 
  left_join(master_grantee_file,
            by = c("provider" = "1a_grantee_provider_name", 
                   "provider_id" = "2grantee_provider_id")) %>% 
  distinct(cps_id, date_completed, .keep_all = TRUE) 

combined_round <- combined_all %>% 
  left_join(discharge_data %>%  
              mutate(provider_system_id = as.character(provider_system_id),
                     client_system_id = as.character(client_system_id)),
            by = c("provider", "provider_id", "provider_system_id", "client_system_id")) 
#%>% 
   #distinct(cps_id, date_completed, .keep_all = TRUE) 

# Question 1 satisfaction by round
pct_4_5_round <- combined_round %>% 
  filter(!is.na(funding_round)) %>% 
  group_by(funding_round) %>% 
  summarise(
    n_surveys = n(),
    pct_4_5 = round((100 * sum(q2 %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q2))), 1),
    .groups = "drop"
  )

pct_4_5_round 
```

#### Combined Survey Satisfaction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

summary_agree <- combined_all %>%
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) 

df2 <- summary_agree %>% 
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4,5), na.rm = TRUE),
    pct_agree = n_agree / n_answered
  ) %>%
  ungroup() %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df2

# CONFIRMED THIS WORKS THE SAME AS ABOVE
df3 <- summary_agree %>%   
  mutate(across(everything(), ~ na_if(., 0))) %>% 
  rowwise() %>% 
  mutate(count_4 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "4", na.rm = TRUE), 
         count_5 = sum(c(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) == "5", na.rm = TRUE),
         n_4_5 = sum(c(count_4, count_5), na.rm = TRUE),
         n_total = sum(!is.na(c_across(starts_with("q")))),
         perc_4_5 = n_4_5 / n_total
         ) %>% 
  ungroup() 

df4 <- df3 %>% 
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(perc_4_5 >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

df4
```

Overall, `{r} df4$pct_meeting_80pct`% of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed, for those who completed a survey on or before November 30, 2025.

The following table shows high survey satisfaction results by funding round, for data where both client_id and `funding_round` were available.

```{r}
#| echo: false
#| warning: false
#| message: false
summary_agree_round <- combined_round %>%
  group_by(`4funding_round`) %>% 
  filter(!is.na(`4funding_round`)) %>% 
  # keep only question columns
  select(q2, q4, q6, q8, q10, q12, q14, q16, q18, q20, q22, q24, q26, q28, q30) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>% 
  # calculate per-person proportion of "Agree"/"Strongly Agree"
  rowwise() %>%
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree = sum(c_across(everything()) %in% c(4, 5)),
    pct_agree = n_agree / n_answered
  ) %>%
  summarise(
    n_total = n(),
    n_meeting_80pct = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round((100 * n_meeting_80pct / n_total), 1)
  )

summary_agree_round

# q_cols <- paste0("q", c(2,4,6,8,10,12,14))
# 
# by_round <- combined_round %>%
#   # ensure numeric just in case
#   select(q2, q4, q6, q8, q10, q12, q14, funding_round) %>% 
#   mutate(across(all_of(q_cols), ~ as.numeric(.))) %>%
#   rowwise() %>%
#   mutate(
#     n_answered = sum(!is.na(c_across(all_of(q_cols)))),
#     n_45       = sum(c_across(all_of(q_cols)) >= 4, na.rm = TRUE),
#     needed     = ceiling(0.80 * n_answered),           # ≥80% threshold
#     high_sat   = n_answered > 0 & n_45 >= needed
#   ) %>%
#   ungroup() %>%
#   group_by(funding_round) %>%
#   summarise(
#     n_resp        = sum(n_answered > 0),               # respondents with ≥1 answer
#     n_high_sat    = sum(high_sat),
#     pct_high_sat  = round(100 * n_high_sat / n_resp, 1)
#   )
# 
# by_round
```

### Successful discharges

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# number of rows with a reason for discharge
count_rows <- nrow(discharge_data %>% filter(!is.na(reason_for_discharge)))
count_rows

# number of rows with successful discharge
# table of positive outcomes 
tab3 <- discharge_data %>%
  filter(
    !is.na(reason_for_discharge),
         reason_for_discharge != "No Reason Selected") %>% 
  tabyl(reason_for_discharge) %>%               
  adorn_pct_formatting(digits = 1)
tab3

# count of successful discharge
yes_count <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(n)
yes_count

# percent of successful discharge
yes_pct3 <- tab3 %>%
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)") %>%
  pull(percent)
yes_pct3
```

Successful discharges were defined as those clients that had "Mutually agreed cessation of treatment (Successful Completion)" as the `reason_for_discharge` in the discharge data from EInsight. Of `{r} nrow(all_data_spss)` rows of discharge data, `{r} count_rows` rows had a reason for discharge stated. Of individuals with a reason for discharge stated, `{r} yes_count` (`{r} yes_pct3`) were successfully discharged.

The results below show % of successful discharges by round, once NA's are removed from the data.

```{r}
#| echo: false
#| warning: false
#| message: false
# total table of successful discharges by funding round
tab4 <- discharge_data %>% 
  filter(!is.na(reason_for_discharge)) %>% 
  tabyl(reason_for_discharge, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() 
tab4
```

How does Rikke's SPSS file compare to the EInsight `funding_round` variable?

```{r}
#| echo: false
#| warning: false
#| message: false
all_data_spss %>% 
  tabyl(funding_round, `4funding_round`)
```

## Final table

|  |  |  |  |  |  |  |  |
|---------|---------|---------|---------|---------|---------|---------|---------|
| Round | Theme | Participants Served | Improved Outcomes | Satisfaction - Q1 | Satisfaction - Overall | Successful Discharges | Quote |
| 1 | Parent/Caregivers Programs and Practices | `{r} part_served[1]` | `{r} tab2[[2]][2]` | `{r} pct_4_5_round[["pct_4_5"]][1]` | `{r} summary_agree_round$pct_meeting_80pct[1]` | `{r} tab4[[2]][3]` |  |
| 2 | Trauma-Informed Programs and Practices | `{r} part_served[2]` | `{r} tab2[[3]][2]` | `{r} pct_4_5_round$pct_4_5[2]` | `{r} summary_agree_round$pct_meeting_80pct[2]` | `{r} tab4[[3]][3]` |  |
| 3 | Early Childhood Wraparound Services | `{r} part_served[3]` | `{r} tab2[[4]][2]` | `{r} pct_4_5_round$pct_4_5[3]` | `{r} summary_agree_round$pct_meeting_80pct[3]` | `{r} tab4[[4]][3]` |  |
| 4 | Youth-Driven Programs | `{r} part_served[4]` | `{r} tab2[[5]][2]` | `{r} pct_4_5_round$pct_4_5[4]` | `{r} summary_agree_round$pct_meeting_80pct[4]` | `{r} tab4[[5]][3]` |  |
| 5 | Early Intervention Programs and Practices | `{r} part_served[5]` | `{r} tab2[[6]][2]` | `{r} pct_4_5_round$pct_4_5[5]` | `{r} summary_agree_round$pct_meeting_80pct[5]` | `{r} tab4[[6]][3]` |  |

## Quarterly Report Results

**For quarterly report:** by round,

-   [x] number and percent enrolled and discharged;

-   [x] overall (not by round) number and percent of client type categories and average age for each category;

-   [ ] overall (not by round) number and percent of cases for caregiver/parent highest level of education

### Number and % enrolled

```{r}
#| echo: false
#| warning: false
#| message: false

discharge_data %>% 
  filter(intervention_start <= "2025-12-31") %>% 
  tabyl(funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Number and % discharged

```{r}
#| echo: false
#| warning: false
#| message: false

discharge_data %>% 
    filter(!is.na(reason_for_discharge), 
         reason_for_discharge != "No Reason Selected",
         intervention_end <= "2025-12-31") %>% 
  tabyl(reason_for_discharge, funding_round) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns()

#NOT unique by client
# subtract round by no reason selected --> NA
discharge_data %>% 
  filter(!is.na(reason_for_discharge), 
         reason_for_discharge != "No Reason Selected",
         intervention_end <= "2025-12-31") %>% 
  tabyl(reason_for_discharge) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)

discharge_data %>% filter(!is.na(intervention_end) & is.na(reason_for_discharge))

#round3 <- all_data_spss %>% 
 # filter(`4funding_round` == "Round 3")

# discharge_data_filtered %>% 
#   filter(!is.na(reason_for_discharge)) %>% 
#   tabyl(reason_for_discharge, funding_round) %>% 
#   adorn_totals("row") %>% 
#   adorn_percentages("col") %>%
#   adorn_pct_formatting(digits = 0) %>%
#   adorn_ns()
```

### Number (%) of client type categories

```{r}
#| echo: false
#| warning: false
#| message: false

all_data_distinct %>% 
  filter(!is.na(client_type)) %>% 
  #distinct(client_system_id, .keep_all = TRUE) %>% 
  tabyl(client_type) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Preferred Language

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
  # filter(primary_language != "Don't Know",
  #        primary_language != "No Primary Language Selected",
  #        primary_language != "Not Applicable",
  #        primary_language != "Prefer Not To Answer",
  #        !is.na(primary_language)) %>%
  tabyl(primary_language) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Housing Status

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
  # filter(housing_status != "Don't Know",
  #        housing_status != "No Housing Status Selected",
  #        housing_status != "Not Applicable",
  #        housing_status != "Prefer Not To Answer",
  #        !is.na(housing_status)) %>%
  tabyl(housing_status) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Number (%) of client type categories with average age for each category

```{r}
#| echo: false
#| warning: false
#| message: false

# first get client type counts
tab_counts <- all_data_distinct %>%
  #filter(client_type != "No Client Type Selected",
   #      !is.na(client_type)) %>% 
  tabyl(client_type) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1)

tab_counts

# then average age for each category
all_data_distinct <- all_data_distinct %>%
  mutate(
    age_start_tx_num = case_when(
      age_at_start_of_treatment_in_years %in% c("No Age Provided", "", " ") ~ NA_character_,
      TRUE ~ age_at_start_of_treatment_in_years
    ) %>% as.numeric())
    
tab_age <- all_data_distinct %>%
  group_by(client_type) %>%
  summarise(
    mean_age = round(mean(age_start_tx_num, na.rm = TRUE), 1),
    n_with_age = sum(!is.na(age_start_tx_num)),
    .groups = "drop"
  )

tab_age

# create final table
final_table <- tab_counts %>%
  left_join(tab_age, by = "client_type") %>%
  select(
    client_type,
    n,
    percent,
    mean_age,
    n_with_age
  )

final_table
```

### Insurance Status

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
  filter(insurance_status != "Don't Know",
         insurance_status != "No Insurance Status Selected",
         insurance_status != "Not Applicable",
         insurance_status != "Prefer Not To Answer",
         !is.na(housing_status)) %>%
  tabyl(insurance_status) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Gender Identity

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
  filter(gender_identity != "Don't Know",
         gender_identity != "No Gender Identify Selected",
         gender_identity != "Not Applicable",
         gender_identity != "Prefer Not To Answer",
         !is.na(gender_identity)) %>%
  tabyl(gender_identity) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Sexual Orientation

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
 filter(sexual_orientation != "Don't Know",
        sexual_orientation != "No Sexual Orientation Selected",
        sexual_orientation != "Not Applicable",
        sexual_orientation != "Prefer Not To Answer",
        !is.na(sexual_orientation)) %>%
  tabyl(sexual_orientation) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Race

1.  Asian

2.  Black or African American

3.  Hispanic or Latino

4.  Middle Eastern or North African

5.  Native American/American Indian or Alaska Native

6.  Native Hawaiian or Pacific Islander

7.  White

8.  2+ races selected

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
# filter(race != "Don't Know",
#        race != "No Race Selected",
#        race != "Not Applicable",
#        race != "Prefer Not To Answer",
#        !is.na(race)) %>%
  tabyl(code_num) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Ethnicity

1.  Chinese
2.  Asian Indian
3.  Filipino
4.  Vietnamese
5.  Korean
6.  Japanese
7.  African American
8.  Jamaican
9.  Haitian
10. Nigerian
11. Somali
12. Mexican
13. Puerto Rican
14. Salvadoran
15. Cuban
16. Dominican
17. Guatemalan
18. Native Hawaiian
19. Samoan
20. Chamorro
21. English
22. German
23. Irish
24. Italian
25. Polish
26. Scottish
27. 2+ ethnicities selected = 88

```{r}
#| echo: false
#| warning: false
#| message: false

# don't know, prefer not to answer, n/a and n/a all summed
all_data_distinct %>% 
# filter(race != "Don't Know",
#        race != "No Race Selected",
#        race != "Not Applicable",
#        race != "Prefer Not To Answer",
#        !is.na(race)) %>%
  tabyl(ethn_num) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```

### Number (%) of cases for caregiver/parent highest level of education

```{r}
#| echo: false
#| warning: false
#| message: false

all_data_new <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv")

all_data_new_distinct <- all_data_new %>% 
  filter(intervention_start >= "2024-01-01",
         intervention_start <= "2025-12-31") %>% 
  distinct(client_system_id, .keep_all = TRUE)

all_data_new_distinct %>% 
  #filter(!is.na(highest_caregiver_education_level)) %>%  
  tabyl(highest_caregiver_education_level) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting(digits = 1)
```
