---
title: "Quarterly Report Results"
subtitle: "`r params$quarter_label`"
author: "Cristin Young"
date: "`r Sys.Date()`"
params:
  quarter_label: "2026 Q2"
  period_start: "2024-01-01"
  period_end: "2026-06-30"
  date_downloaded: "2026-07-17"
  suppression_threshold: 11
  discharge_data_path: "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv"
  demo_data_path: "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv"
  satisfaction_data_path: "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/combined_satisfaction_clean.csv"
  assess_dir: "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"
format:
  docx:
    reference-doc: dhcs-reference-doc.docx
    toc: true
    toc-depth: 2
execute:
  echo: false
  warning: false
  message: false
---

<!--
  READ ME (delete or leave — Quarto ignores HTML comments):

  1. Every quarter, update only the `params:` block above (dates, quarter
     label, file paths) — nothing else in this file should need to change
     unless the underlying data columns change.
  2. This file renders straight to a DHCS-branded .docx using
     dhcs-reference-doc.docx (must live in the same folder as this .qmd).
     There is no template to paste numbers into anymore — the rendered
     .docx *is* the deliverable.
  3. Small-cell suppression (<11) is implemented in code below
     (see `suppress_labeled_counts()` and `dhcs_crosstab()`), per the
     original TODO. This has NOT been validated against real output by
     Anthropic/Claude — spot-check the first render against a few known
     small cells before trusting it for publication.
-->

[ ] create internal report that leaves all the notes 
[ ] create external report that deletes all notes
[ ] Update variable names throughout so they're not raw variable names in the table column headers. If we could specify user-friendly labels for those, it would look nicer?
[ ] update to "don't have this data" 
[ ] wtf claude, I want the tables to have % and total still

```{r}
#| label: setup
library(readr)
library(tidyverse)
library(haven)
library(janitor)
library(flextable)
library(officer)
library(scales)

knitr::opts_chunk$set(rownames.print = FALSE)

threshold <- params$suppression_threshold

# ---- load data -------------------------------------------------------
discharge_data <- read_csv(params$discharge_data_path) %>%
  filter(intervention_start >= params$period_start,
         intervention_start <= params$period_end)

all_data <- read_csv(params$demo_data_path) %>%
  filter(intervention_start >= params$period_start,
         intervention_start <= params$period_end)

discharge_data_distinct <- discharge_data %>%
  distinct(client_system_id, .keep_all = TRUE)

all_data_distinct <- all_data %>%
  distinct(client_system_id, .keep_all = TRUE)

n_providers        <- discharge_data |> distinct(provider_id) |> nrow()
n_discharge_distinct <- discharge_data |> distinct(client_system_id, .keep_all = TRUE) |> nrow()
n_discharge_rows   <- nrow(discharge_data)

# ---- DHCS brand table theme ------------------------------------------
# Applied to every results table so styling lives in one place.
dhcs_table <- function(df, caption = NULL) {
  ft <- flextable(df) |>
    set_table_properties(layout = "autofit", width = 1) |>
    font(fontname = "Segoe UI", part = "all") |>
    fontsize(size = 10, part = "all") |>
    bold(part = "header") |>
    color(color = "white", part = "header") |>
    bg(bg = "#17315A", part = "header") |>
    border_remove() |>
    hline(border = fp_border(color = "#ECEEF0", width = 0.75), part = "body") |>
    hline_bottom(border = fp_border(color = "#17315A", width = 1.5), part = "header") |>
    hline_bottom(border = fp_border(color = "#17315A", width = 1), part = "body") |>
    padding(padding = 4, part = "all") |>
    align(align = "left", part = "all")

  n_rows <- nrow(df)
  if (n_rows > 1) {
    ft <- bg(ft, i = seq(2, n_rows, by = 2), bg = "#ECEEF0", part = "body")
  }
  if (!is.null(caption)) {
    ft <- set_caption(ft, caption = caption, style = "Caption")
  }
  ft
}

# ---- small-cell suppression -------------------------------------------
# DHCS rule: any cell with < threshold individuals is suppressed. If only
# one non-total row in the table is suppressed, the total is suppressed
# too (otherwise it could be back-calculated). NOT independently tested —
# spot-check against real small cells on first render.
suppress_labeled_counts <- function(labels, counts, threshold = 11, total_label = "Total") {
  is_total <- labels == total_label
  is_small <- !is.na(counts) & counts < threshold & !is_total
  any_small <- any(is_small)
  out <- ifelse(!is.na(counts) & counts < threshold,
                paste0("<", threshold), format(counts, big.mark = ","))
  if (any_small) out[is_total] <- paste0("<", threshold)
  list(display = out, suppressed = is_small | (any_small & is_total))
}

# Apply suppression + formatting to a single-variable tabyl() that has
# already had adorn_totals("row") and adorn_pct_formatting() run on it.
apply_suppression <- function(tabyl_df, label_col = 1, n_col = "n", pct_col = "percent") {
  labels <- tabyl_df[[label_col]]
  counts <- tabyl_df[[n_col]]
  sup <- suppress_labeled_counts(labels, counts, threshold = threshold)
  tabyl_df[[n_col]] <- sup$display
  tabyl_df[[pct_col]] <- ifelse(sup$suppressed, "suppressed", tabyl_df[[pct_col]])
  tabyl_df
}

# Build a suppressed, percent-by-column crosstab from raw (unformatted)
# counts — used instead of tabyl()+adorn_ns() so suppression happens
# before any percentage math touches the small cells.
dhcs_crosstab <- function(data, row_var, col_var, threshold = 11) {
  row_var <- rlang::ensym(row_var)
  col_var <- rlang::ensym(col_var)

  counts <- data |>
    filter(!is.na(!!row_var), !is.na(!!col_var)) |>
    count(!!row_var, !!col_var, name = "n") |>
    pivot_wider(names_from = !!col_var, values_from = n, values_fill = 0)

  group_cols <- setdiff(names(counts), rlang::as_string(row_var))

  totals <- counts |>
    summarise(across(all_of(group_cols), sum)) |>
    mutate(!!row_var := "Total")

  full <- bind_rows(counts, totals)
  col_totals <- full[full[[rlang::as_string(row_var)]] == "Total", group_cols] |>
    unlist()

  out <- full
  for (col in group_cols) {
    raw <- full[[col]]
    labels <- full[[rlang::as_string(row_var)]]
    sup <- suppress_labeled_counts(labels, raw, threshold = threshold)
    pct <- percent(raw / col_totals[[col]], accuracy = 1)
    display <- ifelse(sup$suppressed, paste0("<", threshold),
                       paste0(format(raw, big.mark = ","), " (", pct, ")"))
    out[[col]] <- display
  }
  out
}

cat("Unique providers (discharge data):", n_providers, "\n")
cat("Discharge data (distinct by client ID) rows; this is the denominator for most tables below:", n_discharge_distinct, "\n")
cat("Discharge data rows (total):", n_discharge_rows, "\n")
```

```{r load-assessment-data}
adir <- params$assess_dir

pairs_primary_long_child <- read_csv(file.path(adir, "noms_child_pairs_clean.csv"),   show_col_types = FALSE)
pairs_primary_long_self  <- read_csv(file.path(adir, "noms_self_pairs_clean.csv"),    show_col_types = FALSE)
sdq_P24_pairs            <- read_csv(file.path(adir, "sdq_p24_pairs_clean.csv"),      show_col_types = FALSE)
sdq_P410_pairs           <- read_csv(file.path(adir, "sdq_p410_pairs_clean.csv"),     show_col_types = FALSE)
sdq_P1117_pairs          <- read_csv(file.path(adir, "sdq_p1117_pairs_clean.csv"),    show_col_types = FALSE)
sdq_S1117_pairs          <- read_csv(file.path(adir, "sdq_s1117_pairs_clean.csv"),    show_col_types = FALSE)
pss_paired               <- read_csv(file.path(adir, "pss_paired_clean.csv"),         show_col_types = FALSE)
cpss_i_paired            <- read_csv(file.path(adir, "cpss_i_paired_clean.csv"),      show_col_types = FALSE)
cpss_s_paired            <- read_csv(file.path(adir, "cpss_s_paired_clean.csv"),      show_col_types = FALSE)
```

```{r sdq-fn}
analyze_sdq <- function(paired) {
  tab_total <- paired |>
    filter(!is.na(positive_outcome_total)) |>
    tabyl(positive_outcome_total) |>
    adorn_pct_formatting(digits = 0)

  tab_ext <- paired |>
    filter(!is.na(positive_outcome_ext)) |>
    tabyl(positive_outcome_ext) |>
    adorn_pct_formatting(digits = 0)

  list(summary_total = tab_total, summary_ext = tab_ext)
}
```

::: {custom-style="DHCS Data Note"}
Data Source: Scaling EBPs/CDEPs eInsight Data Platform (https://cybhi.einsight.net/nova-login) | Datasets Used: Individual- and assessment-level records, from intake to discharge | Dates Represented: `r params$period_start` - `r params$period_end` | Date downloaded: `r params$date_downloaded` | Prepared by Behavioral Health Evaluation Strategies, LLC.
:::

::: {custom-style="DHCS Callout"}
**A note on suppression.** Any cell representing fewer than `r threshold` individuals is shown as "<`r threshold`" rather than an exact count, to protect client privacy. Where only one cell in a row would otherwise be suppressed, the row's total is suppressed as well so the hidden value cannot be back-calculated.
:::

## Period

`r params$period_start` - `r params$period_end`

## Outcomes Data by Round

### Participants Served

**Data used**: Discharge data, which includes clients that had more than one complete intervention.

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered only for `intervention_start` within the reporting period.

```{r}
#| label: tbl-participants-served-by-round
tbl_served <- discharge_data %>%
  filter(!is.na(intervention_start)) |>
  tabyl(funding_round) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  apply_suppression()

dhcs_table(tbl_served, caption = "Participants served by funding round")
```

### Participants with End Dates

**Data used**: Discharge data, which includes clients that had more than one complete intervention.

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered only for `intervention_start` within the reporting period.

```{r}
#| label: tbl-participants-with-end-date-by-round
tbl_end_date <- discharge_data %>%
  filter(!is.na(intervention_end)) |>
  tabyl(funding_round) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  apply_suppression()

dhcs_table(tbl_end_date, caption = "Participants with an end date, by funding round")
```

### Improved Outcomes

**Data used**: Discharge data

**Nitty gritty cleaning**: took out DEMO data; checked to see which `client_ids` had `pre_measure_s` AND `post_measure_s` per row (i.e., had a complete intervention); left those in; filtered only for `intervention_start` within the reporting period.

```{r}
#| label: n-count-outcomes
n_with_outcomes <- nrow(discharge_data %>% filter(!is.na(positive_outcomes)))
```

A total of [`r format(n_with_outcomes, big.mark = ",")`]{custom-style="DHCS Stat"} clients had a recorded outcome this period.

#### Improved outcomes table

**Assumptions:** Needed to have an intervention start and end date and reason for discharge; data used: discharge_data

```{r}
#| label: tbl-improved-outcomes
tbl_outcomes <- discharge_data %>%
  filter(!is.na(intervention_start),
         !is.na(intervention_end),
         !is.na(reason_for_discharge)) %>%
  tabyl(positive_outcomes) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1) %>%
  apply_suppression()

dhcs_table(tbl_outcomes, caption = "Improved outcomes")
```

#### Improved outcomes by funding round

**Assumptions:** Needed to have an intervention start and end date and reason for discharge; data used: discharge_data

```{r}
#| label: tbl-positive-outcomes-by-round
outcomes_base <- discharge_data %>%
  filter(!is.na(intervention_start),
         !is.na(intervention_end),
         !is.na(reason_for_discharge),
         !is.na(positive_outcomes))

tbl_outcomes_round <- dhcs_crosstab(outcomes_base, positive_outcomes, funding_round)
dhcs_table(tbl_outcomes_round, caption = "Improved outcomes by funding round — n (% of round)")
```

#### Including No Change in Outcomes

```{r}
#| label: setup-outcomes
#| include: false

assessment_rules <- tibble::tribble(
  ~trimmed_pre_assessment_name, ~higher_is_better, ~min_change,
  "Basis-24",  FALSE, 0.001, # weighted 0–4 scale; any positive change counts
  "CATS",      FALSE, 1,
  "GAD-7",     FALSE, 1,
  "NOMS",      TRUE,  1,
  "Pediatric", FALSE, 1,
  "PHQ-9",     FALSE, 1,
  "PSS",       FALSE, 1,
  "SDQ",       FALSE, 1,
  "YOQ",       FALSE, 10
)

discharge_outcome <- discharge_data |>
  left_join(assessment_rules, by = "trimmed_pre_assessment_name") |>
  mutate(
    raw_change    = post_score - pre_score,
    better_change = if_else(higher_is_better, raw_change, -raw_change),
    min_change    = coalesce(min_change, 0),
    outcome = case_when(
      is.na(pre_score) | is.na(post_score) ~ NA_character_,
      better_change >=  min_change ~ "improved",
      better_change <= -min_change ~ "worsened",
      TRUE                         ~ "no_change"
    ),
    outcome = factor(outcome, levels = c("worsened", "no_change", "improved"))
  )
```

#### Including No Change in Outcomes Overall
**Assumptions:** Must have an intervention name and an outcome; data used: discharge_data
```{r}
#| label: tbl-outcome-overall

discharge_outcome |>
  tabyl(outcome) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```


## Satisfaction Surveys

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree." The second is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed. High satisfaction on question 1 is defined as a score of 4 or 5.

**Assumptions:** Distinct by client system ID and joined to satisfaction survey data; must have an intervention named; data used: satisfaction survey and all data (demo+discharge joined)

```{r}
#| label: setup-satisfaction
#| include: false

sat_all <- read_csv(params$satisfaction_data_path, show_col_types = FALSE)

# Unified Q1 score — coalesce across three survey-specific Q1 columns
sat_all <- sat_all |>
  mutate(
    q1_score = coalesce(
      x1_overall_i_am_satisfied_with_the_services_my_child_received_score,
      x1_overall_i_am_satisfied_with_the_services_i_received_score,
      x1_i_like_the_services_that_i_received_here_score
    )
  )

all_data_distinct <- all_data |>
  distinct(client_system_id, .keep_all = TRUE) |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  )

sat_all_data <- sat_all |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  ) |>
  left_join(all_data_distinct,
            by = c("provider", "provider_id", "provider_system_id", "client_system_id")) |>
  filter(!is.na(intervention_name))

score_cols <- names(sat_all_data)[str_detect(names(sat_all_data), "_score$")]

# Overall Q1 high satisfaction rate
pct_4_5 <- sat_all_data |>
  summarise(
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1)
  ) |>
  pull(pct_4_5)

# Overall combined survey high satisfaction rate (>= 80% of questions at 4 or 5)
df2 <- sat_all_data |>
  select(all_of(score_cols)) |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(everything(), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree    = sum(c_across(everything()) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  ungroup() |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )
```

### Question 1 Satisfaction

Overall, [`r pct_4_5`%]{custom-style="DHCS Stat"} of survey respondents were highly satisfied based on question 1. The following table shows high satisfaction by round:

```{r}
#| label: tbl-q1-satisfaction-round
q1_by_round <- sat_all_data |>
  filter(!is.na(funding_round)) |>
  group_by(funding_round) |>
  summarise(
    n_surveys = n(),
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1),
    .groups = "drop"
  ) |>
  adorn_totals("row") %>%
  rename(n = n_surveys) |>
  apply_suppression(label_col = "funding_round", n_col = "n", pct_col = "pct_4_5")

dhcs_table(q1_by_round, caption = "Question 1 high satisfaction, by funding round")
```

### Combined Survey Satisfaction

Overall, [`r df2$pct_meeting_80pct`%]{custom-style="DHCS Stat"} of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed. The following table shows high combined satisfaction by round, for clients where both a client ID and intervention name were available:

**Assumptions:** Must have an intervention name; data used: satisfaction survey and all data (demo+discharge joined)

```{r}
#| label: tbl-combined-satisfaction-round
combined_by_round <- sat_all_data |>
  filter(!is.na(funding_round)) |>
  group_by(funding_round) |>
  select(funding_round, all_of(score_cols)) |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  mutate(across(all_of(score_cols), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(all_of(score_cols)))),
    n_agree    = sum(c_across(all_of(score_cols)) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  ) |>
  adorn_totals("row") %>%
  rename(n = n_total) |>
  apply_suppression(label_col = "funding_round", n_col = "n", pct_col = "pct_meeting_80pct")

dhcs_table(combined_by_round, caption = "Combined survey high satisfaction (>=80% agree/strongly agree), by funding round")
```

## Discharge

### Discharge Reason

**Assumptions:** Needed to have a reason for discharge; data used: discharge_data

```{r}
#| label: tbl-reason-for-discharge
tbl_discharge_reason <- discharge_data |>
  filter(!is.na(reason_for_discharge),
         reason_for_discharge != "No Reason Selected") |>
  tabyl(reason_for_discharge) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_discharge_reason, caption = "Discharge reason")
```

### Discharge Reason by Round

The results below show successful discharges by round, once NA's are removed from the data.

**Assumptions:** Needed to have a reason for discharge and a funding round; data used: discharge data

```{r}
#| label: tbl-discharged-by-reason
discharge_reason_base <- discharge_data |>
  filter(!is.na(reason_for_discharge))

tbl_discharge_by_round <- dhcs_crosstab(discharge_reason_base, reason_for_discharge, funding_round)
dhcs_table(tbl_discharge_by_round, caption = "Discharge reason by funding round — n (% of round)")
```

### Discharge Reason by Funding Round (start date)

**Assumptions:** Needed to have an intervention START date -- this is because the data is broken out by funding round, not the funding round in which the intervention ended; data used: discharge_data

```{r}
#| label: tbl-discharge-reason-by-round
discharge_start_base <- discharge_data |>
  filter(!is.na(intervention_start), !is.na(reason_for_discharge))

tbl_discharge_start_round <- dhcs_crosstab(discharge_start_base, reason_for_discharge, funding_round)
dhcs_table(tbl_discharge_start_round, caption = "Discharge reason by funding round (by intervention start date) — n (% of round)")
```

## Client Type

### Client Type Overall

**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have client_type

```{r}
#| label: tbl-client-type
.d <- all_data |>
  filter(!is.na(client_type)) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_client_type <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_client_type <- .d |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_client_type, caption = "Client type")
```

[`r pct_have_client_type`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Client Type by Funding Round

**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data); must have client type and funding round

```{r}
#| label: tbl-client-type-by-round
.d <- all_data |>
  filter(!is.na(client_type), !is.na(funding_round)) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_client_type_round <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_client_type_round <- dhcs_crosstab(.d, client_type, funding_round)
dhcs_table(tbl_client_type_round, caption = "Client type by funding round — n (% of round)")
```

[`r pct_have_client_type_round`%]{custom-style="DHCS Emphasis"} of clients have this data.

## Client Demographics

### Client Type with Average Age

**Assumptions:** Distinct by client system ID; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-client-type-age
all_data_distinct <- all_data_distinct |>
  mutate(
    age_start_tx_num = case_when(
      age_at_start_of_treatment_in_years %in% c("No Age Provided", "", " ") ~ NA_character_,
      TRUE ~ age_at_start_of_treatment_in_years
    ) |> as.numeric()
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

tab_counts <- all_data_distinct |>
  tabyl(client_type) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

tab_age <- all_data_distinct |>
  group_by(client_type) |>
  summarise(
    mean_age   = round(mean(age_start_tx_num, na.rm = TRUE), 1),
    n_with_age = sum(!is.na(age_start_tx_num)),
    .groups = "drop"
  )

tbl_type_age <- tab_counts |>
  left_join(tab_age, by = "client_type") |>
  select(client_type, n, percent, mean_age, n_with_age)

# suppress rows below threshold; hide mean_age too when the row itself is suppressed
sup <- suppress_labeled_counts(tbl_type_age$client_type, tbl_type_age$n, threshold = threshold)
tbl_type_age <- tbl_type_age |>
  mutate(
    n = sup$display,
    percent = ifelse(sup$suppressed, "suppressed", percent),
    mean_age = ifelse(sup$suppressed, NA, mean_age),
    n_with_age = ifelse(sup$suppressed, NA, n_with_age)
  )

dhcs_table(tbl_type_age, caption = "Client type with average age at start of treatment")
```

### Insurance Status

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-insurance-status
.d <- all_data |>
  filter(
    insurance_status != "Don't Know",
    insurance_status != "No Insurance Status Selected",
    insurance_status != "Not Applicable",
    insurance_status != "Prefer Not To Answer",
    !is.na(insurance_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_insurance <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_insurance <- .d |>
  tabyl(insurance_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_insurance, caption = "Insurance status")
```

[`r pct_have_insurance`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Gender Identity

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-gender-identity
.d <- all_data |>
  filter(
    gender_identity != "Don't Know",
    gender_identity != "No Gender Identify Selected",
    gender_identity != "Not Applicable",
    gender_identity != "Prefer Not To Answer",
    !is.na(gender_identity)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_gender <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_gender <- .d |>
  tabyl(gender_identity) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_gender, caption = "Gender identity")
```

[`r pct_have_gender`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Sexual Orientation

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-sexual-orientation
.d <- all_data |>
  filter(
    sexual_orientation != "Don't Know",
    sexual_orientation != "No Sexual Orientation Selected",
    sexual_orientation != "Not Applicable",
    sexual_orientation != "Prefer Not To Answer",
    !is.na(sexual_orientation)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_orientation <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_orientation <- .d |>
  tabyl(sexual_orientation) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_orientation, caption = "Sexual orientation")
```

[`r pct_have_orientation`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Race

Race is coded numerically as follows:

| Code | Race                                               |
|------|-----------------------------------------------------|
| 1    | Asian                                               |
| 2    | Black or African American                           |
| 3    | Hispanic or Latino                                  |
| 4    | Middle Eastern or North African                     |
| 5    | Native American / American Indian or Alaska Native  |
| 6    | Native Hawaiian or Pacific Islander                 |
| 7    | White                                               |
| 88   | 2+ races selected                                   |

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-race
.d <- all_data |>
  filter(
    race != "Don't Know",
    race != "No Race Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(code_num)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_race <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_race <- .d |>
  tabyl(code_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression(label_col = "code_num")

dhcs_table(tbl_race, caption = "Race")
```

[`r pct_have_race`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Ethnicity

Ethnicity is coded numerically as follows:

| Code | Ethnicity               | Code | Ethnicity       |
|------|--------------------------|------|-----------------|
| 1    | Chinese                  | 14   | Salvadoran      |
| 2    | Asian Indian             | 15   | Cuban           |
| 3    | Filipino                 | 16   | Dominican       |
| 4    | Vietnamese                | 17   | Guatemalan      |
| 5    | Korean                   | 18   | Native Hawaiian |
| 6    | Japanese                 | 19   | Samoan          |
| 7    | African American         | 20   | Chamorro        |
| 8    | Jamaican                 | 21   | English         |
| 9    | Haitian                  | 22   | German          |
| 10   | Nigerian                 | 23   | Irish           |
| 11   | Somali                   | 24   | Italian         |
| 12   | Mexican                  | 25   | Polish          |
| 13   | Puerto Rican             | 26   | Scottish        |
|      |                           | 88   | 2+ ethnicities selected |

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-ethnicity
.d <- all_data |>
  filter(
    race != "Don't Know",
    race != "No Ethnicity Selected",
    race != "Not Applicable",
    race != "Prefer Not To Answer",
    !is.na(ethn_num)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_ethnicity <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_ethnicity <- .d |>
  tabyl(ethn_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression(label_col = "ethn_num")

dhcs_table(tbl_ethnicity, caption = "Ethnicity")
```

[`r pct_have_ethnicity`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Primary Language

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-primary-language
.d <- all_data |>
  filter(
    primary_language != "Don't Know",
    primary_language != "No Primary Language Selected",
    primary_language != "Not Applicable",
    primary_language != "Prefer Not To Answer",
    !is.na(primary_language)) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_language <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_language <- .d |>
  tabyl(primary_language) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_language, caption = "Primary language")
```

[`r pct_have_language`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Housing Status

**Assumptions:** Distinct by client system ID, Don't know/no selection/not applicable/prefer not to answer all grouped as NA; data used: all_data (demo + discharge data)

```{r}
#| label: tbl-housing-status
.d <- all_data |>
  filter(
    housing_status != "Don't Know",
    housing_status != "No Housing Status Selected",
    housing_status != "Not Applicable",
    housing_status != "Prefer Not To Answer",
    !is.na(housing_status)
  ) |>
  distinct(client_system_id, .keep_all = TRUE)

pct_have_housing <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_housing <- .d |>
  tabyl(housing_status) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_housing, caption = "Housing status")
```

[`r pct_have_housing`%]{custom-style="DHCS Emphasis"} of clients have this data.

### Services in Preferred Language

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** The percent of clients that report services received in their preferred language out of the total that completed the satisfaction survey.

**Data used:** satisfaction survey and all data (demo+discharge joined)

```{r}
#| label: tbl-preferred-language
tbl_preferred_lang <- sat_all_data |>
  arrange(date_clean) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label))) |>
  rename(preferred_lang_response = were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |>
  tabyl(preferred_lang_response) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_preferred_lang, caption = "Services received in preferred language")
```

### Caregiver / Parent Highest Level of Education

**Assumptions:** Distinct by client system ID; must have highest level of education information available

```{r}
#| label: tbl-highest-caregiver-education
.d <- all_data |>
  filter(!is.na(highest_caregiver_education_level)) %>%
  distinct(client_system_id, .keep_all = TRUE)

pct_have_education <- round(100 * nrow(.d) / n_distinct(all_data$client_system_id), 1)

tbl_education <- .d |>
  tabyl(highest_caregiver_education_level) %>%
  adorn_totals("row") %>%
  adorn_pct_formatting(digits = 1) |>
  apply_suppression()

dhcs_table(tbl_education, caption = "Caregiver/parent highest level of education")
```

[`r pct_have_education`%]{custom-style="DHCS Emphasis"} of clients have this data.

## Assessment outcomes (questions 3.03–3.16)

**Inclusion criteria:** 1. must have initial time point data; 2. must have a valid assessment at both intake and discharge; 3. limited to tools with a clinical cut-off.

**Assumptions:** Filtered out dates before `r params$period_start`. Paired to the last follow-up. NAs excluded before computing percentages.

---

### 3.03 NOMS Child Q1 — Improved mental health (caregiver-reported)

Range: 1–5.

```{r q303}
q303_data <- pairs_primary_long_child |>
  filter(question == "q1_how_would_you_rate_your_child_s_overall_mental_health_right_now_score")

q303_tabyl <- q303_data |>
  mutate(
    positive_outcome = case_when(
      is.na(follow_up_score) | is.na(initial_score) ~ NA_character_,
      follow_up_score > initial_score ~ "improved",
      follow_up_score < initial_score ~ "worsened",
      TRUE ~ "no_change"
    ),
    positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q303_long <- q303_data |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(c(initial_score, follow_up_score), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_score" = "Initial", "follow_up_score" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q303_stats <- q303_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q303_pct <- q303_data |>
  summarise(initial_4_5  = round(100 * mean(initial_score[!is.na(initial_score)]     %in% c(4, 5), na.rm = TRUE), 2),
            followup_4_5 = round(100 * mean(follow_up_score[!is.na(follow_up_score)] %in% c(4, 5), na.rm = TRUE), 2))

dhcs_table(q303_tabyl, caption = "3.03 NOMS Child Q1 — % with positive outcome")
dhcs_table(q303_stats, caption = "3.03 NOMS Child Q1 — Summary statistics by timepoint")
dhcs_table(q303_pct,   caption = "3.03 NOMS Child Q1 — % scoring 4 or 5")
```

---

### 3.04 NOMS Self Q1 — Improved mental health (youth-reported)

Range: 1–5.

```{r q304}
q304_data <- pairs_primary_long_self |>
  filter(question == "q1_how_would_you_rate_your_overall_mental_health_right_now_score")

q304_tabyl <- q304_data |>
  mutate(
    positive_outcome = case_when(
      is.na(follow_up_score) | is.na(initial_score) ~ NA_character_,
      follow_up_score > initial_score ~ "improved",
      follow_up_score < initial_score ~ "worsened",
      TRUE ~ "no_change"
    ),
    positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q304_long <- q304_data |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(c(initial_score, follow_up_score), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_score" = "Initial", "follow_up_score" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q304_stats <- q304_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q304_pct <- q304_data |>
  summarise(initial_4_5  = round(100 * mean(initial_score[!is.na(initial_score)]     %in% c(4, 5), na.rm = TRUE), 2),
            followup_4_5 = round(100 * mean(follow_up_score[!is.na(follow_up_score)] %in% c(4, 5), na.rm = TRUE), 2))

dhcs_table(q304_tabyl, caption = "3.04 NOMS Self Q1 — % with positive outcome")
dhcs_table(q304_stats, caption = "3.04 NOMS Self Q1 — Summary statistics by timepoint")
dhcs_table(q304_pct,   caption = "3.04 NOMS Self Q1 — % scoring 4 or 5")
```

---

### 3.05 NOMS Child Q2 — Improved daily life functioning (caregiver-reported)

Range: 7–35; Sum = Q2a–Q2g.

```{r q305}
q305_diff <- pairs_primary_long_child |>
  filter(str_detect(question, "^x2.*_score$"),
         !is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(initial_total  = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            positive_outcome = case_when(
              followup_total > initial_total ~ "improved",
              followup_total < initial_total ~ "worsened",
              TRUE ~ "no_change"
            ),
            positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
            .groups = "drop") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

q305_tabyl <- q305_diff |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q305_long <- q305_diff |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(c(initial_total, followup_total), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_total" = "Initial", "followup_total" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q305_stats <- q305_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q305_pct <- q305_diff |>
  summarise(initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
            followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2))

dhcs_table(q305_tabyl, caption = "3.05 NOMS Child Q2 — % with positive outcome")
dhcs_table(q305_stats, caption = "3.05 NOMS Child Q2 — Summary statistics by timepoint")
dhcs_table(q305_pct,   caption = "3.05 NOMS Child Q2 — % scoring 28 or higher")
```

---

### 3.06 NOMS Self Q2 — Improved daily life functioning (youth-reported)

Range: 7–35; Sum = Q2a–Q2g.

```{r q306}
q306_diff <- pairs_primary_long_self |>
  filter(str_detect(question, "^x2.*_score$"),
         !is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(initial_total  = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            positive_outcome = case_when(
              followup_total > initial_total ~ "improved",
              followup_total < initial_total ~ "worsened",
              TRUE ~ "no_change"
            ),
            positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
            .groups = "drop") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

q306_tabyl <- q306_diff |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q306_long <- q306_diff |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(c(initial_total, followup_total), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_total" = "Initial", "followup_total" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q306_stats <- q306_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q306_pct <- q306_diff |>
  summarise(initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
            followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2))

dhcs_table(q306_tabyl, caption = "3.06 NOMS Self Q2 — % with positive outcome")
dhcs_table(q306_stats, caption = "3.06 NOMS Self Q2 — Summary statistics by timepoint")
dhcs_table(q306_pct,   caption = "3.06 NOMS Self Q2 — % scoring 28 or higher")
```

---

### 3.07 NOMS Child Q3 — Improved well-being (caregiver-reported)

Range: 7–35; Sum = Q3a–Q3g.

```{r q307}
q307_diff <- pairs_primary_long_child |>
  filter(str_detect(question, "^x3.*_score$"),
         !is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(initial_total  = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            positive_outcome = case_when(
              followup_total > initial_total ~ "improved",
              followup_total < initial_total ~ "worsened",
              TRUE ~ "no_change"
            ),
            positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
            .groups = "drop") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

q307_tabyl <- q307_diff |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q307_long <- q307_diff |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(c(initial_total, followup_total), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_total" = "Initial", "followup_total" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q307_stats <- q307_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q307_pct <- q307_diff |>
  summarise(initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
            followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2))

dhcs_table(q307_tabyl, caption = "3.07 NOMS Child Q3 — % with positive outcome")
dhcs_table(q307_stats, caption = "3.07 NOMS Child Q3 — Summary statistics by timepoint")
dhcs_table(q307_pct,   caption = "3.07 NOMS Child Q3 — % scoring 28 or higher")
```

---

### 3.08 NOMS Self Q3 — Improved well-being (youth-reported)

Range: 7–35; Sum = Q3a–Q3g.

```{r q308}
q308_diff <- pairs_primary_long_self |>
  filter(str_detect(question, "^x3.*_score$"),
         !is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(initial_total  = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            positive_outcome = case_when(
              followup_total > initial_total ~ "improved",
              followup_total < initial_total ~ "worsened",
              TRUE ~ "no_change"
            ),
            positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
            .groups = "drop") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

q308_tabyl <- q308_diff |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q308_long <- q308_diff |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(c(initial_total, followup_total), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_total" = "Initial", "followup_total" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q308_stats <- q308_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q308_pct <- q308_diff |>
  summarise(initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
            followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2))

dhcs_table(q308_tabyl, caption = "3.08 NOMS Self Q3 — % with positive outcome")
dhcs_table(q308_stats, caption = "3.08 NOMS Self Q3 — Summary statistics by timepoint")
dhcs_table(q308_pct,   caption = "3.08 NOMS Self Q3 — % scoring 28 or higher")
```

---

### 3.09 NOMS Child Q4 — Improved social connectedness (caregiver-reported)

Range: 6–30; Sum = Q4a–Q4f.

```{r q309}
q309_diff <- pairs_primary_long_child |>
  filter(str_detect(question, "^x4.*_score$"),
         !is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(initial_total  = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            positive_outcome = case_when(
              followup_total > initial_total ~ "improved",
              followup_total < initial_total ~ "worsened",
              TRUE ~ "no_change"
            ),
            positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
            .groups = "drop") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 6, 30), between(followup_total, 6, 30))

q309_tabyl <- q309_diff |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q309_long <- q309_diff |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(c(initial_total, followup_total), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_total" = "Initial", "followup_total" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q309_stats <- q309_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q309_pct <- q309_diff |>
  summarise(initial_24  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 24, na.rm = TRUE), 2),
            followup_24 = round(100 * mean(followup_total[!is.na(followup_total)] >= 24, na.rm = TRUE), 2))

dhcs_table(q309_tabyl, caption = "3.09 NOMS Child Q4 — % with positive outcome")
dhcs_table(q309_stats, caption = "3.09 NOMS Child Q4 — Summary statistics by timepoint")
dhcs_table(q309_pct,   caption = "3.09 NOMS Child Q4 — % scoring 24 or higher")
```

---

### 3.10 NOMS Self Q4 — Improved social connectedness (youth-reported)

Range: 6–30; Sum = Q4a–Q4f.

```{r q310}
q310_diff <- pairs_primary_long_self |>
  filter(str_detect(question, "^x4.*_score$"),
         !is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(initial_total  = sum(initial_score, na.rm = TRUE),
            followup_total = sum(follow_up_score, na.rm = TRUE),
            positive_outcome = case_when(
              followup_total > initial_total ~ "improved",
              followup_total < initial_total ~ "worsened",
              TRUE ~ "no_change"
            ),
            positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
            .groups = "drop") |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 6, 30), between(followup_total, 6, 30))

q310_tabyl <- q310_diff |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q310_long <- q310_diff |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(c(initial_total, followup_total), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_total" = "Initial", "followup_total" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q310_stats <- q310_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

q310_pct <- q310_diff |>
  summarise(initial_24  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 24, na.rm = TRUE), 2),
            followup_24 = round(100 * mean(followup_total[!is.na(followup_total)] >= 24, na.rm = TRUE), 2))

dhcs_table(q310_tabyl, caption = "3.10 NOMS Self Q4 — % with positive outcome")
dhcs_table(q310_stats, caption = "3.10 NOMS Self Q4 — Summary statistics by timepoint")
dhcs_table(q310_pct,   caption = "3.10 NOMS Self Q4 — % scoring 24 or higher")
```

---

### 3.11 / 3.16 SDQ — Reductions in disruptive behaviors / improvements in coping

**Scoring:** A decrease in score indicates a positive outcome. Four versions: P2–4; P4–10; P11–17; S11–17.

- **3.11:** Externalizing score (conduct + hyperactivity scales)
- **3.16:** Total difficulties score

```{r q311-316-outcomes}
sdq_subgroup_levels <- c("SDQ Parent (2-4)", "SDQ Parent (4-10)", "SDQ Parent (11-17)", "SDQ Self (11-17)")

res_P24   <- analyze_sdq(sdq_P24_pairs)
res_P410  <- analyze_sdq(sdq_P410_pairs)
res_P1117 <- analyze_sdq(sdq_P1117_pairs)
res_S1117 <- analyze_sdq(sdq_S1117_pairs)

dhcs_table(res_P24$summary_total,   caption = "3.16 SDQ Parent (2–4) — % with positive outcome, total score")
dhcs_table(res_P24$summary_ext,     caption = "3.11 SDQ Parent (2–4) — % with positive outcome, externalizing score")
dhcs_table(res_P410$summary_total,  caption = "3.16 SDQ Parent (4–10) — % with positive outcome, total score")
dhcs_table(res_P410$summary_ext,    caption = "3.11 SDQ Parent (4–10) — % with positive outcome, externalizing score")
dhcs_table(res_P1117$summary_total, caption = "3.16 SDQ Parent (11–17) — % with positive outcome, total score")
dhcs_table(res_P1117$summary_ext,   caption = "3.11 SDQ Parent (11–17) — % with positive outcome, externalizing score")
dhcs_table(res_S1117$summary_total, caption = "3.16 SDQ Self (11–17) — % with positive outcome, total score")
dhcs_table(res_S1117$summary_ext,   caption = "3.11 SDQ Self (11–17) — % with positive outcome, externalizing score")
```

```{r q311-316-stats}
df_from_lists <- bind_rows(
  sdq_P24_pairs   |> transmute(subgroup = "SDQ Parent (2-4)",   initial_total, follow_up_total, initial_ext, follow_up_ext),
  sdq_P410_pairs  |> transmute(subgroup = "SDQ Parent (4-10)",  initial_total, follow_up_total, initial_ext, follow_up_ext),
  sdq_P1117_pairs |> transmute(subgroup = "SDQ Parent (11-17)", initial_total, follow_up_total, initial_ext, follow_up_ext),
  sdq_S1117_pairs |> transmute(subgroup = "SDQ Self (11-17)",   initial_total, follow_up_total, initial_ext, follow_up_ext)
)

sdq_stats_total <- df_from_lists |>
  filter(!is.na(initial_total), !is.na(follow_up_total)) |>
  pivot_longer(c(initial_total, follow_up_total), names_to = "timepoint", values_to = "value") |>
  mutate(subgroup  = factor(subgroup, levels = sdq_subgroup_levels),
         timepoint = recode(timepoint, "initial_total" = "Initial", "follow_up_total" = "Follow-up")) |>
  group_by(subgroup, timepoint) |>
  summarise(mean = round(mean(value, na.rm = TRUE), 2),
            median = median(value, na.rm = TRUE),
            sd = round(sd(value, na.rm = TRUE), 2),
            n = sum(!is.na(value)), .groups = "drop")

sdq_stats_ext <- df_from_lists |>
  filter(!is.na(initial_ext), !is.na(follow_up_ext)) |>
  pivot_longer(c(initial_ext, follow_up_ext), names_to = "timepoint", values_to = "value") |>
  mutate(subgroup  = factor(subgroup, levels = sdq_subgroup_levels),
         timepoint = recode(timepoint, "initial_ext" = "Initial", "follow_up_ext" = "Follow-up")) |>
  group_by(subgroup, timepoint) |>
  summarise(mean = round(mean(value, na.rm = TRUE), 2),
            median = median(value, na.rm = TRUE),
            sd = round(sd(value, na.rm = TRUE), 2),
            n = sum(!is.na(value)), .groups = "drop")

dhcs_table(sdq_stats_total, caption = "3.16 SDQ — Summary statistics, total difficulties score by subgroup and timepoint")
dhcs_table(sdq_stats_ext,   caption = "3.11 SDQ — Summary statistics, externalizing score by subgroup and timepoint")
```

---

### 3.12 PSS — Reduction in parent-related stress

**Scoring:** A decrease in score indicates a positive outcome. Range: 18–90.

```{r q312}
q312_tabyl <- pss_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q312_long <- pss_paired |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(c(initial_score, follow_up_score), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_score" = "Initial", "follow_up_score" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q312_stats <- q312_long |>
  filter(!is.na(score_value)) |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

dhcs_table(q312_tabyl, caption = "3.12 PSS — % with positive outcome (lower score = better)")
dhcs_table(q312_stats, caption = "3.12 PSS — Summary statistics by timepoint")
```

---

### 3.15 CPSS-5 — Reductions in PTSD symptoms

**Scoring:** A decrease in score indicates a positive outcome. Range: 0–80.

#### Interviewer version

```{r q315-i}
q315i_tabyl <- cpss_i_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q315i_long <- cpss_i_paired |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(c(initial_score, follow_up_score), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_score" = "Initial", "follow_up_score" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q315i_stats <- q315i_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

dhcs_table(q315i_tabyl, caption = "3.15 CPSS-5 Interviewer — % with positive outcome")
dhcs_table(q315i_stats, caption = "3.15 CPSS-5 Interviewer — Summary statistics by timepoint")
```

#### Self version

```{r q315-s}
q315s_tabyl <- cpss_s_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

q315s_long <- cpss_s_paired |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(c(initial_score, follow_up_score), names_to = "score_type", values_to = "score_value") |>
  mutate(score_type = recode(score_type, "initial_score" = "Initial", "follow_up_score" = "Follow-up"),
         score_type = factor(score_type, levels = c("Initial", "Follow-up")))

q315s_stats <- q315s_long |>
  group_by(score_type) |>
  summarise(mean = round(mean(score_value, na.rm = TRUE), 2),
            median = median(score_value, na.rm = TRUE),
            sd = round(sd(score_value, na.rm = TRUE), 2),
            n = sum(!is.na(score_value)), .groups = "drop")

dhcs_table(q315s_tabyl, caption = "3.15 CPSS-5 Self — % with positive outcome")
dhcs_table(q315s_stats, caption = "3.15 CPSS-5 Self — Summary statistics by timepoint")
```
