---
title: "Annual Report Data Analysis - Question 1.01 and 1.02"
author: "Cristin Young"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    df_print: paged
  pdf_document: default
  html_notebook: default
---

# 1. Did the EBP/CDEP Workstream increase access to services for targeted populations?

## 1.01 Did improved access to services affect all subpopulations?

Cross-tab of study population by relevant demographic data

Total number of clients served and by demographics

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated timeframe, where "served" means that a client had a new intervention start date attached to them during the designated timeframe.

**Calculation:** Aggregated count of clients served by period with an intervention type and start date listed; intervention does not need to be completed; assumes dates before 2024 are correct and any rows with these dates have not been revised.

**Assumptions:** This is discharge data from EInsight, so clients can appear multiple times (multiple interventions, or the same intervention on multiple dates). Filtered out dates before 2024-01-01; dates were not otherwise cleaned.

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(janitor)
library(gt)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
discharge_data_filtered <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv"
)

all_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv",
  show_col_types = FALSE
)
```

```{r prep_data}
#| echo: false
#| warning: false
#| message: false

# Add year and apply base filters to discharge data
discharge_data_filtered <- discharge_data_filtered |>
  mutate(intervention_start_year = year(intervention_start)) |>
  filter(
    !is.na(intervention_start),
    !is.na(intervention_name)
  )

# Clean all demographic variables in one pass
all_data <- all_data |>
  mutate(
    # --- Insurance ---
    insurance_status_clean = str_squish(insurance_status),
    insurance_status_clean = if_else(
      insurance_status_clean %in% c(
        "Don't Know", "Prefer Not To Answer",
        "Not Applicable", "No Insurance Status Selected"
      ) | insurance_status_clean == "",
      NA_character_,
      insurance_status_clean
    ),
    insurance_num = case_when(
      insurance_status_clean == "Private Insurance"        ~ 1L,
      insurance_status_clean == "Medi-Cal/Medi-Care"       ~ 2L,
      insurance_status_clean == "No insurance/Indigent"    ~ 3L,
      insurance_status_clean == "Other"                    ~ 4L,
      insurance_status_clean == "Veterans Benefits"        ~ 5L,
      TRUE                                                 ~ NA_integer_
    ),

    # --- Sexual orientation ---
    sexual_orientation_clean = str_squish(sexual_orientation),
    sexual_orientation_clean = if_else(
      sexual_orientation_clean %in% c(
        "Don't Know", "No Sexual Orientation Selected",
        "Prefer Not To Answer", "Not Applicable"
      ) | sexual_orientation_clean == "",
      NA_character_,
      sexual_orientation_clean
    ),
    sexual_orientation_num = case_when(
      sexual_orientation_clean == "Asexual"            ~ 1L,
      sexual_orientation_clean == "Bisexual"           ~ 2L,
      sexual_orientation_clean == "Lesbian/Gay"        ~ 3L,
      sexual_orientation_clean == "Other"              ~ 4L,
      sexual_orientation_clean == "Pansexual"          ~ 5L,
      sexual_orientation_clean == "Same Gender Loving" ~ 6L,
      sexual_orientation_clean == "Straight"           ~ 7L,
      TRUE                                             ~ NA_integer_
    ),

    # --- Gender identity ---
    gender_identity_clean = str_squish(gender_identity),
    gender_identity_clean = if_else(
      gender_identity_clean %in% c(
        "Don't Know", "Prefer Not To Answer",
        "Not Applicable", "No Gender Identify Selected"
      ) | gender_identity_clean == "",
      NA_character_,
      gender_identity_clean
    ),
    gender_identity_num = case_when(
      gender_identity_clean == "Genderqueer"            ~ 1L,
      gender_identity_clean == "Man/boy"                ~ 2L,
      gender_identity_clean == "Non-binary"             ~ 3L,
      gender_identity_clean == "Transgender man/boy"    ~ 4L,
      gender_identity_clean == "Transgender woman/girl" ~ 5L,
      gender_identity_clean == "Two-spirit"             ~ 6L,
      gender_identity_clean == "Woman/girl"             ~ 7L,
      TRUE                                              ~ NA_integer_
    ),

    # --- Preferred language ---
    preferred_language_clean = str_squish(preferred_language),
    preferred_language_clean = case_when(
      preferred_language_clean %in% c(
        "Don't Know", "Prefer Not To Answer", "Not Applicable", ""
      )                                                     ~ NA_character_,
      preferred_language_clean %in% c("English", "Spanish") ~ preferred_language_clean,
      TRUE                                                   ~ "Other"
    ),
    preferred_language_num = case_when(
      preferred_language_clean == "English" ~ 1L,
      preferred_language_clean == "Spanish" ~ 2L,
      preferred_language_clean == "Other"   ~ 3L,
      TRUE                                  ~ NA_integer_
    ),

    # --- Primary language ---
    primary_language_clean = str_squish(primary_language),
    primary_language_clean = case_when(
      primary_language_clean %in% c(
        "Don't Know", "Prefer Not To Answer", "Not Applicable", ""
      )                                                    ~ NA_character_,
      primary_language_clean %in% c("English", "Spanish") ~ primary_language_clean,
      TRUE                                                 ~ "Other"
    ),
    primary_language_num = case_when(
      primary_language_clean == "English" ~ 1L,
      primary_language_clean == "Spanish" ~ 2L,
      primary_language_clean == "Other"   ~ 3L,
      TRUE                                ~ NA_integer_
    ),

    # --- Housing status ---
    housing_status_clean = str_squish(housing_status),
    housing_status_clean = case_when(
      housing_status_clean %in% c(
        "Don't Know", "Prefer Not To Answer", "Not Applicable", ""
      )                                                  ~ NA_character_,
      housing_status_clean %in% c("Housed", "Unhoused") ~ housing_status_clean,
      TRUE                                               ~ "Other (i.e. treatment facility)"
    ),
    housing_status_num = case_when(
      housing_status_clean == "Housed"                          ~ 1L,
      housing_status_clean == "Unhoused"                        ~ 2L,
      housing_status_clean == "Other (i.e. treatment facility)" ~ 3L,
      TRUE                                                      ~ NA_integer_
    )
  )

# Base filtered dataset: 2024 onward, with intervention start and name
all_data_filtered <- all_data |>
  filter(
    intervention_start >= "2024-01-01",
    !is.na(intervention_start),
    !is.na(intervention_name)
  )
```

### Total clients served

```{r client_serve}
#| echo: false
discharge_data_filtered |>
  filter(intervention_start >= "2024-01-01") |>
  tabyl(intervention_start_year)
```

### Race

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by race.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and race is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
1. Asian
2. Black or African American
3. Hispanic or Latino
4. Middle Eastern or North African
5. Native American/American Indian or Alaska Native
6. Native Hawaiian or Pacific Islander
7. White
8. 2+ races selected

**Assumptions:** Clients had to have an intervention start, an intervention name, and have race characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r race}
#| echo: false
all_data_filtered |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(code_num) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

all_data_filtered |>
  filter(!is.na(client_type), !is.na(code_num)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, code_num) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()
```

### Ethnicity

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by ethnicity.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and ethnicity is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
`Chinese` = 1, `Asian Indian` = 2, `Filipino` = 3, `Vietnamese` = 4, `Korean` = 5, `Japanese` = 6, `African American` = 7, `Jamaican` = 8, `Haitian` = 9, `Nigerian` = 10, `Somali` = 11, `Mexican` = 12, `Puerto Rican` = 13, `Salvadoran` = 14, `Cuban` = 15, `Dominican` = 16, `Guatemalan` = 17, `Native Hawaiian` = 18, `Samoan` = 19, `Chamorro` = 20, `English` = 21, `German` = 22, `Irish` = 23, `Italian` = 24, `Polish` = 25, `Scottish` = 26, `2+ ethnicities selected` = 88

**Assumptions:** Clients had to have an intervention start, an intervention name, and have race characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r ethnicity}
#| echo: false
all_data_filtered |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(ethn_num) |>
  adorn_pct_formatting(digits = 1) |>
  adorn_totals("row")
```

### Insurance status

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by insurance status.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and insurance is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
1. Private insurance
2. Medi-Cal/Medi-Care
3. No insurance/Indigent
4. Other
5. Veterans Benefits

**Assumptions:** Clients had to have an intervention start, an intervention name, and have insurance status characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r insurance}
#| echo: false
all_data_filtered |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(insurance_status_clean) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 2)

all_data_filtered |>
  filter(!is.na(client_type), !is.na(insurance_status_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, insurance_status_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()
```

### Sexual orientation

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by sexual orientation.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and sexual orientation is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no sexual orientation selected
1. Asexual
2. Bisexual
3. Lesbian/Gay
4. Other
5. Pansexual
6. Same Gender Loving
7. Straight

**Assumptions:** Clients had to have an intervention start, an intervention name, and have sexual orientation characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r sexual_orientation}
#| echo: false
all_data_filtered |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(sexual_orientation_clean) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)

all_data_filtered |>
  filter(!is.na(client_type), !is.na(sexual_orientation_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, sexual_orientation_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()
```

### Gender identity

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by gender identity.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and gender identity is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
1. Genderqueer
2. Man/boy
3. Non-binary
4. Transgender man/boy
5. Transgender woman/girl
6. Two-spirit
7. Woman/girl

**Assumptions:** Clients had to have an intervention start, an intervention name, and have gender identity characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r gender_identity}
#| echo: false
all_data_filtered |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(gender_identity_clean) |>
  adorn_totals("row") |>
  adorn_pct_formatting(digits = 1)
```

### Preferred language

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by preferred language.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and preferred language is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no preferred language selected
1. English
2. Spanish
3. Other

**Assumptions:** Clients had to have an intervention start, an intervention name, and have preferred language characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). Any language with \<200 rows of data is grouped into "Other".

```{r preferred_language}
#| echo: false
all_data_filtered |>
  filter(!is.na(preferred_language_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(intervention_start_quarter, preferred_language_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()

all_data_filtered |>
  filter(!is.na(client_type), !is.na(preferred_language_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, preferred_language_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()
```

### Primary language

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by primary language.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and primary language is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no primary language selected
1. English
2. Spanish
3. Other

**Assumptions:** Clients had to have an intervention start, an intervention name, and have primary language characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). Any language with \<200 rows of data is grouped into "Other".

```{r primary_language}
#| echo: false
all_data_filtered |>
  filter(!is.na(primary_language_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(intervention_start_quarter, primary_language_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()

all_data_filtered |>
  filter(!is.na(client_type), !is.na(primary_language_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, primary_language_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()
```

### Housing status

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by housing status.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and housing status is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no housing status selected
1. Housed
2. Unhoused
3. Other (i.e. treatment facility)

**Assumptions:** Clients had to have an intervention start, an intervention name, and have housing status characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r housing_status}
#| echo: false
all_data_filtered |>
  filter(!is.na(housing_status_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(intervention_start_quarter, housing_status_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()

all_data_filtered |>
  filter(!is.na(client_type), !is.na(housing_status_clean)) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  tabyl(client_type, housing_status_clean) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()
```

## 1.02 How does access differ compared to previous access to services before EBP/CDEP funding? For each subpopulation?

Cross-tab of demographics and grant types: Total number of clients served and by demographic, and by whether grant was start up or expansion

```{r}
#| echo: false

```
