---
title: "Annual Report Data Analysis - Question 4.03"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    df_print: paged
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 4. What percentage of clients reported high satisfaction with services?

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(ggridges)
library(scales)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

# Produced by data cleaning/satisfaction_data_cleaning.qmd
sat_combined <- read_csv(
  file.path(assess_dir, "combined_satisfaction_clean.csv"),
  show_col_types = FALSE
) |>
  filter(
    !is.na(date_clean)
  ) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label)))
```

## 4.03 Within the grantees doing CDEPs, were there higher client satisfaction rates among specific target populations?

**Definition:** 

**Calculation:** 

**Assumptions:** 

**Final analysis:** Regression - outcomes = high satisfaction (4/5 vs other). Same predictors as for 3.01; subset for CDEPs from SPSS file

## Satisfaction Surveys

Satisfaction is defined two ways below. The first is the percent of individuals who answered question 1 as "Agree" or "Strongly Agree." The second is the percent of individuals who selected "Agree" or "Strongly Agree" on 80% or more of the questions completed. High satisfaction on question 1 is defined as a score of 4 or 5.

**Assumptions:** Distinct by client system ID and joined to satisfaction survey data; must have an intervention named; data used: satisfaction survey and all data (demo+discharge joined)

```{r}
#| label: setup-satisfaction
#| include: false

sat_all <- read_csv(params$satisfaction_data_path, show_col_types = FALSE)

# Unified Q1 score — coalesce across three survey-specific Q1 columns
sat_all <- sat_all |>
  mutate(
    q1_score = coalesce(
      x1_overall_i_am_satisfied_with_the_services_my_child_received_score,
      x1_overall_i_am_satisfied_with_the_services_i_received_score,
      x1_i_like_the_services_that_i_received_here_score
    )
  )

all_data_distinct <- all_data |>
  distinct(client_system_id, .keep_all = TRUE) |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  )

sat_all_data <- sat_all |>
  mutate(
    provider_system_id = as.character(provider_system_id),
    client_system_id   = as.character(client_system_id)
  ) |>
  left_join(all_data_distinct,
            by = c("provider", "provider_id", "provider_system_id", "client_system_id")) |>
  filter(!is.na(intervention_name))

score_cols <- names(sat_all_data)[str_detect(names(sat_all_data), "_score$")]

# Overall Q1 high satisfaction rate
pct_4_5 <- sat_all_data |>
  summarise(
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1)
  ) |>
  pull(pct_4_5)

# Overall combined survey high satisfaction rate (>= 80% of questions at 4 or 5)
df2 <- sat_all_data |>
  select(all_of(score_cols)) |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(everything(), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(everything()))),
    n_agree    = sum(c_across(everything()) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  ungroup() |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  )
```

### Question 1 Satisfaction

Overall, [`r pct_4_5`%]{custom-style="DHCS Stat"} of survey respondents were highly satisfied based on question 1. The following table shows high satisfaction by round:

```{r}
#| label: tbl-q1-satisfaction-round
q1_by_round <- sat_all_data |>
  filter(!is.na(funding_round)) |>
  group_by(funding_round) |>
  summarise(
    n_surveys = n(),
    pct_4_5 = round(100 * sum(q1_score %in% c(4, 5), na.rm = TRUE) / sum(!is.na(q1_score)), 1),
    .groups = "drop"
  ) |>
  adorn_totals("row") %>%
  rename(n = n_surveys) |>
  apply_suppression(label_col = "funding_round", n_col = "n", pct_col = "pct_4_5")

dhcs_table(q1_by_round, caption = "Question 1 high satisfaction, by funding round")
```

### Combined Survey Satisfaction

Overall, [`r df2$pct_meeting_80pct`%]{custom-style="DHCS Stat"} of survey respondents were highly satisfied, defined as selecting "Agree" or "Strongly Agree" on 80% or more of the questions completed. The following table shows high combined satisfaction by round, for clients where both a client ID and intervention name were available:

**Assumptions:** Must have an intervention name; data used: satisfaction survey and all data (demo+discharge joined)

```{r}
#| label: tbl-combined-satisfaction-round
combined_by_round <- sat_all_data |>
  filter(!is.na(funding_round)) |>
  group_by(funding_round) |>
  select(funding_round, all_of(score_cols)) |>
  mutate(across(all_of(score_cols), as.numeric)) |>
  mutate(across(all_of(score_cols), ~ na_if(., 0))) |>
  rowwise() |>
  summarise(
    n_answered = sum(!is.na(c_across(all_of(score_cols)))),
    n_agree    = sum(c_across(all_of(score_cols)) %in% c(4, 5), na.rm = TRUE),
    pct_agree  = n_agree / n_answered
  ) |>
  summarise(
    n_total           = n(),
    n_meeting_80pct   = sum(pct_agree >= 0.8, na.rm = TRUE),
    pct_meeting_80pct = round(100 * n_meeting_80pct / n_total, 1)
  ) |>
  adorn_totals("row") %>%
  rename(n = n_total) |>
  apply_suppression(label_col = "funding_round", n_col = "n", pct_col = "pct_meeting_80pct")

dhcs_table(combined_by_round, caption = "Combined survey high satisfaction (>=80% agree/strongly agree), by funding round")
```


