---
title: "Annual Report Data Analysis - Question 3.05 and 3.06"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    df_print: paged  
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

**Inclusion criteria:** 1. must have initial time point data (i.e., be enrolled in an intervention), 2. must have a valid assessment collected at both intake and discharge time points, and 3. limited to tools with a clinical cut-off.

**Scoring:** An **increase** in score indicates a positive outcome. There are no benchmarks or clinical cutoffs established for NOMS scoring — if the follow-up assessment score is higher than the initial score, a positive outcome is indicated.

**Assumptions:** Filtered out dates before 2024-01-01 and after the date the code was run. Filtered data that's \>3 days from initial assessment and paired to the LAST follow-up. Filtered positive outcomes that had NAs before computing %.

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(ggridges)
library(scales)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

# Produced by data cleaning/noms_data_cleaning.qmd
# for 3.06
pairs_primary_long_self <- read_csv(
  file.path(assess_dir, "noms_self_pairs_clean.csv"),
  show_col_types = FALSE
)

# for 3.05
pairs_primary_long_child <- read_csv(
  file.path(assess_dir, "noms_child_pairs_clean.csv"),
  show_col_types = FALSE
)
```

## 3.05 To what extent did parents/caregivers report improved daily life functioning for themselves and their child after receiving the intervention?

Range: 7–35; Sum = Q2a+...+Q2g.

Use the `pairs_primary_long_child` data frame to compute scores. The following had a positive outcome by follow-up quarter:

```{r child_q2_summary}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q2 <- pairs_primary_long_child |>
  filter(str_detect(question, "^x2.*_score$"))

score_diff_child_q2 <- pairs_primary_long_q2 |>
  filter(
    !is.na(initial_score),
    !is.na(follow_up_score),
    !is.na(client_system_id)
  ) |>
  group_by(client_system_id) |>
  summarise(
    initial_total     = sum(initial_score, na.rm = TRUE),
    followup_total    = sum(follow_up_score, na.rm = TRUE),
    score_change      = followup_total - initial_total,
    positive_outcome  = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
    initial_quarter   = paste0("Q", quarter(initial_date), " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(
    between(initial_total,  7, 35),
    between(followup_total, 7, 35)
  )

score_diff_child_q2 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

Plot it.

```{r child_q2_bar_plot}
#| echo: false
#| warning: false
#| message: false
q2_child_score_long <- score_diff_child_q2 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

ggplot(q2_child_score_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8
  ) +
  scale_fill_manual(values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Measure: Improved Daily Life Functioning Score (7-35)",
    y     = "Count",
    title = "Distribution of Initial and Follow-up Scores, NOMS Child Q2"
  ) +
  theme_minimal() +
  theme(
    axis.title           = element_text(size = 9),
    axis.text            = element_text(size = 9),
    panel.grid.major     = element_blank(),
    panel.grid.minor.x   = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position      = "inside",
    legend.justification = c("left", "top"),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 8),
    plot.margin          = margin(5, 5, 5, 5)
  )
```

Ridgeline plot

```{r child_q2_ridgeline}
#| echo: false
#| warning: false
#| message: false
summary_stats_child_q2 <- q2_child_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_child_q2

ggplot(q2_child_score_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha          = 0.7,
    scale          = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_x_continuous(limits = c(5, 37), breaks = c(7, 14, 21, 28, 35)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Measure: Daily Life Functioning Score (7-35)",
    y     = NULL,
    title = "Distribution of Initial and Follow-up Scores, NOMS Child"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x         = element_text(size = 9, hjust = 0.5),
    axis.text.y          = element_blank(),
    axis.text.x          = element_text(size = 9),
    panel.grid           = element_blank(),
    axis.ticks.y         = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position      = "inside",
    legend.justification = c("left", "top"),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 8),
    plot.margin          = margin(5, 5, 5, 5),
    plot.title           = element_text(size = 11)
  ) +
  geom_segment(
    data = summary_stats_child_q2,
    aes(x = median_score, y = 1, xend = median_score, yend = 1.135, color = "grey30"),
    linetype = "dotted"
  ) +
  geom_text(
    data = summary_stats_child_q2,
    aes(x = median_score, y = 1.15,
        label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
    vjust = 2, hjust = 0.5, size = 2
  )
```

#### Summary stats

The percent of those who had total scores of 28 or higher (scores of 4\*7 questions) across Q2 at initial and discharge time points were:

```{r child_q2_pct_28}
#| echo: false
#| warning: false
#| message: false
score_diff_child_q2 |>
  ungroup() |>
  summarise(
    initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
    followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2)
  )
```

## 3.06 To what extent did the youth report improved daily life functioning after receiving the intervention?

Range: 7–35; Sum = Q2a+...+Q2g.

Use the `pairs_primary_long_self` data frame to compute scores. The following had a positive outcome by follow-up quarter:

```{r self_q2_summary}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q2_self <- pairs_primary_long_self |>
  filter(str_detect(question, "^x2.*_score$"))

score_diff_self_q2 <- pairs_primary_long_q2_self |>
  filter(
    !is.na(initial_score),
    !is.na(follow_up_score),
    !is.na(client_system_id)
  ) |>
  group_by(client_system_id) |>
  summarise(
    initial_total     = sum(initial_score, na.rm = TRUE),
    followup_total    = sum(follow_up_score, na.rm = TRUE),
    score_change      = followup_total - initial_total,
    positive_outcome  = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved")),
    initial_quarter   = paste0("Q", quarter(initial_date), " ", year(initial_date)),
    follow_up_quarter = paste0("Q", quarter(follow_up_date), " ", year(follow_up_date))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(
    between(initial_total,  7, 35),
    between(followup_total, 7, 35)
  )

score_diff_self_q2 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

Plot it.

```{r self_q2_bar_plot}
#| echo: false
#| warning: false
#| message: false
q2_self_score_long <- score_diff_self_q2 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

ggplot(q2_self_score_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    position = position_dodge2(preserve = "single", padding = 0.01),
    width = 0.8,
    alpha = 0.8
  ) +
  scale_fill_manual(values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Measure: Improved Daily Life Functioning Score (7-35)",
    y     = "Count",
    title = "Distribution of Initial and Follow-up Scores, NOMS Self Q2"
  ) +
  theme_minimal() +
  theme(
    axis.title           = element_text(size = 9),
    axis.text            = element_text(size = 9),
    panel.grid.major     = element_blank(),
    panel.grid.minor.x   = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position      = "inside",
    legend.justification = c("left", "top"),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 8),
    plot.margin          = margin(5, 5, 5, 5)
  )
```

Ridgeline plot

```{r self_q2_ridgeline}
#| echo: false
#| warning: false
#| message: false
summary_stats_self_q2 <- q2_self_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  ) |>
  mutate(y_text = if_else(score_type == "Initial", 1.155, 1.165))

summary_stats_self_q2

ggplot(q2_self_score_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha          = 0.7,
    scale          = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_x_continuous(limits = c(5, 37), breaks = c(7, 14, 21, 28, 35)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Measure: Daily Life Functioning Score (7-35)",
    y     = NULL,
    title = "Distribution of Initial and Follow-up Scores, NOMS Self"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x         = element_text(size = 9, hjust = 0.5),
    axis.text.y          = element_blank(),
    axis.text.x          = element_text(size = 9),
    panel.grid           = element_blank(),
    axis.ticks.y         = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position      = "inside",
    legend.justification = c("left", "top"),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 8),
    plot.margin          = margin(5, 5, 5, 5),
    plot.title           = element_text(size = 11)
  ) +
  geom_segment(
    data = summary_stats_self_q2,
    aes(x = median_score, y = 1, xend = median_score, yend = 1.135, color = "grey30"),
    linetype = "dotted"
  ) +
  geom_text(
    data = summary_stats_self_q2,
    aes(x = median_score, y = y_text,
        label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
    vjust = 2, hjust = 0.5, size = 2
  )
```

#### Summary stats

The percent of those who had total scores of 28 or higher (scores of 4\*7 questions) across Q2 at initial and discharge time points were:

```{r self_q2_pct_28}
#| echo: false
#| warning: false
#| message: false
score_diff_self_q2 |>
  ungroup() |>
  summarise(
    initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
    followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2)
  )
```

### Summary stats overall

```{r self_summary_dot_arrow}
#| echo: false
#| warning: false
#| message: false
dfq2 <- tribble(
  ~group,   ~initial, ~followup,
  "Child",  35.31,    66.88,
  "Self",   44.91,    57.64
)

ggplot(dfq2, aes(y = group)) +
  geom_segment(
    aes(x = initial, xend = followup, yend = group),
    arrow     = arrow(length = unit(0.2, "cm")),
    color     = "gray40",
    linewidth = 0.8
  ) +
  geom_point(aes(x = initial,  color = "Initial"),   size = 12, show.legend = TRUE) +
  geom_text(aes(x  = initial,  label = paste0(round(initial),  "%")), size = 3.5, color = "grey20") +
  geom_point(aes(x = followup, color = "Follow-up"), size = 12, show.legend = TRUE) +
  geom_text(aes(x  = followup, label = paste0(round(followup), "%")), size = 3.5, color = "grey20") +
  scale_x_continuous(limits = c(0, 100), labels = label_percent(scale = 1)) +
  scale_color_manual(
    name   = "Timepoint",
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49")
  ) +
  labs(
    x     = "Score",
    y     = "",
    title = "Change in direction of total scores for those who scored 28 or higher on NOMS Q2 by subgroup"
  ) +
  theme_minimal() +
  theme(
    axis.title           = element_text(size = 9),
    axis.text            = element_text(size = 9),
    panel.grid.major     = element_blank(),
    panel.grid.minor.x   = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.text          = element_text(size = 8),
    legend.position      = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    plot.margin          = margin(5, 5, 5, 5)
  )
```
