---
title: "Annual report data analysis - Questions 3.03–3.16"
---


# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

**Inclusion criteria:** 1. must have initial time point data (i.e., be enrolled in an intervention), 2. must have a valid assessment collected at both intake and discharge time points, and 3. limited to tools with a clinical cut-off.

**Assumptions:** Filtered out dates before 2024-01-01 and after the date the code was run. Filtered data that's >3 days from initial assessment and paired to the LAST follow-up. Filtered positive outcomes that had NAs before computing %.

```{r load-packages}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(patchwork)
```

```{r load-noms-data}
#| echo: false
#| warning: false
#| message: false
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

# Produced by data cleaning/noms_data_cleaning.qmd
pairs_primary_long_child <- read_csv(
  file.path(assess_dir, "noms_child_pairs_clean.csv"),
  show_col_types = FALSE
)

pairs_primary_long_self <- read_csv(
  file.path(assess_dir, "noms_self_pairs_clean.csv"),
  show_col_types = FALSE
)
```

```{r load-sdq-data}
#| echo: false
#| warning: false
#| message: false
# Produced by data cleaning/sdq_data_cleaning.qmd
sdq_P24_pairs   <- read_csv(file.path(assess_dir, "sdq_p24_pairs_clean.csv"),   show_col_types = FALSE)
sdq_P410_pairs  <- read_csv(file.path(assess_dir, "sdq_p410_pairs_clean.csv"),  show_col_types = FALSE)
sdq_P1117_pairs <- read_csv(file.path(assess_dir, "sdq_p1117_pairs_clean.csv"), show_col_types = FALSE)
sdq_S1117_pairs <- read_csv(file.path(assess_dir, "sdq_s1117_pairs_clean.csv"), show_col_types = FALSE)
```

```{r load-pss-data}
#| echo: false
#| warning: false
#| message: false
# Produced by data cleaning/pss_data_cleaning.qmd
pss_paired <- read_csv(
  file.path(assess_dir, "pss_paired_clean.csv"),
  show_col_types = FALSE
)
```

```{r load-cpss-data}
#| echo: false
#| warning: false
#| message: false
# Produced by data cleaning/cpss5_data_cleaning.qmd
cpss_i_paired <- read_csv(
  file.path(assess_dir, "cpss_i_paired_clean.csv"),
  show_col_types = FALSE
)

cpss_s_paired <- read_csv(
  file.path(assess_dir, "cpss_s_paired_clean.csv"),
  show_col_types = FALSE
)
```

---

## 3.03 NOMS Child Q1 — Improved mental health (caregiver-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 1–5; Sum = Q1.

```{r q303-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q1 <- pairs_primary_long_child |>
  filter(question == "q1_how_would_you_rate_your_child_s_overall_mental_health_right_now_score")

score_diff_child_q1 <- pairs_primary_long_q1 |>
  mutate(
    score_change     = follow_up_score - initial_score,
    positive_outcome = case_when(
      is.na(follow_up_score) | is.na(initial_score) ~ NA_character_,
      follow_up_score > initial_score ~ "improved",
      follow_up_score < initial_score ~ "worsened",
      TRUE                            ~ "no_change"
    ),
    positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  )

score_diff_child_q1 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q303-summary-stats}
#| echo: false
#| warning: false
#| message: false
q1_child_score_long <- pairs_primary_long_q1 |>
  ungroup() |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(
    cols      = c(initial_score, follow_up_score),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score"   = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_child_score_type <- q1_child_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_child_score_type
```

The percent of those who selected ratings of 4 or higher (excellent or very good) across Q1 at initial and discharge time points:

```{r q303-pct-4-5}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q1 |>
  ungroup() |>
  summarise(
    initial_4_5  = round(100 * mean(initial_score[!is.na(initial_score)]     %in% c(4, 5), na.rm = TRUE), 2),
    followup_4_5 = round(100 * mean(follow_up_score[!is.na(follow_up_score)] %in% c(4, 5), na.rm = TRUE), 2)
  )
```

---

## 3.04 NOMS Self Q1 — Improved mental health (youth-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 1–5; Sum = Q1.

```{r q304-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q1_self <- pairs_primary_long_self |>
  filter(question == "q1_how_would_you_rate_your_overall_mental_health_right_now_score")

score_diff_self_q1 <- pairs_primary_long_q1_self |>
  mutate(
    score_change     = follow_up_score - initial_score,
    positive_outcome = case_when(
      is.na(follow_up_score) | is.na(initial_score) ~ NA_character_,
      follow_up_score > initial_score ~ "improved",
      follow_up_score < initial_score ~ "worsened",
      TRUE                            ~ "no_change"
    ),
    positive_outcome = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  )

score_diff_self_q1 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q304-summary-stats}
#| echo: false
#| warning: false
#| message: false
q1_self_score_long <- pairs_primary_long_q1_self |>
  ungroup() |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(
    cols      = c(initial_score, follow_up_score),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score"   = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_self_score_type <- q1_self_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_self_score_type
```

The percent of those who selected ratings of 4 or higher across Q1 at initial and discharge time points:

```{r q304-pct-4-5}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q1_self |>
  ungroup() |>
  summarise(
    initial_4_5  = round(100 * mean(initial_score[!is.na(initial_score)]     %in% c(4, 5), na.rm = TRUE), 2),
    followup_4_5 = round(100 * mean(follow_up_score[!is.na(follow_up_score)] %in% c(4, 5), na.rm = TRUE), 2)
  )
```

---

## 3.05 NOMS Child Q2 — Improved daily life functioning (caregiver-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 7–35; Sum = Q2a+…+Q2g.

```{r q305-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q2 <- pairs_primary_long_child |>
  filter(str_detect(question, "^x2.*_score$"))

score_diff_child_q2 <- pairs_primary_long_q2 |>
  filter(!is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(
    initial_total    = sum(initial_score, na.rm = TRUE),
    followup_total   = sum(follow_up_score, na.rm = TRUE),
    score_change     = followup_total - initial_total,
    positive_outcome = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

score_diff_child_q2 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q305-summary-stats}
#| echo: false
#| warning: false
#| message: false
q2_child_score_long <- score_diff_child_q2 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_child_q2 <- q2_child_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_child_q2
```

The percent of those who had total scores of 28 or higher (4 × 7 questions) across Q2 at initial and discharge time points:

```{r q305-pct-28}
#| echo: false
#| warning: false
#| message: false
score_diff_child_q2 |>
  ungroup() |>
  summarise(
    initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
    followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2)
  )
```

---

## 3.06 NOMS Self Q2 — Improved daily life functioning (youth-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 7–35; Sum = Q2a+…+Q2g.

```{r q306-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q2_self <- pairs_primary_long_self |>
  filter(str_detect(question, "^x2.*_score$"))

score_diff_self_q2 <- pairs_primary_long_q2_self |>
  filter(!is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(
    initial_total    = sum(initial_score, na.rm = TRUE),
    followup_total   = sum(follow_up_score, na.rm = TRUE),
    score_change     = followup_total - initial_total,
    positive_outcome = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

score_diff_self_q2 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q306-summary-stats}
#| echo: false
#| warning: false
#| message: false
q2_self_score_long <- score_diff_self_q2 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_self_q2 <- q2_self_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_self_q2
```

The percent of those who had total scores of 28 or higher across Q2 at initial and discharge time points:

```{r q306-pct-28}
#| echo: false
#| warning: false
#| message: false
score_diff_self_q2 |>
  ungroup() |>
  summarise(
    initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
    followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2)
  )
```

---

## 3.07 NOMS Child Q3 — Improved well-being (caregiver-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 7–35; Sum = Q3a+…+Q3g.

```{r q307-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q3 <- pairs_primary_long_child |>
  filter(str_detect(question, "^x3.*_score$"))

score_diff_child_q3 <- pairs_primary_long_q3 |>
  filter(!is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(
    initial_total    = sum(initial_score, na.rm = TRUE),
    followup_total   = sum(follow_up_score, na.rm = TRUE),
    score_change     = followup_total - initial_total,
    positive_outcome = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

score_diff_child_q3 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q307-summary-stats}
#| echo: false
#| warning: false
#| message: false
q3_child_score_long <- score_diff_child_q3 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_child_q3 <- q3_child_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_child_q3
```

The percent of those who had total scores of 28 or higher across Q3 at initial and discharge time points:

```{r q307-pct-28}
#| echo: false
#| warning: false
#| message: false
score_diff_child_q3 |>
  ungroup() |>
  summarise(
    initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
    followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2)
  )
```

---

## 3.08 NOMS Self Q3 — Improved well-being (youth-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 7–35; Sum = Q3a+…+Q3g.

```{r q308-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q3_self <- pairs_primary_long_self |>
  filter(str_detect(question, "^x3.*_score$"))

score_diff_self_q3 <- pairs_primary_long_q3_self |>
  filter(!is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(
    initial_total    = sum(initial_score, na.rm = TRUE),
    followup_total   = sum(follow_up_score, na.rm = TRUE),
    score_change     = followup_total - initial_total,
    positive_outcome = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 7, 35), between(followup_total, 7, 35))

score_diff_self_q3 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q308-summary-stats}
#| echo: false
#| warning: false
#| message: false
q3_self_score_long <- score_diff_self_q3 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_self_q3 <- q3_self_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_self_q3
```

The percent of those who had total scores of 28 or higher across Q3 at initial and discharge time points:

```{r q308-pct-28}
#| echo: false
#| warning: false
#| message: false
score_diff_self_q3 |>
  ungroup() |>
  summarise(
    initial_28  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 28, na.rm = TRUE), 2),
    followup_28 = round(100 * mean(followup_total[!is.na(followup_total)] >= 28, na.rm = TRUE), 2)
  )
```

---

## 3.09 NOMS Child Q4 — Improved social connectedness (caregiver-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 6–30; Sum = Q4a+…+Q4f.

```{r q309-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q4 <- pairs_primary_long_child |>
  filter(str_detect(question, "^x4.*_score$"))

score_diff_child_q4 <- pairs_primary_long_q4 |>
  filter(!is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(
    initial_total    = sum(initial_score, na.rm = TRUE),
    followup_total   = sum(follow_up_score, na.rm = TRUE),
    score_change     = followup_total - initial_total,
    positive_outcome = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 6, 30), between(followup_total, 6, 30))

score_diff_child_q4 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q309-summary-stats}
#| echo: false
#| warning: false
#| message: false
q4_child_score_long <- score_diff_child_q4 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_child_q4 <- q4_child_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_child_q4
```

The percent of those who had total scores of 24 or higher (6 × 4) across Q4 at initial and discharge time points:

```{r q309-pct-24}
#| echo: false
#| warning: false
#| message: false
score_diff_child_q4 |>
  ungroup() |>
  summarise(
    initial_24  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 24, na.rm = TRUE), 2),
    followup_24 = round(100 * mean(followup_total[!is.na(followup_total)] >= 24, na.rm = TRUE), 2)
  )
```

---

## 3.10 NOMS Self Q4 — Improved social connectedness (youth-reported)

**Scoring:** An **increase** in score indicates a positive outcome. Range: 6–30; Sum = Q4a+…+Q4f.

```{r q310-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pairs_primary_long_q4_self <- pairs_primary_long_self |>
  filter(str_detect(question, "^x4.*_score$"))

score_diff_self_q4 <- pairs_primary_long_q4_self |>
  filter(!is.na(initial_score), !is.na(follow_up_score), !is.na(client_system_id)) |>
  group_by(client_system_id) |>
  summarise(
    initial_total    = sum(initial_score, na.rm = TRUE),
    followup_total   = sum(follow_up_score, na.rm = TRUE),
    score_change     = followup_total - initial_total,
    positive_outcome = case_when(
      followup_total > initial_total ~ "improved",
      followup_total < initial_total ~ "worsened",
      TRUE                           ~ "no_change"
    ),
    positive_outcome  = factor(positive_outcome, levels = c("worsened", "no_change", "improved"))
  ) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  filter(between(initial_total, 6, 30), between(followup_total, 6, 30))

score_diff_self_q4 |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q310-summary-stats}
#| echo: false
#| warning: false
#| message: false
q4_self_score_long <- score_diff_self_q4 |>
  ungroup() |>
  select(client_system_id, initial_total, followup_total) |>
  pivot_longer(
    cols      = c(initial_total, followup_total),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_total"  = "Initial",
                        "followup_total" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_self_q4 <- q4_self_score_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_self_q4
```

The percent of those who had total scores of 24 or higher across Q4 at initial and discharge time points:

```{r q310-pct-24}
#| echo: false
#| warning: false
#| message: false
score_diff_self_q4 |>
  ungroup() |>
  summarise(
    initial_24  = round(100 * mean(initial_total[!is.na(initial_total)]   >= 24, na.rm = TRUE), 2),
    followup_24 = round(100 * mean(followup_total[!is.na(followup_total)] >= 24, na.rm = TRUE), 2)
  )
```

---

## 3.11 / 3.16 SDQ — Reductions in disruptive behaviors / improvements in coping

**Scoring:** A **decrease** in score indicates a positive outcome. Four versions: P2–4; P4–10; P11–17; S11–17.

- **3.11:** Externalizing score (sum of conduct and hyperactivity scales)
- **3.16:** Total difficulties score

**Assumptions:** Filtered out dates before 2024-01-01 and after the date the code was run. Filtered data that's >3 days from initial assessment and paired to the LAST follow-up.

```{r sdq-function}
#| echo: false
#| warning: false
#| message: false
analyze_sdq <- function(paired, label = NULL) {

  message("---- Running analysis for: ", label, " ----")

  tab_total_overall <- paired |>
    filter(!is.na(positive_outcome_total)) |>
    tabyl(positive_outcome_total) |>
    adorn_pct_formatting(digits = 0)

  tab_ext_overall <- paired |>
    filter(!is.na(positive_outcome_ext)) |>
    tabyl(positive_outcome_ext) |>
    adorn_pct_formatting(digits = 0)

  list(
    paired         = paired,
    summary_total  = tab_total_overall,
    summary_ext    = tab_ext_overall
  )
}
```

```{r sdq-run}
#| echo: false
#| warning: false
#| message: false
res_P24   <- analyze_sdq(sdq_P24_pairs,   label = "Parent SDQ (2–4)")
cat("Parent SDQ (2-4) — total\n");   print(res_P24$summary_total)
cat("Parent SDQ (2-4) — ext\n");     print(res_P24$summary_ext)

res_P410  <- analyze_sdq(sdq_P410_pairs,  label = "Parent SDQ (4–10)")
cat("Parent SDQ (4-10) — total\n");  print(res_P410$summary_total)
cat("Parent SDQ (4-10) — ext\n");    print(res_P410$summary_ext)

res_P1117 <- analyze_sdq(sdq_P1117_pairs, label = "Parent SDQ (11–17)")
cat("Parent SDQ (11-17) — total\n"); print(res_P1117$summary_total)
cat("Parent SDQ (11-17) — ext\n");   print(res_P1117$summary_ext)

res_S1117 <- analyze_sdq(sdq_S1117_pairs, label = "Self SDQ (11–17)")
cat("Self SDQ (11-17) — total\n");   print(res_S1117$summary_total)
cat("Self SDQ (11-17) — ext\n");     print(res_S1117$summary_ext)

sdq_list <- list(
  "SDQ Parent (2-4)"   = res_P24,
  "SDQ Parent (4-10)"  = res_P410,
  "SDQ Parent (11-17)" = res_P1117,
  "SDQ Self (11-17)"   = res_S1117
)

df_from_lists <- imap_dfr(sdq_list, ~ {
  .x[[1]] |>
    transmute(
      subgroup        = .y,
      initial_total   = initial_total,
      follow_up_total = follow_up_total,
      initial_ext     = initial_ext,
      follow_up_ext   = follow_up_ext
    )
})
```

### Summary stats — total difficulties score

```{r sdq-summary-stats-total}
#| echo: false
#| warning: false
#| message: false
sdq_subgroup_levels <- c(
  "SDQ Parent (2-4)", "SDQ Parent (4-10)", "SDQ Parent (11-17)", "SDQ Self (11-17)"
)

sdq_long_total <- df_from_lists |>
  pivot_longer(
    cols      = c(initial_total, follow_up_total),
    names_to  = "timepoint",
    values_to = "value"
  ) |>
  mutate(
    subgroup  = factor(subgroup,  levels = sdq_subgroup_levels),
    timepoint = factor(timepoint, levels = c("initial_total", "follow_up_total"))
  )

summary_stats_sdq_total <- sdq_long_total |>
  group_by(subgroup, timepoint) |>
  summarise(
    mean_score   = mean(value, na.rm = TRUE),
    median_score = median(value, na.rm = TRUE),
    sd_score     = sd(value, na.rm = TRUE),
    n            = sum(!is.na(value)),
    .groups      = "drop"
  )

summary_stats_sdq_total
```

### Summary stats — externalizing score

```{r sdq-summary-stats-ext}
#| echo: false
#| warning: false
#| message: false
sdq_long_ext <- df_from_lists |>
  pivot_longer(
    cols      = c(initial_ext, follow_up_ext),
    names_to  = "timepoint",
    values_to = "value"
  ) |>
  mutate(
    subgroup  = factor(subgroup,  levels = sdq_subgroup_levels),
    timepoint = factor(timepoint, levels = c("initial_ext", "follow_up_ext"))
  )

summary_stats_sdq_ext <- sdq_long_ext |>
  group_by(subgroup, timepoint) |>
  summarise(
    mean_score   = mean(value, na.rm = TRUE),
    median_score = median(value, na.rm = TRUE),
    sd_score     = sd(value, na.rm = TRUE),
    n            = sum(!is.na(value)),
    .groups      = "drop"
  )

summary_stats_sdq_ext
```

---

## 3.12 PSS — Reduction in parent-related stress

**Scoring:** A **decrease** in score indicates a positive outcome. Range: 18–90. Higher scores indicate higher perceived stress.

**Assumptions:** Filtered out dates before 2024-01-01 and after the date the code was run. Filtered data that's >3 days from initial assessment and paired to the LAST follow-up. Total score = NA if 2+ items are missing.

```{r q312-positive-outcome}
#| echo: false
#| warning: false
#| message: false
pss_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q312-summary-stats}
#| echo: false
#| warning: false
#| message: false
pss_paired_long <- pss_paired |>
  ungroup() |>
  select(client_system_id, initial_score, follow_up_score) |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  pivot_longer(
    cols      = c(initial_score, follow_up_score),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score"   = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_pss <- pss_paired_long |>
  group_by(score_type) |>
  filter(!is.na(score_value)) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_pss
```

---

## 3.15 CPSS-5 — Reductions in PTSD symptoms

**Scoring:** A **decrease** in score indicates a positive outcome. Range: 0–80. Higher scores indicate greater severity.

**Assumptions:** Used LAST follow-up if 2+ assessments exist; filtered out NAs from follow-up date and positive outcome before calculating %.

### Interviewer version

```{r q315-i-positive-outcome}
#| echo: false
#| warning: false
#| message: false
cpss_i_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q315-i-summary-stats}
#| echo: false
#| warning: false
#| message: false
cpss_i_paired_long <- cpss_i_paired |>
  ungroup() |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(
    cols      = c(initial_score, follow_up_score),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score"   = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_cpss_i <- cpss_i_paired_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_cpss_i
```

### Self version

```{r q315-s-positive-outcome}
#| echo: false
#| warning: false
#| message: false
cpss_s_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)
```

```{r q315-s-summary-stats}
#| echo: false
#| warning: false
#| message: false
cpss_s_paired_long <- cpss_s_paired |>
  ungroup() |>
  filter(!is.na(client_system_id), !is.na(initial_score), !is.na(follow_up_score)) |>
  select(client_system_id, initial_score, follow_up_score) |>
  pivot_longer(
    cols      = c(initial_score, follow_up_score),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score"   = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

summary_stats_cpss_s <- cpss_s_paired_long |>
  group_by(score_type) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  )

summary_stats_cpss_s
```
