---
title: "Annual Report Data Analysis - Question 3.02"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---
# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

## 3.01: Was there meaningful change in intervention scores as assessed by validated outcomes tools, or were scores above clinical cut-offs at discharge?

**Inclusion criteria:** 1. must have initial time point data (i.e., be enrolled in an intervention), 2. must have a valid assessment collected at both intake and discharge time points, and 3. limited to tools with a clinical cut-off. 


```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(gt)
library(gtsummary)
library(openxlsx)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
```
