---
title: "Annual Report Data Analysis - Question 2.01"
author: "Cristin Young and Tristan Burgess"
date: '2025-12-10'
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

This notebook answers Question 2.01: Did successful transition rates out
of EBP/CDEP services differ by location? and Question 2.02: Did
successful transition rates out of EBP/CDEP services differ by
subpopulations?

#Read in data

```{r}
library(tidyverse)
library(janitor)
library(gt)
library(clipr)

all_data <- read_csv("/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
#all_data <- read_csv("/Users/tristan/CWS Dropbox/Tristan Burgess/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv", show_col_types = FALSE)
```

# Question 2.01: Did successful transition rates out of EBP/CDEP services differ by location?

THREE inclusion criteria: 1. must have initial time point data (i.e., be
enrolled in an intervention), 2. must have a valid reason for discharge
selected at time of discharge (i.e., excludes those with "no reason
selected"), 3. client must be attached to a specific grantee location

##Make a table of participation and completion rates by location

```{r}
data201 <- all_data %>% 
  filter(!is.na(location) & 
           !is.na(pre_score) & 
           !is.na(post_score) & 
           !is.na(reason_for_discharge))

data201$positive_outcomes <- factor(data201$positive_outcomes, levels = c("No", "Yes"))
```

## % reported Mutually agreed upon cessation of Tx/successful completion at Discharge by location

```{r}
perc_successful_table <- data201 %>% 
  group_by(location) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100 * successful / total, 1)
  )
              
perc_successful_table %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

openxlsx::write.xlsx(perc_successful_table,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.01/perc_successful_table.xlsx")
```

## % of those who met Tx goals (% successful completion + improved score) by location"

mutually agreed cessation of tx = 3782

clinically referred out = 63

```{r}
table(data201$reason_for_discharge)

data201_txgoals <- data201 %>% 
  filter(reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)" | reason_for_discharge == "Clinically referred out",
         positive_outcomes == "Yes")

perc_successful_table_pos_outcome <- data201 %>% 
  group_by(location) %>% 
  summarise(
    total = n(),
    successful_discharge = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success_discharge = round(100 * successful_discharge / total, 1),
    successful_outcome = sum(positive_outcomes %in% c("Yes"), na.rm = TRUE),
    perc_success_outcome = round(100 * successful_outcome / total, 1)
  )
              
perc_successful_table_pos_outcome %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

xlsx::write.xlsx(perc_successful_table,"/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.01/perc_successful_table_pos_outcome.xlsx")

#or
p_yes = function(x) round(x[2]/sum(x), 2) # 1 in this line was doing p_no!

table <- addmargins(xtabs(data=data201, ~location + positive_outcomes, subset = data201$reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)"), FUN = list(p_yes))

table %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

xlsx::write.xlsx(perc_successful_table,"/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.01/Mutually agreed cessation of treatment (Successful Completion).xlsx")

```

#Question 2.02: Did successful transition rates out of EBP/CDEP services
differ by subpopulations? look at data201 by demographics

##Race

```{r}
detailed_race <- data201 %>% 
  group_by(code_num) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

detailed_race %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

chisq.test(data201$positive_outcomes, data201$code_num)

model <- glm(positive_outcomes ~ factor(code_num), data = data201, family = "binomial")
summary(model)

openxlsx::write.xlsx(summary(model)$coefficients,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/detailed race (model).xlsx")
```

##Race summary Roll detailed race variables, based on first answer, up
into: - Asian - Black or African American - Don't Know - Hispanic or
Latino - Middle Eastern or North African - Native American/American
Indian or Alaska Native - Native Hawaiian or Pacific Islander - Not
Applicable - Prefer Not to Answer - White

```{r}
#make new column of only the first race recorded
data201 <- data201 %>% 
  separate(race, into = c("first_race", NA), sep = ",", extra = "drop", remove = FALSE)

summary_race <- data201 %>% 
  group_by(first_race) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

summary_race %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

openxlsx::write.xlsx(summary_race,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/summary_race.xlsx")

chisq.test(data201$positive_outcomes, data201$first_race)

model <- glm(positive_outcomes ~ first_race, data = data201, family = "binomial")
summary(model)
table(data201$first_race)

xlsx::write.xlsx(summary(model)$coefficients,"/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/summary race (model).xlsx")

```

##Ethnicity

```{r}
detailed_ethnicity <- data201 %>% 
  group_by(ethnicity) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

detailed_ethnicity %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

xlsx::write.xlsx(detailed_ethnicity,"/Users/cristin/Dropbox/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/detailed_ethnicity.xlsx")

chisq.test(data201$positive_outcomes, data201$ethnicity)
```

##Gender Identity

```{r}
gender_identity <- data201 %>% 
  group_by(gender_identity) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

gender_identity %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

openxlsx::write.xlsx(gender_identity,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/gender_identity.xlsx")

chisq.test(data201$gender_identity, data201$positive_outcomes)

model <- glm(positive_outcomes ~ gender_identity, data = data201, family = "binomial")

summary(model)

openxlsx::write.xlsx(summary(model)$coefficients,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/gender_identity(model).xlsx")
```

##Sexual orientation

```{r}
sex_orientation <- data201 %>% 
  group_by(sexual_orientation) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

sex_orientation %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

openxlsx::write.xlsx(sex_orientation,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/sex_orientation.xlsx")

chisq.test(data201$positive_outcomes, data201$sexual_orientation)

model <- glm(positive_outcomes ~ sexual_orientation, data = data201, family = "binomial")

summary(model)

```

##Insurance status

```{r}
insurance <- data201 %>% 
  group_by(insurance_status) %>% 
  summarise(
    total = n(),
    successful = sum(reason_for_discharge %in% c("Mutually agreed cessation of treatment (Successful Completion)","Clinically referred out"), na.rm = TRUE),
    perc_success = round(100* successful / total, 1)
  )

insurance %>% 
  knitr::kable() %>% 
  clipr::write_clip(allow_non_interactive = TRUE)

openxlsx::write.xlsx(insurance,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/insurance.xlsx")

chisq.test(data201$positive_outcomes, data201$insurance_status)

model <- glm(positive_outcomes ~ insurance_status, data = data201, family = "binomial")
summary(model)

openxlsx::write.xlsx(summary(model)$coefficients,"/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/Annual Report Development/Results Tables/Q2.02/insurance(model).xlsx")

```
