---
title: "Annual Report Data Analysis - Question 2.01 and 2.02"
author: "Cristin Young and Tristan Burgess"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    df_print: paged
  word_document: default
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 2. Did the EBP/CDEP Workstream increase access to services for targeted populations?

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(gt)
library(svglite)
library(haven)
library(sf)
library(tigris)
library(ggrepel)
library(openxlsx)
options(tigris_use_cache = TRUE)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
all_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv",
  show_col_types = FALSE
)

discharge_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv",
  show_col_types = FALSE
)

demographic_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv"
)

spss <- haven::read_spss(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Data/MASTER Grantee Files/Quant Team File/MasterGranteeSpreadsheet_forBHESQuantTeam_1-20-26_NEW!.sav"
) |>
  zap_missing() |>
  mutate(across(where(is.labelled), as.character)) |>
  mutate(across(where(is.character), ~ na_if(str_trim(.x), ""))) |>
  rename_with(~ gsub("^@", "", .x)) |>
  janitor::clean_names() |>
  rename_with(~ gsub("^x", "", .x))
```

```{r prep_data201}
#| echo: false
#| warning: false
#| message: false

# Base dataset for Q2.01/2.02: requires location, scores, and discharge reason
data201 <- all_data |>
  filter(
    !is.na(location),
    !is.na(pre_score),
    !is.na(post_score),
    !is.na(reason_for_discharge)
  ) |>
  mutate(positive_outcomes = factor(positive_outcomes, levels = c("No", "Yes")))

# Join discharge data to demographic data for region analysis
key <- c("client_system_id", "provider_system_id", "provider_id")

discharge_unique <- discharge_data |>
  select(all_of(key), provider, intervention_name) |>
  distinct(across(all_of(key)), .keep_all = TRUE)

demo_discharge <- left_join(demographic_data, discharge_unique, by = key)

demo_spss <- left_join(
  demo_discharge,
  spss,
  by = c("provider_id" = "2grantee_provider_id")
) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  rename(region = `8c_regions`)

# Join data201 to SPSS for region-level outcomes
data201_spss <- left_join(
  data201,
  spss,
  by = c("provider" = "1a_grantee_provider_name", "provider_id" = "2grantee_provider_id")
) |>
  distinct(client_system_id, .keep_all = TRUE) |>
  rename(region = `8c_regions`)
```

# Question 2.01: Did successful transition rates out of EBP/CDEP services differ by location?

THREE inclusion criteria: 1. must have initial time point data (i.e., be enrolled in an intervention), 2. must have a valid reason for discharge selected at time of discharge (i.e., excludes those with "no reason selected"), 3. client must be attached to a specific grantee location.

## % successful completion at discharge by location

```{r success_by_location}
#| echo: false
perc_successful_table <- data201 |>
  group_by(location) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * (successful / total), 1)
  )

perc_successful_table
```

## % successful completion at discharge by region

```{r success_by_region}
#| echo: false
region_success <- data201_spss |>
  group_by(region) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * (successful / total), 1)
  )

region_success
```

### Mapping by region

```{r map_by_region}
#| echo: false
#| warning: false
#| message: false
ca_counties <- counties("CA", cb = TRUE, class = "sf")

region_lookup <- tibble::tribble(
  ~county,           ~region,
  # Region 1
  "Butte",           1,
  "Colusa",          1,
  "El Dorado",       1,
  "Glenn",           1,
  "Lassen",          1,
  "Modoc",           1,
  "Nevada",          1,
  "Placer",          1,
  "Plumas",          1,
  "Sacramento",      1,
  "Shasta",          1,
  "Sierra",          1,
  "Siskiyou",        1,
  "Sutter",          1,
  "Tehama",          1,
  "Yolo",            1,
  "Yuba",            1,
  # Region 2
  "Del Norte",       2,
  "Humboldt",        2,
  "Lake",            2,
  "Mendocino",       2,
  "Napa",            2,
  "Sonoma",          2,
  "Trinity",         2,
  # Region 3
  "Alameda",         3,
  "Contra Costa",    3,
  "Marin",           3,
  "San Francisco",   3,
  "San Mateo",       3,
  "Santa Clara",     3,
  "Solano",          3,
  # Region 4
  "Alpine",          4,
  "Amador",          4,
  "Calaveras",       4,
  "Madera",          4,
  "Mariposa",        4,
  "Merced",          4,
  "Mono",            4,
  "San Joaquin",     4,
  "Stanislaus",      4,
  "Tuolumne",        4,
  # Region 5
  "Monterey",        5,
  "San Benito",      5,
  "San Luis Obispo", 5,
  "Santa Barbara",   5,
  "Santa Cruz",      5,
  "Ventura",         5,
  # Region 6
  "Fresno",          6,
  "Inyo",            6,
  "Kern",            6,
  "Kings",           6,
  "Tulare",          6,
  # Region 7
  "Riverside",       7,
  "San Bernardino",  7,
  # Region 8
  "Los Angeles",     8,
  # Region 9
  "Orange",          9,
  # Region 10
  "Imperial",        10,
  "San Diego",       10
)

ca_counties_region <- ca_counties |>
  mutate(county = str_remove(NAME, " County$")) |>
  left_join(region_lookup, by = "county")

ca_census_regions <- ca_counties_region |>
  group_by(region) |>
  summarise(geometry = sf::st_union(geometry), .groups = "drop") |>
  arrange(region)

# Confirm regional polygons look correct
ca_census_regions |>
  ggplot() +
  geom_sf(aes(fill = factor(region)), color = "white", linewidth = 0.3) +
  scale_fill_viridis_d(name = "Region") +
  theme_minimal() +
  labs(title = "California Census Regions (1–10)")

# Merge outcomes and client counts with region polygons
ca_census_regions <- ca_census_regions |>
  mutate(region = as.character(region))

region_success <- region_success |>
  mutate(region = as.character(region))

ca_regions_data <- ca_census_regions |>
  left_join(region_success, by = "region") |>
  mutate(label_color = if_else(perc_success < 90, "white", "black"))

clients_by_region <- demo_spss |>
  group_by(region) |>
  summarise(clients_served = n_distinct(client_system_id), .groups = "drop")

ca_regions_demo_data <- ca_regions_data |>
  left_join(clients_by_region, by = "region") |>
  mutate(label_color = if_else(clients_served < 2200, "white", "black"))

# Map: % successful outcomes by region
ggplot(ca_regions_data) +
  geom_sf(aes(fill = perc_success), color = "white", linewidth = 0.3) +
  geom_sf_text(
    aes(
      label = paste0("Region ", region, "\n", perc_success, "%"),
      color = label_color
    ),
    size = 3,
    lineheight = 0.9
  ) +
  scale_color_identity() +
  scale_fill_gradientn(
    colours = c("#003C71", "#8BD3DD", "#BBDB49"),
    name = "% successful",
    limits = c(87, 100)
  ) +
  labs(title = "% Successful Outcomes by Region", fill = "% success") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )

# Map: clients served by region
ggplot(ca_regions_demo_data) +
  geom_sf(aes(fill = clients_served), color = "white", linewidth = 0.3) +
  geom_sf_text(
    aes(
      label = paste0("Region ", region, "\n", scales::comma(clients_served)),
      color = label_color
    ),
    size = 3,
    lineheight = 0.9
  ) +
  scale_color_identity() +
  scale_fill_gradientn(
    colours = c("#003C71", "#8BD3DD", "#BBDB49"),
    name = "Clients served",
    limits = c(1200, 5500)
  ) +
  labs(title = "Clients Served by Region", fill = "Clients") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )

#ggsave("/Users/cristin/Downloads/map_clients.svg", width = 8, height = 5)
```

### Mapping by primary county

```{r map_by_county}
  geom_sf(aes(fill = factor(name)), color = "white", linewidth = 0.3) +
  scale_fill_viridis_d(name = "Region") +
  theme_minimal() +
  labs(title = "California Census Regions (1–10)")
#| echo: false
#| warning: false
#| message: false
ca_counties_clean <- counties("CA", cb = TRUE, class = "sf") |>
  janitor::clean_names()

county_success <- data201_spss |>
  rename(primary_county = `8a_primary_county`) |>
  group_by(primary_county) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

county_map_data <- ca_counties_clean |>
  left_join(county_success, by = c(name = "primary_county"))

x_push <- diff(st_bbox(county_map_data)[c("xmin", "xmax")]) * 0.45

# NOTE: `threshold` must be defined before running (e.g., threshold <- 75)
ggplot(county_map_data) +
  geom_sf(aes(fill = perc_success), color = "white", linewidth = 0.3) +
  geom_text_repel(
    data = county_map_data |> filter(perc_success < threshold),
    aes(
      geometry = geometry,
      label = paste0(name, "\n", perc_success, "%")
    ),
    color = "black",
    stat = "sf_coordinates",
    nudge_x = -x_push,
    direction = "y",
    hjust = 1,
    box.padding = 0.6,
    point.padding = 0.4,
    max.overlaps = Inf,
    min.segment.length = 0,
    segment.size = 0.4,
    size = 3.2
  ) +
  scale_fill_viridis_c(
    option = "C",
    direction = -1,
    name = "% successful",
    limits = c(50, 100)
  ) +
  labs(
    title = "% Successful Outcomes by Primary County",
    subtitle = "Primary counties with <75% successful completion labeled",
    fill = "% success"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  ) +
  coord_sf(clip = "off")
```

## % of those who met treatment goals (successful completion + improved score) by location

```{r tx_goals}
#| echo: false
table(data201$reason_for_discharge)

data201 |>
  filter(positive_outcomes == "Yes") |>
  tabyl(reason_for_discharge) |>
  adorn_pct_formatting(digits = 1) |>
  adorn_totals("row")

p_yes <- function(x) round(x[2] / sum(x), 2)

addmargins(
  xtabs(
    data = data201,
    ~ location + positive_outcomes,
    subset = data201$reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)"
  ),
  FUN = list(p_yes)
) |>
  knitr::kable()
```

# Question 2.02: Did successful transition rates out of EBP/CDEP services differ by subpopulations?

## Race

```{r race_detailed}
#| echo: false
detailed_race <- data201 |>
  group_by(code_num) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

detailed_race

chisq.test(data201$positive_outcomes, data201$code_num)
```

## Race summary

Roll detailed race variables up based on first answer into: Asian, Black or African American, Don't Know, Hispanic or Latino, Middle Eastern or North African, Native American/American Indian or Alaska Native, Native Hawaiian or Pacific Islander, Not Applicable, Prefer Not to Answer, White.

```{r race_summary}
#| echo: false
data201 <- data201 |>
  separate(race, into = c("first_race", NA), sep = ",", extra = "drop", remove = FALSE)

summary_race <- data201 |>
  group_by(first_race) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

summary_race

chisq.test(data201$positive_outcomes, data201$first_race)

model_race <- glm(positive_outcomes ~ first_race, data = data201, family = "binomial")
summary(model_race)
table(data201$first_race)

summary(model_race)
```

## Ethnicity

Codebook: `Chinese` = 1, `Asian Indian` = 2, `Filipino` = 3, `Vietnamese` = 4, `Korean` = 5, `Japanese` = 6, `African American` = 7, `Jamaican` = 8, `Haitian` = 9, `Nigerian` = 10, `Somali` = 11, `Mexican` = 12, `Puerto Rican` = 13, `Salvadoran` = 14, `Cuban` = 15, `Dominican` = 16, `Guatemalan` = 17, `Native Hawaiian` = 18, `Samoan` = 19, `Chamorro` = 20, `English` = 21, `German` = 22, `Irish` = 23, `Italian` = 24, `Polish` = 25, `Scottish` = 26, `2+ ethnicities selected` = 88

```{r ethnicity}
#| echo: false
detailed_ethnicity <- data201 |>
  group_by(ethn_num) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

detailed_ethnicity

chisq.test(data201$positive_outcomes, data201$ethnicity)
```

## Gender identity

```{r gender_identity}
#| echo: false
gender_identity <- data201 |>
  group_by(gender_identity) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

gender_identity 

chisq.test(data201$gender_identity, data201$positive_outcomes)

data201 <- data201 |>
  mutate(gender_identity = relevel(as.factor(gender_identity), ref = "Woman/girl"))

model_gender <- glm(positive_outcomes ~ gender_identity, data = data201, family = "binomial")

summary(model_gender)
```

## Sexual orientation

```{r sexual_orientation}
#| echo: false
sex_orientation <- data201 |>
  group_by(sexual_orientation) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

sex_orientation 

chisq.test(data201$positive_outcomes, data201$sexual_orientation)

model_sexori <- glm(positive_outcomes ~ sexual_orientation, data = data201, family = "binomial")

summary(model_sexori)
```

## Insurance status

```{r insurance}
#| echo: false
insurance <- data201 |>
  group_by(insurance_status) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

insurance 

chisq.test(data201$positive_outcomes, data201$insurance_status)

model_insurance <- glm(positive_outcomes ~ insurance_status, data = data201, family = "binomial")

summary(model_insurance)
```