---
title: "Annual Report Data Analysis - Question 3.01"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---
# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

## 3.01: Are there factors such as race, ethnicity, age, gender identity, sexual orientation, language barriers, etc. that are associated with differences in clinical and functional outcomes after the intervention?

**Inclusion criteria:** 1. must have initial time point data (i.e., be enrolled in an intervention), 2. must have a valid assessment collected at both intake and discharge time points, and 3. must have reported information in the demographic variables of interest.

### TO DO FOR JAKE

- [ ] add in more info about filtering/assumptions
- [ ] throw in codebooks for race/ethnicity
- [ ] write down variables used
- [ ] ask rikke for intervention type from SPSS file, care setting, 

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(gt)
library(gtsummary)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
all_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv",
  show_col_types = FALSE
)

sat_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"
combined_all <- read_csv(
  file.path(sat_dir, "combined_satisfaction_clean.csv"),
  show_col_types = FALSE
) 
```

```{r prep_data301}
#| echo: false
#| warning: false
#| message: false

# Base dataset: requires location and pre/post scores
data301 <- all_data |>
  filter(
    !is.na(location),
    !is.na(pre_score),
    !is.na(post_score)
  ) |>
  mutate(
    positive_outcomes = factor(positive_outcomes, levels = c("No", "Yes")),

    # Race: collapse to key groups; everything else is "Other"
    simple_race = case_when(
      is.na(cleaned_race) ~ NA_character_,
      cleaned_race %in% c(
        "Asian", "White", "Hispanic or Latino",
        "Black or African American",
        "Native American/American Indian or Alaska Native"
      ) ~ cleaned_race,
      TRUE ~ "Other"
    ),
    simple_race = relevel(factor(simple_race), ref = "White"),

    # Age/client type: treat "No Client Type Selected" as missing
    client_type_clean = if_else(
      client_type == "No Client Type Selected", NA_character_, client_type
    ),

    # Sexual orientation: consolidate non-responses and merge Lesbian/Gay/SGL
    sexual_orientation_clean = case_when(
      is.na(sexual_orientation) ~ "No Response",
      sexual_orientation %in% c(
        "Not Applicable", "Prefer Not To Answer",
        "Don't Know", "No Sexual Orientation Selected"
      ) ~ "No Response",
      sexual_orientation %in% c("Same Gender Loving", "Lesbian/Gay") ~ "Lesbian/Gay/SGL",
      TRUE ~ sexual_orientation
    ),
    sexual_orientation_clean = relevel(factor(sexual_orientation_clean), ref = "Straight"),

    # Gender identity: consolidate non-responses
    gender_identity_clean = case_when(
      is.na(gender_identity) ~ "No Response",
      gender_identity %in% c(
        "Not Applicable", "Prefer Not To Answer",
        "Don't Know", "No Gender Identify Selected"
      ) ~ "No Response",
      TRUE ~ gender_identity
    ),
    gender_identity_clean = relevel(factor(gender_identity_clean), ref = "Woman/girl")
  )

# For clients with multiple survey responses, keep only the most recent
combined_lang <- combined_all |>
  group_by(client_system_id) |>
  slice_max(date_completed, n = 1, with_ties = FALSE) |>
  ungroup()

data301 <- data301 |>
  left_join(select(combined_lang, client_system_id, lang), by = "client_system_id") |>
  mutate(lang = relevel(factor(lang), ref = "Yes"))
```

## Race model

```{r race_model}
#| echo: false
detailed_race <- data301 |>
  group_by(code_num) |>
  summarise(
    total      = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

chisq.test(data301$positive_outcomes, data301$code_num)
chisq.test(data301$positive_outcomes, data301$simple_race)

model_race <- glm(positive_outcomes ~ simple_race, data = data301, family = "binomial")
summary(model_race)

tab_race <- table(data301$simple_race, data301$positive_outcomes) |>
  addmargins() |>
  as.data.frame.matrix()

modtab_race <- tbl_regression(model_race, exponentiate = TRUE)
```

## Ethnicity model

```{r ethnicity_model}
#| echo: false

# Build a crosswalk: groups with n > 60 keep their label; all others become "Other";
# non-responses and administrative values are set to NA
ethnicity_xwalk <- data.frame(table(as.factor(data301$ethnicity))) |>
  rename(old = Var1) |>
  mutate(
    old = as.character(old),
    new = if_else(Freq > 60, old, "Other"),
    new = if_else(
      grepl("unknow|2019|dont|don't|report|prefer|2_text|refuse|specify",
            old, ignore.case = TRUE) | old == "na",
      NA_character_,
      new
    )
  )

data301 <- data301 |>
  mutate(
    simple_ethnicity = ethnicity_xwalk$new[match(ethnicity, ethnicity_xwalk$old)],
    simple_ethnicity  = relevel(as.factor(simple_ethnicity), ref = "Mexican")
  )

chisq.test(data301$positive_outcomes, data301$simple_ethnicity)

model_ethnicity <- glm(positive_outcomes ~ simple_ethnicity, data = data301, family = "binomial")
summary(model_ethnicity)

tab_ethnicity <- table(data301$simple_ethnicity, data301$positive_outcomes) |>
  addmargins() |>
  as.data.frame.matrix()

modtab_ethnicity <- tbl_regression(model_ethnicity, exponentiate = TRUE)
```

## Age model

```{r age_model}
#| echo: false
model_age <- glm(positive_outcomes ~ client_type_clean, data = data301, family = "binomial")
summary(model_age)

tab_age <- table(data301$client_type_clean, data301$positive_outcomes) |>
  addmargins() |>
  as.data.frame.matrix()

modtab_age <- tbl_regression(model_age, exponentiate = TRUE)
```

## Sexual orientation model

```{r sexual_orientation_model}
#| echo: false
chisq.test(data301$positive_outcomes, data301$sexual_orientation_clean)

model_sexori <- glm(positive_outcomes ~ sexual_orientation_clean, data = data301, family = "binomial")
summary(model_sexori)

tab_sexori <- table(data301$sexual_orientation_clean, data301$positive_outcomes) |>
  addmargins() |>
  as.data.frame.matrix()

modtab_sexori <- tbl_regression(model_sexori, exponentiate = TRUE)
```

## Gender identity model

```{r gender_identity_model}
#| echo: false
chisq.test(data301$gender_identity_clean, data301$positive_outcomes)

model_gender <- glm(positive_outcomes ~ gender_identity_clean, data = data301, family = "binomial")
summary(model_gender)

tab_gender <- table(data301$gender_identity_clean, data301$positive_outcomes) |>
  addmargins() |>
  as.data.frame.matrix()

modtab_gender <- tbl_regression(model_gender, exponentiate = TRUE)
```

## Preferred language model

```{r lang_model}
#| echo: false
summary(factor(data301$preferred_language))
summary(factor(data301$lang))

model_lang <- glm(positive_outcomes ~ lang, data = data301, family = "binomial")
summary(model_lang)

tab_lang <- table(data301$lang, data301$positive_outcomes) |>
  addmargins() |>
  as.data.frame.matrix()

modtab_lang <- tbl_regression(model_lang, exponentiate = TRUE)
```
