---
title: "Annual Report Data Analysis - Question 3.12"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

## 3.12 To what extent did parents/caregivers report a reduction in parent-related stress after receiving the intervention?

**Definition:** PSS item(s) measuring reduction in parent/caregiver-related stress after the intervention. Assessments showing improvement from pre- to post-time point for clients who discharged from an intervention during the designated time frame.

**Scoring:** If follow-up is LOWER than initial, that indicates a positive outcome. To calculate a total PSS score, items 1, 2, 5, 6, 7, 8, 17, and 18 should be reverse scored as follows: (1=5) (2=4) (3=3) (4=2) (5=1). The PSS score is then obtained by summing across all items. Higher scores indicate higher levels of perceived stress. If follow-up is LOWER than initial, that indicates a positive outcome.

-   1 = Strongly disagree

-   2 = Disagree

-   3 = Undecided

-   4 = Agree

-   5 = Strongly Agree

**Assumptions:** Filtered out dates before 2024-01-01 and after the date the code was run. Filtered data that's \>3 days from initial assessment and paired to the LAST follow-up. Filtered positive outcomes that had NAs before computing %. Total score = NA if 2+ items are missing; used LAST follow-up if 2+ assessments exist; filtered out NA's from follow-up date and positive outcome variable before calculating %.

**Range:** 18 - 90

**Inclusion Criteria:** 1. must have initial time point data (i.e., be enrolled in an intervention); 2. must have a valid assessment collected at both initial and discharge comparison time points

**Final analysis:** Regression - outcome = binary (4/5 vs other), predictors as for 3.01

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(ggridges)
library(scales)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

# Produced by data cleaning/pss_data_cleaning.qmd
pss_paired <- read_csv(
  file.path(assess_dir, "pss_paired_clean.csv"),
  show_col_types = FALSE
)
```

```{r pss_summary}
#| echo: false
#| warning: false
#| message: false
tab <- pss_paired |>
  filter(!is.na(positive_outcome)) |>
  tabyl(positive_outcome) |>
  adorn_pct_formatting(digits = 0)

tab
```

Plot it.

```{r pss_bar_plot}
#| echo: false
#| warning: false
#| message: false
pss_paired_long <- pss_paired |>
  ungroup() |>
  select(client_system_id, initial_score, follow_up_score) |>
  filter(
    !is.na(client_system_id),
    !is.na(initial_score),
    !is.na(follow_up_score)
  ) |>
  pivot_longer(
    cols      = c(initial_score, follow_up_score),
    names_to  = "score_type",
    values_to = "score_value"
  ) |>
  mutate(
    score_type = recode(score_type,
                        "initial_score"   = "Initial",
                        "follow_up_score" = "Follow-up"),
    score_type = factor(score_type, levels = c("Initial", "Follow-up"))
  )

ggplot(pss_paired_long, aes(x = score_value, fill = score_type)) +
  geom_bar(
    position = position_dodge2(preserve = "single", padding = 0.01),
    width    = 0.8,
    alpha    = 0.8
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Score",
    y     = "Count",
    title = "Distribution of Initial and Follow-up Scores, PSS"
  ) +
  theme_minimal() +
  theme(
    axis.title           = element_text(size = 9),
    axis.text            = element_text(size = 9),
    panel.grid.major     = element_blank(),
    panel.grid.minor.x   = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position      = "inside",
    legend.justification = c("left", "top"),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 8),
    plot.margin          = margin(5, 5, 5, 5)
  )
```

Ridgeline plot

```{r pss_ridgeline}
#| echo: false
#| warning: false
#| message: false
summary_stats_pss <- pss_paired_long |>
  group_by(score_type) |>
  filter(!is.na(score_value)) |>
  summarise(
    mean_score   = mean(score_value, na.rm = TRUE),
    median_score = median(score_value, na.rm = TRUE),
    sd_score     = sd(score_value, na.rm = TRUE),
    n            = sum(!is.na(score_value))
  ) |>
  mutate(y_text = if_else(score_type == "Follow-up", 1.08, 1.085))

summary_stats_pss

ggplot(pss_paired_long, aes(x = score_value, y = 1, fill = score_type)) +
  geom_density_ridges(
    aes(color = score_type),
    alpha          = 0.7,
    scale          = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("Initial" = "#8A8C8E", "Follow-up" = "#BBDB49"),
    breaks = c("Initial", "Follow-up"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "PSS Score",
    y     = NULL,
    title = "Distribution of Initial and Follow-up Scores, PSS (lower is better)"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x         = element_text(size = 9, hjust = 0.5),
    axis.text.y          = element_blank(),
    axis.text.x          = element_text(size = 9),
    panel.grid           = element_blank(),
    axis.ticks.y         = element_blank(),
    axis.line.x          = element_line(color = "lightgrey", linewidth = 0.5),
    legend.position      = "inside",
    legend.justification = c("left", "top"),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 8),
    plot.margin          = margin(5, 5, 5, 5),
    plot.title           = element_text(size = 11)
  ) +
  geom_segment(
    data     = summary_stats_pss,
    aes(x = median_score, y = 1, xend = median_score, yend = 1.07),
    color    = "grey30",
    linetype = "dotted"
  ) +
  geom_text(
    data = summary_stats_pss,
    aes(x     = median_score,
        y     = y_text,
        label = paste0(score_type, "\nmedian = ", round(median_score, 2))),
    vjust = 2,
    hjust = 0.5,
    size  = 2
  )
```
