---
title: "Annual Report Data Analysis - Question 4.01"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 4. What percentage of clients reported high satisfaction with services?

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(ggridges)
library(scales)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

# Produced by data cleaning/satisfaction_data_cleaning.qmd
sat_combined <- read_csv(
  file.path(assess_dir, "combined_satisfaction_clean.csv"),
  show_col_types = FALSE
) |>
  filter(
    !is.na(date_clean)
  ) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label)))
```

## 4.01 What percentage of clients reported receiving services in their preferred language?

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** This KPI represents the percent of clients that report services received in their preferred language on the satisfaction survey out of the total that completed the satisfaction survey by period(s).

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (<= "2025-12-31"). I filtered out NAs for both the quarter label and the question first before calculating %.

The following table shows the percent of clients who received services in their preferred language on the satisfaction survey by quarter:

```{r preferred_language}
#| echo: false
#| warning: false
#| message: false
client_lang_all <- sat_combined |>
  rename(lang_pref = were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response) |>
  filter(
    !is.na(quarter_label),
    !is.na(lang_pref)
  )

# By quarter
client_lang_all |>
  tabyl(quarter_label, lang_pref) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()

# By year
client_lang_all |>
  tabyl(year, lang_pref) |>
  adorn_totals("row") |>
  adorn_percentages("row") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns()

# Overall
client_lang_all |>
  tabyl(lang_pref) |>
  adorn_pct_formatting(digits = 1)
```
