---
title: "Final Report Data Analysis: Study Questions 1.01 and 1.02"
author: "Cristin Young"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-tools: false
execute:
  echo: false
  warning: false   # warning text can quote raw data values, so it stays off in every render
  message: false
params:
  shareable: true
  suppress_n: 60
---

```{r}
#| label: shareable-banner
#| echo: false
#| results: asis
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
if (exists("params") && isTRUE(as.logical(params$shareable))) {
  k0 <- if (!is.null(params$suppress_n)) as.integer(params$suppress_n) else 60L
  cat(sprintf(paste0(
    "\n\n::: {.callout-note}\n## Shareable render\n\n",
    "Counts from 1 to %1$s are shown as \"Suppr.\", along with any other cell, total, or percentage ",
    "that would let a suppressed count be recovered by subtraction. Zeros are also hidden in any row ",
    "or column with fewer than %2$s clients, and a row whose total is suppressed shows counts without ",
    "percentages. Raw category values that are not in ",
    "the codebook and were recorded for fewer than %2$s clients are masked, and data-quality counts below ",
    "%2$s are shown as \"<%2$s\". Render with `shareable: false` to see exact counts locally.\n:::\n\n"),
    k0 - 1L, k0))
}
```

# 1. Did the EBP/CDEP Workstream increase access to services for targeted populations?

## 1.01 Did improved access to services affect all subpopulations?

Cross-tab of study population by relevant demographic data

Total number of clients served and by demographics

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated timeframe, where "served" means that a client had a new intervention start date attached to them during the designated timeframe.

**Calculation:** Aggregated count of clients served by period with an intervention type and start date listed; intervention does not need to be completed; assumes dates before 2024 are correct and any rows with these dates have not been revised.

**Assumptions:** This is discharge data from EInsight, so clients can appear multiple times (multiple interventions, or the same intervention on multiple dates). Filtered out dates before 2024-01-01; dates were not otherwise cleaned.

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(tidyverse)
  library(knitr)
})
options(dplyr.summarise.inform = FALSE, knitr.kable.NA = "", scipen = 999)
```

```{r}
#| label: config
# Defaults are used when running chunks interactively; YAML params take precedence when rendering.
defaults <- list(
  path_discharge   = "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv",
  path_demographic = "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv",
  output_dir       = "output/sq101",
  start_date       = "2024-01-01",
  suppress_n       = 60,
  shareable        = FALSE
)
cfg <- modifyList(defaults, if (exists("params")) as.list(params) else list())

# Shareable render: counts of 1 to SUPPRESS_N - 1 are suppressed in every results table (with
# complementary suppression), raw values outside the codebook seen for fewer than SUPPRESS_N clients
# are masked, and QA counts of 1 to SUPPRESS_N - 1 display as "<SUPPRESS_N". The default (internal)
# render shows exact counts and raw values and must not be shared.
SUPPRESS_N <- as.integer(cfg$suppress_n)
SHAREABLE  <- isTRUE(as.logical(cfg$shareable))
RUN_MODE   <- if (SHAREABLE) "SHAREABLE" else "INTERNAL"
SUPPR      <- "Suppr."
MASKED     <- "Unrecognized value (masked)"
NOT_REPORTED <- "Not reported / missing"

# ---- Codebooks (label = code). The *_num columns and the "known value" lists both come from here. ----
INSURANCE_CODES <- c(
  "Private Insurance" = 1L, "Medi-Cal/Medi-Care" = 2L, "No insurance/Indigent" = 3L,
  "Other" = 4L, "Veterans Benefits" = 5L
)
SEXUAL_ORIENTATION_CODES <- c(
  "Asexual" = 1L, "Bisexual" = 2L, "Lesbian/Gay" = 3L, "Other" = 4L,
  "Pansexual" = 5L, "Same Gender Loving" = 6L, "Straight" = 7L
)
GENDER_IDENTITY_CODES <- c(
  "Genderqueer" = 1L, "Man/boy" = 2L, "Non-binary" = 3L, "Transgender man/boy" = 4L,
  "Transgender woman/girl" = 5L, "Two-spirit" = 6L, "Woman/girl" = 7L
)
LANGUAGE_CODES <- c("English" = 1L, "Spanish" = 2L, "Other" = 3L)
HOUSING_CODES  <- c("Housed" = 1L, "Unhoused" = 2L, "Other (i.e. treatment facility)" = 3L)

# Race and ethnicity arrive as numeric codes (code_num, ethn_num). Code 0 is a non-response.
RACE_LABELS <- c(
  "1" = "Asian", "2" = "Black or African American", "3" = "Hispanic or Latino",
  "4" = "Middle Eastern or North African", "5" = "Native American/American Indian or Alaska Native",
  "6" = "Native Hawaiian or Pacific Islander", "7" = "White", "8" = "2+ races selected"
)
ETHN_LABELS <- c(
  "1" = "Chinese", "2" = "Asian Indian", "3" = "Filipino", "4" = "Vietnamese", "5" = "Korean",
  "6" = "Japanese", "7" = "African American", "8" = "Jamaican", "9" = "Haitian", "10" = "Nigerian",
  "11" = "Somali", "12" = "Mexican", "13" = "Puerto Rican", "14" = "Salvadoran", "15" = "Cuban",
  "16" = "Dominican", "17" = "Guatemalan", "18" = "Native Hawaiian", "19" = "Samoan", "20" = "Chamorro",
  "21" = "English", "22" = "German", "23" = "Irish", "24" = "Italian", "25" = "Polish",
  "26" = "Scottish", "88" = "2+ ethnicities selected"
)
CODE_0_LABEL <- "Not reported (code 0)"

# ---- Non-substantive responses, recoded to NA ----
NR_BASE <- c("Don't Know", "Prefer Not To Answer", "Not Applicable")
NR_INSURANCE          <- c(NR_BASE, "No Insurance Status Selected")
NR_SEXUAL_ORIENTATION <- c(NR_BASE, "No Sexual Orientation Selected")
# Both spellings: the earlier script used "Identify"; the interim report cites "No Gender Identity Selected".
NR_GENDER_IDENTITY    <- c(NR_BASE, "No Gender Identify Selected", "No Gender Identity Selected")

dir.create(cfg$output_dir, recursive = TRUE, showWarnings = FALSE)
```

```{r}
#| label: helpers
fmt_n <- function(n) formatC(as.numeric(n), format = "f", digits = 0, big.mark = ",")
fmt_n_qa <- function(n) {
  ifelse(SHAREABLE & !is.na(n) & n > 0 & n < SUPPRESS_N, paste0("<", SUPPRESS_N), fmt_n(n))
}
fmt_pct <- function(p, digits = 1) {
  ifelse(is.na(p) | is.nan(p), "", sprintf(paste0("%.", digits, "f%%"), 100 * p))
}

# Squish, send non-substantive responses and blanks to NA, and optionally collapse everything
# outside `keep` into `other`. Missing values stay missing (they are not collapsed into `other`).
clean_category <- function(x, non_response = NR_BASE, keep = NULL, other = NULL) {
  x <- str_squish(as.character(x))
  x <- if_else(x %in% non_response | x == "", NA_character_, x)
  if (!is.null(keep)) x <- if_else(is.na(x) | x %in% keep, x, other)
  x
}

code_to_label <- function(code, labels) {
  code_chr <- str_squish(as.character(code))
  case_when(
    is.na(code_chr) | code_chr == "" ~ NA_character_,
    code_chr == "0"                  ~ CODE_0_LABEL,
    code_chr %in% names(labels)      ~ unname(labels[code_chr]),
    TRUE                             ~ paste0("Code ", code_chr, " (not in codebook)")
  )
}

# Shareable render only: a value outside `known` that fewer than SUPPRESS_N distinct clients carry is
# the most likely to be free text, so it is replaced with MASKED. Counting clients (not rows) keeps a
# single client's repeated rows from pushing a free-text value over the threshold.
mask_rare_values <- function(x, id, known = character()) {
  if (!SHAREABLE) return(x)
  unknown <- !is.na(x) & !(x %in% known)
  if (!any(unknown)) return(x)
  n_clients <- tapply(id[unknown], x[unknown], n_distinct)
  rare <- names(n_clients)[n_clients < SUPPRESS_N]
  if_else(x %in% rare, MASKED, x)
}

# ---- Primary and complementary suppression ----
# `x` holds every published or derivable number in a table (cells and totals). Each element of `lines`
# indexes entries tied by one additive constraint, with the total LAST (a row's cells, then its total).
# Counts of 1 to k - 1 are suppressed, then more entries are suppressed until:
#   (a) no line has exactly one suppressed entry, which subtraction would recover;
#   (b) in a line whose total is shown, the suppressed cells add up to at least k, so their combined
#       count (total minus the shown cells) is not itself a small count; and
#   (c) no suppressed count below k, and no sum of two suppressed counts that is below k, can be
#       solved for exactly from all of the table's constraints taken together (checked by matrix rank).
# Extra suppressions go to the entry with the lowest `weight` (by default, the smallest count).
# An entry with weight Inf is never chosen, e.g. because another table already shows it.
pick_entry <- function(cand, weight) {
  cand <- cand[is.finite(weight[cand])]
  if (length(cand) == 0L) NA_integer_ else cand[which.min(weight[cand])]
}

recoverable_entries <- function(x, lines, s, k) {
  sup <- which(s)
  if (length(sup) == 0L) return(integer())
  A <- do.call(rbind, lapply(lines, function(ln) {
    row <- numeric(length(x))
    row[ln] <- c(rep(1, length(ln) - 1L), -1)
    row[sup]
  }))
  A <- A[rowSums(A != 0) > 0, , drop = FALSE]
  if (nrow(A) == 0L) return(integer())
  r <- qr(A)$rank
  solvable <- function(v) qr(rbind(A, v))$rank == r
  unit <- function(j) as.numeric(seq_along(sup) == j)
  small <- which(x[sup] < k)
  hits <- sup[small[vapply(small, function(j) solvable(unit(j)), logical(1))]]
  if (length(hits) == 0L && length(small) > 1L) {
    for (pr in combn(small, 2, simplify = FALSE)) {
      if (sum(x[sup[pr]]) < k && solvable(unit(pr[1]) + unit(pr[2]))) {
        hits <- sup[pr]
        break
      }
    }
  }
  hits
}

suppress_lines <- function(x, lines, k = SUPPRESS_N, init = NULL, weight = x) {
  s <- !is.na(x) & x > 0 & x < k
  if (!is.null(init)) s <- s | init
  open_entries <- function(idx) idx[!s[idx] & x[idx] > 0]
  repeat {
    j <- NA_integer_
    for (ln in lines) {
      tot <- ln[length(ln)]
      cells <- ln[-length(ln)]
      small_sum <- !s[tot] && any(s[cells]) && sum(x[cells][s[cells]]) < k
      if (sum(s[ln]) == 1L || small_sum) {
        j <- pick_entry(open_entries(ln), weight)
        if (is.na(j)) j <- pick_entry(open_entries(seq_along(x)), weight)
        if (!is.na(j)) break
      }
    }
    if (is.na(j)) {
      rec <- recoverable_entries(x, lines, s, k)
      if (length(rec) > 0L) {
        near <- unique(unlist(lines[vapply(lines, function(ln) any(rec %in% ln), logical(1))]))
        j <- pick_entry(open_entries(near), weight)
        if (is.na(j)) j <- pick_entry(open_entries(seq_along(x)), weight)
      }
    }
    if (is.na(j)) break
    s[j] <- TRUE
  }
  s
}

# One-way frequency table in the interim report's layout: N, % of all, and % excluding N/As.
# `not_reported` lists labels (besides NA) that count as non-substantive. The result carries the
# categories it shows and hides, so a cross-tab of the same variable can match them (see `like`).
tab_one_way <- function(x, label = "Category", order = character(), not_reported = character(),
                        show_valid = TRUE, digits = 1) {
  if (length(x) == 0L) return(tibble(!!label := "No records"))
  x <- if_else(is.na(x), NOT_REPORTED, as.character(x))
  nr_labels <- c(not_reported, NOT_REPORTED)
  counts <- count(tibble(category = x), category)
  present <- counts$category
  lv <- c(intersect(setdiff(order, nr_labels), present),
          sort(setdiff(present, c(order, nr_labels))),
          intersect(nr_labels, present))
  counts <- counts[match(lv, counts$category), ]

  n <- counts$n
  is_sub <- !(counts$category %in% nr_labels)
  m <- length(n)
  i_sub <- m + 1L; i_nr <- m + 2L; i_tot <- m + 3L
  x_all <- c(n, sum(n[is_sub]), sum(n[!is_sub]), sum(n))
  lines <- list(c(which(is_sub), i_sub), c(which(!is_sub), i_nr), c(i_sub, i_nr, i_tot))
  s <- if (SHAREABLE) suppress_lines(x_all, lines) else rep(FALSE, length(x_all))
  sc <- s[seq_len(m)]

  body <- tibble(
    category = counts$category,
    N = if_else(sc, SUPPR, fmt_n(n)),
    `%` = if_else(sc | s[i_tot], SUPPR, fmt_pct(n / x_all[i_tot], digits)),
    `% (excluding N/As)` = case_when(
      !is_sub         ~ "N/A",
      sc | s[i_sub]   ~ SUPPR,
      TRUE            ~ fmt_pct(n / x_all[i_sub], digits)
    )
  )
  subtotal <- tibble(
    category = "Subtotal: substantive responses",
    N = if_else(s[i_sub], SUPPR, fmt_n(x_all[i_sub])),
    `%` = if_else(s[i_sub] | s[i_tot], SUPPR, fmt_pct(x_all[i_sub] / x_all[i_tot], digits)),
    `% (excluding N/As)` = if_else(s[i_sub], SUPPR, "100.0%")
  )
  total <- tibble(category = "Total", N = if_else(s[i_tot], SUPPR, fmt_n(x_all[i_tot])),
                  `%` = if_else(s[i_tot], SUPPR, "100.0%"), `% (excluding N/As)` = "")

  out <- if (show_valid) {
    bind_rows(body[is_sub, ], subtotal, body[!is_sub, ], total)
  } else {
    bind_rows(body, total) |> select(-`% (excluding N/As)`)
  }
  out <- rename(out, !!label := category)
  attr(out, "hidden") <- counts$category[sc]
  attr(out, "shown")  <- counts$category[!sc]
  out
}

# Two-way table with row percentages and counts, e.g. "45% (1,234)", plus a Total row and an N column.
# Every row and column (including the Total row and N column) is a suppression line. Extra
# suppressions go to inner cells first and to a row total only as a last resort. A row whose total is
# hidden shows counts without percentages, because its percentages would reveal that total.
# `like`: an earlier table of the same column variable (one-way or two-way). Its hidden categories
# start hidden in this table's Total row, and categories it shows are never hidden here, so the two
# tables cannot be combined to recover a suppressed count. When `like` is given, the grand total is
# also treated as shown elsewhere and is never used as a complementary suppression.
tab_two_way <- function(rows, cols, row_label = "Group", col_order = character(), digits = 0,
                        like = NULL) {
  if (length(rows) == 0L) return(tibble(!!row_label := "No records"))
  row_lv <- as.character(sort(unique(rows[!is.na(rows)])))
  rows <- if_else(is.na(rows), NOT_REPORTED, as.character(rows))
  cols <- if_else(is.na(cols), NOT_REPORTED, as.character(cols))
  if (any(rows == NOT_REPORTED)) row_lv <- c(row_lv, NOT_REPORTED)
  present <- unique(cols)
  col_lv <- c(intersect(setdiff(col_order, NOT_REPORTED), present),
              sort(setdiff(present, c(col_order, NOT_REPORTED))),
              intersect(NOT_REPORTED, present))

  m <- unclass(table(factor(rows, row_lv), factor(cols, col_lv)))
  R <- nrow(m); C <- ncol(m)
  ext <- rbind(cbind(m, rowSums(m)), c(colSums(m), sum(m)))
  idx <- matrix(seq_along(ext), nrow(ext))
  lines <- c(lapply(seq_len(nrow(ext)), function(i) idx[i, ]),
             lapply(seq_len(ncol(ext)), function(j) idx[, j]))
  x <- as.vector(ext)

  s <- rep(FALSE, length(x))
  if (SHAREABLE) {
    weight <- x
    weight[idx[seq_len(R), C + 1]] <- x[idx[seq_len(R), C + 1]] + 1e9   # row totals: last resort
    total_row <- idx[R + 1, seq_len(C)]
    if (!is.null(like)) {
      s[total_row[col_lv %in% attr(like, "hidden")]] <- TRUE
      weight[total_row[col_lv %in% attr(like, "shown")]] <- Inf
      weight[idx[R + 1, C + 1]] <- Inf
    }
    s <- suppress_lines(x, lines, init = s, weight = weight)
  }
  s_mat <- matrix(s, nrow(ext))
  shown_s <- s_mat
  if (SHAREABLE) {
    # In a row or column whose total is below k, zeros are hidden too: otherwise they would show
    # which categories the small group falls into (e.g., a row that is zero everywhere but one cell).
    small_row <- ext[, C + 1] > 0 & ext[, C + 1] < SUPPRESS_N
    small_col <- ext[R + 1, ] > 0 & ext[R + 1, ] < SUPPRESS_N
    shown_s <- s_mat | (ext == 0 & (outer(small_row, rep(TRUE, C + 1)) | outer(rep(TRUE, R + 1), small_col)))
  }

  pct <- ext / ext[, C + 1]
  cell <- matrix(paste0(fmt_pct(pct, digits), " (", fmt_n(ext), ")"), nrow(ext))
  no_pct <- s_mat[, C + 1]                      # rows (including Total) whose total is hidden
  cell[no_pct, ] <- matrix(fmt_n(ext[no_pct, , drop = FALSE]), sum(no_pct))
  cell[shown_s] <- SUPPR
  cell[, C + 1] <- ifelse(s_mat[, C + 1], SUPPR, fmt_n(ext[, C + 1]))
  colnames(cell) <- c(col_lv, "N")
  out <- as_tibble(cell) |> mutate(!!row_label := c(row_lv, "Total"), .before = 1)
  attr(out, "hidden") <- union(attr(like, "hidden"), col_lv[s_mat[R + 1, seq_len(C)]])
  attr(out, "shown")  <- col_lv[!s_mat[R + 1, seq_len(C)]]
  out
}

# Show a table and keep a copy for the CSV export at the end.
EXPORT <- list()
show_table <- function(tbl, id, caption = NULL) {
  EXPORT[[id]] <<- tbl
  kable(tbl, caption = caption, align = c("l", rep("r", ncol(tbl) - 1)))
}

# Unique clients (first row per client) among rows that pass the given filters.
clients_with <- function(...) {
  all_data_filtered |> filter(...) |> distinct(client_system_id, .keep_all = TRUE)
}
```

```{r}
#| label: load_data
discharge_data_filtered <- read_csv(cfg$path_discharge, show_col_types = FALSE, progress = FALSE)
all_data <- read_csv(cfg$path_demographic, show_col_types = FALSE, progress = FALSE)

qa_flow <- tibble(step = "Discharge file: rows read", n = nrow(discharge_data_filtered))
qa_flow <- bind_rows(qa_flow, tibble(step = "Demographic file: rows read", n = nrow(all_data)))
```

```{r}
#| label: prep_data
# Add year and apply base filters to discharge data
discharge_data_filtered <- discharge_data_filtered |>
  mutate(intervention_start_year = year(intervention_start)) |>
  filter(
    !is.na(intervention_start),
    !is.na(intervention_name)
  )

# Clean all demographic variables in one pass
all_data <- all_data |>
  mutate(
    insurance_status_clean   = clean_category(insurance_status, NR_INSURANCE),
    insurance_num            = unname(INSURANCE_CODES[insurance_status_clean]),

    sexual_orientation_clean = clean_category(sexual_orientation, NR_SEXUAL_ORIENTATION),
    sexual_orientation_num   = unname(SEXUAL_ORIENTATION_CODES[sexual_orientation_clean]),

    gender_identity_clean    = clean_category(gender_identity, NR_GENDER_IDENTITY),
    gender_identity_num      = unname(GENDER_IDENTITY_CODES[gender_identity_clean]),

    preferred_language_clean = clean_category(preferred_language, NR_BASE, c("English", "Spanish"), "Other"),
    preferred_language_num   = unname(LANGUAGE_CODES[preferred_language_clean]),

    primary_language_clean   = clean_category(primary_language, NR_BASE, c("English", "Spanish"), "Other"),
    primary_language_num     = unname(LANGUAGE_CODES[primary_language_clean]),

    housing_status_clean     = clean_category(housing_status, NR_BASE, c("Housed", "Unhoused"),
                                              "Other (i.e. treatment facility)"),
    housing_status_num       = unname(HOUSING_CODES[housing_status_clean]),

    race_label = code_to_label(code_num, RACE_LABELS),
    ethn_label = code_to_label(ethn_num, ETHN_LABELS)
  )

# Base filtered dataset: start date onward, with intervention start and name
all_data_filtered <- all_data |>
  filter(
    intervention_start >= cfg$start_date,
    !is.na(intervention_start),
    !is.na(intervention_name)
  )

# Values each field is allowed to take without masking (client_type has no codebook, so any of its
# values seen for fewer than SUPPRESS_N clients is masked in a shareable render).
KNOWN_VALUES <- list(
  insurance_status_clean   = names(INSURANCE_CODES),
  sexual_orientation_clean = names(SEXUAL_ORIENTATION_CODES),
  gender_identity_clean    = names(GENDER_IDENTITY_CODES),
  preferred_language_clean = names(LANGUAGE_CODES),
  primary_language_clean   = names(LANGUAGE_CODES),
  housing_status_clean     = names(HOUSING_CODES),
  race_label               = c(unname(RACE_LABELS), CODE_0_LABEL),
  ethn_label               = c(unname(ETHN_LABELS), CODE_0_LABEL)
)
all_data_filtered <- all_data_filtered |>
  mutate(
    across(all_of(names(KNOWN_VALUES)),
           ~ mask_rare_values(.x, client_system_id, KNOWN_VALUES[[cur_column()]])),
    client_type = mask_rare_values(as.character(client_type), client_system_id)
  )

clients <- all_data_filtered |> distinct(client_system_id, .keep_all = TRUE)

qa_flow <- bind_rows(
  qa_flow,
  tibble(step = paste0("Discharge file: rows with intervention start and name, from ", cfg$start_date),
         n = sum(discharge_data_filtered$intervention_start >= cfg$start_date)),
  tibble(step = paste0("Demographic file: rows with intervention start and name, from ", cfg$start_date),
         n = nrow(all_data_filtered)),
  tibble(step = "Demographic file: unique clients (client_system_id)", n = nrow(clients))
)

RACE_ORDER <- c(unname(RACE_LABELS), CODE_0_LABEL)
ETHN_ORDER <- c(unname(ETHN_LABELS), CODE_0_LABEL)
```

### Total clients served

Counts are intervention records by start year (a client with more than one intervention is counted once per record).

```{r}
#| label: client_serve
discharge_data_filtered |>
  filter(intervention_start >= cfg$start_date) |>
  pull(intervention_start_year) |>
  tab_one_way("Intervention start year", show_valid = FALSE) |>
  rename(`Intervention records (N)` = N) |>
  show_table("total_records_by_year")
```

### Race

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by race.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and race is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
1. Asian
2. Black or African American
3. Hispanic or Latino
4. Middle Eastern or North African
5. Native American/American Indian or Alaska Native
6. Native Hawaiian or Pacific Islander
7. White
8. 2+ races selected

**Assumptions:** Clients had to have an intervention start, an intervention name, and have race characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). The "% (excluding N/As)" column excludes both code 0 and blank values.

```{r}
#| label: race
t1 <- tab_one_way(clients$race_label, "Race", RACE_ORDER, not_reported = CODE_0_LABEL)
show_table(t1, "race")

d <- clients_with(!is.na(client_type), !is.na(race_label))
tab_two_way(d$client_type, d$race_label, "Client type", RACE_ORDER, like = t1) |>
  show_table("race_by_client_type")
```

### Ethnicity

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by ethnicity.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and ethnicity is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
`Chinese` = 1, `Asian Indian` = 2, `Filipino` = 3, `Vietnamese` = 4, `Korean` = 5, `Japanese` = 6, `African American` = 7, `Jamaican` = 8, `Haitian` = 9, `Nigerian` = 10, `Somali` = 11, `Mexican` = 12, `Puerto Rican` = 13, `Salvadoran` = 14, `Cuban` = 15, `Dominican` = 16, `Guatemalan` = 17, `Native Hawaiian` = 18, `Samoan` = 19, `Chamorro` = 20, `English` = 21, `German` = 22, `Irish` = 23, `Italian` = 24, `Polish` = 25, `Scottish` = 26, `2+ ethnicities selected` = 88

**Assumptions:** Clients had to have an intervention start, an intervention name, and have race characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). The "% (excluding N/As)" column excludes both code 0 and blank values.

```{r}
#| label: ethnicity
tab_one_way(clients$ethn_label, "Ethnicity", ETHN_ORDER, not_reported = CODE_0_LABEL) |>
  show_table("ethnicity")
```

### Insurance status

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by insurance status.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and insurance is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
1. Private insurance
2. Medi-Cal/Medi-Care
3. No insurance/Indigent
4. Other
5. Veterans Benefits

**Assumptions:** Clients had to have an intervention start, an intervention name, and have insurance status characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| label: insurance
t1 <- tab_one_way(clients$insurance_status_clean, "Insurance status", names(INSURANCE_CODES))
show_table(t1, "insurance")

d <- clients_with(!is.na(client_type), !is.na(insurance_status_clean))
tab_two_way(d$client_type, d$insurance_status_clean, "Client type", names(INSURANCE_CODES), like = t1) |>
  show_table("insurance_by_client_type")
```

### Sexual orientation

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by sexual orientation.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and sexual orientation is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no sexual orientation selected
1. Asexual
2. Bisexual
3. Lesbian/Gay
4. Other
5. Pansexual
6. Same Gender Loving
7. Straight

**Assumptions:** Clients had to have an intervention start, an intervention name, and have sexual orientation characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| label: sexual_orientation
t1 <- tab_one_way(clients$sexual_orientation_clean, "Sexual orientation", names(SEXUAL_ORIENTATION_CODES))
show_table(t1, "sexual_orientation")

d <- clients_with(!is.na(client_type), !is.na(sexual_orientation_clean))
tab_two_way(d$client_type, d$sexual_orientation_clean, "Client type", names(SEXUAL_ORIENTATION_CODES),
            like = t1) |>
  show_table("sexual_orientation_by_client_type")
```

### Gender identity

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by gender identity.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and gender identity is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer
1. Genderqueer
2. Man/boy
3. Non-binary
4. Transgender man/boy
5. Transgender woman/girl
6. Two-spirit
7. Woman/girl

**Assumptions:** Clients had to have an intervention start, an intervention name, and have gender identity characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| label: gender_identity
tab_one_way(clients$gender_identity_clean, "Gender identity", names(GENDER_IDENTITY_CODES)) |>
  show_table("gender_identity")
```

### Preferred language

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by preferred language.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and preferred language is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no preferred language selected
1. English
2. Spanish
3. Other

**Assumptions:** Clients had to have an intervention start, an intervention name, and have preferred language characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). Any language other than English or Spanish is grouped into "Other".

```{r}
#| label: preferred_language
d <- clients_with(!is.na(preferred_language_clean))
t1 <- tab_two_way(d$intervention_start_quarter, d$preferred_language_clean, "Start quarter", names(LANGUAGE_CODES))
show_table(t1, "preferred_language_by_quarter")

d <- clients_with(!is.na(client_type), !is.na(preferred_language_clean))
tab_two_way(d$client_type, d$preferred_language_clean, "Client type", names(LANGUAGE_CODES), like = t1) |>
  show_table("preferred_language_by_client_type")
```

### Primary language

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by primary language.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and primary language is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no primary language selected
1. English
2. Spanish
3. Other

**Assumptions:** Clients had to have an intervention start, an intervention name, and have primary language characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients). Any language other than English or Spanish is grouped into "Other".

```{r}
#| label: primary_language
d <- clients_with(!is.na(primary_language_clean))
t1 <- tab_two_way(d$intervention_start_quarter, d$primary_language_clean, "Start quarter", names(LANGUAGE_CODES))
show_table(t1, "primary_language_by_quarter")

d <- clients_with(!is.na(client_type), !is.na(primary_language_clean))
tab_two_way(d$client_type, d$primary_language_clean, "Client type", names(LANGUAGE_CODES), like = t1) |>
  show_table("primary_language_by_client_type")
```

### Housing status

**Definition:** The number of new clients served by grantees within the primary evaluation in the designated time frame, where "served" means that a client had a new intervention, grouped by housing status.

**Calculation:** This KPI represents an ***aggregated count*** of clients served by quarter and uses client demographic data, where NA values are removed and housing status is recoded as follows:

**Codebook:**

0. NA, don't know, prefer not to answer, no housing status selected
1. Housed
2. Unhoused
3. Other (i.e. treatment facility)

**Assumptions:** Clients had to have an intervention start, an intervention name, and have housing status characterized (filtered out NAs). These results are distinct by client_system_id (no duplicate rows of clients).

```{r}
#| label: housing_status
d <- clients_with(!is.na(housing_status_clean))
t1 <- tab_two_way(d$intervention_start_quarter, d$housing_status_clean, "Start quarter", names(HOUSING_CODES))
show_table(t1, "housing_by_quarter")

d <- clients_with(!is.na(client_type), !is.na(housing_status_clean))
tab_two_way(d$client_type, d$housing_status_clean, "Client type", names(HOUSING_CODES), like = t1) |>
  show_table("housing_by_client_type")
```

## 1.02 How does access differ compared to previous access to services before EBP/CDEP funding? For each subpopulation?

Cross-tab of demographics and grant types: Total number of clients served and by demographic, and by whether grant was start up or expansion

```{r}
#| label: q102
# Placeholder. Once the start-up vs. expansion column is available, the same helpers apply, e.g.:
# d <- clients_with(!is.na(grant_type), !is.na(race_label))
# tab_two_way(d$grant_type, d$race_label, "Grant type", RACE_ORDER) |> show_table("race_by_grant_type")
```

# Data quality checks

These counts are for review, not reporting.

```{r}
#| label: qa
#| results: asis
qa_flow |>
  mutate(n = fmt_n_qa(n)) |>
  rename(Step = step, N = n) |>
  show_table("qa_sample_flow", caption = "Sample flow")

unrecognized <- map_dfr(names(KNOWN_VALUES), function(f) {
  tibble(field = f, value = all_data_filtered[[f]], id = all_data_filtered$client_system_id) |>
    filter(!is.na(value), !(value %in% KNOWN_VALUES[[f]])) |>
    summarise(clients = n_distinct(id), records = n(), .by = c(field, value))
})

if (nrow(unrecognized) == 0) {
  cat("Every non-missing value matched the codebook.\n")
} else {
  unrecognized |>
    arrange(field, desc(clients)) |>
    mutate(clients = fmt_n_qa(clients), records = fmt_n_qa(records)) |>
    rename(Field = field, `Value (not in codebook)` = value, Clients = clients, Records = records) |>
    show_table("qa_unrecognized_values",
               caption = "Values kept as their own category because no codebook entry matched them")
}
```

```{r}
#| label: export
export_long <- imap_dfr(EXPORT, function(tbl, id) {
  first <- names(tbl)[1]
  tbl |>
    mutate(across(everything(), as.character), row_order = row_number()) |>
    rename(row = all_of(first)) |>
    pivot_longer(-c(row, row_order), names_to = "column", values_to = "value") |>
    mutate(table = id, row_header = first, .before = 1)
}) |>
  mutate(render_mode = RUN_MODE, suppress_n = SUPPRESS_N, .before = 1)

export_path <- file.path(cfg$output_dir, paste0("sq101_tables_", RUN_MODE, ".csv"))
write_csv(export_long, export_path, na = "")
```
