---
title: "Annual Report Data Analysis - Question 3.13"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

## 3.13 To what extent did children/youth report reductions in overall distress after receiving the intervention?

**Definition:** This KPI represents the percent of assessments that show clients who have improvement from pre-post time points out of the total number of assessments from clients who discharge from the intervention and complete pre-post measures by period(s).

The YOQ total score reflects total distress in a client's life, with a high score reflecting increased distress related to experiencing a high number of symptoms, interpersonal difficulties, and decreased satisfaction and quality of life.

**Range:** 0 - 120

**Calculation:** Sum of 30 items (0-4 scale); invalid if 3+ items missing; total score only calculated if 28+ items completed; including those doing reliably better / total who took YOQ by quarter RCI = baseline score - follow-up score - RCI ≤ -10 doing reliably worse - RCI ≥ 10 doing reliably better

**Assumptions:** A positive outcome was if a client was doing reliably better as defined by the YOQ RCI. I used follow-up quarter as the date. I used the LAST follow-up date if there were \>1 assessments. I filtered out NA's from "reliably better"; Filtered out dates before 2024-01-01 and after the date the code was run. 

**Final analysis:** Regression - outcome = binary (positive or clincal cutoff met). Predictors as for 3.01

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(ggridges)
library(scales)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
assess_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

# Produced by data cleaning/yoq_data_cleaning.qmd
yoq_caregiver_paired <- read_csv(
  file.path(assess_dir, "yoq_caregiver_paired_clean.csv"),
  show_col_types = FALSE
) |>
  mutate(rci_col = as.factor(as.integer(rci > 10 | rci <= -10)))

yoq_self_paired <- read_csv(
  file.path(assess_dir, "yoq_self_paired_clean.csv"),
  show_col_types = FALSE
) |>
  mutate(rci_col = as.factor(as.integer(rci > 10 | rci <= -10)))
```

### Caregiver

```{r yoq_caregiver_summary}
#| echo: false
#| warning: false
#| message: false
yoq_caregiver_paired |>
  filter(!is.na(reliably_better), rci != 0) |>
  tabyl(reliably_better) |>
  adorn_pct_formatting(digits = 0)
```

Of those with meaningful change, the following did reliably better:

```{r yoq_caregiver_reliable_better}
#| echo: false
#| warning: false
#| message: false
yoq_caregiver_paired |>
  filter(!is.na(reliably_better), rci_col == 1) |>
  tabyl(reliably_better) |>
  adorn_pct_formatting(digits = 0)
```

```{r yoq_caregiver_plot}
#| echo: false
#| warning: false
#| message: false
yoq_caregiver_paired |>
  filter(rci != 0) |>
  ggplot(aes(rci)) +
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 1) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Reliable Change Index",
    y     = "",
    title = "Distribution of reliable change scores for YOQ Caregiver"
  ) +
  scale_fill_manual(
    "Reliable Change Index",
    labels = c("no meaningful change (n=77)", "meaningful change (n=114)"),
    values = c("#ABDDDE", "#046C9A")
  )
```

### Self

```{r yoq_self_summary}
#| echo: false
#| warning: false
#| message: false
yoq_self_paired |>
  filter(!is.na(reliably_better), rci != 0) |>
  tabyl(reliably_better) |>
  adorn_pct_formatting(digits = 0)
```

Of those with meaningful change, the following did reliably better:

```{r yoq_self_reliable_better}
#| echo: false
#| warning: false
#| message: false
yoq_self_paired |>
  filter(!is.na(reliably_better), rci_col == 1) |>
  tabyl(reliably_better) |>
  adorn_pct_formatting(digits = 0)
```

```{r yoq_self_plot}
#| echo: false
#| warning: false
#| message: false
yoq_self_paired |>
  filter(rci != 0) |>
  ggplot(aes(rci)) +
  geom_histogram(aes(fill = rci_col), color = "white", binwidth = 1) +
  theme_classic() +
  scale_x_continuous(breaks = seq(-25, 50, 5), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x     = "Reliable Change Index",
    y     = "",
    title = "Distribution of reliable change scores for YOQ Self"
  ) +
  scale_fill_manual(
    "Reliable Change Index",
    labels = c("no meaningful change (n=199)", "meaningful change (n=206)"),
    values = c("#ABDDDE", "#046C9A")
  )
```
