---
title: "Annual Report Data Analysis - Question 2.01 and 2.02"
author: "Cristin Young and Tristan Burgess"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    df_print: paged
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 2. Did the EBP/CDEP Workstream increase access to services for targeted populations?

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(gt)
library(svglite)
library(haven)
library(sf)
library(tigris)
library(ggrepel)
library(openxlsx)
options(tigris_use_cache = TRUE)
```

```{r load_data}
#| echo: false
#| warning: false
#| message: false
all_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_discharge_clean.csv",
  show_col_types = FALSE
)

# clean county column for joining 
all_data <- all_data |>
  mutate(
    county_clean = county |>
      str_trim() |>                          # trailing spaces (e.g., "Butte ")
      str_remove("\\s+County$") |>           # "Los Angeles County" -> "Los Angeles"
      str_to_title() |>                      # "SACRAMENTO", "los Angeles" -> title case
      recode(
        "La"           = "Los Angeles",      # "LA County" -> "La" after steps above
        "San Bernadino" = "San Bernardino",  # typo
        "San Deigo"     = "San Diego"        # typo
      ),
    # statewide entries aren't a county — no region to assign
    county_clean = if_else(county_clean %in% c("Ca", "California"),
                           "Statewide", county_clean)
  )

discharge_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/discharge_data_clean.csv",
  show_col_types = FALSE
)

demographic_data <- read_csv(
  "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data/demographic_data_clean.csv"
)

regions <- tibble::tribble(~Region, ~Counties,
                                                                                                                                         "1",
  "Butte, Colusa, El Dorado, Glenn, Lassen, Modoc, Nevada, Placer, Plumas, Sacramento, Shasta, Sierra, Siskiyou, Sutter, Tehama, Yolo, Yuba",
                                                                  "2",
                                                                  "Del Norte, Humboldt, Lake, Mendocino, Napa, Sonoma, Trinity",                                                                       "3",
                                                               "Alameda, Contra Costa, Marin, San Francisco, San Mateo, Santa Clara, Solano",
                                                                                                                                         "4",
                                              "Alpine, Amador, Calaveras, Madera, Mariposa, Merced, Mono, San Joaquin, Stanislaus, Tuolumne",
                                                                                                                                         "5",
                                                                 "Monterey, San Benito, San Luis Obispo, Santa Barbara, Santa Cruz, Ventura",
                                                                                                                                         "6",
                                                                                                         "Fresno, Inyo, Kern, Kings, Tulare",
                                                                                                                                         "7",
                                                                                                                 "Riverside, San Bernardino",
                                                                                                                                         "8",
                                                                                                                               "Los Angeles",
                                                                                                                                         "9",
                                                                                                                                    "Orange",
                                                                                                                                        "10",
                                                                                                                       "Imperial, San Diego", "11", "Statewide"
  )

# Unnest comma-separated counties into one row per county
county_region <- regions |>
  separate_rows(Counties, sep = ",\\s*") |>
  rename(county = Counties, region = Region)

# Join onto all_data
all_data <- all_data |>
  left_join(county_region, by = c("county_clean" = "county"))

# Check to see if any unmatched rows
all_data |>
  filter(is.na(region), !is.na(county))
```

```{r prep_data201}
#| echo: false
#| warning: false
#| message: false

# Base dataset for Q2.01/2.02: requires location, scores, and discharge reason
data201 <- all_data |>
  filter(
    !is.na(location),
    !is.na(pre_score),
    !is.na(post_score),
    !is.na(reason_for_discharge)
  ) |>
  mutate(positive_outcomes = factor(positive_outcomes, levels = c("No", "Yes")))
```

# Question 2.01: Did successful transition rates out of EBP/CDEP services differ by location?

THREE inclusion criteria: 1. must have initial time point data (i.e., be enrolled in an intervention), 2. must have a valid reason for discharge selected at time of discharge (i.e., excludes those with "no reason selected"), 3. client must be attached to a specific grantee location.

## % successful completion at discharge by location

```{r success_by_location}
#| echo: false
perc_successful_table <- data201 |>
  group_by(location) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * (successful / total), 1)
  )

perc_successful_table
```

## % successful completion at discharge by region

Note: this is by episode of care, not by distinct client 
```{r success_by_region}
#| echo: false
region_success <- data201 |>
  group_by(region) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * (successful / total), 1)
  )

region_success
```

### Mapping by region

```{r map_by_region}
#| echo: false
#| warning: false
#| message: false
ca_counties <- counties("CA", cb = TRUE, class = "sf")

ca_counties_region <- ca_counties |>
  mutate(county = str_remove(NAME, " County$")) |>
  left_join(county_region, by = "county")

ca_census_regions <- ca_counties_region |>
  group_by(region) |>
  summarise(geometry = sf::st_union(geometry), .groups = "drop") |>
  arrange(region)

# Confirm regional polygons look correct
ca_census_regions |>
  ggplot() +
  geom_sf(aes(fill = factor(region)), color = "white", linewidth = 0.3) +
  scale_fill_viridis_d(name = "Region") +
  theme_minimal() +
  labs(title = "California Census Regions (1–10)")

# Merge outcomes and client counts with region polygons
ca_census_regions <- ca_census_regions |>
  mutate(region = as.character(region))

region_success <- region_success |>
  mutate(region = as.character(region))

ca_regions_data <- ca_census_regions |>
  left_join(region_success, by = "region") |>
  mutate(label_color = if_else(perc_success < 84, "white", "black"))

clients_by_region <- data201 |>
  group_by(region) |>
  summarise(clients_served = n_distinct(client_system_id), .groups = "drop")

ca_regions_demo_data <- ca_regions_data |>
  left_join(clients_by_region, by = "region") |>
  mutate(label_color = if_else(clients_served < 1000, "white", "black"))

statewide <- data201 |>
  filter(!is.na(region), region == "11") |>
  summarise(
    total = n(),
    perc_success = round(100 * sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ), na.rm = TRUE
    ) / total, 1)
  )

statewide_clients <- data201 |>
  filter(!is.na(region), region == "11") |>
  summarise(clients_served = n_distinct(client_system_id))

# Map: % successful outcomes by region
ggplot(ca_regions_data) +
  geom_sf(aes(fill = perc_success), color = "white", linewidth = 0.3) +
  geom_sf_text(
    aes(
      label = paste0("Region ", region, "\n", perc_success, "%"),
      color = label_color
    ),
    size = 3,
    lineheight = 0.9
  ) +
  scale_color_identity() +
  scale_fill_gradientn(
    colours = c("#003C71", "#8BD3DD", "#BBDB49"),
    name = "% successful",
    limits = c(75, 100)
  ) +
  labs(title = "% Successful Outcomes by Region", subtitle = paste0("Statewide: ", statewide$perc_success, "%"), fill = "% success") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )

# Map: clients served by region
ggplot(ca_regions_demo_data) +
  geom_sf(aes(fill = clients_served), color = "white", linewidth = 0.3) +
  geom_sf_text(
    aes(
      label = paste0("Region ", region, "\n", scales::comma(clients_served)),
      color = label_color
    ),
    size = 3,
    lineheight = 0.9
  ) +
  scale_color_identity() +
  scale_fill_gradientn(
    colours = c("#003C71", "#8BD3DD", "#BBDB49"),
    name = "Clients served",
    limits = c(0, 3000)
  ) +
  labs(title = "Clients Served by Region", subtitle = paste0("Statewide: ", scales::comma(statewide_clients$clients_served), " clients"), fill = "Clients") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  ) 
# + 
#   annotate(
#   "text", x = -124, y = 33.5,
#   label = paste0("Statewide\n", statewide$perc_success, "%"),
#   fontface = "bold", size = 3
# )

#ggsave("/Users/cristin/Downloads/map_clients.svg", width = 8, height = 5)
```

### Mapping by primary county

```{r map_by_county}
#| echo: false
#| warning: false
#| message: false
ca_counties_clean <- counties("CA", cb = TRUE, class = "sf") |>
  janitor::clean_names()

county_success <- data201 |>
  group_by(county) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

county_map_data <- ca_counties_clean |>
  left_join(county_success, by = c(name = "county"))

x_push <- diff(st_bbox(county_map_data)[c("xmin", "xmax")]) * 0.45

threshold <- 75
ggplot(county_map_data) +
  geom_sf(aes(fill = perc_success), color = "white", linewidth = 0.3) +
  geom_text_repel(
    data = county_map_data |> filter(perc_success < threshold),
    aes(
      geometry = geometry,
      label = paste0(name, "\n", perc_success, "%")
    ),
    color = "black",
    stat = "sf_coordinates",
    nudge_x = -x_push,
    direction = "y",
    hjust = 1,
    box.padding = 0.6,
    point.padding = 0.4,
    max.overlaps = Inf,
    min.segment.length = 0,
    segment.size = 0.4,
    size = 3.2
  ) +
  scale_fill_viridis_c(
    option = "C",
    direction = -1,
    name = "% successful",
    limits = c(0, 100)
  ) +
  labs(
    title = "% Successful Outcomes by County",
    subtitle = "Counties with <75% successful completion labeled",
    fill = "% success"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  ) +
  coord_sf(clip = "off")
```

## % of those who met treatment goals (successful completion + improved score) by location

```{r tx_goals}
#| echo: false
table(data201$reason_for_discharge)

data201 |>
  filter(positive_outcomes == "Yes") |>
  tabyl(reason_for_discharge) |>
  adorn_pct_formatting(digits = 1) |>
  adorn_totals("row")

p_yes <- function(x) round(x[2] / sum(x), 2)

addmargins(
  xtabs(
    data = data201,
    ~ location + positive_outcomes,
    subset = data201$reason_for_discharge == "Mutually agreed cessation of treatment (Successful Completion)"
  ),
  FUN = list(p_yes)
) |>
  knitr::kable()
```

# Question 2.02: Did successful transition rates out of EBP/CDEP services differ by subpopulations?

## Race

```{r race_detailed}
#| echo: false
detailed_race <- data201 |>
  group_by(code_num) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

detailed_race

chisq.test(data201$positive_outcomes, data201$code_num)
```

## Race summary

Roll detailed race variables up based on first answer into: Asian, Black or African American, Don't Know, Hispanic or Latino, Middle Eastern or North African, Native American/American Indian or Alaska Native, Native Hawaiian or Pacific Islander, Not Applicable, Prefer Not to Answer, White.

```{r race_summary}
#| echo: false
data201 <- data201 |>
  separate(race, into = c("first_race", NA), sep = ",", extra = "drop", remove = FALSE)

summary_race <- data201 |>
  group_by(first_race) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

summary_race

chisq.test(data201$positive_outcomes, data201$first_race)

model_race <- glm(positive_outcomes ~ first_race, data = data201, family = "binomial")
summary(model_race)
table(data201$first_race)
```

## Ethnicity

Codebook: `Chinese` = 1, `Asian Indian` = 2, `Filipino` = 3, `Vietnamese` = 4, `Korean` = 5, `Japanese` = 6, `African American` = 7, `Jamaican` = 8, `Haitian` = 9, `Nigerian` = 10, `Somali` = 11, `Mexican` = 12, `Puerto Rican` = 13, `Salvadoran` = 14, `Cuban` = 15, `Dominican` = 16, `Guatemalan` = 17, `Native Hawaiian` = 18, `Samoan` = 19, `Chamorro` = 20, `English` = 21, `German` = 22, `Irish` = 23, `Italian` = 24, `Polish` = 25, `Scottish` = 26, `2+ ethnicities selected` = 88

```{r ethnicity}
#| echo: false
detailed_ethnicity <- data201 |>
  group_by(ethn_num) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

detailed_ethnicity

chisq.test(data201$positive_outcomes, data201$ethnicity)
```

## Gender identity

```{r gender_identity}
#| echo: false
gender_identity <- data201 |>
  group_by(gender_identity) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

gender_identity 

chisq.test(data201$gender_identity, data201$positive_outcomes)

data201 <- data201 |>
  mutate(gender_identity = relevel(as.factor(gender_identity), ref = "Woman/girl"))

model_gender <- glm(positive_outcomes ~ gender_identity, data = data201, family = "binomial")

summary(model_gender)
```

## Sexual orientation

```{r sexual_orientation}
#| echo: false
sex_orientation <- data201 |>
  group_by(sexual_orientation) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

sex_orientation 

chisq.test(data201$positive_outcomes, data201$sexual_orientation)

model_sexori <- glm(positive_outcomes ~ sexual_orientation, data = data201, family = "binomial")

summary(model_sexori)
```

## Insurance status

```{r insurance}
#| echo: false
insurance <- data201 |>
  group_by(insurance_status) |>
  summarise(
    total = n(),
    successful = sum(
      reason_for_discharge %in% c(
        "Mutually agreed cessation of treatment (Successful Completion)",
        "Clinically referred out"
      ),
      na.rm = TRUE
    ),
    perc_success = round(100 * successful / total, 1)
  )

insurance 

chisq.test(data201$positive_outcomes, data201$insurance_status)

model_insurance <- glm(positive_outcomes ~ insurance_status, data = data201, family = "binomial")

summary(model_insurance)
```