---
title: "Annual Report Data Analysis - Question 3.11_3.16"
author: "Tristan Burgess and Cristin Young"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# 3. Did EBP/CDEP services impact clinical and functional outcomes for clients who participated?

**Inclusion criteria:** 1. must have initial time point data (i.e., be
enrolled in an intervention); 2. must have a valid assessment collected
at both initial and discharge comparison time points

**Scoring:** An **increase** in score indicates a positive outcome.
There are no benchmarks or clinical cutoffs established for NOMS scoring
— if the follow-up assessment score is higher than the initial score, a
positive outcome is indicated. Valid assessment: Two studies (here and
here) support the exclusion of cases with a total of more than one
missing value on all SDQ items.

**Assumptions:** Filtered out dates before 2024-01-01 and after the date
the code was run. Filtered data that's \>3 days from initial assessment
and paired to the LAST follow-up. Filtered positive outcomes that had
NAs before computing %.

**TO DO**:

-   Display type: Table of average, median, SD, ranges at initial and
    discharge time points, percent of clients within 0-80 percentile,
    81-90 percentile, \>90 percentile for each of the four groups (P2-4;
    P4-10; P11-17; S11-17). Sankey diagram of those in each percentile
    range at initial and at discharge. Couple example codes on how to
    create this: here and here (I haven't tried either, I've only used
    this manual online creation tool).

-   Outcome tool: The Strengths and Difficulties Questionnaire (SDQ);
    four versions: P2-4; P4-10; P11-17; S11-17

-   Reporting indicators: Report average, median, SD, ranges at initial
    and discharge time points

-   Final report analysis: Report percent of clients within 0-80
    percentile, 81-90 percentile, \>90 percentile. " Regression -
    outcome = SDQ score, predictors as for 3.01.

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(janitor)
library(ggridges)
library(patchwork)
```

# SDQ scoring

The Strengths and Difficulties Questionnaire (SDQ) is a brief behavioral
screening questionnaire about 2-17 year-olds. It exists in several
versions to meet the needs of researchers, clinicians and educators.
Each version includes between one and three of the following
components: 1) Items on psychological attributes; 2) An impact
supplement; 3) Follow-up questions.

\
An **increase** in score indicates a negative outcome. Published
cut-points are available for both total score and sub-scales and a range
of normative data distributions are also available.

```{r load_data}
#| echo: false
#| warning: false
#| message: false

# Produced by data cleaning/sdq_data_cleaning.qmd
in_dir <- "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data"

sdq_P24_pairs   <- read_csv(file.path(in_dir, "sdq_p24_pairs_clean.csv"),   show_col_types = FALSE)
sdq_P410_pairs  <- read_csv(file.path(in_dir, "sdq_p410_pairs_clean.csv"),  show_col_types = FALSE)
sdq_P1117_pairs <- read_csv(file.path(in_dir, "sdq_p1117_pairs_clean.csv"), show_col_types = FALSE)
sdq_S1117_pairs <- read_csv(file.path(in_dir, "sdq_s1117_pairs_clean.csv"), show_col_types = FALSE)
```

The following data answers the follow study sub-questions:

-   **3.11: "**To what extent did parents/caregivers report reductions
    in disruptive behaviors after receiving the intervention?"

-   **3.16: "**To what extent did children/youth report improvements in
    coping/handling overall difficulties after the intervention?**"**

## Function to summarize SDQ paired data

```{r creating_summary_tables}
#| echo: false
#| warning: false
#| message: false
analyze_sdq <- function(paired, label = NULL) {

  message("---- Running analysis for: ", label, " ----")

  # Summary table: TOTAL by quarter
  tab_total <- paired |>
    filter(!is.na(positive_outcome_total)) |>
    tabyl(follow_up_quarter, positive_outcome_total) |>
    adorn_totals("row") |>
    adorn_percentages("row") |>
    adorn_pct_formatting(digits = 0) |>
    adorn_ns()

  # Summary table: EXTERNALIZING by quarter
  tab_ext <- paired |>
    filter(!is.na(positive_outcome_ext)) |>
    tabyl(follow_up_quarter, positive_outcome_ext) |>
    adorn_totals("row") |>
    adorn_percentages("row") |>
    adorn_pct_formatting(digits = 0) |>
    adorn_ns()

  # Summary table: TOTAL overall
  tab_total_overall <- paired |>
    filter(!is.na(positive_outcome_total)) |>
    tabyl(positive_outcome_total) |>
    adorn_pct_formatting(digits = 0)

  # Summary table: EXTERNALIZING overall
  tab_ext_overall <- paired |>
    filter(!is.na(positive_outcome_ext)) |>
    tabyl(positive_outcome_ext) |>
    adorn_pct_formatting(digits = 0)

  message("---- Total score outcomes by follow-up quarter ----")
  print(tab_total)
  message("---- Externalizing outcomes by follow-up quarter ----")
  print(tab_ext)
  message("---- Total score outcomes ----")
  print(tab_total_overall)
  message("---- Externalizing outcomes score ----")
  print(tab_ext_overall)

  list(
    paired                = paired,
    summary_total         = tab_total,
    summary_externalizing = tab_ext
  )
}
```

## Scoring the datasets

```{r running_for_each_df}
#| echo: false
#| warning: false
#| message: false
res_P24   <- analyze_sdq(sdq_P24_pairs,   label = "Parent SDQ (2–4)")
res_P24$summary_total
res_P24$summary_externalizing

res_P410  <- analyze_sdq(sdq_P410_pairs,  label = "Parent SDQ (4–10)")
res_P410$summary_total
res_P410$summary_externalizing

res_P1117 <- analyze_sdq(sdq_P1117_pairs, label = "Parent SDQ (11–17)")
res_P1117$summary_total
res_P1117$summary_externalizing

res_S1117 <- analyze_sdq(sdq_S1117_pairs, label = "Self SDQ (11–17)")
res_S1117$summary_total
res_S1117$summary_externalizing

sdq_list <- list(
  "SDQ Parent (2-4)"   = res_P24,
  "SDQ Parent (4-10)"  = res_P410,
  "SDQ Parent (11-17)" = res_P1117,
  "SDQ Self (11-17)"   = res_S1117
)

df_from_lists <- imap_dfr(sdq_list, ~ {
  .x[[1]] |>
    transmute(
      subgroup        = .y,
      initial_total   = initial_total,
      follow_up_total = follow_up_total,
      initial_ext     = initial_ext,
      follow_up_ext   = follow_up_ext
    )
})
```

```{r plotting_total}
#| echo: false
#| warning: false
#| message: false
sdq_subgroup_levels <- c(
  "SDQ Parent (2-4)", "SDQ Parent (4-10)", "SDQ Parent (11-17)", "SDQ Self (11-17)"
)

sdq_long_total <- df_from_lists |>
  pivot_longer(
    cols      = c(initial_total, follow_up_total),
    names_to  = "timepoint",
    values_to = "value"
  ) |>
  mutate(
    subgroup  = factor(subgroup,  levels = sdq_subgroup_levels),
    timepoint = factor(timepoint, levels = c("initial_total", "follow_up_total"))
  )

# Summary stats
summary_stats <- sdq_long_total |>
  group_by(subgroup, timepoint) |>
  summarise(
    mean_score   = mean(value, na.rm = TRUE),
    median_score = median(value, na.rm = TRUE),
    sd_score     = sd(value, na.rm = TRUE),
    n            = sum(!is.na(value))
  )

summary_stats
#clipr::write_clip(summary_stats)

# Mean difference table
mean_differences <- sdq_long_total |>
  group_by(subgroup, timepoint) |>
  summarise(mean_value = mean(value, na.rm = TRUE)) |>
  pivot_wider(names_from = timepoint, values_from = mean_value) |>
  mutate(difference = follow_up_total - initial_total)

gap_table <- ggplot(mean_differences, aes(x = 1, y = subgroup)) +
  geom_text(aes(label = round(difference, 2)), size = 3, hjust = 0) +
  theme_void() +
  labs(title = "Mean Diff.\n(Follow-up - Initial)\ndecrease indicates\npositive outcome") +
  theme(
    plot.title  = element_text(hjust = 0.5, size = 7, margin = margin(t = 8, b = -15)),
    axis.text.y = element_blank(),
    plot.margin = margin(t = 5, b = -5, l = 5, r = 5)
  ) +
  scale_y_discrete(limits = sdq_subgroup_levels)

ridge_plot <- ggplot(sdq_long_total, aes(x = value, y = subgroup, fill = timepoint)) +
  geom_density_ridges(
    aes(color = timepoint),
    alpha          = 0.7,
    scale          = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("initial_total" = "#8A8C8E", "follow_up_total" = "#BBDB49"),
    breaks = c("initial_total", "follow_up_total"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("initial_total" = "#8A8C8E", "follow_up_total" = "#BBDB49"),
    breaks = c("initial_total", "follow_up_total"),
    labels = c("Initial", "Follow-up")
  ) +
  labs(
    x     = "SDQ total score",
    y     = NULL,
    title = "Distribution of initial and follow-up total scores by subgroup, SDQ"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x      = element_text(size = 9, hjust = 0),
    axis.text.y       = element_text(size = 9, hjust = 1),
    axis.text.x       = element_text(size = 9),
    legend.position   = "top",
    legend.title      = element_blank(),
    legend.text       = element_text(size = 8),
    legend.box.margin = margin(b = -5),
    plot.margin       = margin(5, 5, 5, 5)
  )

(ridge_plot | gap_table) +
  plot_layout(widths = c(5, 1)) +
  plot_annotation(title = NULL, subtitle = NULL)
```

```{r plotting_ext}
#| echo: false
#| warning: false
#| message: false
sdq_long_ext <- df_from_lists |>
  pivot_longer(
    cols      = c(initial_ext, follow_up_ext),
    names_to  = "timepoint",
    values_to = "value"
  ) |>
  mutate(
    subgroup  = factor(subgroup,  levels = sdq_subgroup_levels),
    timepoint = factor(timepoint, levels = c("initial_ext", "follow_up_ext"))
  )

# Summary stats
summary_stats <- sdq_long_ext |>
  group_by(subgroup, timepoint) |>
  summarise(
    mean_score   = mean(value, na.rm = TRUE),
    median_score = median(value, na.rm = TRUE),
    sd_score     = sd(value, na.rm = TRUE),
    n            = sum(!is.na(value))
  )

summary_stats
#clipr::write_clip(summary_stats)

# Mean difference table
mean_differences <- sdq_long_ext |>
  group_by(subgroup, timepoint) |>
  summarise(mean_value = mean(value, na.rm = TRUE)) |>
  pivot_wider(names_from = timepoint, values_from = mean_value) |>
  mutate(difference = follow_up_ext - initial_ext)

gap_table <- ggplot(mean_differences, aes(x = 1, y = subgroup)) +
  geom_text(aes(label = round(difference, 2)), size = 3, hjust = 0) +
  theme_void() +
  labs(title = "Mean Diff.\n(Follow-up - Initial)\ndecrease indicates\npositive outcome") +
  theme(
    plot.title  = element_text(hjust = 0.5, size = 7, margin = margin(t = 8, b = -15)),
    axis.text.y = element_blank(),
    plot.margin = margin(t = 5, b = -5, l = 5, r = 5)
  ) +
  scale_y_discrete(limits = sdq_subgroup_levels)

ridge_plot <- ggplot(sdq_long_ext, aes(x = value, y = subgroup, fill = timepoint)) +
  geom_density_ridges(
    aes(color = timepoint),
    alpha          = 0.7,
    scale          = 1.2,
    rel_min_height = 0.01
  ) +
  scale_fill_manual(
    values = c("initial_ext" = "#8A8C8E", "follow_up_ext" = "#BBDB49"),
    breaks = c("initial_ext", "follow_up_ext"),
    labels = c("Initial", "Follow-up")
  ) +
  scale_color_manual(
    values = c("initial_ext" = "#8A8C8E", "follow_up_ext" = "#BBDB49"),
    breaks = c("initial_ext", "follow_up_ext"),
    labels = c("Initial", "Follow-up")
  ) +
  labs(
    x     = "SDQ externalizing score",
    y     = NULL,
    title = "Distribution of initial and follow-up externalizing scores by subgroup, SDQ"
  ) +
  theme_ridges(font_family = "Arial") +
  theme(
    axis.title.x      = element_text(size = 9, hjust = 0),
    axis.text.y       = element_text(size = 9, hjust = 1),
    axis.text.x       = element_text(size = 9),
    legend.position   = "top",
    legend.title      = element_blank(),
    legend.text       = element_text(size = 8),
    legend.box.margin = margin(b = -5),
    plot.margin       = margin(5, 5, 5, 5)
  )

(ridge_plot | gap_table) +
  plot_layout(widths = c(5, 1)) +
  plot_annotation(title = NULL, subtitle = NULL)
```

## Reductions in disruptive behaviors (SDQ)

### 2-4 (Provider)

**'Internalising' and 'externalising' scores**: The internalising score
ranges from 0 to 20 and is the sum of the emotional and peer
difficulties scales. The externalising score ranges from 0 to 20 and is
the sum of the conduct and hyperactivity scales. We will be using the
externalising score for this KPI.

**Calculation:** Hyperactivity scale (2, 10, 15, 21, 25) + conduct scale
(5, 7, 12, 18, 22) Cut off bands

## Improvements in coping and handling overall difficulties (SDQ)

### 2-4 (Provider)

**Scoring:** In most respects, the 2-4 year old version of the SDQ is
scored in an identical way to the 4-17 year old version of the SDQ, and
we therefore suggest that you start by reading those instructions. The
only differences are:

-   The wording of two items on the 'conduct' scales is markedly
    different in the questionnaire for 2-4 year olds compared to the
    questionnaire for 4-17 (but the scoring stays the same). One further
    hyperactivity item is softened slightly in the 2-4 year old
    questionnaire. These three items are highlighted in Table 1.

-   Large population-based surveys in the UK suggest slightly different
    population norms for 2-4 year olds compared to older children. The
    banding of SDQ scores for the 2-4 year old version is shown in Table
    3.

**Total difficulties score**: This is generated by summing scores from
all the scales except the prosocial scale. The resultant score ranges
from 0 to 40, and is counted as missing if one of the 4 component scores
is missing.

don't sum: 1, 4, 9, 17, 20

reverse code: 7, 21, 25, 11, 14

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered
out NA's from follow-up date and positive outcome variable before
calculating %.

#### Cut off bands

We used the SDQ distribution data to propose the following banding of
SDQ scores for 2-4 year olds. We have sought cut-points such that around
80% of children are 'close to average', 12% 'slightly raised', 4% 'high'
and 4% 'very high' (or, for the prosocial scale, 80% are close to
average, 12% 'slightly lowered', 4% 'low' and 4% 'very low').

Similar cut-points were observed for parent SDQs collected in a smaller
American sample (N=307) of 2-4 year olds, and for parent SDQs collected
in the English 'Mental Health of Children And Young People, 2017' study.

|   | Close to average | Slightly raised (/slightly lowered) | High (/Low) | Very high (/very low) |
|---------------|:-------------:|:-------------:|:-------------:|:-------------:|
| Parented completed SDQ total score | 0-12 | 13-15 | 16-18 | 19-40 |
| Teacher completed SDQ total score | 0-10 | 11-14 | 15-17 | 18-40 |

### 4-10 (Provider)

**Scoring:** The 25 items in the SDQ comprise 5 scales of 5 items each.
It is usually easiest to score all 5 scales first before working out the
total difficulties score. 'Somewhat True' is always scored as 1, but the
scoring of 'Not True' and 'Certainly True' varies with the item, as
shown below scale by scale. For each of the 5 scales the score can range
from 0 to 10 if all items were completed. These scores can be scaled up
pro-rata if at least 3 items were completed, e.g. a score of 4 based on
3 completed items can be scaled up to a score of 7 (6.67 rounded up) for
5 items.

**Total difficulties score**: This is generated by summing scores from
all the scales except the prosocial scale. The resultant score ranges
from 0 to 40, and is counted as missing of one of the 4 component scores
is missing.

don't sum: 1, 4, 9, 17, 20

reverse code: 7, 21, 25, 11, 14

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered
out NA's from follow-up date and positive outcome variable before
calculating %.

#### Cut off bands

|   | Close to average | Slightly raised (/slightly lowered) | High (/Low) | Very high (/very low) |
|---------------|:-------------:|:-------------:|:-------------:|:-------------:|
| Parented completed SDQ total score | 0-13 | 14-16 | 17-19 | 20-40 |
| Teacher completed SDQ total score | 0-11 | 12-15 | 16-18 | 19-40 |
| Self completed SDQ total score | 0-14 | 15-17 | 18-19 | 20-40 |

### 11-17 (Provider)

**Scoring:** The 25 items in the SDQ comprise 5 scales of 5 items each.
It is usually easiest to score all 5 scales first before working out the
total difficulties score. 'Somewhat True' is always scored as 1, but the
scoring of 'Not True' and 'Certainly True' varies with the item, as
shown below scale by scale. For each of the 5 scales the score can range
from 0 to 10 if all items were completed. These scores can be scaled up
pro-rata if at least 3 items were completed, e.g. a score of 4 based on
3 completed items can be scaled up to a score of 7 (6.67 rounded up) for
5 items.

**Total difficulties score**: This is generated by summing scores from
all the scales except the prosocial scale. The resultant score ranges
from 0 to 40, and is counted as missing of one of the 4 component scores
is missing.

don't sum: 1, 4, 9, 17, 20

reverse code: 7, 21, 25, 11, 14

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered
out NA's from follow-up date and positive outcome variable before
calculating %.

#### Cut off bands

|   | Close to average | Slightly raised (/slightly lowered) | High (/Low) | Very high (/very low) |
|---------------|:-------------:|:-------------:|:-------------:|:-------------:|
| Parented completed SDQ total score | 0-13 | 14-16 | 17-19 | 20-40 |
| Teacher completed SDQ total score | 0-11 | 12-15 | 16-18 | 19-40 |
| Self completed SDQ total score | 0-14 | 15-17 | 18-19 | 20-40 |

### 11-17 (Self)

**Scoring:** The 25 items in the SDQ comprise 5 scales of 5 items each.
It is usually easiest to score all 5 scales first before working out the
total difficulties score. 'Somewhat True' is always scored as 1, but the
scoring of 'Not True' and 'Certainly True' varies with the item, as
shown below scale by scale. For each of the 5 scales the score can range
from 0 to 10 if all items were completed. These scores can be scaled up
pro-rata if at least 3 items were completed, e.g. a score of 4 based on
3 completed items can be scaled up to a score of 7 (6.67 rounded up) for
5 items.

**Total difficulties score**: This is generated by summing scores from
all the scales except the prosocial scale. The resultant score ranges
from 0 to 40, and is counted as missing of one of the 4 component scores
is missing.

don't sum: 1, 4, 9, 17, 20

reverse code: 7, 21, 25, 11, 14

**Assumptions:** used LAST follow-up if 2+ assessments exist; filtered
out NA's from follow-up date and positive outcome variable before
calculating %.

#### Cut off bands

|   | Close to average | Slightly raised (/slightly lowered) | High (/Low) | Very high (/very low) |
|---------------|:-------------:|:-------------:|:-------------:|:-------------:|
| Parented completed SDQ total score | 0-13 | 14-16 | 17-19 | 20-40 |
| Teacher completed SDQ total score | 0-11 | 12-15 | 16-18 | 19-40 |
| Self completed SDQ total score | 0-14 | 15-17 | 18-19 | 20-40 |
