---
title: "Final Report Data Analysis: Study Question 4.01"
author: "Tristan Burgess and Cristin Young"
date: today
format:
  html:
    embed-resources: true
    toc: true
    code-tools: false
execute:
  echo: false
  warning: false   # warning text can quote raw data values, so it stays off in every render
  message: false
params:
  shareable: false
  suppress_n: 60
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

```{r}
#| label: shareable-banner
#| echo: false
#| results: asis
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
if (exists("params") && isTRUE(as.logical(params$shareable))) {
  k0 <- if (!is.null(params$suppress_n)) as.integer(params$suppress_n) else 60L
  cat(sprintf(paste0(
    "\n\n::: {.callout-note}\n## Shareable render\n\n",
    "Counts from 1 to %1$s are shown as \"Suppr.\", along with any other cell, total, or percentage ",
    "that would let a suppressed count be recovered by subtraction. Zeros are also hidden in any row ",
    "or column with fewer than %2$s surveys, and a row whose total is suppressed shows counts without ",
    "percentages. Response values that are not in the expected list and were given by fewer than ",
    "%2$s respondents are masked, and data-quality counts below %2$s are shown as \"<%2$s\". ",
    "Render with `shareable: false` to see exact counts locally.\n:::\n\n"),
    k0 - 1L, k0))
}
```

# 4. What percentage of clients reported high satisfaction with services?

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(tidyverse)
  library(knitr)
})
options(dplyr.summarise.inform = FALSE, knitr.kable.NA = "", scipen = 999)
```

```{r}
#| label: config
# Defaults are used when running chunks interactively; YAML params take precedence when rendering.
defaults <- list(
  assess_dir        = "/Users/cristin/CWS Dropbox/Cristin Young/CYBHI Project/Rebecca and Tristan and Cristin/EInsight Data",
  file_satisfaction = "combined_satisfaction_clean.csv",   # produced by data cleaning/satisfaction_data_cleaning.qmd
  output_dir        = "output/sq401",
  suppress_n        = 60,
  shareable         = FALSE
)
cfg <- modifyList(defaults, if (exists("params")) as.list(params) else list())

# Shareable render: counts of 1 to SUPPRESS_N - 1 are suppressed in every results table (with
# complementary suppression), response values outside KNOWN_LANG_RESPONSES given by fewer than
# SUPPRESS_N respondents are masked, and QA counts of 1 to SUPPRESS_N - 1 display as "<SUPPRESS_N".
# The default (internal) render shows exact counts and raw values and must not be shared.
SUPPRESS_N   <- as.integer(cfg$suppress_n)
SHAREABLE    <- isTRUE(as.logical(cfg$shareable))
RUN_MODE     <- if (SHAREABLE) "SHAREABLE" else "INTERNAL"
SUPPR        <- "Suppr."
MASKED       <- "Unrecognized value (masked)"
NOT_REPORTED <- "Not reported / missing"

LANG_COL <- "were_written_documents_and_or_the_services_you_received_provided_in_the_language_you_prefer_response"
# Expected answers to the preferred-language item. Other values are kept as their own category
# (as before); in a shareable render, any of them given by fewer than SUPPRESS_N respondents is masked.
KNOWN_LANG_RESPONSES <- c("Yes", "No", "S\u00ed", "Si")   # "S\u00ed" is "Sí"
ID_COL <- "client_system_id"   # counts respondents for masking; rows are used if the column is absent

dir.create(cfg$output_dir, recursive = TRUE, showWarnings = FALSE)
```

```{r}
#| label: helpers
fmt_n <- function(n) formatC(as.numeric(n), format = "f", digits = 0, big.mark = ",")
fmt_n_qa <- function(n) {
  ifelse(SHAREABLE & !is.na(n) & n > 0 & n < SUPPRESS_N, paste0("<", SUPPRESS_N), fmt_n(n))
}
fmt_pct <- function(p, digits = 1) {
  ifelse(is.na(p) | is.nan(p), "", sprintf(paste0("%.", digits, "f%%"), 100 * p))
}

# Squish, send non-substantive responses and blanks to NA, and optionally collapse everything
# outside `keep` into `other`. Missing values stay missing (they are not collapsed into `other`).
clean_category <- function(x, non_response = NR_BASE, keep = NULL, other = NULL) {
  x <- str_squish(as.character(x))
  x <- if_else(x %in% non_response | x == "", NA_character_, x)
  if (!is.null(keep)) x <- if_else(is.na(x) | x %in% keep, x, other)
  x
}

code_to_label <- function(code, labels) {
  code_chr <- str_squish(as.character(code))
  case_when(
    is.na(code_chr) | code_chr == "" ~ NA_character_,
    code_chr == "0"                  ~ CODE_0_LABEL,
    code_chr %in% names(labels)      ~ unname(labels[code_chr]),
    TRUE                             ~ paste0("Code ", code_chr, " (not in codebook)")
  )
}

# Shareable render only: a value outside `known` that fewer than SUPPRESS_N distinct clients carry is
# the most likely to be free text, so it is replaced with MASKED. Counting clients (not rows) keeps a
# single client's repeated rows from pushing a free-text value over the threshold.
mask_rare_values <- function(x, id, known = character()) {
  if (!SHAREABLE) return(x)
  unknown <- !is.na(x) & !(x %in% known)
  if (!any(unknown)) return(x)
  n_clients <- tapply(id[unknown], x[unknown], n_distinct)
  rare <- names(n_clients)[n_clients < SUPPRESS_N]
  if_else(x %in% rare, MASKED, x)
}

# ---- Primary and complementary suppression ----
# `x` holds every published or derivable number in a table (cells and totals). Each element of `lines`
# indexes entries tied by one additive constraint, with the total LAST (a row's cells, then its total).
# Counts of 1 to k - 1 are suppressed, then more entries are suppressed until:
#   (a) no line has exactly one suppressed entry, which subtraction would recover;
#   (b) in a line whose total is shown, the suppressed cells add up to at least k, so their combined
#       count (total minus the shown cells) is not itself a small count; and
#   (c) no suppressed count below k, and no sum of two suppressed counts that is below k, can be
#       solved for exactly from all of the table's constraints taken together (checked by matrix rank).
# Extra suppressions go to the entry with the lowest `weight` (by default, the smallest count).
# An entry with weight Inf is never chosen, e.g. because another table already shows it.
pick_entry <- function(cand, weight) {
  cand <- cand[is.finite(weight[cand])]
  if (length(cand) == 0L) NA_integer_ else cand[which.min(weight[cand])]
}

recoverable_entries <- function(x, lines, s, k) {
  sup <- which(s)
  if (length(sup) == 0L) return(integer())
  A <- do.call(rbind, lapply(lines, function(ln) {
    row <- numeric(length(x))
    row[ln] <- c(rep(1, length(ln) - 1L), -1)
    row[sup]
  }))
  A <- A[rowSums(A != 0) > 0, , drop = FALSE]
  if (nrow(A) == 0L) return(integer())
  r <- qr(A)$rank
  solvable <- function(v) qr(rbind(A, v))$rank == r
  unit <- function(j) as.numeric(seq_along(sup) == j)
  small <- which(x[sup] < k)
  hits <- sup[small[vapply(small, function(j) solvable(unit(j)), logical(1))]]
  if (length(hits) == 0L && length(small) > 1L) {
    for (pr in combn(small, 2, simplify = FALSE)) {
      if (sum(x[sup[pr]]) < k && solvable(unit(pr[1]) + unit(pr[2]))) {
        hits <- sup[pr]
        break
      }
    }
  }
  hits
}

suppress_lines <- function(x, lines, k = SUPPRESS_N, init = NULL, weight = x) {
  s <- !is.na(x) & x > 0 & x < k
  if (!is.null(init)) s <- s | init
  open_entries <- function(idx) idx[!s[idx] & x[idx] > 0]
  repeat {
    j <- NA_integer_
    for (ln in lines) {
      tot <- ln[length(ln)]
      cells <- ln[-length(ln)]
      small_sum <- !s[tot] && any(s[cells]) && sum(x[cells][s[cells]]) < k
      if (sum(s[ln]) == 1L || small_sum) {
        j <- pick_entry(open_entries(ln), weight)
        if (is.na(j)) j <- pick_entry(open_entries(seq_along(x)), weight)
        if (!is.na(j)) break
      }
    }
    if (is.na(j)) {
      rec <- recoverable_entries(x, lines, s, k)
      if (length(rec) > 0L) {
        near <- unique(unlist(lines[vapply(lines, function(ln) any(rec %in% ln), logical(1))]))
        j <- pick_entry(open_entries(near), weight)
        if (is.na(j)) j <- pick_entry(open_entries(seq_along(x)), weight)
      }
    }
    if (is.na(j)) break
    s[j] <- TRUE
  }
  s
}

# One-way frequency table in the interim report's layout: N, % of all, and % excluding N/As.
# `not_reported` lists labels (besides NA) that count as non-substantive. The result carries the
# categories it shows and hides, so a cross-tab of the same variable can match them (see `like`).
tab_one_way <- function(x, label = "Category", order = character(), not_reported = character(),
                        show_valid = TRUE, digits = 1) {
  if (length(x) == 0L) return(tibble(!!label := "No records"))
  x <- if_else(is.na(x), NOT_REPORTED, as.character(x))
  nr_labels <- c(not_reported, NOT_REPORTED)
  counts <- count(tibble(category = x), category)
  present <- counts$category
  lv <- c(intersect(setdiff(order, nr_labels), present),
          sort(setdiff(present, c(order, nr_labels))),
          intersect(nr_labels, present))
  counts <- counts[match(lv, counts$category), ]

  n <- counts$n
  is_sub <- !(counts$category %in% nr_labels)
  m <- length(n)
  i_sub <- m + 1L; i_nr <- m + 2L; i_tot <- m + 3L
  x_all <- c(n, sum(n[is_sub]), sum(n[!is_sub]), sum(n))
  lines <- list(c(which(is_sub), i_sub), c(which(!is_sub), i_nr), c(i_sub, i_nr, i_tot))
  s <- if (SHAREABLE) suppress_lines(x_all, lines) else rep(FALSE, length(x_all))
  sc <- s[seq_len(m)]

  body <- tibble(
    category = counts$category,
    N = if_else(sc, SUPPR, fmt_n(n)),
    `%` = if_else(sc | s[i_tot], SUPPR, fmt_pct(n / x_all[i_tot], digits)),
    `% (excluding N/As)` = case_when(
      !is_sub         ~ "N/A",
      sc | s[i_sub]   ~ SUPPR,
      TRUE            ~ fmt_pct(n / x_all[i_sub], digits)
    )
  )
  subtotal <- tibble(
    category = "Subtotal: substantive responses",
    N = if_else(s[i_sub], SUPPR, fmt_n(x_all[i_sub])),
    `%` = if_else(s[i_sub] | s[i_tot], SUPPR, fmt_pct(x_all[i_sub] / x_all[i_tot], digits)),
    `% (excluding N/As)` = if_else(s[i_sub], SUPPR, "100.0%")
  )
  total <- tibble(category = "Total", N = if_else(s[i_tot], SUPPR, fmt_n(x_all[i_tot])),
                  `%` = if_else(s[i_tot], SUPPR, "100.0%"), `% (excluding N/As)` = "")

  out <- if (show_valid) {
    bind_rows(body[is_sub, ], subtotal, body[!is_sub, ], total)
  } else {
    bind_rows(body, total) |> select(-`% (excluding N/As)`)
  }
  out <- rename(out, !!label := category)
  attr(out, "hidden") <- counts$category[sc]
  attr(out, "shown")  <- counts$category[!sc]
  out
}

# Two-way table with row percentages and counts, e.g. "45% (1,234)", plus a Total row and an N column.
# Every row and column (including the Total row and N column) is a suppression line. Extra
# suppressions go to inner cells first and to a row total only as a last resort. A row whose total is
# hidden shows counts without percentages, because its percentages would reveal that total.
# `like`: an earlier table of the same column variable (one-way or two-way). Its hidden categories
# start hidden in this table's Total row, and categories it shows are never hidden here, so the two
# tables cannot be combined to recover a suppressed count. When `like` is given, the grand total is
# also treated as shown elsewhere and is never used as a complementary suppression.
tab_two_way <- function(rows, cols, row_label = "Group", col_order = character(), digits = 0,
                        like = NULL) {
  if (length(rows) == 0L) return(tibble(!!row_label := "No records"))
  row_lv <- as.character(sort(unique(rows[!is.na(rows)])))
  rows <- if_else(is.na(rows), NOT_REPORTED, as.character(rows))
  cols <- if_else(is.na(cols), NOT_REPORTED, as.character(cols))
  if (any(rows == NOT_REPORTED)) row_lv <- c(row_lv, NOT_REPORTED)
  present <- unique(cols)
  col_lv <- c(intersect(setdiff(col_order, NOT_REPORTED), present),
              sort(setdiff(present, c(col_order, NOT_REPORTED))),
              intersect(NOT_REPORTED, present))

  m <- unclass(table(factor(rows, row_lv), factor(cols, col_lv)))
  R <- nrow(m); C <- ncol(m)
  ext <- rbind(cbind(m, rowSums(m)), c(colSums(m), sum(m)))
  idx <- matrix(seq_along(ext), nrow(ext))
  lines <- c(lapply(seq_len(nrow(ext)), function(i) idx[i, ]),
             lapply(seq_len(ncol(ext)), function(j) idx[, j]))
  x <- as.vector(ext)

  s <- rep(FALSE, length(x))
  if (SHAREABLE) {
    weight <- x
    weight[idx[seq_len(R), C + 1]] <- x[idx[seq_len(R), C + 1]] + 1e9   # row totals: last resort
    total_row <- idx[R + 1, seq_len(C)]
    if (!is.null(like)) {
      s[total_row[col_lv %in% attr(like, "hidden")]] <- TRUE
      weight[total_row[col_lv %in% attr(like, "shown")]] <- Inf
      weight[idx[R + 1, C + 1]] <- Inf
    }
    s <- suppress_lines(x, lines, init = s, weight = weight)
  }
  s_mat <- matrix(s, nrow(ext))
  shown_s <- s_mat
  if (SHAREABLE) {
    # In a row or column whose total is below k, zeros are hidden too: otherwise they would show
    # which categories the small group falls into (e.g., a row that is zero everywhere but one cell).
    small_row <- ext[, C + 1] > 0 & ext[, C + 1] < SUPPRESS_N
    small_col <- ext[R + 1, ] > 0 & ext[R + 1, ] < SUPPRESS_N
    shown_s <- s_mat | (ext == 0 & (outer(small_row, rep(TRUE, C + 1)) | outer(rep(TRUE, R + 1), small_col)))
  }

  pct <- ext / ext[, C + 1]
  cell <- matrix(paste0(fmt_pct(pct, digits), " (", fmt_n(ext), ")"), nrow(ext))
  no_pct <- s_mat[, C + 1]                      # rows (including Total) whose total is hidden
  cell[no_pct, ] <- matrix(fmt_n(ext[no_pct, , drop = FALSE]), sum(no_pct))
  cell[shown_s] <- SUPPR
  cell[, C + 1] <- ifelse(s_mat[, C + 1], SUPPR, fmt_n(ext[, C + 1]))
  colnames(cell) <- c(col_lv, "N")
  out <- as_tibble(cell) |> mutate(!!row_label := c(row_lv, "Total"), .before = 1)
  attr(out, "hidden") <- union(attr(like, "hidden"), col_lv[s_mat[R + 1, seq_len(C)]])
  attr(out, "shown")  <- col_lv[!s_mat[R + 1, seq_len(C)]]
  out
}

# Show a table and keep a copy for the CSV export at the end.
EXPORT <- list()
show_table <- function(tbl, id, caption = NULL) {
  EXPORT[[id]] <<- tbl
  kable(tbl, caption = caption, align = c("l", rep("r", ncol(tbl) - 1)))
}
```

```{r}
#| label: load_data
sat_combined <- read_csv(file.path(cfg$assess_dir, cfg$file_satisfaction),
                         show_col_types = FALSE, progress = FALSE)
qa_flow <- tibble(step = "Satisfaction file: rows read", n = nrow(sat_combined))

sat_combined <- sat_combined |>
  filter(
    !is.na(date_clean)
  ) |>
  mutate(quarter_label = factor(quarter_label, levels = unique(quarter_label)))

qa_flow <- bind_rows(qa_flow, tibble(step = "Rows with a cleaned survey date", n = nrow(sat_combined)))
```

## 4.01 What percentage of clients reported receiving services in their preferred language?

**Definition:** Clients from the primary evaluation who indicate services received in their preferred language on the satisfaction survey divided by the total number of clients served by the grantees who completed a satisfaction survey after an intervention during the designated time frame.

**Calculation:** This KPI represents the percent of clients that report services received in their preferred language on the satisfaction survey out of the total that completed the satisfaction survey by period(s).

**Assumptions:** I cleaned dates with "002X" and "292X" and replaced them with "202X." I also filtered dates in the future (e.g, <= SysDate()). I filtered out NAs for both the quarter label and the question first before calculating %.

The following table shows the percent of clients who received services in their preferred language on the satisfaction survey by quarter:

```{r}
#| label: prep_lang
client_lang_all <- sat_combined |>
  rename(lang_pref = all_of(LANG_COL)) |>
  filter(
    !is.na(quarter_label),
    !is.na(lang_pref)
  ) |>
  mutate(lang_pref = as.character(lang_pref))

respondent_id <- if (ID_COL %in% names(client_lang_all)) {
  as.character(client_lang_all[[ID_COL]])
} else {
  as.character(seq_len(nrow(client_lang_all)))
}
client_lang_all$lang_pref <- mask_rare_values(client_lang_all$lang_pref, respondent_id, KNOWN_LANG_RESPONSES)

qa_flow <- bind_rows(
  qa_flow,
  tibble(step = "Rows with a quarter and a preferred-language response (analysis rows)",
         n = nrow(client_lang_all))
)

# Overall is built first so the quarter and year tables can match what it shows (see `like`).
t_overall <- tab_one_way(client_lang_all$lang_pref, "Response", KNOWN_LANG_RESPONSES, show_valid = FALSE)
t_quarter <- tab_two_way(client_lang_all$quarter_label, client_lang_all$lang_pref, "Quarter",
                         KNOWN_LANG_RESPONSES, like = t_overall)
t_year    <- tab_two_way(client_lang_all$year, client_lang_all$lang_pref, "Year",
                         KNOWN_LANG_RESPONSES, like = t_quarter)
```

### By quarter

```{r}
#| label: by_quarter
show_table(t_quarter, "by_quarter")
```

### By year

```{r}
#| label: by_year
show_table(t_year, "by_year")
```

### Overall

```{r}
#| label: overall
show_table(t_overall, "overall")
```

# Data quality checks

These counts are for review, not reporting.

```{r}
#| label: qa
qa_flow |>
  mutate(n = fmt_n_qa(n)) |>
  rename(Step = step, N = n) |>
  show_table("qa_sample_flow", caption = "Sample flow")

tibble(value = client_lang_all$lang_pref, id = respondent_id) |>
  summarise(respondents = n_distinct(id), rows = n(), .by = value) |>
  mutate(expected = if_else(value %in% KNOWN_LANG_RESPONSES, "Yes", "No")) |>
  arrange(expected, desc(rows)) |>
  mutate(respondents = fmt_n_qa(respondents), rows = fmt_n_qa(rows)) |>
  select(`Response value` = value, `In expected list` = expected, Respondents = respondents, Rows = rows) |>
  show_table("qa_response_values", caption = "Preferred-language response values")
```

```{r}
#| label: export
export_long <- imap_dfr(EXPORT, function(tbl, id) {
  first <- names(tbl)[1]
  tbl |>
    mutate(across(everything(), as.character), row_order = row_number()) |>
    rename(row = all_of(first)) |>
    pivot_longer(-c(row, row_order), names_to = "column", values_to = "value") |>
    mutate(table = id, row_header = first, .before = 1)
}) |>
  mutate(render_mode = RUN_MODE, suppress_n = SUPPRESS_N, .before = 1)

write_csv(export_long, file.path(cfg$output_dir, paste0("sq401_tables_", RUN_MODE, ".csv")), na = "")
```
